### 2019-03-25 research/paper/draft.tex
periodic backup

\documentclass[iop]{emulateapj}

\usepackage{tikz}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{graphicx}
\graphicspath{{./}{./figures/}}
\bibliographystyle{apj}

\usetikzlibrary{shapes.geometric, arrows}
\usetikzlibrary{fit}

\tikzstyle{hyper} = [circle, text centered, draw=black]
\tikzstyle{param} = [circle, text centered, draw=black]
\tikzstyle{data} = [circle, text centered, draw=black, line width=2pt]
\tikzstyle{arrow} = [thick,->,>=stealth]

\usepackage{soul}
\usepackage[title]{appendix}

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}[1]{\underline{#1}}
\newcommand{\chippr}{CHIPPR}
\newcommand{\nz}{$n(z)$}
\newcommand{\pz}{photo-$z$}
\newcommand{\pzpdf}{photo-$z$ PDF}
\newcommand{\lsst}{\textsc{lsst}}

\newcommand{\aim}[1]{\textcolor{blue}{#1}}

\begin{document}

\title{How to obtain the redshift distribution from probabilistic redshift estimates}

\author{Alex Malz\altaffilmark{1}}
\author{David W. Hogg\altaffilmark{1,2,3,4}}
\email{aimalz@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\begin{abstract}
A trustworthy estimate of the redshift distribution $n(z)$ is crucial for weak-lensing cosmology as we know it.
Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be inaccessible, making photometric redshifts (photo-$z$s) the next-best alternative.
The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
We introduce a mathematically rigorous probabilistic graphical model (PGM) of hierarchical inference of $n(z)$, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
The novel Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator produced by traditional procedures.
We conclude by propagating these effects to constraints in the space of cosmological parameters.
\end{abstract}

\keywords{cosmology: cosmological parameters --- galaxies: statistics --- gravitational lensing: weak --- methods: data analysis --- methods: statistical}

\maketitle

\section{Introduction}
\label{sec:introduction}

\aim{Brief literature review:\\
Major data releases including \pzpdf s\\
The cosmology papers by those missions that stack PDFs\\
Recent papers simultaneously inferring \pzpdf s and \nz}

\aim{Q: What is the status quo?\\
A: The future of galaxy surveys is \pz s, and \pzpdf s rightfully replace \pz s; however, usage of \pzpdf s is limited.}

\aim{Q: Why should we question existing methods?\\
A: Stacking is bad and only looks like it works because of assumptions that don't hold when the data is as bad as we anticipate.  Cite pedantic doc?}

\begin{itemize}
	\item
	Q: How can we improve the effectiveness of using \pzpdf s in inference?\\
	A: Hierarchical inference is the only self-consistent way; see flow chart for outline of how it works in practice.
	\item
	Q: How does the result of \chippr compare to established estimators in terms of the accuracy of \nz ?\\
	A: \chippr\ yields the best possible \nz , conditional on the accuracy of the \pzpdf s used.
	\item
	Q: How does the result of \chippr compare to established estimators in terms of the precision on cosmological constraints?
\end{itemize}

\section{Methods}
\label{sec:methods}

\textcolor{blue}{Q: How can we improve the effectiveness of using \pzpdf s in inference?\\
A: Hierarchical inference is the only self-consistent way.}

\subsection{Model generalities}
\label{sec:model}

We begin by reframing the redshift distribution \nz\ from a probabilistic perspective.
Here we define a redshift distribution \nz\ as the normalized probability density
\begin{equation}
\label{eqn:nz}
\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ \int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
\end{equation}
of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
We may without loss of generality impose a parametrization
\begin{equation}
\label{eqn:fz}
f_{\phi}(z)\ \equiv\ n(z)
\end{equation}
in terms of some parameter vector $\phi$.
We believe that galaxy redshifts are indeed drawn from \nz, making it a probability density over redshift; this fact can also be confirmed from dimensional analysis of Equation~\ref{eqn:nz}.
Therefore, it can be rewritten as
\begin{equation}
\label{eqn:pz}
z_{j}\ \sim\ \Pr(z \mid \phi)\ \equiv\ f_{\phi}(z),
\end{equation}
a probability density over redshift conditioned on the parameters defining \nz.
Note that $z_{j}$ does not depend on $z_{j'}$, a statement of the causal independence of galaxy redshifts from one another.

In addition to believing \nz\ is a PDF from which redshifts are drawn, we also believe that there is some PDF from which photometric data $d$, which may be any combination of fluxes, magnitudes, colors, and their observational errors, are drawn.
Such a PDF over data is a likelihood
\begin{equation}
\label{eqn:pzpdf}
d\ \sim\ \Pr(d \mid z)
\end{equation}
conditioned on redshift.
This assumption that the data are drawn from some function of the redshift forms the foundation upon which \pz\ estimation is based.
Note that galaxies may have different observational data $d$ despite sharing the same redshift and that the data $d_{j}$ of one galaxy is causally independent of the redshifts $z_{j'}$ and data $d_{j'}$ of other galaxies.

This description of the physical system corresponds to a forward model by which we actually believe photometry is generated:
\begin{enumerate}
	\item There exists a redshift distribution \nz\ with parameters $\phi$.
	\item Galaxy redshifts $\{z_{j}\}$ are independent draws from $\Pr(z 
\mid \phi)$.
	\item Galaxy photometry $d_{j}$ is drawn from the likelihoods $\Pr(d | 
z_{j})$.
\end{enumerate}
A forward model such as this corresponds to a probabilistic graphical model (PGM), represented by a directed acyclic graph (DAG) as in Figure~\ref{fig:pgm}.
\begin{figure}
	\begin{center}
		\includegraphics[width=0.25\textwidth]{fig/pgm.png}
		\caption{The PGM}
		\label{fig:pgm}
	\end{center}
\end{figure}
A DAG conveys the causal relationships between physical parameters and, like a Feynman diagram in the context of particle physics, is a shorthand for mathematical relationships between variables.

\subsection{Application to cosmology}
\label{sec:application}

The problem facing cosmologists is to determine the true parameters $\phi_{0}$ of \nz\ from observing the photometry $\{d_{j}\}$ of a large sample of galaxies $j$.
To self-consistently propagate the uncertainty in the redshifts, however, it is more appropriate to estimate the posterior $\Pr(\phi \mid \{d_{j}\})$ over all possible parameters $\phi$ (and thus potential redshift distributions \nz ) conditioned on all the observed data $\{d_{j}\}$ available in a generic catalog.

In order to use the DAG of Figure~\ref{fig:pgm} to derive an expression for $\Pr(\phi \mid \{d_{j}\})$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the implicit prior and the prior probability density.

When we constrain the redshift of a galaxy using its observed photometric data $d_{j}$, we are effectively estimating a posterior $\Pr(z \mid d_{j})$.
However, to do this, we must have a model for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
Such a relationship is defined in the space of probability density over redshift, so it must be able to be parameterized by the same functional form $f_{\phi}(z)$ as \nz .
It is thus natural to write it as $\Pr(z \mid \phi^{*})$, where $\phi^{*}$ is the parameters for this relationship under some generic \pzpdf\ method.
We call $\Pr(z \mid \phi^{*})$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher; for template-based methods where it can be chosen to be ``realistic,'' it may be appropriate to call it an \textit{interim prior}.
Because the implicit prior is unavoidable, the \pzpdf s reported by any method are really \textit{implicit-weighted posteriors} $\Pr(z \mid d, \phi^{*})$.

The prior probability density $\Pr(\phi)$ is a more familiar concept in astronomy; to progress, we will have to choose a prior probability density over all possible parameters $\phi$.
This prior need not be excessively proscriptive; for example, it may be chosen to enforce smoothness at physically motivated scales in redshift without imposing any particular region as over- or under-dense.

With these definitions, we obtain the desired expression for $\Pr(\phi \mid 
\{d_{j}\})$,
\begin{align}
\begin{split}
  \label{eqn:fullpost}
  \ln[\Pr(\phi \mid \{d_{j}\})] & \propto \ln[\Pr(\phi)] + \ln \left[\int dz 
\right.\\
  & \left. \exp \left[\sum_{j=1}^{J} \left(\ln[\Pr(z \mid d_{j}, \phi^{*})] 
\right. \right. \right.\\
  & \left. \left. \left. + \ln[\Pr(z \mid \phi)] - \ln[\Pr(z \mid \phi^{*})]\ 
\right)\ \right]\ \right] ,
\end{split}
\end{align}
which is the heart of \chippr.
The entire derivation of Equation~\ref{eqn:fullpost} is provided in Appendix~\ref{app:math}.

\subsection{Implementation}
\label{sec:implementation}

In this study, we compare the results of Equation~\ref{eqn:fullpost} to those of the two most common approaches to estimating \nz\ from a catalog of \pzpdf s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior probability (i.e. the modes of the \pzpdf s), and the stacked estimator
\begin{align}
  \label{eqn:stacked}
  \hat{n}(z) &\equiv \frac{1}{J}\ \sum_{j=1}^{J}\ \Pr(z \mid d_{j}, \phi^{*}) .
\end{align}

We perform this comparison on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed above and presented in detail in Appendix~\ref{app:data}.
The mock data emulates the three sources of error of highest concern to the \pz\ community: intrinsic scatter, catastrophic outliers, and systematic bias.
Figure~\ref{fig:mega_scatter} illustrates these three effects individually at ten times the tolerance of the upcoming Large Synoptic Survey Telescope (LSST).
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.3\textwidth]{fig/hivarsigmas_scatter.png}
    	\includegraphics[width=0.3\textwidth]{fig/uouthi_scatter.png}
    	\includegraphics[width=0.3\textwidth]{fig/neghivarbias_scatter.png}
		\caption{intrinsic scatter (left), uniform outliers (center), bias (right);
   		Not sure how best to broach the subject of nonuniform outliers and bias as model misspecification
    	}
		\label{fig:mega_scatter}
	\end{center}
\end{figure*}
Tests including all three effects at the tolerance levels of LSST are presented in Section~\ref{sec:results}.

\subsection{Limitations}
\label{sec:limitations}

Finally, we explicitly review the assumptions made by this approach.
\begin{itemize}
  \item prior $\Pr(\phi)$
  \item implicit prior $\phi^{*}$ known and shared among all PDFs
	\item \pzpdf s are accurate posteriors
	\end{itemize}

\section{Results}
\label{sec:results}

\textcolor{blue}{Q: How does the result of \chippr\ compare to established estimators in terms of the accuracy of $n(z)$?\\
A: \chippr\ yields the best possible $n(z)$, conditional on the accuracy of the photo-$z$ PDFs used.}

\subsection{LSST Requirements}
\label{sec:lsstdemo}

\textcolor{blue}{LSST specs: biased to 0.003, 10\% uniform outliers, z-dependent scatter of 0.05}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.45\textwidth]{fig/lsst_scatter.png}
		\includegraphics[width=0.45\textwidth]{fig/lsst_log_estimators.png}
		\caption{Using LSST numbers}
		\label{fig:lsstdemo}
	\end{center}
\end{figure*}

\subsection{Violations of the model}
\label{sec:violations}

\textcolor{blue}{mischaracterized interim prior, with idealized or realistic data?}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.45\textwidth]{fig/wrong_scatter.png}
		\includegraphics[width=0.45\textwidth]{fig/wrong_log_estimators.png}
		\caption{this is for wrong implicit prior, not bias, haven't done that one yet}
		\label{fig:mischaracterized}
	\end{center}
\end{figure*}

\section{Discussion}
\label{sec:discussion}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.99\textwidth]{fig/final_plot.png}
		\caption{propagation to cosmology}
		\label{fig:cornerplot}
	\end{center}
\end{figure*}

\section{Conclusion}
\label{sec:conclusion}

\section*{Appendices}

\begin{appendices}

\section{Mathematical derivation}
\label{app:math}

The derivation is most straightforward in the space of log probabilities.

According to Bayes' rule, the posterior distribution of interest is
\begin{align}
\label{eq:hyper_bayes}
\log[p(\phi | \{d_{i}\})] &\propto \log[p(\{d_{i}\} | \phi)] + \log[p(\phi)] ,
\end{align}
where $p(\phi)$ is the \textit{hyperprior} probability distribution over possible values of the hyperparameters $\phi$ defining the redshift distribution.  
The hyperprior is a choice made by those performing the inference and is not to be confused with the implicit prior parameters $\phi^{*}$, which are a specific instance of $\phi$.

Focusing on the likelihood term $p(\{d_{i}\} | \phi)$, we employ the power of the hierarchical model through \textit{marginalization} of the unobserved redshift parameters by integrating over them to obtain 
\begin{align}
\label{eq:marginalize}
\begin{split}
\log[p(\{d_{i}\} | \phi)] = & \log\left[\int\exp\left[\log[p(\{d_{i}\} | \{z_{i}\})]\right.\right. \\
& \left.\left.+ \log[p(\{z_{i}\} | \phi)]\right]\ d\{z_{i}\}\right] .
\end{split}
\end{align}
The redshifts $\{z_{i}\}$ are statistically independent of each other, and each galaxy's photometry depends solely on its own redshift, so we may invoke statistical independence to get
\begin{align}
\label{eq:independence_hyper}
\log[p(\{z_{i}\} |\phi)] &= \sum_{i}^{N} \log[p(z_{i} | \phi)]
\end{align}
and
\begin{align}
\label{eq:independence}
\log[p(\{d_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(d_{i} | z_{i})] .
\end{align}
Having chosen the functional form mapping $\phi$ to $n(z)$, the terms on the righthand side of Eq. \ref{eq:independence_hyper} are known.  
However, the terms on the righthand side of Eq. \ref{eq:independence} are likelihoods, and the photo-$z$ PDFs we have are implicit posteriors.

To transform the implicit posteriors we have into the likelihoods we need, we start with the vacuously true statement
\begin{align}
\begin{split}
\label{eq:unity}
\log[p(d_{i} | z_{i})] =& \log[p(d_{i} | z_{i})] + \log[p(z_{i} | d_{i}, \phi^{*})]\\
& - \log[p(z_{i} | \vec{d}_{i}, \phi^{*})] .
\end{split}
\end{align}
We expand the last term using Bayes' Rule to get
\begin{align}
\begin{split}
\label{eq:bayes}
\log[p(d_{i} | z_{i})] =& \log[p(d_{i} | z_{i})] + \log[p(z_{i} | d_{i}, \phi^{*})]\\
& - \log[p(d_{i} | z_{i}, \phi^{*})]\\
& + \log[p(d_{i} | \phi^{*})] - \log[p(z_{i} | \phi^{*})] .
\end{split}
\end{align}
Fig. \ref{fig:pgm} has no direct connection between $d_{i}$ and $\phi$, so we can break up $\log[p(d_{i} | z_{i}, \phi^{*})]$ as follows:
\begin{align}
\begin{split}
\label{eq:split}
\log[p(d_{i} | z_{i})] =& \log[p(d_{i} | z_{i})] + \log[p(z_{i} | d_{i}, \phi^{*})]\\
& - \log[p(d_{i} | z_{i})] - \log[p(d_{i} | \phi^{*})]\\
& + \log[p(d_{i} | \phi^{*})] - \log[p(z_{i} | \phi^{*})] .
\end{split}
\end{align}
Canceling terms, we find
\begin{align}
\label{eq:cancel}
\log[p(d_{i} | z_{i})] =& \log[p(z_{i} | d_{i}, \phi^{*})] - \log[p(z_{i} | \phi^{*})].
\end{align}

Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, \ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to arrive at 
\begin{align}
\begin{split}
\label{eq:final}
\log[p(\phi | \{d_{i}\})] \propto& \log[p(\phi)] + \log\left[\int \exp\left[\sum_{i}^{N} \right.\right.\\
& \left.\left.\left(\log[p(z_{i} | \phi)] - \log[p(z_{i} | \phi^{*})]\right.\right.\right.\\
& \left.\left.\left. + \log[p(z_{i} | d_{i}, \phi^{*})]
\right)\right] d\{z_{i}\}\right].
\end{split}
\end{align}
Thus, if we have a catalog of photo-$z$ implicit posteriors $\{p(z_{i} | d_{i}, \phi^{*})\}$, the interim prior parameters $\phi^{*}$, and a prior distribution $p(\phi)$, we may find the posterior $\log[p(\phi | \{d_{i}\})]$ of the redshift distribution $n(z)$.
Furthermore, since all galaxies share the same range of possible redshifts, the unknown $z_{i}$ may be shortened to $z$.

\section{Mock data generation}
\label{app:data}

\textul{Plots: flow chart of forward model}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.75\textwidth]{fig/flowchart.png}
		\caption{In appendix?}
		\label{fig:flowchart}
	\end{center}
\end{figure*}

\end{appendices}

\begin{acknowledgements}
AIM thanks Elisabeth Krause for assistance with the \texttt{CosmoLike} code, Phil Marshall for instruction in software carpentry, Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for advice on debugging, and Boris Leistedt for helpful comments provided in the preparation of this paper.
This work was completed under the generous nutritional support of the Center for Computational Astrophysics.
\end{acknowledgements}

\bibliographystyle{apj}

\end{document}

### 2019-03-25 chippr/__init__.py

@@ -1,17 +1,19 @@
-from defaults import *
+from __future__ import absolute_import
 
-from utils import *
-from stat_utils import *
-from plot_utils import *
+from .defaults import *
 
-from sim_utils import *
-from discrete import *
-from gauss import *
-from gamma import *
-from gmix import *
-from mvn import *
-from multi_dist import *
-from catalog import *
+from .utils import *
+from .stat_utils import *
+from .plot_utils import *
 
-from log_z_dens import *
-from log_z_dens_plots import *
+from .sim_utils import *
+from .discrete import *
+from .gauss import *
+from .gamma import *
+from .gmix import *
+from .mvn import *
+from .multi_dist import *
+from .catalog import *
+
+from .log_z_dens import *
+from .log_z_dens_plots import *

### 2019-03-25 chippr/plot_utils.py

@@ -10,6 +10,7 @@ from chippr import defaults as d
 
 # cmap = np.linspace(0., 1., d.plot_colors)
 # colors = [cm.viridis(i) for i in cmap]
+# colors = ['#14B03D', '#540D6E', '#EB811B', '#EE4266']#
 colors = np.array([(213, 94, 0), (0, 158, 115), (0, 114, 178), (86, 180, 233), (230, 159, 0), (204, 121, 167), (240, 228, 66)])/256.
 # ['Vermilion', 'Bluish green', 'Blue', 'Sky blue', 'Orange', 'Reddish purple', 'Yellow']
 

### 2019-03-25 research/scripts/basic_inf_script.py
periodic backup

@@ -176,7 +176,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    test_name = 'single_lsst'
+    test_name = 'single_trpr'
     all_tests = {}
     test_info = {}
     test_info['name'] = test_name

### 2019-03-25 research/scripts/basic_replot_script.py

@@ -100,7 +100,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    test_name = 'single_varbias'
+    test_name = 'single_trpr'
 
     all_tests = {}
     test_info = {}

### 2019-03-25 research/scripts/basic_sim_script.py

@@ -258,7 +258,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    test_name = 'single_lsst'
+    test_name = 'single_trpr'
     all_tests = {}
     test_info = {}
     test_info['name'] = test_name

### 2019-03-25 research/scripts/multi_replot_script.py

@@ -100,7 +100,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    name_file = 'which_inf_tests.txt'
+    name_file = 'which_plt_tests.txt'
 
     with open(name_file) as tests_to_run:
         all_tests = {}

### 2019-03-25 research/scripts/multi_sim_script.py

@@ -167,9 +167,9 @@ def make_catalog(given_key):
 
     test_dir = os.path.join(result_dir, test_name)
     test_info['dir'] = test_dir
-    if os.path.exists(test_dir):
-        shutil.rmtree(test_dir)
-    os.makedirs(test_dir)
+#    if os.path.exists(test_dir):
+#        shutil.rmtree(test_dir)
+#    os.makedirs(test_dir)
 
     param_file_name = test_name + '.txt'
     params = chippr.utils.ingest(param_file_name)

### 2019-03-25 research/scripts/cosmolike_inf_script.py
want to run the cosmolike tests on broiler

@@ -1,0 +1,203 @@
+def check_prob_params(params):
+    """
+    Sets parameter values pertaining to components of probability
+
+    Parameters
+    ----------
+    params: dict
+        dictionary containing key/value pairs for probability
+
+    Returns
+    -------
+    params: dict
+        dictionary containing key/value pairs for probability
+    """
+    if 'prior_mean' not in params:
+        params['prior_mean'] = 'interim'
+    else:
+        params['prior_mean'] = params['prior_mean'][0]
+    if 'no_prior' not in params:
+        params['no_prior'] = 0
+    else:
+        params['no_prior'] = int(params['no_prior'][0])
+    if 'no_data' not in params:
+        params['no_data'] = 0
+    else:
+        params['no_data'] = int(params['no_data'][0])
+    return params
+
+# def make_true_nz(test_name):
+#     """
+#     Function to create true redshift distribution to be shared among several
+#     test cases
+#
+#     Parameters
+#     ----------
+#     test_name: string
+#         name used to look up parameters for making true_nz
+#
+#     Returns
+#     -------
+#     true_nz: chippr.gmix object
+#         gaussian mixture probability distribution
+#
+#     Notes
+#     -----
+#     test_name is currently ignored but will soon be used to load parameters for making true_nz instead of hardcoded values.
+#     """
+#     true_amps = np.array([0.20, 0.35, 0.55])
+#     true_means = np.array([0.5, 0.2, 0.75])
+#     true_sigmas = np.array([0.4, 0.2, 0.1])
+#
+#     true_nz = chippr.gmix(true_amps, true_means, true_sigmas, limits=(0., 1.))
+#
+#     return true_nz
+
+def set_up_prior(data, params):
+    """
+    Function to create prior distribution from data
+
+    Parameters
+    ----------
+    data: dict
+        catalog dictionary containing bin endpoints, log interim prior, and log
+        interim posteriors
+    params: dict
+        dictionary of parameter values for creation of prior
+
+    Returns
+    -------
+    prior: chippr.mvn object
+        prior distribution as multivariate normal
+    """
+    zs = data['bin_ends']
[…]

### 2019-03-25 research/scripts/cosmolike_sim_script.py

@@ -1,0 +1,247 @@
+def check_extra_params(params):
+    """
+    Sets parameter values pertaining to true n(z)
+
+    Parameters
+    ----------
+    params: dict
+        dictionary containing key/value pairs for simulation
+
+    Returns
+    -------
+    params: dict
+        dictionary containing key/value pairs for simulation
+    """
+    # if 'smooth_truth' not in params:
+    #     params['smooth_truth'] = 0
+    # else:
+    #     params['smooth_truth'] = int(params['smooth_truth'][0])
+    #     # params['true_shape'] = int(params['true_shape'][0])
+    #     params['true_loc'] = float(params['true_loc'][0])
+    #     params['true_scale'] = 1./float(params['true_scale'][0])
+
+    if 'interim_prior' not in params:
+        params['interim_prior'] = 'flat'
+    else:
+        params['interim_prior'] = str(params['interim_prior'][0])
+    if 'wrong_prior' not in params:
+        params['wrong_prior'] = 0
+    else:
+        params['wrong_prior'] = int(params['wrong_prior'][0])
+        if 'bias_model' not in params:
+            params['bias_model'] = 'flat'
+        else:
+            params['bias_model'] = str(params['bias_model'][0])
+    if 'n_galaxies' not in params:
+        params['n_galaxies'] = 10**4
+    else:
+        params['n_galaxies'] = 10**int(params['n_galaxies'][0])
+
+    return(params)
+
+def read_cosmolike(bin_ends, loc='.', fn='nz_histo.txt'):
+    """
+    Turns cosmolike n(z) file into true n(z)s
+
+    Parameters
+    ----------
+    given_key: string
+        name of test case for which true n(z) is to be made
+    loc: string
+        path to file with n(z) evaluations
+    fn: string
+        filename
+
+    Returns
+    -------
+    true_nzs: list, chippr.discrete objects
+        per-bin true_nzs
+    """
+    bin_mids = (bin_ends[1:] + bin_ends[:-1]) / 2.
+    cl_input = np.genfromtxt(loc+'/'+fn).T
+    n_tomobins = len(cl_input) - 1
+    zs = cl_input[0]
+    nzs = cl_input[1:]
+    true_nzs = []
+    writeout = []
+    for b in range(n_tomobins):
+        f = spi.interp1d(zs, nzs[b])
+        bin_vals = f(bin_mids)
+        writeout.append(bin_vals)
+        true_func = discrete(bin_ends, bin_vals)
[…]

### 2019-03-25 research/scripts/cosmolike_sim_script.py
fixing path

@@ -179,7 +179,7 @@ def make_catalog_pre_split(given_key):
 
     # replace make_true with read_cosmolike in loop
     # test_info, true_nz = make_true(given_key)
-    true_nzs = read_cosmolike(test_info['bin_ends'], loc='../../../CosmoLike/chippr_forecasts', fn='nz_histo.txt')
+    true_nzs = read_cosmolike(test_info['bin_ends'], loc='.', fn='nz_histo.txt')
 
     interim_prior = make_interim_prior(given_key)
 

### 2019-03-25 chippr/log_z_dens_plots.py
plotting intermediate output for cosmolike tests

@@ -17,6 +17,8 @@ from chippr import stat_utils as s
 
 lnz, nz = '', ''#r'$\ln[n(z)]$', r'$n(z)$'
 
+
+# TODO Fix this horrible mess of repeated code, use a dict of namedtuples for per-curve params!
 s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '-', 2., 1., 'k', [(0, (1, 0.001))], 'Underlying Truth '
 s_tbp, w_tbp, a_tbp, c_tbp, d_tbp, l_tbp = ':', 2., 0.75, 'k', [(0, (1, 0.001))], 'Binned Truth '
 s_int, w_int, a_int, c_int, d_int, l_int = '-', 1., 0.5, 'k', [(0, (1, 0.001))], 'Implicit Prior '
@@ -259,9 +261,9 @@ def plot_sampler_progress(plot_information, sampler_output, full_chain, burn_ins
     return plot_information
 
 def make_err_txt(info, key):
-    rms = "{0:.3e}".format(info['stats']['rms']['true_nz'+ '__' + key[4:]])
+    # rms = "{0:.3e}".format(info['stats']['rms']['true_nz'+ '__' + key[4:]])
     kld = "{0:.3e}".format(info['stats']['kld'][key])
-    plot_txt = r'(KLD='+kld+', RMSE='+rms+')'
+    plot_txt = r'(KLD='+kld+')'#', RMSE='+rms+')'
     return plot_txt
 
 def plot_estimators(info, plot_dir, log=True, prepend='', metrics=True, mini=True, both=False):
@@ -337,7 +339,7 @@ def plot_estimators(info, plot_dir, log=True, prepend='', metrics=True, mini=Tru
         sps_log.set_ylim(-4., 1.)
         sps_log.set_ylabel(r'$\ln[n(z)]$')
     else:
-        sps_log.set_ylim(-0.15, 1.5)
+        sps_log.set_ylim(0., 3.)
         sps_log.set_ylabel(r'$n(z)$')
     sps_log.set_xlim(info['bin_ends'][0], info['bin_ends'][-1])
 

### 2019-05-17 research/paper/draft.tex
setting up some things for revising the paer

@@
 \usepackage{amsmath}
-\usepackage{graphicx}
-\graphicspath{{./}{./figures/}}
-\bibliographystyle{apj}
 
@@
 
+\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
+\newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
 \newcommand{\myemail}{aimalz@nyu.edu}
 \newcommand{\mathul}[1]{\underline{#1}}
-\newcommand{\chippr}{CHIPPR}
+\newcommand{\Sect}[1]{Section~\ref{#1}}
+\newcommand{\Eq}[1]{Equation~\ref{#1}}
+\newcommand{\Fig}[1]{Figure~\ref{#1}}
+\newcommand{\project}[1]{{\textsc{#1}}~}
+\newcommand{\lsst}{\project{LSST}}
+\newcommand{\desc}{\lsst-\project{DESC}}
+\newcommand{\sdss}{\project{SDSS}}
+\newcommand{\boss}{\project{BOSS}}
+\newcommand{\des}{\project{DES}}
+\newcommand{\Chippr}{\project{CHIPPR}}
+\newcommand{\repo}[1]{{\texttt{#1}}~}
+\newcommand{\qp}{\repo{qp}}
+\newcommand{\chippr}{\repo{chippr}}
+\newcommand{\cosmolike}{\repo{CosmoLike}}
+\newcommand{\emcee}{\repo{emcee}}
+\newcommand{\github}{\href{https://github.com}{GitHub}}
+\newcommand{\python}{\textit{Python}}
 \newcommand{\nz}{$n(z)$}
-\newcommand{\pz}{photo-$z$}
-\newcommand{\pzpdf}{photo-$z$ PDF}
-\newcommand{\lsst}{\textsc{lsst}}
-
-\newcommand{\aim}[1]{\textcolor{blue}{#1}}
+\newcommand{\data}{\ensuremath{\vec{d}}}
+\newcommand{\like}{\mathscr{L}}
+\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
+\newcommand{\expect}[1]{\left<#1\right>}
+\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
+\newcommand{\gvn}{\mid}% could use | or \vert
+\newcommand{\integral}[2]{\ensuremath{\int\ #1\ \mathrm{d} #2}}
+\newcommand{\sz}{spec-$z$~}
+\newcommand{\Sz}{Spec-$z$~}
+\newcommand{\pz}{photo-$z$~}
+\newcommand{\Pz}{Photo-$z$~}
+\newcommand{\pzpdf}{\pz PDF}
+\newcommand{\Pzpdf}{\Pz PDF}
+\newcommand{\zpdf}{redshift posterior}
+\newcommand{\pzip}{\pz implicit posterior}
+\newcommand{\Nz}{$N(z)$~}
+\newcommand{\stack}{$\hat{N}(z)$~}
+\newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
+\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
+\newcommand{\ndphi}{\bvec{\phi}}
+\newcommand{\mmle}{marginalized maximum likelihood estimate}
 
@@
 
-\title{How to obtain the redshift distribution from probabilistic redshift estimates}
+\title{How to obtain the redshift distribution from probabilistic redshift 
+estimates}
 
@@
 
-\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
-\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
-\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
-\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
+Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
+\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
+Avenue, 7th floor, New York, NY 10010, USA}
+\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
+7th floor, New York, NY 10003, USA}
+\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
+Heidelberg, Germany}
 
 \begin{abstract}
-A trustworthy estimate of the redshift distribution $n(z)$ is crucial for weak-lensing cosmology as we know it.
-Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be inaccessible, making photometric redshifts (photo-$z$s) the next-best alternative.
-The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
-However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
-The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
-We introduce a mathematically rigorous probabilistic graphical model (PGM) of hierarchical inference of $n(z)$, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
-The novel Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator produced by traditional procedures.
-We conclude by propagating these effects to constraints in the space of cosmological parameters.
+A trustworthy estimate of the redshift distribution \nz\ is crucial for using 
+galaxy catalogs to study cosmology via weak gravitational lensing and 
+large-scale structure.
+Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing 
+surveys are expected to be inaccessible, making photometric redshifts 
+(photo-$z$s) the next-best alternative.
+The nontrivial systematics affecting photo-$z$ estimation have motivated the 
+weak-lensing community to favor photo-$z$ probability density functions (PDFs) 
+as a more comprehensive alternative to photo-$z$ point estimates.
+However, analytic methods for utilizing these new data products in cosmological 
+inference are still evolving.
+The ubiquitous methodology known as stacking produces a systematically biased 
+estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very 
+regime where photo-$z$ PDFs are most necessary.
+We introduce a mathematically rigorous probabilistic graphical model (PGM) of 
+hierarchical inference of $n(z)$, which is provably the only self-consistent 
+way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
+The novel Cosmological Hierarchical Inference with Probabilistic Photometric 
+Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by 
+correctly propagating the redshift uncertainty information beyond the best-fit 
+estimator produced by traditional procedures.
+We conclude by propagating these effects to constraints in the space of 
+cosmological parameters.
 \end{abstract}
 
-\keywords{cosmology: cosmological parameters --- galaxies: statistics --- gravitational lensing: weak --- methods: data analysis --- methods: statistical}
+\keywords{cosmology: cosmological parameters --- galaxies: statistics --- 
+gravitational lensing: weak --- methods: data analysis --- methods: statistical}
 
@@
 
-\aim{Brief literature review:\\
-Major data releases including \pzpdf s\\
-The cosmology papers by those missions that stack PDFs\\
-Recent papers simultaneously inferring \pzpdf s and \nz}
+\aim{TODO: Pull in content from global intro to thesis.}
 
-\aim{Q: What is the status quo?\\
-A: The future of galaxy surveys is \pz s, and \pzpdf s rightfully replace \pz s; however, usage of \pzpdf s is limited.}
+Though their potential to improve estimates of physical parameters is 
+tremendous, \pzpdf s have been applied only to a limited extent.  
+They have been used to form selection criteria of samples from galaxy surveys 
+without propagation 
+through the calculations of physical parameters 
+\citep{VanBreukelen2009,Viironen2015}.  
+Probability cuts on Bayesian quantities are not uncommon \citep{Leung2015, 
+DiPompeo2015a}, but that procedure does not fully take advantage of all 
+information contained in a probability distribution for parameter inference.
 
-\aim{Q: Why should we question existing methods?\\
-A: Stacking is bad and only looks like it works because of assumptions that don't hold when the data is as bad as we anticipate.  Cite pedantic doc?}
+Despite the growing prevalence of \pzpdf\ production, no implementation of 
+inference using \pzpdf s has yet been presented with a mathematically 
+consistent methodology.  
+I present and validate a hierarchical Bayesian technique for the use of \pzpdf\ 
+in inference of arbitrary statistics relevant to cosmology, large-scale 
+structure, and galaxy evolution.  
+For simplicity, we consider only one-point statistics, though future work will 
+extend this methodology to higher-order statistics.
 
-\begin{itemize}
-	\item
-	Q: How can we improve the effectiveness of using \pzpdf s in inference?\\
-	A: Hierarchical inference is the only self-consistent way; see flow chart for outline of how it works in practice.
-	\item
-	Q: How does the result of \chippr compare to established estimators in terms of the accuracy of \nz ?\\
-	A: \chippr\ yields the best possible \nz , conditional on the accuracy of the \pzpdf s used.
-	\item
-	Q: How does the result of \chippr compare to established estimators in terms of the precision on cosmological constraints?
-\end{itemize}
+The \textit{redshift distribution function \Nz}, or, almost interchangably, its 
+normalized cousin the \textit{redshift density function \nz}, serves as an 
+ideal statistic upon which to demonstrate this novel approach.  
+\Nz\ is necessary for calculations of two-point correlation functions of weak 
+gravitational lensing and counting statistics that are used to probe dark 
+energy \citep{Masters2015}.  
+The observed \Nz\ also been used to validate survey selection functions 
+usnfenjed in generation of realistic, multi-purpose mock catalogs 
+\citep{Norberg2002}.  
+Additionally, \Nz\ has been the subject of inference using \pzpdf s before 
+\citep{sheldon_photometric_2012, Hildebrandt2012, Kelly2014, Benjamin2013, 
+bonnett_using_2015, Viironen2015, Asorey2016, leistedt_hierarchical_2016}, so 
+comparisons to the literature may easily be made.
 
-\section{Methods}
-\label{sec:methods}
+The prevailing way to estimate \nz\ for a sample of $\ntot$ galaxies $i$ is to 
+``stack'' \pzpdf s according to Equation~\ref{intro:eq:eqn:stack}.
+The stacked estimator of the redshift density function is effectively an 
+average of the \pzpdf\ catalog and, as I will show in this paper, not in 
+general mathematically valid.
+Stacking is nonetheless considered the preferred method for obtaining \Nz\ from 
+a \pzpdf\ catalog \citep{sheldon_photometric_2012, Kelly2014, Benjamin2013, 
+bonnett_using_2015, Viironen2015, Asorey2016}.  
+However, it must be noted here that Equation~\ref{intro:eq:eqn:stack} is not in 
+general mathematically valid.  
+(See \citet{Hogg2012} for a complete discussion.)
 
-\textcolor{blue}{Q: How can we improve the effectiveness of using \pzpdf s in inference?\\
-A: Hierarchical inference is the only self-consistent way.}
+I aim to develop a clear methodology guiding the use of \pzpdf s in inference 
+so they may be utilized effectively by the cosmology community.
+Though others have approached the problem before 
+\citep{leistedt_hierarchical_2016}, the method presented here differs in that 
+it makes use of any existing catalog of \pzpdf s, rather than requiring a 
+simultaneous derivation of the \pzpdf s and the redshift distribution, making 
+it preferable to ongoing surveys that may be resistant to restructuring their 
+analysis pipelines.
 
-\subsection{Model generalities}
-\label{sec:model}
+In Section~\ref{sec:meth}, I present the \Chippr\ model and \chippr\ 
+implementation for characterizing the full posterior probability landscape of 
+\Nz\ using \pzpdf s. 
+In Section~\ref{sec:alldata}, I describe the experimental design for testing 
+the fully probabilistic approach to mock and real datasets, the results of 
+which are found in Section~\ref{sec:results}.
+In Section~\ref{sec:results}, I stress-test the \Chippr\ model against stacking 
+and its other competitors in the context of cosmology.
 
-We begin by reframing the redshift distribution \nz\ from a probabilistic perspective.
-Here we define a redshift distribution \nz\ as the normalized probability density
+\section{Method}
+\label{sec:meth}
+
+Consider a survey of $J$ galaxies $j$, each with photometric data $\data_{j}$; 
+thus the entire survey over some solid angle produces the ensemble of 
+photometric magnitudes (or colors) and their associated observational errors 
+$\{\data_{j}\}$.  
+Each galaxy $j$ has a redshift $z_{j}$ that we would like to learn; redshift is 
+a parameter in this case.  
+The distribution of the ensemble of redshifts $\{z_{j}\}$ may be described by 
+the hyperparameters defining the redshift distribution function \nz\ that we 
+would like to quantify.  
+This situation may be considered to be a probabilistic generative model, 
+illustrated by the directed acyclic graph of \Fig{fig:pgm}.
+
+The redshift distribution function \nz\ is the number of galaxies per unit 
+redshift, effectively defining the evolution in the number of galaxies 
+\citep{Menard2013}.  
+In the following sections, I present and compare methods for estimating \nz\ 
+from \pzpdf s.  
+\Sect{sec:prob} contains the mathematical derivation of a probabilistic model 
+for $n(z)$ dependent on photo-$z$ probability distribution functions, and 
+\Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.
+
+\subsection{Forward Model}
+\label{sec:forward}
+
+We begin by reframing the redshift distribution \nz\ from a probabilistic 
+perspective.
+Here we define a redshift density \nz\ as the normalized probability density
 \begin{equation}
 \label{eqn:nz}
-\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ \int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
+\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ 
+\int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
 \end{equation}
 of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
-We may without loss of generality impose a parametrization
+We believe that galaxy redshifts are indeed drawn from \nz, making it a 
+probability density over redshift; this fact can also be confirmed by 
+dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{Hogg2012}.
+
+We may without loss of generality impose a parameterization
 \begin{equation}
 \label{eqn:fz}
-f_{\phi}(z)\ \equiv\ n(z)
+f(z; \ndphi)\ \equiv\ n(z)
 \end{equation}
-in terms of some parameter vector $\phi$.
-We believe that galaxy redshifts are indeed drawn from \nz, making it a probability density over redshift; this fact can also be confirmed from dimensional analysis of Equation~\ref{eqn:nz}.
-Therefore, it can be rewritten as
+in terms of some parameter vector $\ndphi$.
+At this point, the parameter vector is quite general and may represent 
+coefficients in a high-order polynomial as a function of redshift, a set of 
+means and variances defining Gaussians that sum to the desired distribution, a 
+set of histogram heights that describe a binned version of the redshift 
+distribution function, etc.
+Upon doing so, we may rewrite \Eq{eqn:fz} as 
 \begin{equation}
 \label{eqn:pz}
-z_{j}\ \sim\ \Pr(z \mid \phi)\ \equiv\ f_{\phi}(z),
+z_{j}\ \sim\ \pr{z \gvn \ndphi}\ \equiv\ f(z; \ndphi),
 \end{equation}
-a probability density over redshift conditioned on the parameters defining \nz.
-Note that $z_{j}$ does not depend on $z_{j'}$, a statement of the causal independence of galaxy redshifts from one another.
+a probability density over redshift conditioned on the parameters $\ndphi$ 
+specifying \nz.
+Note that $z_{j}$ does not depend on the redshift $z_{j'}$ of some other galaxy 
+$j' \neq j$, a statement of the causal independence of galaxy redshifts from 
+one another.
 
-In addition to believing \nz\ is a PDF from which redshifts are drawn, we also believe that there is some PDF from which photometric data $d$, which may be any combination of fluxes, magnitudes, colors, and their observational errors, are drawn.
-Such a PDF over data is a likelihood
+In addition to believing \nz\ is a PDF from which redshifts are drawn, we also 
+believe that there is some higher dimensional probability space $\pr{z, \data}$ 
+of redshift and photometric data vectors $\data$, which may be any combination 
+of fluxes, magnitudes, colors, and their observational errors.
+In that sense \nz\ is equivalent to an integral
+\begin{equation}
+\label{eqn:integral}
+n(z)\ =\ \integral{\pr{z, \data}}{\data}
+\end{equation}
+over the dimension of data in that joint probability space.
+Note that galaxies may have different observational data despite sharing the 
+same redshift, and that galaxies at different redshifts may have identical 
+photometry; the space $\pr{z, \data}$ need not be one-to-one.
+We assume a stronger version of statistical independence here, that draws 
+$(z_{j}, \data_{j})$ are independent of draws $(z_{j'}, \data_{j'})$ in this 
+space; the data and redshift of each galaxy are independent of those of other 
+galaxies.
+
+However, this problem has additional causal structure that we can acknowledge.
+The photometry results from the redshifts, not the other way around.
+This is the fundamental assumption upon which \pz\ estimation is based.
+The forward model corresponds to first drawing redshifts according to 
+\Eq{eqn:pz} and then drawing data from the likelihood
 \begin{equation}
 \label{eqn:pzpdf}
-d\ \sim\ \Pr(d \mid z)
+\data_{j}\ \sim\ \pr{\data \gvn z_{j}}
 \end{equation}
-conditioned on redshift.
-This assumption that the data are drawn from some function of the redshift forms the foundation upon which \pz\ estimation is based.
-Note that galaxies may have different observational data $d$ despite sharing the same redshift and that the data $d_{j}$ of one galaxy is causally independent of the redshifts $z_{j'}$ and data $d_{j'}$ of other galaxies.
+of photometry conditioned on redshift, illustrated in 
+Figure~\ref{intro:fig:fig:pedagogical_scatter}.
 
-This description of the physical system corresponds to a forward model by which we actually believe photometry is generated:
+This description of the physical system corresponds to a forward model by which 
+we actually believe photometry is generated:
 \begin{enumerate}
-	\item There exists a redshift distribution \nz\ with parameters $\phi$.
-	\item Galaxy redshifts $\{z_{j}\}$ are independent draws from $\Pr(z 
-\mid \phi)$.
-	\item Galaxy photometry $d_{j}$ is drawn from the likelihoods $\Pr(d | 
-z_{j})$.
+	\item There exists a redshift distribution \nz\ with parameters 
+$\ndphi$.
+	\item Galaxy redshifts $\{z_{j}\}$ are independent draws from $\pr{z 
+\gvn \ndphi}$.
+	\item Galaxy photometry $\data_{j}$ is drawn from the likelihoods 
+$\pr{\data_{j} \gvn z}$.
 \end{enumerate}
-A forward model such as this corresponds to a probabilistic graphical model (PGM), represented by a directed acyclic graph (DAG) as in Figure~\ref{fig:pgm}.
+
+\subsection{Probabilistic Model}
+\label{sec:prob}
+
+A forward model such as that of \Sect{sec:forward} corresponds to a 
+probabilistic graphical model (PGM), represented by a directed acyclic graph 
+(DAG) as in \Fig{fig:pgm}.
+A DAG conveys the causal relationships between physical parameters and, like a 
+Feynman diagram in the context of particle physics, is a shorthand for 
+mathematical relationships between variables.
+The photometric data $\data_{j}$ of a galaxy is drawn from some function of its 
+redshift $z_{j}$, independent of other galaxies' data and redshift.
+Both data and redshift are random variables, but data is the one that we 
+observe and redshift is not directly observable.
+In this problem, we don't care about further constraining the redshifts of 
+individual galaxies, only the redshift distribution, so we consider redshift to 
+be a \textit{latent variable}.
+Because the parameters $\ndphi$ that we seek are causally separated from the 
+data by the latent variable of redshift, we call them \textit{hyperparameters}.
+
 \begin{figure}
 	\begin{center}
-		\includegraphics[width=0.25\textwidth]{fig/pgm.png}
-		\caption{The PGM}
+		\includegraphics[width=0.2\textwidth]{figures/chippr/pgm.png}
+		\caption{The directed acyclic graph of the CHIPPR model, where 
+circles indicate random variables and arrows indicate causal relationships.
+			The redshift distribution function parameterized by 
+hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, 
+indicated as a box.  
+			The redshifts $\{z_{j}\}$ of all galaxies in the survey 
+are latent variables independently drawn from the redshift distribution, which 
+is a function of $\ndphi$. 
+			The photometric data $\data_{j}$ for each galaxy is 
+drawn from a function of its redshift $z_{j}$ and observed, indicated by a 
+shaded circle.}
 		\label{fig:pgm}
@@
 \end{figure}
-A DAG conveys the causal relationships between physical parameters and, like a Feynman diagram in the context of particle physics, is a shorthand for mathematical relationships between variables.
 
-\subsection{Application to cosmology}
-\label{sec:application}
+The problem facing cosmologists is to determine the true value of $\ndphi$ from 
+observing the photometry $\{\data_{j}\}$ of a large sample of $J$ galaxies $j$.
+To self-consistently propagate the uncertainty in the redshifts, however, it is 
+more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ 
+over all possible values of $\ndphi$ conditioned on all the observed data 
+$\{\data_{j}\}$ available in a generic catalog.
+In order to use the DAG of \Fig{fig:pgm} to derive an expression for 
+$\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two 
+more concepts, confusingly named the implicit prior and the prior probability 
+density.
 
-The problem facing cosmologists is to determine the true parameters $\phi_{0}$ of \nz\ from observing the photometry $\{d_{j}\}$ of a large sample of galaxies $j$.
-To self-consistently propagate the uncertainty in the redshifts, however, it is more appropriate to estimate the posterior $\Pr(\phi \mid \{d_{j}\})$ over all possible parameters $\phi$ (and thus potential redshift distributions \nz ) conditioned on all the observed data $\{d_{j}\}$ available in a generic catalog.
+When we constrain the redshift of a galaxy using its observed photometric data 
+$\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$.
+However, to do this, we must have a model for the general relationship between 
+redshifts and photometry, whether empirical, as is the case for machine 
+learning \pzpdf\ methods, or analytic, as is the case for template-based 
+\pzpdf\ methods.
+Such a relationship is defined in the space of probability density over 
+redshift, so it must be able to be parameterized by the same functional form 
+$f(z; \ndphi)$ as \nz .
+It is thus natural to write it as $\pr{z \gvn \ndphi^{*}}$, where $\ndphi^{*}$ 
+is the parameters for this relationship under some generic \pzpdf\ method.
+We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely 
+explicitly known nor chosen by the researcher; for template-based methods where 
+it can be chosen to be ``realistic,'' it may be appropriate to call it an 
+\textit{interim prior}.
+Because the implicit prior is unavoidable, the \pzpdf s reported by any method 
+are really \textit{implicit-weighted posteriors} $\pr{z \gvn \data, 
+\ndphi^{*}}$.
 
-In order to use the DAG of Figure~\ref{fig:pgm} to derive an expression for $\Pr(\phi \mid \{d_{j}\})$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the implicit prior and the prior probability density.
+The implicit prior $\ndphi^{*}$ may be thought of as an initial guess for the 
+parameters contained in $\ndphi$, inspired by the generative model for 
+photometry from the redshift distribution functions and including some 
+parameters defining intrinsic galaxy spectra and instrumental effects. 
+(See \citealt{benitez_bayesian_2000} for more detail.)  
+For statistical purposes, we would like any interim prior to be uninformative, 
+but this is rarely achievable.  
+In the case of estimating \nz\ photometrically, it is common to use 
+$\ndphi^{*}$ corresponding to \nz\ derived from some different, 
+spectroscopically confirmed sample or from a cosmological simulation.
 
-When we constrain the redshift of a galaxy using its observed photometric data $d_{j}$, we are effectively estimating a posterior $\Pr(z \mid d_{j})$.
-However, to do this, we must have a model for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
-Such a relationship is defined in the space of probability density over redshift, so it must be able to be parameterized by the same functional form $f_{\phi}(z)$ as \nz .
-It is thus natural to write it as $\Pr(z \mid \phi^{*})$, where $\phi^{*}$ is the parameters for this relationship under some generic \pzpdf\ method.
-We call $\Pr(z \mid \phi^{*})$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher; for template-based methods where it can be chosen to be ``realistic,'' it may be appropriate to call it an \textit{interim prior}.
-Because the implicit prior is unavoidable, the \pzpdf s reported by any method are really \textit{implicit-weighted posteriors} $\Pr(z \mid d, \phi^{*})$.
+The prior probability density $\pr{\ndphi}$ is a more familiar concept in 
+astronomy; to progress, we will have to choose a prior probability density over 
+all possible parameters $\ndphi$.
+This prior need not be excessively proscriptive; for example, it may be chosen 
+to enforce smoothness at physically motivated scales in redshift without 
+imposing any particular region as over- or under-dense.
 
-The prior probability density $\Pr(\phi)$ is a more familiar concept in astronomy; to progress, we will have to choose a prior probability density over all possible parameters $\phi$.
-This prior need not be excessively proscriptive; for example, it may be chosen to enforce smoothness at physically motivated scales in redshift without imposing any particular region as over- or under-dense.
-
-With these definitions, we obtain the desired expression for $\Pr(\phi \mid 
-\{d_{j}\})$,
+With these definitions, we obtain the desired expression for $\pr{\ndphi \gvn 
+\{\data_{j}\}}$,
 \begin{align}
@@
 \label{eqn:fullpost}
-  \ln[\Pr(\phi \mid \{d_{j}\})] & \propto \ln[\Pr(\phi)] + \ln \left[\int dz 
-\right.\\
-  & \left. \exp \left[\sum_{j=1}^{J} \left(\ln[\Pr(z \mid d_{j}, \phi^{*})] 
-\right. \right. \right.\\
-  & \left. \left. \left. + \ln[\Pr(z \mid \phi)] - \ln[\Pr(z \mid \phi^{*})]\ 
-\right)\ \right]\ \right] ,
+\pr{\ndphi \gvn \{\data_{j}\}} & \propto 
+\pr{\ndphi}\left[\integral{\prod_{j=1}^{J} \left(\pr{z \gvn \data_{j}, 
+\ndphi^{*}}\pr{z \gvn \ndphi}(\pr{z \gvn \ndphi^{*}})^{-1}\right)}{z}\right] ,
 \end{split}
 \end{align}
-which is the heart of \chippr.
-The entire derivation of Equation~\ref{eqn:fullpost} is provided in Appendix~\ref{app:math}.
+the heart of \Chippr.
+The derivation is provided below in \Sect{app:math}.
+
+\subsubsection{Derivation}
+\label{app:math}
+
+In this paper, we work exclusively with log-probabilities.  
+What we wish to estimate is then the full log-posterior probability 
+distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ 
+given the catalog of photometry $\{\data_{j}\}$.
+
+By Bayes' Rule, the full log-posterior
+\begin{equation}
+\label{eqn:basicbayes}
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] = \ln[\pr{\{\data_{j}\} \gvn \ndphi}] + 
+\ln[\pr{\ndphi}] - \ln[\pr{\{\data_{j}\}}]
+\end{equation}
+may be expressed in terms of the full log-likelihood probability distribution 
+(hereafter the full log-likelihood) $\ln[\pr{\{\data_{j}\} \gvn \ndphi}]$ by 
+way of a hyperprior log-probability distribution (hereafter the hyperprior) 
+$\ln[\pr{\ndphi}]$ over the hyperparameters and the log-evidence probability of 
+the data $\ln[\pr{\{\data_{j}\}}]$.
+However, the evidence is rarely known, so we probe the full log-posterior 
+modulo an unknown constant of proportionality.
+
+The full log-likelihood may be expanded in terms of a marginalization over the 
+redshifts as parameters, as in 
+\begin{equation}
+\label{eqn:marginalize}
+\ln[\pr{\{\data_{j}\} \gvn \ndphi}] = \ln\left[\integral{\pr{\{\data_{j}\} \gvn 
+\{z_{j}\}} \pr{\{z_{j}\} \gvn \ndphi}}{\{z_{j}\}}\right].
+\end{equation}
+
+We shall make two assumptions of independence in order to make the problem 
+tractable; their limitations are be discussed below.  
+First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ 
+individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn 
+z_{j}}]$, as in 
+\begin{equation}
+\label{eqn:indiedat}
+\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn 
+z_{j}}],
+\end{equation}
+a result of the definition of probabilistic independence encoded by the box in 
+\Fig{fig:pgm}.
+Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent 
+draws from the true $\pr{z \gvn \ndphi}$.  
+Additionally, $J$ itself is a Poisson random variable.  
+The combination of these assumptions is given by 
+\begin{equation}
+\label{eqn:indie}
+\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ 
+\ln[\pr{z_{j} \gvn \ndphi}].
+\end{equation}
+The derivation differs when $J$ is not known, say, when we want to learn about 
+a distribution in nature rather than a distribution specific to data in hand, 
+but for a photometric galaxy catalog where the desired quantity is $n(z)$ for 
+the galaxies entering a larger cosmology calculation, it is a fixed quantity.
+A detailed discussion of this matter may be found in 
+\citet{Foreman-Mackey2014}.  
+Applying Bayes' Rule, we may combine terms to obtain 
+\begin{equation}
+\label{eqn:posterior}
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto \ln[\pr{\ndphi}] - \integral{f(z; 
+\ndphi)}{z} + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn 
+\ndphi}}{z}\right].
+\end{equation}
+
+Since we only have access to implicit \pz\ posteriors, we must be able to write 
+the full log-posterior in terms of implicit \pz\ log-posteriors rather than the 
+log-likelihoods of \Eq{eqn:posterior}.
+To do so, we will need an explicit statement of this implicit prior 
+$\ndphi^{*}$ for whatever method is chosen to produce the implicit \pz\ 
+posteriors.
+
+To perform the necessary transformation from likelihoods to posteriors, we 
+follow the reasoning of \citet{Foreman-Mackey2014}.  
+Let us consider the probability of the parameters conditioned on the data and 
+an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} 
+as 
+\begin{equation}
+\label{eqn:trick}
+\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn 
+\data_{j}, \ndphi^{*}}] - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
+\end{equation}
+
+Once the implicit prior $\ndphi^{*}$ is explicitly introduced, we may expand 
+the last term in \Eq{eqn:trick} according to Bayes' Rule to get 
+\begin{equation}
+\label{eqn:expand}
+\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn 
+\data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn 
+\ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
+\end{equation}
+Because there is no direct dependence of the data upon the hyperparameters, we 
+may again expand the term $\ln[\pr{\data_{j} \gvn z, \ndphi^{*}}]$ to obtain 
+\begin{equation}
+\label{eqn:indterm}
+\ln[\pr{\vec{d}_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn 
+\data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn 
+\ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] - \ln[\pr{\data_{j} \gvn \ndphi^{*}}].
+\end{equation}
+Canceling the undesirable terms for the inaccessible likelihood 
+$\ln[\pr{\data_{j} \gvn z}]$ and trivial $\ln[\pr{\data_{j} \gvn \ndphi^{*}}]$ 
+yields
+\begin{equation}
+\label{eqn:cancel}
+\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]  - 
+\ln[\pr{z \gvn \ndphi^{*}}].
+\end{equation}
+We put this all together to get the full log-posterior probability distribution 
+of 
+\begin{equation}
+\label{eqn:final}
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto \ln[\pr{\ndphi}] + \ln 
+\left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, 
+\ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn 
+\ndphi^{*}}]\right)\right]}{z}\right] ,
+\end{equation}
+
+The argument of the integral in the log-posterior of \Eq{eqn:final} depends 
+solely on knowable quantities (and those we must explicitly assume) and can be 
+calculated for a given sample of \pz\ log-posteriors $\{\ln[\pr{z \gvn 
+\data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with 
+which they were obtained, noting the relation of 
+\begin{equation}
+\label{eqn:params}
+\pr{z \gvn \ndphi} = \frac{f(z; \ndphi)}{\integral{f(z; \ndphi)}{z}}.
+\end{equation}
+Since we cannot know constant of proportionality, we sample the desired full 
+log-posterior $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$ using Monte Carlo-Markov 
+chain (MCMC) methods.  
+The method outlined here is valid regardless of how the implicit \pz\ 
+log-posteriors are obtained so the many approaches to producing \pzpdf s will 
+not be rehashed; though the matter is outside the scope of this paper, reviews 
+of various methods have been presented in the literature 
+\citep{sheldon_photometric_2012, Ball2008, carrasco_kind_tpz:_2013, 
+carrasco_kind_exhausting_2014}, and will be briefly reviewed in \aim{cite PZDC1 
+paper here}.
+
+\subsubsection{Model Limitations}
+\label{sec:limitations}
+
+Finally, we explicitly review the assumptions made by this approach, which are 
+as follows:
+
+\begin{enumerate}
+	\item Photometric measurements of galaxies are statistically 
+independent Poisson draws from the set of all galaxies such that 
+\Eq{eqn:indiedat} and \Eq{eqn:indie} hold.
+	\item We take the reported \pzip s to be accurate, free of model 
+misspecification; draws thereof must not be inconsistent with the distribution 
+of photometry and redshifts.
+	Furthermore, we must be given the implicit prior $\ndphi^{*}$ used to 
+produce the \pzip s.
+	\item We must assume a hyperprior distribution $\pr{\ndphi}$ 
+constraining the underlying probability distribution of the hyperparameters, 
+which is informed by our prior beliefs about the true redshift distribution 
+function.
+\end{enumerate}
+
+These assumptions have known limitations.  
+First, the photometric data are not a set of independent measurements; the data 
+are correlated not only by the conditions of the experiment under which they 
+were observed (instrument and observing conditions) but also by redshift 
+covariances resulting from physical processes governing underlying galaxy 
+spectra and their relation to the redshift distribution function.
+Second, the reported \pzip s may not be trustworthy; there is not yet agreement 
+on the best technique to obtain \pzpdf s, and the implicit prior may not be 
+appropriate or even known to us as consumers of \pzip s.  
+Third, the hyperprior may be quite arbitrary and poorly motivated if the 
+underlying physics is complex, and it can only be appropriate if our prior 
+beliefs about \nz\ are accurate.
 
 \subsection{Implementation}
-\label{sec:implementation}
+\label{sec:exp}
 
-In this study, we compare the results of Equation~\ref{eqn:fullpost} to those of the two most common approaches to estimating \nz\ from a catalog of \pzpdf s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior probability (i.e.
-the modes of the \pzpdf s), and the stacked estimator
-\begin{align}
+I implemented the \Chippr\ model in code in order to perform tests of its 
+validity and to compare its performance to that of traditional alternatives.
+In \Sect{sec:mcmc}, I describe the publicly available \chippr\ library.
+In \Sect{sec:acorr}, I outline how \chippr\ can be used to sample the full 
+log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
+
+\subsubsection{Code}
+\label{sec:mcmc}
+
+\chippr\ is a \python\ 2 
+library\footnote{\url{https://github.com/aimalz/chippr}} that includes an 
+implementation of the \Chippr\ model as well as an extensive suite of tools for 
+comparing \Chippr\ to other approaches.
+
+Though there are plans for future expansion to more flexible parameterizations, 
+the current version of \chippr\ uses a log-space piecewise constant 
+parameterization
+\begin{equation}
+\label{eqn:logstepfunc}
+f(z; \ndphi) = \exp[\phi^{k}]\ \mathrm{if}\ z^{k} < z < z^{k+1}
+\end{equation}
+for \nz\ and every \pzpdf, satisfying
+\begin{equation}
+\label{eqn:logstepfuncnorm}
+\sum_{k=1}^{K} \exp[\phi^{k}] \delta z^{k} = 1
+\end{equation}
+with $K$ bins of width $\delta z^{1}, \dots, \delta z^{K}$ defined by endpoints 
+$z^{0}, \dots, z^{K}$.
+Thus each $\pr{z \gvn \data_{j}} = f(z; \ndphi_{j})$ has parameters 
+$\ndphi_{j}$ that are defined in the same basis as those of \nz.
+To infer the full log-posterior distribution $\ln[\pr{\ndphi \gvn 
+\{\data_{j}\}}]$, one must provide a plaintext file with $K+1$ redshift bin 
+endpoints $\{z_{k}\}$, the parameters $\ndphi^{*}$ of the implicit log-prior, 
+and the parameters $\{\ndphi_{j}\}$ of the log-posteriors $\ln[\pr{z \gvn 
+\data_{j}, \ndphi^{*})}$.
+
+The \emcee \citep{Foreman-Mackey2013} implementation of ensemble sampling is 
+used to sample the full log-posterior of \Eq{eqn:final}. 
+\chippr\ accepts a configuration file of user-specified parameters, among them 
+the number $W$ of walkers.
+At each iteration $i$ and for each walker, a proposal distribution 
+$\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for 
+acceptance to or rejection from the full log-posterior distribution.
+
+Two threshold conditions are defined, one designating all previous samples to 
+be ignored as as products of a burn-in phase and another indicating when a 
+sufficient number of post-burn samples have been accepted.  
+In this case, the first threshold (described in \Sect{sec:acorr}) is defined in 
+terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an 
+accumulation of $10^{4}$ samples.
+
+Though previous versions used \texttt{HDF5} for the primary I/O format due to 
+its efficiency for large quantities of data, it was abandoned in favor of 
+\texttt{pickle} in the working release due to the instability of the \python\ 
+implementation of the format on high-performance computing systems.  
+The resulting output is a set of $I$ ordered \texttt{hickle} files enumerated 
+by $\rho$ containing the state information after each sub-run.  
+The state information includes $\frac{I_{0}}{s}$ actual samples $\ndphi_{i}$ 
+for a pre-specified chain thinning factor $s$ and their full posterior 
+probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$ as well as the 
+autocorrelation times and acceptance fractions calculated for each element of 
+$\ndphi$ over the entire sub-run.
+
+\subsubsection{Convergence Criteria}
+\label{sec:acorr}
+
+In addition to qualitative visual inspection of the chains, two quantities that 
+probe the convergence of the sampler are used in this study, the 
+autocorrelation time and the Gelman-Rubin convergence criterion.
+
+The autocorrelation time is effectively a measure of the efficiency of the 
+method and can be described as the expected number of iterations necessary to 
+accept a new sample independent of the current accepted sample.  
+A sampler that converges faster will have a smaller autocorrelation time, and 
+smaller autocorrelation times are preferable because it means fewer iterations 
+are wasted on non-independent samples when independent samples are desired.  
+See \citet{Foreman-Mackey2013} for a more complete exploration of the 
+autocorrelation time.  
+In all tests discussed here, autocorrelation times across walkers and 
+parameters were approximately 20, meaning two samples 20 or more iterations 
+apart were independent, a satisfactory level of efficiency.  
+Low autocorrelation times are a necessary but not always sufficient convergence 
+condition, as the autocorrelation times calculated for tests in this paper were 
+constant across all sub-runs, even those that were obviously burning in.
+
+The Gelman-Rubin statistic
+\begin{equation}
+\label{eqn:gr}
+R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} 
+b_{k}}{w_{k}}},
+\end{equation}
+a weighted sum of the mean $w_{k}$ of the variances within individual walkers' 
+chains and the variance $b_{k}$ between chains of different walkers $m$, is 
+calculated over each sub-run $i$ to determine the duration of the burn-in 
+period.  
+Convergence is achieved when the statistic approaches unity.
+
+\subsection{Comparison with Alternative Approaches}
+\label{sec:sheldon}
+
+In this study, we compare the results of \Eq{eqn:fullpost} to those of the two 
+most common approaches to estimating \nz\ from a catalog of \pzpdf s: the 
+distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior 
+probability
+\begin{equation}
+\label{eqn:mmap}
+f^{MMAP}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \delta(z, \mathrm{mode}[\pr{z \gvn 
+\data_{j}, \ndphi^{*}}])
+\end{equation}
+(i.e. the distribution of modes of the \pzpdf s) and the stacked estimator of 
+\Eq{eqn:stacked}, which can be rewritten as 
+\begin{equation}
 \label{eqn:stacked}
-  \hat{n}(z) &\equiv \frac{1}{J}\ \sum_{j=1}^{J}\ \Pr(z \mid d_{j}, \phi^{*}) .
-\end{align}
+f^{stack}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \pr{z \gvn \data_{j}, \ndphi^{*}}
+\end{equation}
+in terms of the implicit \pz\ posteriors we have.
+These two approaches have been compared to one another by 
+\citet{Hildebrandt2012}, \citet{Benjamin2013}, and \citet{Asorey2016} in the 
+past but not to \Chippr.
 
-We perform this comparison on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed above and presented in detail in Appendix~\ref{app:data}.
-The mock data emulates the three sources of error of highest concern to the \pz\ community: intrinsic scatter, catastrophic outliers, and systematic bias.
-Figure~\ref{fig:mega_scatter} illustrates these three effects individually at ten times the tolerance of the upcoming Large Synoptic Survey Telescope (LSST).
+Point estimation converts the implicit photo-$z$ posteriors $\pr{z \gvn 
+\data_{j}, \ndphi^{*}}$ into delta functions with all probability at a single 
+estimated redshift.  
+Some variants of point estimation choose this single redshift to be that of 
+maximum a posteriori probability $\mathrm{mode}[\pr{z \gvn \data_{j}, 
+\ndphi^{*}}]$ or the expected value of redshift $\langle z \rangle = 
+\integral{z \pr{z \gvn \data_{j}, \ndphi^{*}}}{z}$.
+\citet{tanaka_photometric_2018} directs attention to deriving an optimal point 
+estimate reduction of a \pzpdf, but since the purpose of this paper is to 
+compare against the most established alternative estimators of \nz, its use 
+will be postponed until a future study.
+Stacking these modified \pzpdf s leads to the marginalized maximum a posteriori 
+(MMAP) estimator and the marginalized expected value (MExp) estimator, though 
+only the former is included in this study since the latter has fallen out of 
+favor in recent years\footnote{And for good reason!  Consider a bimodal \pzpdf; 
+its expected value may very well fall in a region of very low probability, 
+yielding a less probable point estimate than the point at which either peak 
+achieves its maximum.}.
+
+It is worth discussing the relationship between point estimation and stacking.  
+When the point estimator of redshift is equal to the true redshift, stacking 
+delta function \pzpdf s will indeed lead to an accurate recovery of the true 
+redshift distribution function.  
+However, stacking is in general applied indiscriminately to broader \pzpdf s 
+and imperfect point estimators of redshift.  
+It is for these reasons that alternatives are considered here.
+
+A final estimator of the hyperparameters is the maximum marginalized likelihood 
+estimator, the value of $\ndphi$ maximizing the log posterior given by 
+\Eq{eqn:final} using any optimization code.  
+To compare with sampling, the MMLE also depends on the choice of the hyperprior 
+distribution, and it does not produce a full posterior probability distribution 
+over the parameters of interest, only point estimators.  
+It must be noted that computation of the maximum marginalized likelihood 
+estimator may be unstable depending on the strengths and weaknesses of the 
+optimizer.  
+In general, derivatives will not be available for the full posterior 
+distribution, restricting optimization methods used.
+
+In \Sect{sec:diag} we outline the measures used to evaluate the performance of 
+the method.
+
+\subsubsection{Performance metrics}
+\label{sec:diag}
+
+The results of the computation described in \Sect{sec:mcmc} are evaluated for 
+accuracy on the basis of some quantitative measures.  
+Beyond visual inspection of samples, we calculate summary statistics to 
+quantitatively compare different estimators' precision and accuracy.  
+Since MCMC samples of hyperparameters are Gaussian distributions, we can 
+quantify the breadth of the distribution for each hyperparameter using the 
+standard deviation regardless of whether the true values are known.
+
+In simulated cases where the true parameter values are known, we calculate the 
+Kullback-Leibler divergence (KLD), given by 
+\begin{equation}
+\label{eqn:kl}
+KL_{',\ddagger} = \integral{\pr{z \gvn \ndphi'} \ln \left[ \frac{\pr{z \gvn 
+\ndphi'}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
+\end{equation}
+which measures a distance from parameter values $\ndphi'$ to true parameter 
+values $\ndphi^{\dagger}$, which is invariant under changes of variables.  
+We note that $KL_{',\dagger} \neq KL_{\dagger,'}$ and is only interpretable 
+when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than 
+$\ndphi'$.
+The KLD is explored in detail in the Appendix to \aim{cite qp paper here}.
+In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that 
+produced by one of the methods in question.
+
+\section{Validation}
+\label{sec:alldata}
+
+We compare the results of \Chippr\ to those of stacking and the histogram of 
+\pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf 
+s generated via the forward model discussed in \Sect{sec:forward}.
+\Fig{fig:flowchart} illustrates the implementation of this forward model used 
+in this paper.
+
 \begin{figure*}
 	\begin{center}
-		\includegraphics[width=0.3\textwidth]{fig/hivarsigmas_scatter.png}
-    	\includegraphics[width=0.3\textwidth]{fig/uouthi_scatter.png}
-    	\includegraphics[width=0.3\textwidth]{fig/neghivarbias_scatter.png}
-		\caption{intrinsic scatter (left), uniform outliers (center), bias (right);
-   		Not sure how best to broach the subject of nonuniform outliers and bias as model misspecification
+		\includegraphics[width=\textwidth]{figures/chippr/flowchart.png}
+		\caption{A flow chart illustrating the forward model used to 
+generate mock data in the validation of \Chippr, as described in 
+\Sect{sec:forward}.
+			Ovals indicate a quantity that must be chosen in order 
+to generate the data, rectangles indicate an operation I perform, and rounded 
+rectangles indicate a quantity created by the forward model.
+			Arrows indicate the inputs and outputs of each 
+operation performed to simulate mock \pzpdf\ catalogs.}
+		\label{fig:flowchart}
+	\end{center}
+\end{figure*}
+
+The true redshift distribution used in these tests is a particular instance of 
+the gamma function
+\begin{equation}
+\label{eqn:gamma}
+n^{\dagger}(z) = \frac{1}{2 c_{z}} \left(\frac{z}{c_{z}}\right)^{2}\ 
+\exp\left[-\frac{z}{c_{z}}\right]
+\end{equation}
+with $c_{z} = 0.3$, because it has been used in forecasting studies for \des\ 
+and \lsst.
+
+The mock data emulates the three sources of error of highest concern to the 
+\pz\ community that are explored in detail later in this section: intrinsic 
+scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and 
+systematic bias (\Sect{sec:bias}).
+Tests including all three effects at the tolerance levels of \lsst\ (see 
+Table~\ref{intro:tab:tab:lsstsrd}) are presented in \Sect{sec:results}.
+\Fig{fig:mega_scatter} illustrates these three effects individually at twice 
+the tolerance of \lsst\ for demonstrative purposes, hearkening back to 
+Figure~\ref{intro:fig:fig:pedagogical_scatter}.
+We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to 
+be a priority for the \pz\ community.
+
+\begin{figure}
+	\begin{center}
+
+\includegraphics[width=0.3\textwidth]{figures/chippr/scatter_scatplot.png}
+
+\includegraphics[width=0.3\textwidth]{figures/chippr/outlier_scatplot.png}
+
+\includegraphics[width=0.3\textwidth]{figures/chippr/bias_scatplot.png}
+		\caption{The joint probability space of true and estimated 
+redshift for the three concerning \pz\ systematics at twice the level of the 
+\lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and 
+bias (right).
+			The main panel of each shows samples (black points) in 
+the space of mock data and redshift, akin to the standard scatterplots of true 
+and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal 
+(thin red line), and posterior probabilities evaluated at the given estimated 
+redshift (colored step functions).
+			The insets show marginal histograms (gray) in each 
+dimension, that can be compared with the true \nz\ used to make the figure 
+(blue curve) to see the effect of the isolated systematic, as well as the 
+implicit prior (red line).
 		}
@@
 	\end{center}
-\end{figure*}
-Tests including all three effects at the tolerance levels of LSST are presented in Section~\ref{sec:results}.
+\end{figure}
 
-\subsection{Limitations}
-\label{sec:limitations}
+The hyperprior distribution chosen for these tests is a multivariate normal 
+distribution with mean $\vec{\mu}$ equal to the implicit prior $\ndphi^{*}$ and 
+covariance
+\begin{equation}
+\label{eqn:priorcov}
+\Sigma_{k,k'} = q\ \exp[-\frac{e}{2}\ (\bar{z}_{k}-\bar{z}_{k'})^{2}]\ +\ 
+t\delta(k,k')
+\end{equation}
+inspired by one used in Gaussian processes, where $k$ and $k'$ are indices 
+ranging from $1$ to $K$ and $q=1.0$, $e=100.0$, and $t=q\cdot10^{-5}$ are 
+constants chosen to permit draws from this prior distribution to produce shapes 
+similar to that of a true $\tilde{\ndphi}$.  
+We adapt the full log-posterior of \Eq{eqn:final} to the chosen binning of 
+redshift space.
 
-Finally, we explicitly review the assumptions made by this approach.
-\begin{itemize}
-  \item prior $\Pr(\phi)$
-  \item implicit prior $\phi^{*}$ known and shared among all PDFs
-	\item \pzpdf s are accurate posteriors
-	\end{itemize}
+The sampler is initialized with $W=100$ walkers each with a value chosen from a 
+Gaussian distribution of identity covariance around a sample from the 
+hyperprior distribution.
 
-\section{Results}
+\subsection{Intrinsic scatter}
+\label{sec:scatter}
+
+\Fig{fig:pzs-scatter} shows some examples of \pzpdf s generated with only the 
+systematic of intrinsic scatter, at the level of the \lsst\ requirements on the 
+left and twice that on the right.
+One can see that the histogram of redshift estimates is broader than that of 
+true redshifts, and that the effect is substantially more pronounced by just 
+doubling the intrinsic scatter from the level of the \lsst\ requirements.
+
+\begin{figure}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_scatter1.png}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_scatter2.png}
+	\caption{
+		Examples of mock \pzpdf s (colored lines) generated with 
+intrinsic scatter at the \lsst\ requirements (left) and twice the \lsst\ 
+requirements (right), including samples from the probability space of true and 
+observed redshift (black points), \pzpdf s (colored lines), the true redshifts 
+of the example \pzpdf s (black vertical lines).
+		A histogram (gray) of points in each dimension is shown in the 
+respective inset, with the true redshift distribution (blue curve) and implicit 
+prior (red curve).
+	}
+	\label{fig:pzs-scatter}
+\end{figure}
+
+\Fig{fig:results-scatter} shows the \nz\ recovered by \Chippr\ and the 
+alternative approaches.
+As expected, the estimates of \nz\ based on the modes of the \pzpdf s and 
+stacking are broader than the marginalized maximum likelihood estimator from 
+\chippr, with more broadening as the intrinsic scatter increases.
+\Chippr's \mmle\ is robust to intrinsic scatter and is unaffected by increased 
+intrinsic scatter, though the \Chippr\ posterior distribution on the redshift 
+distribution is itself broader for the higher intrinsic scatter case than for 
+the \lsst\ requirements.
+The broadening of the alternative estimators corresponds to a loss of 3-4 times 
+as many nats of information about \nz\ for the \lsst\ requirements relative to 
+the \mmle\ of \Chippr.
+
+\begin{figure}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/results_scatter1.png}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/results_scatter2.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization 
+in dark blue) and the alternative approaches (the stacked estimator in red and 
+the histogram of modes in yellow) on \pzpdf s with intrinsic scatter of the 
+\lsst\ requirements (left) and twice that (right), with the true redshift 
+density (black curve) and implicit prior (gray curve).
+		\Chippr\ is robust to intrinsic scatter, but the alternatives 
+suffer from overly broad \nz\ estimates that worsen with increasing intrinsic 
+scatter.
+	}
+	\label{fig:results-scatter}
+\end{figure}
+
+\subsection{Catastrophic outliers}
+\label{sec:outliers}
+
+As was covered in \Sect{sec:intro}, catastrophic outliers tend to be 
+distributed non-uniformly across the space of observed and true redshift.
+However, the \lsst\ requirements do not specify details for a distribution of 
+outliers to which they were tuned, and it is still instructive to examine the 
+impact of uniform outliers on the inference of \nz, so we begin by addressing 
+uniformly distributed outliers before considering more realistic outlier 
+distributions.
+
+A uniformly distributed population of outliers was simulated by giving every 
+sample in true redshift a $10\%$ chance of having an observed redshift drawn 
+from a uniform distribution rather than the Gaussian about the true redshift.
+Though this results in slightly less than the $10\%$ catastrophic outlier rate, 
+it can be done independently of the definition of the standard deviation so was 
+implemented for demonstrative purposes.
+\Fig{fig:uniform-outliers} shows examples of \pzpdf s from a uniformly 
+distributed outlier population at the level of the \lsst\ requirements (left) 
+as well as the results of \Chippr\ and other \nz\ estimation methods (right).
+The intrinsic scatter of the tests in this section does not increase with 
+redshift as indicated in Table~\ref{intro:tab:tab:lsstsrd}, in order to isolate 
+the effect of outliers, and is instead held at a constant $\sigma_{z} = 0.02$.
+
+\begin{figure}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/single_uout_mega_scatter.p
+ng}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/single_uout_log_estimators
+.png}
+	\caption{
+		Left: Examples of \pzpdf s with a uniform catastrophic outlier 
+population at the level of the \lsst\ requirements, including samples from the 
+probability space of true and observed redshift (black points), \pzpdf s 
+(colored curves), and the true redshifts of the example \pzpdf s (black 
+vertical lines), with marginal histograms (gray) for each dimension with the 
+true redshift distribution (blue curve) and implicit prior (red curve) in the 
+insets.
+		Right: The results of \Chippr\ (samples in light blue, 
+optimization in dark blue) and the alternative approaches (the stacked 
+estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly 
+distributed catastrophic outliers, with the true redshift density (black curve) 
+and implicit prior (gray curve).
+		The presence of the catastrophic outlier population broadens 
+the histogram of modes and stacked estimator of the redshift distribution, but 
+the result of \Chippr\ is unbiased.
+	}
+	\label{fig:uniform-outliers}
+\end{figure}
+
+\Fig{fig:uniform-outliers} shows that at the level of the \lsst\ requirements, 
+the alternative estimators are overly broad, whereas \Chippr's \mmle\ yields an 
+unbiased estimate of \nz.
+Further, the result of stacking is even broader than that of the histogram of 
+modes, corresponding to ten times the information loss of \Chippr's \mmle, 
+making it worse than the most naive reduction of \pzpdf s to point estimates.
+
+When one thinks of the \pzpdf s of catastrophic outliers, however, what comes 
+to mind is multimodal \pzpdf s, wherein reducing \pzpdf s to point estimates to 
+make a standard scatterplot of the true and observed redshifts leads to 
+substantial probability density off the diagonal.
+These coordinated catastrophic outliers may be emulated in the joint 
+probability space of true and estimated redshifts by using a mixture of the 
+unbiased diagonal defined by the intrinsic scatter and an additional Gaussian 
+in one dimension, with constant observed redshift for a template-fitting code 
+and constant true redshift for a machine learning code.
+
+In the case of a catastrophic outlier population like that anticipated of 
+template-fitting codes, $10\%$ of all galaxies have their observed redshift at 
+a particular value unrelated to their true redshift, illustrated in the left 
+panel of \Fig{fig:nonuniform-outliers-data}.
+This case is subject to the same caveat as the uniformly distributed outliers 
+when it comes to the \lsst\ requirement.
+It is less straightforward to emulate catastrophic outliers like those 
+anticipated of a machine learning code, those that are truly multimodal.
+The testing conditions here, illustrated in the right panel of 
+\Fig{fig:nonuniform-outliers-data}, gives $10\%$ of galaxies at the redshift 
+affected by outliers an observed redshift that is uniformly distributed 
+relative to the true redshift, meaning that far fewer than $10\%$ of all 
+galaxies in the sample are catastrophic outliers.
+
+\begin{figure}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_eout_mega_scatter.p
+ng}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_rout_mega_scatter.p
+ng}
+	\caption{
+		Examples of \pzpdf s with a catastrophic outlier population 
+like that seen in template-fitting \pzpdf\ codes (left) and machine learning 
+\pzpdf\ codes (right), including samples from the probability space of true and 
+observed redshift (black points), \pzpdf s (colored curves), and the true 
+redshifts of the example \pzpdf s (black vertical lines), with marginal 
+histograms (gray) for each dimension with the true redshift distribution (blue 
+curve) and implicit prior (red curve) in the insets.
+	}
+	\label{fig:nonuniform-outliers-data}
+\end{figure}
+
+The results of \Chippr\ and the alternative estimators of \nz\ are presented in 
+\Fig{fig:nonuniform-outliers-results}.
+The most striking feature is that the histogram of modes is highly sensitive to 
+both outlier populations, producing a severe overestimate in the case of an 
+outlier population like those seen in template-fitting codes and a severe 
+underestimate in the case of an outlier population like those seen in machine 
+learning codes, corresponding to a twenty-fold loss of information compared to 
+the \Chippr\ \mmle\ in both cases.
+The effect on the stacked estimator of \nz\ is more subtle though still 
+concerning.
+In the case of outliers like those resulting from template-fitting, the stacked 
+estimator is overly broad even without realistic intrinsic scatter, resulting 
+in ten times the information loss compared to the \Chippr\ \mmle, and in the 
+case of outliers like those resulting from machine learning, the stacked 
+estimator features an overestimate at the redshift affected by the outlier 
+population, resulting in about five times the information loss as the \Chippr\ 
+\mmle.
+The \Chippr\ \mmle, however, appears unbiased and withstands these effects, and 
+the breadth of the distribution of samples of \nz\ is invariant.
+
+\begin{figure}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_eout_log_estimators
+.png}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_rout_log_estimators
+.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization 
+in dark blue) and the alternative approaches (the stacked estimator in red, the 
+histogram of modes in yellow) on \pzpdf s with catastrophic outliers like those 
+seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ 
+codes (right) to the \lsst\ requirements, with the true redshift density (black 
+curve) and implicit prior (gray curve).
+		Though the histogram of modes is most sensitive to a 
+catastrophic outlier population, the stacked estimator also overestimates \nz\ 
+under (machine learning-like outliers) and beyond (template fitting-like 
+outliers).
+	}
+	\label{fig:nonuniform-outliers-results}
+\end{figure}
+
+\subsection{Systematic bias}
+\label{sec:bias}
+
+Systematic bias in \pz\ point estimates is a concern for \lsst's cosmology 
+results, for the same reasons explored in \citet{hoyle_dark_2017}.
+However, in the context of \pzpdf s, the notion of redshift bias is a form of 
+model misspecification.
+Consider that if bias were included in the framework of 
+Figure~\ref{intro:fig:fig:pedagogical_scatter}, a simple linear transformation 
+of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + 
+z_{\mathrm{phot}})$ would eliminate the bias.
+Regardless, for completeness, a test at ten times the bias of the \lsst\ 
+requirements, with no redshift-dependent intrinsic scatter nor catastrophic 
+outliers, is provided in \Fig{fig:bias}.
+
+\begin{figure}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_neghibias_mega_scat
+ter.png}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_neghibias_log_estim
+ators.png}
+	\caption{
+		Left: Examples of \pzpdf s with ten times the bias of the 
+\lsst\ requirements, including samples from the probability space of true and 
+observed redshift (black points), \pzpdf s (colored curves), and the true 
+redshifts of the example \pzpdf s (black vertical lines), with marginal 
+histograms (gray) for each dimension with the true redshift distribution (blue 
+curve) and implicit prior (red curve) in the insets.
+		Right: The results of \Chippr\ (samples in light blue, 
+optimization in dark blue) and the alternative approaches (the stacked 
+estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly 
+distributed catastrophic outliers, with the true redshift density (black curve) 
+and implicit prior (gray curve).
+		The impact of bias at even ten times the level of the \lsst\ 
+requirements is almost imperceptible on all estimators, though the \Chippr\ 
+\mmle\ minimizes the information loss regardless.
+	}
+	\label{fig:bias}
+\end{figure}
+
+As expected based on consistency with the forward model, \Chippr\ is completely 
+resistant to bias, and the alternative estimators are only weakly affected, 
+with information loss two and four times greater than that of the \Chippr\ 
+\mmle\ for the histogram of modes and stacked estimator respectively.
+
+\subsection{Implicit prior}
+\label{sec:interim}
+
+\chippr\ can handle any implicit prior with support over the redshift range 
+where \nz\ is defined, but some archetypes of implicit prior are more likely to 
+be encountered in the wilds of \pzpdf\ codes.
+Ideally, an uninformative implicit prior would be used, although it may be 
+complicated to compute from the covariances of the raw data.  
+Template-fitting codes have an explicit prior input formed by redshifting a 
+small number of templates, leading to a highly nonuniform but 
+physically-motivated interim prior.
+Machine learning approaches tend to be trained on previously observed data sets 
+that are biased towards low redshift, which biases the implicit prior towards 
+low redshift.
+Some efforts have been made to modify an observationally informed implicit 
+prior so that it is more representative of the photometric data for which 
+redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal 
+to the true \nz, it will propagate to the results of traditional \nz\ 
+estimation methods.
+
+\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring 
+implicit prior emulating that of a machine learning approach to \pz\ estimation 
+(left panel) and a more complex interim prior emulating that of a 
+template-fitting \pz\ method (right panel).
+One can see that the \pzpdf s take different shapes from one another even 
+though the marginal histograms of the points are identical.
+The machine learning-like implicit prior has been modified to have nonzero 
+value at high redshift because the implicit prior must be strictly positive 
+definite for the \Chippr\ model to be valid.
+
+\begin{figure}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_trpr.png}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_tmpr.png}
+	\caption{
+		Examples of mock \pzpdf s (colored lines) generated with a 
+machine learning-like implicit prior (left) and a template-fitting-like 
+implicit prior (right), including samples from the probability space of true 
+and observed redshift (black points), \pzpdf s (colored lines), the true 
+redshifts of the example \pzpdf s (black vertical lines).
+		A histogram (gray) of points in each dimension is shown in the 
+respective inset, with the true redshift distribution (blue curve) and implicit 
+prior (red curve).
+	}
+	\label{fig:pzs-priors}
+\end{figure}
+
+\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional 
+methods on \pzpdf s generated with nontrivial implicit priors.
+In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift 
+distribution, and the distribution of \nz\ parameter values reflects higher 
+uncertainty where the implicit prior undergoes large changes in derivative.
+The alternatives, on the other hand, are biased by the implicit prior except 
+where it is flat, in the case of high redshifts for the machine learning-like 
+implicit prior, resulting in over $1,000$ times the information loss on \nz\ 
+for the machine learning-like implicit prior and some $5-20$ times the 
+information loss for the template fitting-like implicit prior, relative to the 
+\Chippr\ \mmle.
+
+\begin{figure}
+	\includegraphics[width=0.49\textwidth]{figures/chippr/results_trpr.png}
+	\includegraphics[width=0.49\textwidth]{figures/chippr/results_tmpr.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization 
+in dark blue) and the alternative approaches (the stacked estimator in red and 
+the histogram of modes in yellow) on \pzpdf s with an implicit prior like that 
+of machine learning \pzpdf\ approaches (left) and an implicit prior like that 
+of template-fitting \pzpdf\ codes (right), with the true redshift density 
+(black curve) and implicit prior (gray curve).
+		\Chippr\ is robust to a nontrivial implicit prior, but the 
+alternatives are biased toward the implicit prior.
+	}
+	\label{fig:results-priors}
+\end{figure}
+
+The main implication of the response of \nz\ estimates to a nontrivial implicit 
+prior is that the implicit prior must be accounted for when using \pzpdf\ 
+catalogs.
+
+\section{Discussion}
 \label{sec:results}
 
-\textcolor{blue}{Q: How does the result of \chippr\ compare to established estimators in terms of the accuracy of $n(z)$?\\
-A: \chippr\ yields the best possible $n(z)$, conditional on the accuracy of the photo-$z$ PDFs used.}
+The experiments of \Sect{sec:alldata} isolate the potential sources of error in 
+\nz\ estimation one at a time.
+Now, we stress-test \Chippr\ by investigating two realistically complex cases, 
+one in which the \nz\ estimates are made tomographically as in a modern 
+cosmological analysis (\Sect{sec:lsstdemo}) and one in which the \nz\ 
+estimators are not provided with the same implicit prior used to generate the 
+\pzpdf\ catalog (\Sect{sec:violations}).
 
@@
 
-\textcolor{blue}{LSST specs: biased to 0.003, 10\% uniform outliers, z-dependent scatter of 0.05}
+It is of interest to explore the impact of incorrectly estimated \nz\ on the 
+cosmological inference to answer the question of how wrong we will be in our 
+understanding of the universe if we incorrectly constrain \nz.
+To test the impact of these uncertainties, I simulated mock data with all three 
+effects with which \lsst\ is concerned at the levels of 
+Table~\ref{intro:tab:tab:lsstsrd} and propagated the results of \Chippr\ and 
+the other estimators to a Fisher matrix forecast using \cosmolike\ 
+\citep{krause_cosmolike_2017}, a publicly available cosmological forecasting 
+code.
+Though redshift tomography is non-physical, as redshift is a continuous random 
+variable, and binning in estimated redshift introduces poorly understood 
+systematic error, I perform this analysis as an example of how it affects the 
+current standard in how cosmological parameters are constrained by galaxy 
+surveys, rather than how I think they ought to be constrained.
 
@@
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{fig/lsst_scatter.png}
-		\includegraphics[width=0.45\textwidth]{fig/lsst_log_estimators.png}
-		\caption{Using LSST numbers}
-		\label{fig:lsstdemo}
+
+\includegraphics[width=0.74\textwidth]{figures/chippr/cosmolike_inputs.png}
+		\caption{
+			The \lsst-like tomographic binning and true redshift 
+distribution, where the truth (solid) is a PDF evaluated on a fine grid of 
+$350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn 
+(dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced 
+bins, for four different tomographic bins (colors).
+		}
+		\label{fig:tomobins}
+	\end{center}
+\end{figure*}
+
+I considered a set of tomographically binned \nz\ and cosmological parameter 
+covariance matrices used for \desc\ forecasting, for which the true \nz\ in 
+each pre-defined bin is already provided in the form of an evaluation of a 
+function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+First, I binned them down to a piecewise constant parameterization with a 
+manageable $35$ parameters for \chippr's sampling capabilities.
+Next, I drew $10^{4}$ true redshifts from the binned true \nz\ for each 
+tomographic bin.
+The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
+I emulated \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in 
+each bin using the procedure of \Fig{fig:flowchart} with all three effects of 
+Table~\ref{intro:tab:tab:lsstsrd} at their given levels.
+Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
+
+\begin{figure*}
+	\begin{center}
+
+\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.
+png}
+
+\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.
+png}		
+\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.
+png}
+
+\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.
+png}
+		\caption{As in \Fig{fig:mega_scatter}, with a different 
+tomographic bin in each panel and the three effects of intrinsic scatter, 
+uniformly distributed catastrophic outliers, and bias at the levels of the 
+\lsst\ SRD, given in Table~\ref{intro:tab:tab:lsstsrd}.
+		}
+		\label{fig:per-bin-scatter}
+	\end{center}
+\end{figure*}
+
+I then made a point estimate of \nz\ using \Chippr\ as well as the alternative 
+methods on the \pzpdf\ catalog for each tomographic bin, shown in 
+\Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from 
+a single \nz\ result, rather than samples from the full posterior probability 
+density of possible \nz.
+Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability 
+units, unlike all other plots in this paper, to better show the behavior at low 
+probability.
+The excessive breadth of the alternative estimators can be seen quite plainly.
+
+\begin{figure*}
+	\begin{center}
+
+\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimator
+s.png}
+
+\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimator
+s.png}		
+\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimator
+s.png}
+
+\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimator
+s.png}
+		\caption{
+			The \chippr-derived and other estimators of \nz\ in 
+each tomographic bin, with the true \nz\ (black), the implicit prior (gray), 
+stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ 
+(blue).
+			The result of stacking is far too broad for \lsst-like 
+\pzpdf s, even moreso than the simplistic histogram of modes.
+		}
+		\label{fig:per-bin-ests}
+	\end{center}
+\end{figure*}
+
+I then used the different estimators of \nz\ in a cosmological forecasting 
+procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, 
+$w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
+Though there are also slight differences in the angle of the error ellipses, 
+most of the differences are due to broadening of the contours under the 
+alternative estimators relative to \Chippr, which are almost indistinguishable 
+from those derived by using the true redshift distribution in each bin.
+The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all 
+parameters except $\Omega_{b}$ and $H_{0}$.
+Stacking, however, outperforms the histogram of modes for all parameters except 
+$\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
+Though the true values of the parameters themselves were not accessible with 
+the Fisher matrix-based framework, I calculated the seven-dimensional KLD for 
+the three \nz\ estimators relative to the constraints derived from the true 
+\nz, showing that \Chippr\ preserves information $200-800$ times better than 
+the alternatives, with the histogram of modes doing about four times better 
+than stacking.
+
+\begin{figure*}
+	\begin{center}
+
+\includegraphics[width=0.99\textwidth]{figures/chippr/final_plot.png}
+		\caption{
+			The result of propagating the estimators of \nz\ by 
+stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true 
+\nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
+			For all parameters considered, \Chippr\ yields contours 
+no broader than those corresponding to the true \nz, whereas for most 
+parameters, stacking and the histogram of modes yield broader contours.
+		}
+		\label{fig:cornerplot}
 	\end{center}
@@
 
-\textcolor{blue}{mischaracterized interim prior, with idealized or realistic data?}
+In this test, without tomographic binning, the \pzip s are made to the \lsst\ 
+requirements but the implicit prior used for the inference is not the same as 
+the implicit prior used for generating the data.
+\Pzpdf\ codes do not generally provide their implicit prior, with the exception 
+of some template-fitting techniques for which it is a known input.
+If we naively used the \pzpdf\ catalog produced by a generic machine learning 
+code and assumed a flat implicit prior, we would observe the contents of 
+\Fig{fig:mischaracterized}.
 
@@
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{fig/wrong_scatter.png}
-		\includegraphics[width=0.45\textwidth]{fig/wrong_log_estimators.png}
-		\caption{this is for wrong implicit prior, not bias, haven't done that one yet}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/misspecified_mega_scatter.
+png}
+
+\includegraphics[width=0.49\textwidth]{figures/chippr/misspecified_log_estimator
+s.png}
+		\caption{
+			Left: Examples of \pzpdf s with systematics at the 
+level of the \lsst\ requirements with a machine learning-like implicit prior, 
+including samples from the probability space of true and observed redshift 
+(black points), \pzpdf s (colored curves), and the true redshifts of the 
+example \pzpdf s (black vertical lines), with marginal histograms (gray) for 
+each dimension with the true redshift distribution (blue curve) and implicit 
+prior (red curve) in the insets.
+			Right: The results of \Chippr\ (samples in light blue, 
+optimization in dark blue) and the alternative approaches (the stacked 
+estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly 
+distributed catastrophic outliers, with the true redshift density (black curve) 
+and the uniform implicit prior given to \chippr\ (gray curve).
+			When the incorrect implicit prior is provided to 
+\chippr, even Bayesian inference cannot recover the true \nz.
+		}
 		\label{fig:mischaracterized}
@@
 
-\section{Discussion}
-\label{sec:discussion}
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.99\textwidth]{fig/final_plot.png}
-		\caption{propagation to cosmology}
-		\label{fig:cornerplot}
-	\end{center}
-\end{figure*}
+The results of using a mischaracterized implicit prior are disastrous, causing 
+every estimator, including \Chippr, to be strongly biased.
+The stacked estimator and histogram of modes don't make use of the implicit 
+prior so do no worse than when the implicit prior is accurately provided, but 
+\Chippr\ is sensitive to prior misspecification, which violates the model upon 
+which it is based.
+It is thus crucial that \pzpdf\ methods always characterize and provide the 
+implicit prior.
 
 \section{Conclusion}
-\label{sec:conclusion}
+\label{sec:con}
+
+This study derives and demonstrates a mathematically consistent inference of a 
+one-point statistic, the redshift density function \nz, based on an arbitrary 
+catalog of \pzpdf s.  
+The fully Bayesian method, based in the fundamental laws of probability, begins 
+with a probabilistic graphical model corresponding to equations for the full 
+posterior distribution over the parameters for \nz.  
+The method is validated on mock data and tested in the regime of \lsst\ with 
+promising results, outperforming the traditional stacking estimator at the 
+level of \nz\ as well as in terms of constraining power on the cosmological 
+parameters.
+Not only is this the only mathematically correct approach to the problem, it 
+also recovers the true parameter values better than popular alternatives, as 
+measured by the loss of information in \nz\ and the size of error ellipses in 
+the space of cosmological parameters.
+
+The following conclusions and recommendations can be made with confidence:
+
+\begin{enumerate}
+	\item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are 
+good point estimators of \nz, whereas the histogram of modes is very sensitive 
+to outliers and the stacked estimator is always excessively broad.
+	\item The error bars on the posterior distribution over \nz\ 
+hyperparameters are interpretable and arise naturally under \Chippr, unlike 
+those that may be assumed for the conventional point estimators.
+	\item When the implicit prior is known to be a poor match to the data, 
+only the results of \Chippr\ are satisfactory estimators of the redshift 
+distribution function because they are the only methods that can account for 
+the bias induced on the \pzpdf\ catalog by the method that produces it; this is 
+the most compelling case for the sampler because of the ubiquity of 
+inappropriate interim priors.
+\end{enumerate}
+
+By showing that \Chippr\ is effective in recovering the true redshift 
+distribution function and posterior distributions on its parameters from 
+catalogs of \pzpdf s, this work supports the production of \pzpdf s by upcoming 
+photometric surveys such as \lsst\ to enable more accurate inference of the 
+cosmological parameters.  
+We discourage researchers from co-adding \pzpdf s or converting them into point 
+estimates of redshift and instead recommend the use of Bayesian probability to 
+guide the usage of \pzpdf s.  
+We emphasize to those who produce \pzpdf s from data that it is essential to 
+release the implicit prior used in generating this data product in order for 
+proper inference to be conducted by consumers of this information.
+
+The technique herein developed is applicable with minimal modification to other 
+one-point statistics of redshift to which we will apply this method in the 
+future, such as the redshift-dependent luminosity function and weak lensing 
+mean distance ratio.  
+Future work will also include the extension of this fully probabilistic 
+approach to higher-order statistics of redshift such as the two-point 
+correlation function.
 
@@
 
-\section{Mathematical derivation}
-\label{app:math}
-
-The derivation is most straightforward in the space of log probabilities.
-
-According to Bayes' rule, the posterior distribution of interest is
-\begin{align}
-\label{eq:hyper_bayes}
-\log[p(\phi | \{d_{i}\})] &\propto \log[p(\{d_{i}\} | \phi)] + \log[p(\phi)] ,
-\end{align}
-where $p(\phi)$ is the \textit{hyperprior} probability distribution over possible values of the hyperparameters $\phi$ defining the redshift distribution.  
-The hyperprior is a choice made by those performing the inference and is not to be confused with the implicit prior parameters $\phi^{*}$, which are a specific instance of $\phi$.
-
-Focusing on the likelihood term $p(\{d_{i}\} | \phi)$, we employ the power of the hierarchical model through \textit{marginalization} of the unobserved redshift parameters by integrating over them to obtain
-\begin{align}
-\label{eq:marginalize}
-\begin{split}
-\log[p(\{d_{i}\} | \phi)] = & \log\left[\int\exp\left[\log[p(\{d_{i}\} | \{z_{i}\})]\right.\right. \\
-& \left.\left.+ \log[p(\{z_{i}\} | \phi)]\right]\ d\{z_{i}\}\right] .
-\end{split}
-\end{align}
-The redshifts $\{z_{i}\}$ are statistically independent of each other, and each galaxy's photometry depends solely on its own redshift, so we may invoke statistical independence to get
-\begin{align}
-\label{eq:independence_hyper}
-\log[p(\{z_{i}\} |\phi)] &= \sum_{i}^{N} \log[p(z_{i} | \phi)]
-\end{align}
-and
-\begin{align}
-\label{eq:independence}
-\log[p(\{d_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(d_{i} | z_{i})] .
-\end{align}
-Having chosen the functional form mapping $\phi$ to $n(z)$, the terms on the righthand side of Eq. \ref{eq:independence_hyper} are known.  
-However, the terms on the righthand side of Eq. \ref{eq:independence} are likelihoods, and the photo-$z$ PDFs we have are implicit posteriors.
-
-To transform the implicit posteriors we have into the likelihoods we need, we start with the vacuously true statement
-\begin{align}
-\begin{split}
-\label{eq:unity}
-\log[p(d_{i} | z_{i})] =& \log[p(d_{i} | z_{i})] + \log[p(z_{i} | d_{i}, \phi^{*})]\\
-& - \log[p(z_{i} | \vec{d}_{i}, \phi^{*})] .
-\end{split}
-\end{align}
-We expand the last term using Bayes' Rule to get
-\begin{align}
-\begin{split}
-\label{eq:bayes}
-\log[p(d_{i} | z_{i})] =& \log[p(d_{i} | z_{i})] + \log[p(z_{i} | d_{i}, \phi^{*})]\\
-& - \log[p(d_{i} | z_{i}, \phi^{*})]\\
-& + \log[p(d_{i} | \phi^{*})] - \log[p(z_{i} | \phi^{*})] .
-\end{split}
-\end{align}
-Fig. \ref{fig:pgm} has no direct connection between $d_{i}$ and $\phi$, so we can break up $\log[p(d_{i} | z_{i}, \phi^{*})]$ as follows:
-\begin{align}
-\begin{split}
-\label{eq:split}
-\log[p(d_{i} | z_{i})] =& \log[p(d_{i} | z_{i})] + \log[p(z_{i} | d_{i}, \phi^{*})]\\
-& - \log[p(d_{i} | z_{i})] - \log[p(d_{i} | \phi^{*})]\\
-& + \log[p(d_{i} | \phi^{*})] - \log[p(z_{i} | \phi^{*})] .
-\end{split}
-\end{align}
-Canceling terms, we find
-\begin{align}
-\label{eq:cancel}
-\log[p(d_{i} | z_{i})] =& \log[p(z_{i} | d_{i}, \phi^{*})] - \log[p(z_{i} | \phi^{*})].
-\end{align}
-
-Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, \ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to arrive at 
-\begin{align}
-\begin{split}
-\label{eq:final}
-\log[p(\phi | \{d_{i}\})] \propto& \log[p(\phi)] + \log\left[\int \exp\left[\sum_{i}^{N} \right.\right.\\
-& \left.\left.\left(\log[p(z_{i} | \phi)] - \log[p(z_{i} | \phi^{*})]\right.\right.\right.\\
-& \left.\left.\left. + \log[p(z_{i} | d_{i}, \phi^{*})]
-\right)\right] d\{z_{i}\}\right].
-\end{split}
-\end{align}
-Thus, if we have a catalog of photo-$z$ implicit posteriors $\{p(z_{i} | d_{i}, \phi^{*})\}$, the interim prior parameters $\phi^{*}$, and a prior distribution $p(\phi)$, we may find the posterior $\log[p(\phi | \{d_{i}\})]$ of the redshift distribution $n(z)$.
-Furthermore, since all galaxies share the same range of possible redshifts, the unknown $z_{i}$ may be shortened to $z$.
-
-\section{Mock data generation}
-\label{app:data}
-
-\textul{Plots: flow chart of forward model}
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.75\textwidth]{fig/flowchart.png}
-		\caption{In appendix?}
-		\label{fig:flowchart}
-	\end{center}
-\end{figure*}
-
 \end{appendices}
@@
 \begin{acknowledgements}
-AIM thanks Elisabeth Krause for assistance with the \texttt{CosmoLike} code, Phil Marshall for instruction in software carpentry, Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for advice on debugging, and Boris Leistedt for helpful comments provided in the preparation of this paper.
-This work was completed under the generous nutritional support of the Center for Computational Astrophysics.
+AIM was supported by National Science Foundation grant AST-1517237 and the U.S. 
+Department of Energy, Office of Science, Office of Workforce Development for 
+Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) 
+program, administered by the Oak Ridge Institute for Science and Education for 
+the DOE under contract number DE‐SC0014664.
+The authors thank Phil Marshall for advice on relevant examples, Elisabeth 
+Krause for assistance with the \cosmolike code, Mohammadjavad Vakili for 
+statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt 
+for other helpful comments in the development of \Chippr.
+This work was completed with generous nutritional support from the Center for 
+Computational Astrophysics.
 \end{acknowledgements}
@@
 \bibliographystyle{apj}
+\bibliography{thesis}
 

### 2019-05-17 chippr/catalog_plots.py

@@ -141,6 +141,7 @@ def plot_mega_scatter(zs, pfs, z_grid, grid_ends, truth=None, plot_loc='', prepe
     scatplot.set_ylim([min(grid_ends)-limval, max(grid_ends)+limval])
     scatplot.set_xlabel(r'$z_{spec}$', fontsize=20)
     scatplot.set_ylabel(r'$z_{phot}$', fontsize=20)
+    scatplot.text(0.5, 3., r'mock $p(z\mid \mathrm{``data"})$ data', rotation=0, size=16)
 
     # scatplot.set_aspect(1.)
     divider = make_axes_locatable(scatplot)

### 2019-05-17 chippr/log_z_dens_plots.py

@@ -500,6 +500,7 @@ def plot_estimators(info, plot_dir, log=True, prepend='', metrics=True, mini=Tru
 
     # sps_log.legend(handles=color_plots[:-1], fontsize='x-small', loc='lower center', frameon=False)
     sps_log.legend(fontsize='small', loc='upper right', frameon=False)
+    sps_log.text(0.5, 0., r'inferred $n(z)$ results', rotation=0, size=16)
     f.subplots_adjust(hspace=0, wspace=0)
     f.savefig(os.path.join(plot_dir, prepend+'estimators.png'), bbox_inches='tight', pad_inches = 0, dpi=d.dpi)
     print(info['stats'])

### 2019-05-17 research/paper/cleaner.sh

@@ -1,7 +1,7 @@
-infile='draft-descope.tex'
+infile='from_thesis.tex'
 outfile='draft.tex'
 
-sed -e "s/{\\textbackslash}'\\\\{i\\\\}/\\'{i}/g" references.bib | grep -E -v '^\W*(urldate|abstract|file|keywords)' > references.bib.tmp
-mv references.bib.tmp references.bib
+sed -e "s/{\\textbackslash}'\\\\{i\\\\}/\\'{i}/g" from_thesis.bib | grep -E -v '^\W*(doi|url|urldate|abstract|file|keywords|annote|note)' > from_thesis.bib.tmp
+mv from_thesis.bib.tmp thesis.bib
 
-grep -E -v '^(%|[[:blank:]]*%|\\COMMENT)' "$infile" | fold -w80 -s > "$outfile"
+#grep -E -v '^(%|[[:blank:]]*%|\\COMMENT|\includegraphics*)' "$infile" | fold -w80 -s > "$outfile"

### 2019-05-20 research/paper/draft.tex
updating files

@@
+\RequirePackage[]{lineno}
+
 \documentclass[iop]{emulateapj}
@@
 \usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{lineno}
 
@@
 \usepackage{soul}
-\usepackage[title]{appendix}
+%\usepackage[title]{appendix}
 
@@
 \newcommand{\Fig}[1]{Figure~\ref{#1}}
+\newcommand{\Tab}[1]{Table~\ref{#1}}
 \newcommand{\project}[1]{{\textsc{#1}}~}
@@
 \newcommand{\data}{\ensuremath{\vec{d}}}
+% could change to bold
 \newcommand{\like}{\mathscr{L}}
 \newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
+% could change to Prob or Pr 
 \newcommand{\expect}[1]{\left<#1\right>}
@@
 \newcommand{\pzpdf}{\pz PDF}
+% could change to posterior
 \newcommand{\Pzpdf}{\Pz PDF}
+% could change to posterior
 \newcommand{\zpdf}{redshift posterior}
+% could change to implicit posterior
 \newcommand{\pzip}{\pz implicit posterior}
@@
 \newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
-\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
+\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
+% could change to \vec
 \newcommand{\ndphi}{\bvec{\phi}}
 \newcommand{\mmle}{marginalized maximum likelihood estimate}
+% really marginalized maximum a posteriori estimate, may still change this
 
 \begin{document}
-
-\title{How to obtain the redshift distribution from probabilistic redshift 
-estimates}
-
-\author{Alex Malz\altaffilmark{1}}
+\linenumbers
+
+\title{How to obtain the redshift distribution from probabilistic redshift estimates}
+
+\author{Alex I. Malz\altaffilmark{1}}
 \author{David W. Hogg\altaffilmark{1,2,3,4}}
@@
 
-\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
-Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
-\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
-Avenue, 7th floor, New York, NY 10010, USA}
-\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
-7th floor, New York, NY 10003, USA}
-\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
-Heidelberg, Germany}
+\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
+\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
+\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
+\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 
 \begin{abstract}
-A trustworthy estimate of the redshift distribution \nz\ is crucial for using 
-galaxy catalogs to study cosmology via weak gravitational lensing and 
-large-scale structure.
-Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing 
-surveys are expected to be inaccessible, making photometric redshifts 
-(photo-$z$s) the next-best alternative.
-The nontrivial systematics affecting photo-$z$ estimation have motivated the 
-weak-lensing community to favor photo-$z$ probability density functions (PDFs) 
-as a more comprehensive alternative to photo-$z$ point estimates.
-However, analytic methods for utilizing these new data products in cosmological 
-inference are still evolving.
-The ubiquitous methodology known as stacking produces a systematically biased 
-estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very 
-regime where photo-$z$ PDFs are most necessary.
-We introduce a mathematically rigorous probabilistic graphical model (PGM) of 
-hierarchical inference of $n(z)$, which is provably the only self-consistent 
-way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
-The novel Cosmological Hierarchical Inference with Probabilistic Photometric 
-Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by 
-correctly propagating the redshift uncertainty information beyond the best-fit 
-estimator produced by traditional procedures.
-We conclude by propagating these effects to constraints in the space of 
-cosmological parameters.
+A trustworthy estimate of the redshift distribution \nz\ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
+Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
+The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
+However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
+The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
+We introduce a mathematically rigorous probabilistic graphical model (PGM) of $n(z)$ for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
+Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator produced by traditional procedures.
+The CHIPPR approach is applicable to any one-point statistic of any random variable.
+We conclude by propagating these effects to constraints in the space of cosmological parameters, .
+% be clear that rigor comes at a cost, chippr makes explicit the requirements that the pzpdfs be posteriors under a known prior (which would be required by any justifiable method), be honest that we're not ready
 \end{abstract}
 
-\keywords{cosmology: cosmological parameters --- galaxies: statistics --- 
-gravitational lensing: weak --- methods: data analysis --- methods: statistical}
+\keywords{cosmology: cosmological parameters --- galaxies: statistics --- gravitational lensing: weak --- methods: data analysis --- methods: statistical}
 
 \maketitle
+
+\aim{Some of the figures refer to left/right but are shown as top/bottom \textemdash I'm still working through a LaTeX bug.}
+\aim{The language is very thesis-centric (``me, me, me''), but I'll change it to the standard ``we'' style soon.}
 
@@
 
-\aim{TODO: Pull in content from global intro to thesis.}
-
-Though their potential to improve estimates of physical parameters is 
-tremendous, \pzpdf s have been applied only to a limited extent.  
-They have been used to form selection criteria of samples from galaxy surveys 
-without propagation 
-through the calculations of physical parameters 
-\citep{VanBreukelen2009,Viironen2015}.  
-Probability cuts on Bayesian quantities are not uncommon \citep{Leung2015, 
-DiPompeo2015a}, but that procedure does not fully take advantage of all 
-information contained in a probability distribution for parameter inference.
-
-Despite the growing prevalence of \pzpdf\ production, no implementation of 
-inference using \pzpdf s has yet been presented with a mathematically 
-consistent methodology.  
-I present and validate a hierarchical Bayesian technique for the use of \pzpdf\ 
-in inference of arbitrary statistics relevant to cosmology, large-scale 
-structure, and galaxy evolution.  
-For simplicity, we consider only one-point statistics, though future work will 
-extend this methodology to higher-order statistics.
-
-The \textit{redshift distribution function \Nz}, or, almost interchangably, its 
-normalized cousin the \textit{redshift density function \nz}, serves as an 
-ideal statistic upon which to demonstrate this novel approach.  
-\Nz\ is necessary for calculations of two-point correlation functions of weak 
-gravitational lensing and counting statistics that are used to probe dark 
-energy \citep{Masters2015}.  
-The observed \Nz\ also been used to validate survey selection functions 
-usnfenjed in generation of realistic, multi-purpose mock catalogs 
-\citep{Norberg2002}.  
-Additionally, \Nz\ has been the subject of inference using \pzpdf s before 
-\citep{sheldon_photometric_2012, Hildebrandt2012, Kelly2014, Benjamin2013, 
-bonnett_using_2015, Viironen2015, Asorey2016, leistedt_hierarchical_2016}, so 
-comparisons to the literature may easily be made.
-
-The prevailing way to estimate \nz\ for a sample of $\ntot$ galaxies $i$ is to 
-``stack'' \pzpdf s according to Equation~\ref{intro:eq:eqn:stack}.
-The stacked estimator of the redshift density function is effectively an 
-average of the \pzpdf\ catalog and, as I will show in this paper, not in 
-general mathematically valid.
-Stacking is nonetheless considered the preferred method for obtaining \Nz\ from 
-a \pzpdf\ catalog \citep{sheldon_photometric_2012, Kelly2014, Benjamin2013, 
-bonnett_using_2015, Viironen2015, Asorey2016}.  
-However, it must be noted here that Equation~\ref{intro:eq:eqn:stack} is not in 
-general mathematically valid.  
-(See \citet{Hogg2012} for a complete discussion.)
-
-I aim to develop a clear methodology guiding the use of \pzpdf s in inference 
-so they may be utilized effectively by the cosmology community.
-Though others have approached the problem before 
-\citep{leistedt_hierarchical_2016}, the method presented here differs in that 
-it makes use of any existing catalog of \pzpdf s, rather than requiring a 
-simultaneous derivation of the \pzpdf s and the redshift distribution, making 
-it preferable to ongoing surveys that may be resistant to restructuring their 
-analysis pipelines.
-
-In Section~\ref{sec:meth}, I present the \Chippr\ model and \chippr\ 
-implementation for characterizing the full posterior probability landscape of 
-\Nz\ using \pzpdf s. 
-In Section~\ref{sec:alldata}, I describe the experimental design for testing 
-the fully probabilistic approach to mock and real datasets, the results of 
-which are found in Section~\ref{sec:results}.
-In Section~\ref{sec:results}, I stress-test the \Chippr\ model against stacking 
-and its other competitors in the context of cosmology.
+\aim{TODO: Pull in content from global intro of thesis.}
+
+Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
+They have been used to form selection criteria of samples from galaxy surveys without propagation 
+through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
+Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.
+
+Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
+I present and validate a hierarchical Bayesian technique for the use of \pzpdf\ in inference of arbitrary statistics relevant to cosmology, large-scale structure, and galaxy evolution.  
+For simplicity, we consider only one-point statistics, though future work will extend this methodology to higher-order statistics.
+
+The \textit{redshift distribution function \Nz}, or, almost interchangably, its normalized cousin the \textit{redshift density function \nz}, serves as an ideal statistic upon which to demonstrate this novel approach.
+\Nz\ is necessary for calculations of two-point correlation functions of weak gravitational lensing and counting statistics that are used to probe dark energy \citep{masters_mapping_2015}.  
+The observed \Nz\ also been used to validate survey selection functions usnfenjed in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.  
+Additionally, \Nz\ has been the subject of inference using \pzpdf s before \citep{sheldon_photometric_2012, hildebrandt_cfhtlens:_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016, leistedt_hierarchical_2016}, so comparisons to the literature may easily be made.
+
+The prevailing way to estimate \nz\ for a sample of $\ntot$ galaxies $i$ is to ``stack'' \pzpdf s according to \aim{fix reference}.
+% Equation~\ref{intro:eq:eqn:stack}.
+The stacked estimator of the redshift density function is effectively an average of the \pzpdf\ catalog and, as I will show in this paper, not in general mathematically valid.
+Stacking is nonetheless considered the preferred method for obtaining \Nz\ from a \pzpdf\ catalog \citep{sheldon_photometric_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016}.
+However, it must be noted here that \aim{fix reference}
+%Equation~\ref{intro:eq:eqn:stack} 
+is not in general mathematically valid.  
+(See \citet{hogg_data_2012} for a complete discussion.)
+
+%\aim{Motivate how well we need to know \Nz\ and why it matters.
+%	New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
+%	Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}
+
+I aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
+Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys that may be resistant to restructuring their analysis pipelines.
+
+In Section~\ref{sec:meth}, I present the \Chippr\ model and \chippr\ implementation for characterizing the full posterior probability landscape of \Nz\ using \pzpdf s. 
+In Section~\ref{sec:alldata}, I describe the experimental design for testing the fully probabilistic approach to mock and real datasets, the results of which are found in Section~\ref{sec:results}.
+In Section~\ref{sec:results}, I stress-test the \Chippr\ model against stacking and its other competitors in the context of cosmology.
 
@@
 
-Consider a survey of $J$ galaxies $j$, each with photometric data $\data_{j}$; 
-thus the entire survey over some solid angle produces the ensemble of 
-photometric magnitudes (or colors) and their associated observational errors 
-$\{\data_{j}\}$.  
-Each galaxy $j$ has a redshift $z_{j}$ that we would like to learn; redshift is 
-a parameter in this case.  
-The distribution of the ensemble of redshifts $\{z_{j}\}$ may be described by 
-the hyperparameters defining the redshift distribution function \nz\ that we 
-would like to quantify.  
-This situation may be considered to be a probabilistic generative model, 
-illustrated by the directed acyclic graph of \Fig{fig:pgm}.
-
-The redshift distribution function \nz\ is the number of galaxies per unit 
-redshift, effectively defining the evolution in the number of galaxies 
-\citep{Menard2013}.  
-In the following sections, I present and compare methods for estimating \nz\ 
-from \pzpdf s.  
-\Sect{sec:prob} contains the mathematical derivation of a probabilistic model 
-for $n(z)$ dependent on photo-$z$ probability distribution functions, and 
-\Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.
+Consider a survey of $J$ galaxies $j$, each with photometric data $\data_{j}$; thus the entire survey over some solid angle produces the ensemble of photometric magnitudes (or colors) and their associated observational errors $\{\data_{j}\}$.
+Each galaxy $j$ has a redshift $z_{j}$ that we would like to learn; redshift is a parameter in this case.  
+The distribution of the ensemble of redshifts $\{z_{j}\}$ may be described by the hyperparameters defining the redshift distribution function \nz\ that we would like to quantify.  
+This situation may be considered to be a probabilistic generative model, illustrated by the directed acyclic graph of \Fig{fig:pgm}.
+
+The redshift distribution function \nz\ is the number of galaxies per unit redshift, effectively defining the evolution in the number of galaxies \citep{menard_clustering-based_2013}.  
+% \aim{also selection function}
+In the following sections, I present and compare methods for estimating \nz\ from \pzpdf s.  
+\Sect{sec:prob} contains the mathematical derivation of a probabilistic model for $n(z)$ dependent on photo-$z$ probability distribution functions, and \Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.
 
@@
 
-We begin by reframing the redshift distribution \nz\ from a probabilistic 
-perspective.
+We begin by reframing the redshift distribution \nz\ from a probabilistic perspective.
 Here we define a redshift density \nz\ as the normalized probability density
@@
 \label{eqn:nz}
-\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ 
-\int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
+\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ \int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
 \end{equation}
 of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
-We believe that galaxy redshifts are indeed drawn from \nz, making it a 
-probability density over redshift; this fact can also be confirmed by 
-dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{Hogg2012}.
+We believe that galaxy redshifts are indeed drawn from \nz, making it a probability density over redshift; this fact can also be confirmed by dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{hogg_data_2012}.
 
@@
 in terms of some parameter vector $\ndphi$.
-At this point, the parameter vector is quite general and may represent 
-coefficients in a high-order polynomial as a function of redshift, a set of 
-means and variances defining Gaussians that sum to the desired distribution, a 
-set of histogram heights that describe a binned version of the redshift 
-distribution function, etc.
+At this point, the parameter vector is quite general and may represent coefficients in a high-order polynomial as a function of redshift, a set of means and variances defining Gaussians that sum to the desired distribution, a set of histogram heights that describe a binned version of the redshift distribution function, etc.
 Upon doing so, we may rewrite \Eq{eqn:fz} as 
@@
 \end{equation}
-a probability density over redshift conditioned on the parameters $\ndphi$ 
-specifying \nz.
-Note that $z_{j}$ does not depend on the redshift $z_{j'}$ of some other galaxy 
-$j' \neq j$, a statement of the causal independence of galaxy redshifts from 
-one another.
-
-In addition to believing \nz\ is a PDF from which redshifts are drawn, we also 
-believe that there is some higher dimensional probability space $\pr{z, \data}$ 
-of redshift and photometric data vectors $\data$, which may be any combination 
-of fluxes, magnitudes, colors, and their observational errors.
+% \aim{explain "is drawn from"}
+a probability density over redshift conditioned on the parameters $\ndphi$ specifying \nz.
+Note that $z_{j}$ does not depend on the redshift $z_{j'}$ of some other galaxy $j' \neq j$, a statement of the causal independence of galaxy redshifts from one another.
+
+In addition to believing \nz\ is a PDF from which redshifts are drawn, we also believe that there is some higher dimensional probability space $\pr{z, \data}$ of redshift and photometric data vectors $\data$, which may be any combination of fluxes, magnitudes, colors, and their observational errors.
 In that sense \nz\ is equivalent to an integral
@@
 over the dimension of data in that joint probability space.
-Note that galaxies may have different observational data despite sharing the 
-same redshift, and that galaxies at different redshifts may have identical 
-photometry; the space $\pr{z, \data}$ need not be one-to-one.
-We assume a stronger version of statistical independence here, that draws 
-$(z_{j}, \data_{j})$ are independent of draws $(z_{j'}, \data_{j'})$ in this 
-space; the data and redshift of each galaxy are independent of those of other 
-galaxies.
+Note that galaxies may have different observational data despite sharing the same redshift, and that galaxies at different redshifts may have identical photometry; the space $\pr{z, \data}$ need not be one-to-one.
+We assume a stronger version of statistical independence here, that draws $(z_{j}, \data_{j})$ are independent of draws $(z_{j'}, \data_{j'})$ in this space; the data and redshift of each galaxy are independent of those of other galaxies.
 
@@
 This is the fundamental assumption upon which \pz\ estimation is based.
-The forward model corresponds to first drawing redshifts according to 
-\Eq{eqn:pz} and then drawing data from the likelihood
+The forward model corresponds to first drawing redshifts according to \Eq{eqn:pz} and then drawing data from the likelihood
 \begin{equation}
@@
 \end{equation}
-of photometry conditioned on redshift, illustrated in 
-Figure~\ref{intro:fig:fig:pedagogical_scatter}.
-
-This description of the physical system corresponds to a forward model by which 
-we actually believe photometry is generated:
+of photometry conditioned on redshift, illustrated in \aim{fix reference}.
+%Figure~\ref{intro:fig:fig:pedagogical_scatter}.
+
+This description of the physical system corresponds to a forward model by which we actually believe photometry is generated:
 \begin{enumerate}
-	\item There exists a redshift distribution \nz\ with parameters 
-$\ndphi$.
-	\item Galaxy redshifts $\{z_{j}\}$ are independent draws from $\pr{z 
-\gvn \ndphi}$.
-	\item Galaxy photometry $\data_{j}$ is drawn from the likelihoods 
-$\pr{\data_{j} \gvn z}$.
+	\item There exists a redshift distribution \nz\ with parameters $\ndphi$.
+	\item Galaxy redshifts $\{z_{j}\}$ are independent draws from $\pr{z \gvn \ndphi}$.
+	\item Galaxy photometry $\data_{j}$ is drawn from the likelihoods $\pr{\data_{j} \gvn z}$.
 \end{enumerate}
@@
 
-A forward model such as that of \Sect{sec:forward} corresponds to a 
-probabilistic graphical model (PGM), represented by a directed acyclic graph 
-(DAG) as in \Fig{fig:pgm}.
-A DAG conveys the causal relationships between physical parameters and, like a 
-Feynman diagram in the context of particle physics, is a shorthand for 
-mathematical relationships between variables.
-The photometric data $\data_{j}$ of a galaxy is drawn from some function of its 
-redshift $z_{j}$, independent of other galaxies' data and redshift.
-Both data and redshift are random variables, but data is the one that we 
-observe and redshift is not directly observable.
-In this problem, we don't care about further constraining the redshifts of 
-individual galaxies, only the redshift distribution, so we consider redshift to 
-be a \textit{latent variable}.
-Because the parameters $\ndphi$ that we seek are causally separated from the 
-data by the latent variable of redshift, we call them \textit{hyperparameters}.
+A forward model such as that of \Sect{sec:forward} corresponds to a probabilistic graphical model (PGM), represented by a directed acyclic graph (DAG) as in \Fig{fig:pgm}.
+A DAG conveys the causal relationships between physical parameters and, like a Feynman diagram in the context of particle physics, is a shorthand for mathematical relationships between variables.
+The photometric data $\data_{j}$ of a galaxy is drawn from some function of its redshift $z_{j}$, independent of other galaxies' data and redshift.
+Both data and redshift are random variables, but data is the one that we observe and redshift is not directly observable.
+In this problem, we don't care about further constraining the redshifts of individual galaxies, only the redshift distribution, so we consider redshift to be a \textit{latent variable}.
+Because the parameters $\ndphi$ that we seek are causally separated from the data by the latent variable of redshift, we call them \textit{hyperparameters}.
+%\aim{I'm not explaining the distinction between hyperparameters and parameters adequately.}
 
@@
 	\begin{center}
-		\includegraphics[width=0.2\textwidth]{figures/chippr/pgm.png}
-		\caption{The directed acyclic graph of the CHIPPR model, where 
-circles indicate random variables and arrows indicate causal relationships.
-			The redshift distribution function parameterized by 
-hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, 
-indicated as a box.  
-			The redshifts $\{z_{j}\}$ of all galaxies in the survey 
-are latent variables independently drawn from the redshift distribution, which 
-is a function of $\ndphi$. 
-			The photometric data $\data_{j}$ for each galaxy is 
-drawn from a function of its redshift $z_{j}$ and observed, indicated by a 
-shaded circle.}
+		\includegraphics[width=0.15\textwidth]{figures/chippr/pgm.png}
+		\caption{The directed acyclic graph of the CHIPPR model, where circles indicate random variables and arrows indicate causal relationships.
+			The redshift distribution function parameterized by hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, indicated as a box.  
+			The redshifts $\{z_{j}\}$ of all galaxies in the survey are latent variables independently drawn from the redshift distribution, which is a function of $\ndphi$. 
+			The photometric data $\data_{j}$ for each galaxy is drawn from a function of its redshift $z_{j}$ and observed, indicated by a shaded circle.}
 		\label{fig:pgm}
@@
 
-The problem facing cosmologists is to determine the true value of $\ndphi$ from 
-observing the photometry $\{\data_{j}\}$ of a large sample of $J$ galaxies $j$.
-To self-consistently propagate the uncertainty in the redshifts, however, it is 
-more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ 
-over all possible values of $\ndphi$ conditioned on all the observed data 
-$\{\data_{j}\}$ available in a generic catalog.
-In order to use the DAG of \Fig{fig:pgm} to derive an expression for 
-$\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two 
-more concepts, confusingly named the implicit prior and the prior probability 
-density.
-
-When we constrain the redshift of a galaxy using its observed photometric data 
-$\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$.
-However, to do this, we must have a model for the general relationship between 
-redshifts and photometry, whether empirical, as is the case for machine 
-learning \pzpdf\ methods, or analytic, as is the case for template-based 
-\pzpdf\ methods.
-Such a relationship is defined in the space of probability density over 
-redshift, so it must be able to be parameterized by the same functional form 
-$f(z; \ndphi)$ as \nz .
-It is thus natural to write it as $\pr{z \gvn \ndphi^{*}}$, where $\ndphi^{*}$ 
-is the parameters for this relationship under some generic \pzpdf\ method.
-We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely 
-explicitly known nor chosen by the researcher; for template-based methods where 
-it can be chosen to be ``realistic,'' it may be appropriate to call it an 
-\textit{interim prior}.
-Because the implicit prior is unavoidable, the \pzpdf s reported by any method 
-are really \textit{implicit-weighted posteriors} $\pr{z \gvn \data, 
-\ndphi^{*}}$.
-
-The implicit prior $\ndphi^{*}$ may be thought of as an initial guess for the 
-parameters contained in $\ndphi$, inspired by the generative model for 
-photometry from the redshift distribution functions and including some 
-parameters defining intrinsic galaxy spectra and instrumental effects. 
-(See \citealt{benitez_bayesian_2000} for more detail.)  
-For statistical purposes, we would like any interim prior to be uninformative, 
-but this is rarely achievable.  
-In the case of estimating \nz\ photometrically, it is common to use 
-$\ndphi^{*}$ corresponding to \nz\ derived from some different, 
-spectroscopically confirmed sample or from a cosmological simulation.
-
-The prior probability density $\pr{\ndphi}$ is a more familiar concept in 
-astronomy; to progress, we will have to choose a prior probability density over 
-all possible parameters $\ndphi$.
-This prior need not be excessively proscriptive; for example, it may be chosen 
-to enforce smoothness at physically motivated scales in redshift without 
-imposing any particular region as over- or under-dense.
-
-With these definitions, we obtain the desired expression for $\pr{\ndphi \gvn 
-\{\data_{j}\}}$,
+The problem facing cosmologists is to determine the true value of $\ndphi$ from observing the photometry $\{\data_{j}\}$ of a large sample of $J$ galaxies $j$.
+To self-consistently propagate the uncertainty in the redshifts, however, it is more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ over all possible values of $\ndphi$ conditioned on all the observed data $\{\data_{j}\}$ available in a generic catalog.
+In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the implicit prior and the prior probability density.
+
+When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$.
+However, to do this, we must have a model for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
+Such a relationship is defined in the space of probability density over redshift, so it must be able to be parameterized by the same functional form $f(z; \ndphi)$ as \nz .
+It is thus natural to write it as $\pr{z \gvn \ndphi^{*}}$, where $\ndphi^{*}$ is the parameters for this relationship under some generic \pzpdf\ method.
+We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher; for template-based methods where it can be chosen to be ``realistic,'' it may be appropriate to call it an \textit{interim prior}.
+Because the implicit prior is unavoidable, the \pzpdf s reported by any method are really \textit{implicit-weighted posteriors} $\pr{z \gvn \data, \ndphi^{*}}$.
+
+%Posteriors differ from likelihoods by way of a prior distribution, so we cannot simply assume that the available data products are \pz\ posteriors $\pr{z \gvn \data_{j}}$.  
+%Rather, we have a catalog of implicit-prior weighted \pz\ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$.  
+%There must have been some interim prior probability distribution $p(z|\vec{\theta}^{*})$ defined in terms of the interim prior parameter values (hereafter the interim prior) $\vec{\theta}^{*}$ explicitly chosen or implicitly made to perform the calculation of the probabilistic photo-$z$s.
+%If it is implicit, it may not be representable in the parametrization we have chosen, and furthermore it may not be known at all; a method that produces interim photo-$z$ posteriors of this kind is not suitable for inference.
+%However, so long as the implicit prior is known, hierarchical inference is possible.
+
+The implicit prior $\ndphi^{*}$ may be thought of as an initial guess for the parameters contained in $\ndphi$, inspired by the generative model for photometry from the redshift distribution functions and including some parameters defining intrinsic galaxy spectra and instrumental effects.
+(See \citet{benitez_bayesian_2000} for more detail.)  
+For statistical purposes, we would like any interim prior to be uninformative, but this is rarely achievable.  
+In the case of estimating \nz\ photometrically, it is common to use $\ndphi^{*}$ corresponding to \nz\ derived from some different, spectroscopically confirmed sample or from a cosmological simulation.
+
+The prior probability density $\pr{\ndphi}$ is a more familiar concept in astronomy; to progress, we will have to choose a prior probability density over all possible parameters $\ndphi$.
+This prior need not be excessively proscriptive; for example, it may be chosen to enforce smoothness at physically motivated scales in redshift without imposing any particular region as over- or under-dense.
+
+% this is what we want to compute
+% what we're given is implicit posteriors
+% this turns out to be the correct answer
+% derivation in appendix
+% human-readable description of ratios replaces the implict prior with model at each sample step, converting posterior into likelihood
+With these definitions, we obtain the desired expression for $\pr{\ndphi \gvn \{\data_{j}\}}$,
 \begin{align}
@@
 \label{eqn:fullpost}
-\pr{\ndphi \gvn \{\data_{j}\}} & \propto 
-\pr{\ndphi}\left[\integral{\prod_{j=1}^{J} \left(\pr{z \gvn \data_{j}, 
-\ndphi^{*}}\pr{z \gvn \ndphi}(\pr{z \gvn \ndphi^{*}})^{-1}\right)}{z}\right] ,
+\pr{\ndphi \gvn \{\data_{j}\}} \propto & \pr{\ndphi} \left[\integral{\prod_{j=1}^{J} \right.\\
+	& \left. \left(\pr{z \gvn \data_{j}, \ndphi^{*}} \pr{z \gvn \ndphi}(\pr{z \gvn \ndphi^{*}})^{-1}\right)}{z}\right] ,
 \end{split}
@@
 
+\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
+
+%We begin by parametrizing $N(z)$ in terms of $\vec{\theta}$, comprising some set of hyperparameters that define the form $N(z)$ may take in whatever basis we choose.  
+%We define a function $f_{\vec{\theta}}(z)=N(z)$ that transforms these hyperparameters into the redshift distribution function $N(z)$.  
+%Because 
+%\begin{equation}
+%\eqlabel{eq:definition}
+%N(z) \propto p(z \gvn \vec{\theta}),
+%\end{equation}
+%we may discontinue discussion of $N(z)$ in favor of the likelihood $p(z|\vec{\theta})$.
+
 In this paper, we work exclusively with log-probabilities.  
-What we wish to estimate is then the full log-posterior probability 
-distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ 
-given the catalog of photometry $\{\data_{j}\}$.
+What we wish to estimate is then the full log-posterior probability distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ given the catalog of photometry $\{\data_{j}\}$.
 
@@
 \label{eqn:basicbayes}
-\ln[\pr{\ndphi \gvn \{\data_{j}\}}] = \ln[\pr{\{\data_{j}\} \gvn \ndphi}] + 
-\ln[\pr{\ndphi}] - \ln[\pr{\{\data_{j}\}}]
-\end{equation}
-may be expressed in terms of the full log-likelihood probability distribution 
-(hereafter the full log-likelihood) $\ln[\pr{\{\data_{j}\} \gvn \ndphi}]$ by 
-way of a hyperprior log-probability distribution (hereafter the hyperprior) 
-$\ln[\pr{\ndphi}]$ over the hyperparameters and the log-evidence probability of 
-the data $\ln[\pr{\{\data_{j}\}}]$.
-However, the evidence is rarely known, so we probe the full log-posterior 
-modulo an unknown constant of proportionality.
-
-The full log-likelihood may be expanded in terms of a marginalization over the 
-redshifts as parameters, as in 
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] = \ln[\pr{\{\data_{j}\} \gvn \ndphi}] + \ln[\pr{\ndphi}] - \ln[\pr{\{\data_{j}\}}]
+\end{equation}
+may be expressed in terms of the full log-likelihood probability distribution (hereafter the full log-likelihood) $\ln[\pr{\{\data_{j}\} \gvn \ndphi}]$ by way of a hyperprior log-probability distribution (hereafter the hyperprior) $\ln[\pr{\ndphi}]$ over the hyperparameters and the log-evidence probability of the data $\ln[\pr{\{\data_{j}\}}]$.
+However, the evidence is rarely known, so we probe the full log-posterior modulo an unknown constant of proportionality.
+
+The full log-likelihood may be expanded in terms of a marginalization over the redshifts as parameters, as in 
 \begin{equation}
 \label{eqn:marginalize}
-\ln[\pr{\{\data_{j}\} \gvn \ndphi}] = \ln\left[\integral{\pr{\{\data_{j}\} \gvn 
-\{z_{j}\}} \pr{\{z_{j}\} \gvn \ndphi}}{\{z_{j}\}}\right].
-\end{equation}
-
-We shall make two assumptions of independence in order to make the problem 
-tractable; their limitations are be discussed below.  
-First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ 
-individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn 
-z_{j}}]$, as in 
+\ln[\pr{\{\data_{j}\} \gvn \ndphi}] = \ln\left[\integral{\pr{\{\data_{j}\} \gvn \{z_{j}\}} \pr{\{z_{j}\} \gvn \ndphi}}{\{z_{j}\}}\right].
+\end{equation}
+
+We shall make two assumptions of independence in order to make the problem tractable; their limitations are be discussed below.  
+First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn z_{j}}]$, as in 
 \begin{equation}
 \label{eqn:indiedat}
-\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn 
-z_{j}}],
-\end{equation}
-a result of the definition of probabilistic independence encoded by the box in 
-\Fig{fig:pgm}.
-Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent 
-draws from the true $\pr{z \gvn \ndphi}$.  
+\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn z_{j}}],
+\end{equation}
+a result of the definition of probabilistic independence encoded by the box in \Fig{fig:pgm}.
+Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent draws from the true $\pr{z \gvn \ndphi}$.  
 Additionally, $J$ itself is a Poisson random variable.  
@@
 \label{eqn:indie}
-\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ 
-\ln[\pr{z_{j} \gvn \ndphi}].
-\end{equation}
-The derivation differs when $J$ is not known, say, when we want to learn about 
-a distribution in nature rather than a distribution specific to data in hand, 
-but for a photometric galaxy catalog where the desired quantity is $n(z)$ for 
-the galaxies entering a larger cosmology calculation, it is a fixed quantity.
-A detailed discussion of this matter may be found in 
-\citet{Foreman-Mackey2014}.  
+\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ \ln[\pr{z_{j} \gvn \ndphi}].
+\end{equation}
+%It is important to note that the integral $\integral{n(z)}{z} N(z)\ dz$ is not constrained to equal the variable defining the Poisson distribution but instead $J$ by \Eq{eq:definition}, which can be thought of as another parameter.
+The derivation differs when $J$ is not known, say, when we want to learn about a distribution in nature rather than a distribution specific to data in hand, but for a photometric galaxy catalog where the desired quantity is $n(z)$ for the galaxies entering a larger cosmology calculation, it is a fixed quantity.
+A detailed discussion of this matter may be found in \citet{foreman-mackey_exoplanet_2014}.  
 Applying Bayes' Rule, we may combine terms to obtain 
-\begin{equation}
+\begin{align}
+\begin{split}
 \label{eqn:posterior}
-\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto \ln[\pr{\ndphi}] - \integral{f(z; 
-\ndphi)}{z} + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn 
-\ndphi}}{z}\right].
-\end{equation}
-
-Since we only have access to implicit \pz\ posteriors, we must be able to write 
-the full log-posterior in terms of implicit \pz\ log-posteriors rather than the 
-log-likelihoods of \Eq{eqn:posterior}.
-To do so, we will need an explicit statement of this implicit prior 
-$\ndphi^{*}$ for whatever method is chosen to produce the implicit \pz\ 
-posteriors.
-
-To perform the necessary transformation from likelihoods to posteriors, we 
-follow the reasoning of \citet{Foreman-Mackey2014}.  
-Let us consider the probability of the parameters conditioned on the data and 
-an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} 
-as 
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] - \integral{f(z; \ndphi)}{z}\\
+& + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn \ndphi}}{z}\right].
+\end{split}
+\end{align}
+
+%\Eq{eq:posterior} contains two quantities that merit further discussion, the prior distribution $p(\vec{\theta})$ discussed further in \Sect{sec:exp} and the photo-$z$ log-likelihoods $\ln[p(\vec{d}_{j}|z_{j})]$ that have not been mentioned since \Eq{eq:marginalize}.
+%Though photo-$z$ log-likelihoods would be desirable for use in these equations, they are not generally the product of either empirical and data-driven methods for obtaining photo-$z$ probability distributions.
+%Though probabilistic photo-$z$s are typically reported as generic probability distributions $p(z_{j})$, the methods that produce them may be understood to always yield posteriors, probability distributions conditioned on the data we believe to be true.
+%If they were not based in this assumption, they would require a sum over an infinite space of possible datasets.
+
+Since we only have access to implicit \pz\ posteriors, we must be able to write the full log-posterior in terms of implicit \pz\ log-posteriors rather than the log-likelihoods of \Eq{eqn:posterior}.
+To do so, we will need an explicit statement of this implicit prior $\ndphi^{*}$ for whatever method is chosen to produce the implicit \pz\ posteriors.
+
+To perform the necessary transformation from likelihoods to posteriors, we follow the reasoning of \citet{foreman-mackey_exoplanet_2014}.  
+Let us consider the probability of the parameters conditioned on the data and an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} as 
 \begin{equation}
 \label{eqn:trick}
-\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn 
-\data_{j}, \ndphi^{*}}] - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
-\end{equation}
-
-Once the implicit prior $\ndphi^{*}$ is explicitly introduced, we may expand 
-the last term in \Eq{eqn:trick} according to Bayes' Rule to get 
-\begin{equation}
+\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
+\end{equation}
+
+Once the implicit prior $\ndphi^{*}$ is explicitly introduced, we may expand the last term in \Eq{eqn:trick} according to Bayes' Rule to get 
+\begin{align}
+\begin{split}
 \label{eqn:expand}
-\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn 
-\data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn 
-\ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
-\end{equation}
-Because there is no direct dependence of the data upon the hyperparameters, we 
-may again expand the term $\ln[\pr{\data_{j} \gvn z, \ndphi^{*}}]$ to obtain 
-\begin{equation}
+\ln[\pr{\data_{j} \gvn z}] & = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}]\\
+& - \ln[\pr{z \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
+\end{split}
+\end{align}
+Because there is no direct dependence of the data upon the hyperparameters, we may again expand the term $\ln[\pr{\data_{j} \gvn z, \ndphi^{*}}]$ to obtain 
+\begin{align}
+\begin{split}
 \label{eqn:indterm}
-\ln[\pr{\vec{d}_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn 
-\data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn 
-\ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] - \ln[\pr{\data_{j} \gvn \ndphi^{*}}].
-\end{equation}
-Canceling the undesirable terms for the inaccessible likelihood 
-$\ln[\pr{\data_{j} \gvn z}]$ and trivial $\ln[\pr{\data_{j} \gvn \ndphi^{*}}]$ 
-yields
+\ln[\pr{\vec{d}_{j} \gvn z}] & = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] \\
+& - \ln[\pr{z \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] - \ln[\pr{\data_{j} \gvn \ndphi^{*}}].
+\end{split}
+\end{align}
+Canceling the undesirable terms for the inaccessible likelihood $\ln[\pr{\data_{j} \gvn z}]$ and trivial $\ln[\pr{\data_{j} \gvn \ndphi^{*}}]$ yields
 \begin{equation}
 \label{eqn:cancel}
-\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]  - 
-\ln[\pr{z \gvn \ndphi^{*}}].
-\end{equation}
-We put this all together to get the full log-posterior probability distribution 
-of 
-\begin{equation}
+\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]  - \ln[\pr{z \gvn \ndphi^{*}}].
+\end{equation}
+We put this all together to get the full log-posterior probability distribution of 
+\begin{align}
+\begin{split}
 \label{eqn:final}
-\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto \ln[\pr{\ndphi}] + \ln 
-\left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, 
-\ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn 
-\ndphi^{*}}]\right)\right]}{z}\right] ,
-\end{equation}
-
-The argument of the integral in the log-posterior of \Eq{eqn:final} depends 
-solely on knowable quantities (and those we must explicitly assume) and can be 
-calculated for a given sample of \pz\ log-posteriors $\{\ln[\pr{z \gvn 
-\data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with 
-which they were obtained, noting the relation of 
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \right.\right.\\
+& \left.\left.\left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
+\end{split}
+\end{align}
+
+The argument of the integral in the log-posterior of \Eq{eqn:final} depends solely on knowable quantities (and those we must explicitly assume) and can be calculated for a given sample of \pz\ log-posteriors $\{\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with which they were obtained, noting the relation of
 \begin{equation}
@@
 \end{equation}
-Since we cannot know constant of proportionality, we sample the desired full 
-log-posterior $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$ using Monte Carlo-Markov 
-chain (MCMC) methods.  
-The method outlined here is valid regardless of how the implicit \pz\ 
-log-posteriors are obtained so the many approaches to producing \pzpdf s will 
-not be rehashed; though the matter is outside the scope of this paper, reviews 
-of various methods have been presented in the literature 
-\citep{sheldon_photometric_2012, Ball2008, carrasco_kind_tpz:_2013, 
-carrasco_kind_exhausting_2014}, and will be briefly reviewed in \aim{cite PZDC1 
-paper here}.
+Since we cannot know constant of proportionality, we sample the desired full log-posterior $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$ using Monte Carlo-Markov chain (MCMC) methods.  
+The method outlined here is valid regardless of how the implicit \pz\ log-posteriors are obtained so the many approaches to producing \pzpdf s will not be rehashed; though the matter is outside the scope of this paper, reviews of various methods have been presented in the literature \citep{sheldon_photometric_2012, ball_robust_2008, carrasco_kind_tpz:_2013, carrasco_kind_exhausting_2014}, and will be briefly reviewed in \aim{cite PZDC1 paper here}.
+%
+%\begin{align}
+%\begin{split}
+%\eqlabel{eqn:fullpost}
+%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
+%\end{split}
+%\end{align}
 
@@
 
-Finally, we explicitly review the assumptions made by this approach, which are 
-as follows:
+Finally, we explicitly review the assumptions made by this approach, which are as follows:
 
 \begin{enumerate}
-	\item Photometric measurements of galaxies are statistically 
-independent Poisson draws from the set of all galaxies such that 
-\Eq{eqn:indiedat} and \Eq{eqn:indie} hold.
-	\item We take the reported \pzip s to be accurate, free of model 
-misspecification; draws thereof must not be inconsistent with the distribution 
-of photometry and redshifts.
-	Furthermore, we must be given the implicit prior $\ndphi^{*}$ used to 
-produce the \pzip s.
-	\item We must assume a hyperprior distribution $\pr{\ndphi}$ 
-constraining the underlying probability distribution of the hyperparameters, 
-which is informed by our prior beliefs about the true redshift distribution 
-function.
+	\item Photometric measurements of galaxies are statistically independent Poisson draws from the set of all galaxies such that \Eq{eqn:indiedat} and \Eq{eqn:indie} hold.
+	\item We take the reported \pzip s to be accurate, free of model misspecification; draws thereof must not be inconsistent with the distribution of photometry and redshifts.
+	Furthermore, we must be given the implicit prior $\ndphi^{*}$ used to produce the \pzip s.
+	\item We must assume a hyperprior distribution $\pr{\ndphi}$ constraining the underlying probability distribution of the hyperparameters, which is informed by our prior beliefs about the true redshift distribution function.
 \end{enumerate}
@@
 These assumptions have known limitations.  
-First, the photometric data are not a set of independent measurements; the data 
-are correlated not only by the conditions of the experiment under which they 
-were observed (instrument and observing conditions) but also by redshift 
-covariances resulting from physical processes governing underlying galaxy 
-spectra and their relation to the redshift distribution function.
-Second, the reported \pzip s may not be trustworthy; there is not yet agreement 
-on the best technique to obtain \pzpdf s, and the implicit prior may not be 
-appropriate or even known to us as consumers of \pzip s.  
-Third, the hyperprior may be quite arbitrary and poorly motivated if the 
-underlying physics is complex, and it can only be appropriate if our prior 
-beliefs about \nz\ are accurate.
-
+First, the photometric data are not a set of independent measurements; the data are correlated not only by the conditions of the experiment under which they were observed (instrument and observing conditions) but also by redshift covariances resulting from physical processes governing underlying galaxy spectra and their relation to the redshift distribution function.
+Second, the reported \pzip s may not be trustworthy; there is not yet agreement on the best technique to obtain \pzpdf s, and the implicit prior may not be appropriate or even known to us as consumers of \pzip s.
+Third, the hyperprior may be quite arbitrary and poorly motivated if the underlying physics is complex, and it can only be appropriate if our prior beliefs about \nz\ are accurate.
+
+%\clearpage
 \subsection{Implementation}
@@
 
-I implemented the \Chippr\ model in code in order to perform tests of its 
-validity and to compare its performance to that of traditional alternatives.
+I implemented the \Chippr\ model in code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
 In \Sect{sec:mcmc}, I describe the publicly available \chippr\ library.
-In \Sect{sec:acorr}, I outline how \chippr\ can be used to sample the full 
-log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
-
+In \Sect{sec:acorr}, I outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
+
+%\clearpage
 \subsubsection{Code}
@@
 
-\chippr\ is a \python\ 2 
-library\footnote{\url{https://github.com/aimalz/chippr}} that includes an 
-implementation of the \Chippr\ model as well as an extensive suite of tools for 
-comparing \Chippr\ to other approaches.
-
-Though there are plans for future expansion to more flexible parameterizations, 
-the current version of \chippr\ uses a log-space piecewise constant 
-parameterization
+\chippr\ is a \python\ 2 library\footnote{\url{https://github.com/aimalz/chippr}} that includes an implementation of the \Chippr\ model as well as an extensive suite of tools for comparing \Chippr\ to other approaches.
+
+Though there are plans for future expansion to more flexible parameterizations, the current version of \chippr\ uses a log-space piecewise constant parameterization
 \begin{equation}
@@
 \end{equation}
-with $K$ bins of width $\delta z^{1}, \dots, \delta z^{K}$ defined by endpoints 
-$z^{0}, \dots, z^{K}$.
-Thus each $\pr{z \gvn \data_{j}} = f(z; \ndphi_{j})$ has parameters 
-$\ndphi_{j}$ that are defined in the same basis as those of \nz.
-To infer the full log-posterior distribution $\ln[\pr{\ndphi \gvn 
-\{\data_{j}\}}]$, one must provide a plaintext file with $K+1$ redshift bin 
-endpoints $\{z_{k}\}$, the parameters $\ndphi^{*}$ of the implicit log-prior, 
-and the parameters $\{\ndphi_{j}\}$ of the log-posteriors $\ln[\pr{z \gvn 
-\data_{j}, \ndphi^{*})}$.
-
-The \emcee \citep{Foreman-Mackey2013} implementation of ensemble sampling is 
-used to sample the full log-posterior of \Eq{eqn:final}. 
-\chippr\ accepts a configuration file of user-specified parameters, among them 
-the number $W$ of walkers.
-At each iteration $i$ and for each walker, a proposal distribution 
-$\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for 
-acceptance to or rejection from the full log-posterior distribution.
-
-Two threshold conditions are defined, one designating all previous samples to 
-be ignored as as products of a burn-in phase and another indicating when a 
-sufficient number of post-burn samples have been accepted.  
-In this case, the first threshold (described in \Sect{sec:acorr}) is defined in 
-terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an 
-accumulation of $10^{4}$ samples.
-
-Though previous versions used \texttt{HDF5} for the primary I/O format due to 
-its efficiency for large quantities of data, it was abandoned in favor of 
-\texttt{pickle} in the working release due to the instability of the \python\ 
-implementation of the format on high-performance computing systems.  
-The resulting output is a set of $I$ ordered \texttt{hickle} files enumerated 
-by $\rho$ containing the state information after each sub-run.  
-The state information includes $\frac{I_{0}}{s}$ actual samples $\ndphi_{i}$ 
-for a pre-specified chain thinning factor $s$ and their full posterior 
-probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$ as well as the 
-autocorrelation times and acceptance fractions calculated for each element of 
-$\ndphi$ over the entire sub-run.
-
+with $K$ bins of width $\delta z^{1}, \dots, \delta z^{K}$ defined by endpoints $z^{0}, \dots, z^{K}$.
+%\aim{Maybe include an equation for a general piecewise constant function here (which is currently in \Chap{qp})?}
+Thus each $\pr{z \gvn \data_{j}} = f(z; \ndphi_{j})$ has parameters $\ndphi_{j}$ that are defined in the same basis as those of \nz.
+To infer the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$, one must provide a plaintext file with $K+1$ redshift bin endpoints $\{z_{k}\}$, the parameters $\ndphi^{*}$ of the implicit log-prior, and the parameters $\{\ndphi_{j}\}$ of the log-posteriors $\ln[\pr{z \gvn \data_{j}, \ndphi^{*})}$.
+
+The \emcee \citep{foreman-mackey_emcee_2013} implementation of ensemble sampling is used to sample the full log-posterior of \Eq{eqn:final}. 
+\chippr\ accepts a configuration file of user-specified parameters, among them the number $W$ of walkers.
+At each iteration $i$ and for each walker, a proposal distribution $\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for acceptance to or rejection from the full log-posterior distribution.
+
+Two threshold conditions are defined, one designating all previous samples to be ignored as as products of a burn-in phase and another indicating when a sufficient number of post-burn samples have been accepted.
+In this case, the first threshold (described in \Sect{sec:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
+
+Though previous versions used \texttt{HDF5} for the primary I/O format due to its efficiency for large quantities of data, it was abandoned in favor of \texttt{pickle} in the working release due to the instability of the \python\ implementation of the format on high-performance computing systems.
+The resulting output is a set of $I$ ordered \texttt{hickle} files enumerated by $\rho$ containing the state information after each sub-run.  
+The state information includes $\frac{I_{0}}{s}$ actual samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$ as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$ over the entire sub-run.
+
+%\clearpage
 \subsubsection{Convergence Criteria}
@@
 
-In addition to qualitative visual inspection of the chains, two quantities that 
-probe the convergence of the sampler are used in this study, the 
-autocorrelation time and the Gelman-Rubin convergence criterion.
-
-The autocorrelation time is effectively a measure of the efficiency of the 
-method and can be described as the expected number of iterations necessary to 
-accept a new sample independent of the current accepted sample.  
-A sampler that converges faster will have a smaller autocorrelation time, and 
-smaller autocorrelation times are preferable because it means fewer iterations 
-are wasted on non-independent samples when independent samples are desired.  
-See \citet{Foreman-Mackey2013} for a more complete exploration of the 
-autocorrelation time.  
-In all tests discussed here, autocorrelation times across walkers and 
-parameters were approximately 20, meaning two samples 20 or more iterations 
-apart were independent, a satisfactory level of efficiency.  
-Low autocorrelation times are a necessary but not always sufficient convergence 
-condition, as the autocorrelation times calculated for tests in this paper were 
-constant across all sub-runs, even those that were obviously burning in.
+\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
+
+In addition to qualitative visual inspection of the chains, two quantities that probe the convergence of the sampler are used in this study, the autocorrelation time and the Gelman-Rubin convergence criterion.
+%\Fig{fig:chains} shows the %evolution of the values of one parameter of one walker over the course of all %iterations of the sampler.
+
+%\begin{figure}
+%%\includegraphics[width=0.5\textwidth]{figs/null/chain0.pdf}
+%\caption{This figure shows the evolution of one walker's parameter values for 
+%one element of the parameter vector $\vec{\theta}$ as a function of iteration 
+%number, demonstrating the completion of the burn-in phase.}
+%\label{fig:chains}
+%\end{figure}
+
+The autocorrelation time is effectively a measure of the efficiency of the method and can be described as the expected number of iterations necessary to accept a new sample independent of the current accepted sample.
+A sampler that converges faster will have a smaller autocorrelation time, and smaller autocorrelation times are preferable because it means fewer iterations are wasted on non-independent samples when independent samples are desired.
+See \citet{foreman-mackey_emcee_2013} for a more complete exploration of the autocorrelation time.  
+In all tests discussed here, autocorrelation times across walkers and parameters were approximately 20, meaning two samples 20 or more iterations apart were independent, a satisfactory level of efficiency.
+Low autocorrelation times are a necessary but not always sufficient convergence condition, as the autocorrelation times calculated for tests in this paper were constant across all sub-runs, even those that were obviously burning in.
 
@@
 \label{eqn:gr}
-R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} 
-b_{k}}{w_{k}}},
-\end{equation}
-a weighted sum of the mean $w_{k}$ of the variances within individual walkers' 
-chains and the variance $b_{k}$ between chains of different walkers $m$, is 
-calculated over each sub-run $i$ to determine the duration of the burn-in 
-period.  
+R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} b_{k}}{w_{k}}},
+\end{equation}
+a weighted sum of the mean $w_{k}$ of the variances within individual walkers' chains and the variance $b_{k}$ between chains of different walkers $m$, is calculated over each sub-run $i$ to determine the duration of the burn-in period.
 Convergence is achieved when the statistic approaches unity.
@@
 
-In this study, we compare the results of \Eq{eqn:fullpost} to those of the two 
-most common approaches to estimating \nz\ from a catalog of \pzpdf s: the 
-distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior 
-probability
+In this study, we compare the results of \Eq{eqn:fullpost} to those of the two most common approaches to estimating \nz\ from a catalog of \pzpdf s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior probability
 \begin{equation}
 \label{eqn:mmap}
-f^{MMAP}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \delta(z, \mathrm{mode}[\pr{z \gvn 
-\data_{j}, \ndphi^{*}}])
-\end{equation}
-(i.e. the distribution of modes of the \pzpdf s) and the stacked estimator of 
-\Eq{eqn:stacked}, which can be rewritten as 
+f^{MMAP}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \delta(z, \mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}])
+\end{equation}
+(i.e. the distribution of modes of the \pzpdf s) and the stacked estimator of \Eq{eqn:stacked}, which can be rewritten as 
 \begin{equation}
@@
 in terms of the implicit \pz\ posteriors we have.
-These two approaches have been compared to one another by 
-\citet{Hildebrandt2012}, \citet{Benjamin2013}, and \citet{Asorey2016} in the 
-past but not to \Chippr.
-
-Point estimation converts the implicit photo-$z$ posteriors $\pr{z \gvn 
-\data_{j}, \ndphi^{*}}$ into delta functions with all probability at a single 
-estimated redshift.  
-Some variants of point estimation choose this single redshift to be that of 
-maximum a posteriori probability $\mathrm{mode}[\pr{z \gvn \data_{j}, 
-\ndphi^{*}}]$ or the expected value of redshift $\langle z \rangle = 
-\integral{z \pr{z \gvn \data_{j}, \ndphi^{*}}}{z}$.
-\citet{tanaka_photometric_2018} directs attention to deriving an optimal point 
-estimate reduction of a \pzpdf, but since the purpose of this paper is to 
-compare against the most established alternative estimators of \nz, its use 
-will be postponed until a future study.
-Stacking these modified \pzpdf s leads to the marginalized maximum a posteriori 
-(MMAP) estimator and the marginalized expected value (MExp) estimator, though 
-only the former is included in this study since the latter has fallen out of 
-favor in recent years\footnote{And for good reason!  Consider a bimodal \pzpdf; 
-its expected value may very well fall in a region of very low probability, 
-yielding a less probable point estimate than the point at which either peak 
-achieves its maximum.}.
+These two approaches have been compared to one another by \citet{hildebrandt_cfhtlens:_2012}, \citet{benjamin_cfhtlens_2013}, and \citet{asorey_galaxy_2016} in the past but not to \Chippr.
+
+Point estimation converts the implicit photo-$z$ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$ into delta functions with all probability at a single estimated redshift.  
+Some variants of point estimation choose this single redshift to be that of maximum a posteriori probability $\mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}]$ or the expected value of redshift $\langle z \rangle = \integral{z \pr{z \gvn \data_{j}, \ndphi^{*}}}{z}$.
+\aim{cite tanaka 2018 here} directs attention to deriving an optimal point estimate reduction of a \pzpdf, but since the purpose of this paper is to compare against the most established alternative estimators of \nz, its use will be postponed until a future study.
+Stacking these modified \pzpdf s leads to the marginalized maximum a posteriori (MMAP) estimator and the marginalized expected value (MExp) estimator, though only the former is included in this study since the latter has fallen out of favor in recent years\footnote{And for good reason!
+Consider a bimodal \pzpdf; its expected value may very well fall in a region of very low probability, yielding a less probable point estimate than the point at which either peak achieves its maximum.}.
 
 It is worth discussing the relationship between point estimation and stacking.  
-When the point estimator of redshift is equal to the true redshift, stacking 
-delta function \pzpdf s will indeed lead to an accurate recovery of the true 
-redshift distribution function.  
-However, stacking is in general applied indiscriminately to broader \pzpdf s 
-and imperfect point estimators of redshift.  
+When the point estimator of redshift is equal to the true redshift, stacking delta function \pzpdf s will indeed lead to an accurate recovery of the true redshift distribution function.  
+However, stacking is in general applied indiscriminately to broader \pzpdf s and imperfect point estimators of redshift.  
 It is for these reasons that alternatives are considered here.
 
-A final estimator of the hyperparameters is the maximum marginalized likelihood 
-estimator, the value of $\ndphi$ maximizing the log posterior given by 
-\Eq{eqn:final} using any optimization code.  
-To compare with sampling, the MMLE also depends on the choice of the hyperprior 
-distribution, and it does not produce a full posterior probability distribution 
-over the parameters of interest, only point estimators.  
-It must be noted that computation of the maximum marginalized likelihood 
-estimator may be unstable depending on the strengths and weaknesses of the 
-optimizer.  
-In general, derivatives will not be available for the full posterior 
-distribution, restricting optimization methods used.
-
-In \Sect{sec:diag} we outline the measures used to evaluate the performance of 
-the method.
-
+A final estimator of the hyperparameters is the maximum marginalized likelihood estimator, the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.  
+To compare with sampling, the MMLE also depends on the choice of the hyperprior distribution, and it does not produce a full posterior probability distribution over the parameters of interest, only point estimators.
+It must be noted that computation of the maximum marginalized likelihood estimator may be unstable depending on the strengths and weaknesses of the optimizer.  
+In general, derivatives will not be available for the full posterior distribution, restricting optimization methods used.
+%\begin{equation}
+%\label{eq:mmle}
+%\ln[p(\{\vec{d}_{j}\}|\vec{\theta})] \propto -\int\ f_{\vec{\theta}}(z)\ 
+%dz+\sum_{j=1}^{J}\ln\left[\int\ 
+%\exp\left[\ln[p(z_{j}|\vec{d}_{j},\vec{\theta}^{*})]+\ln[f_{\vec{\theta}}(z)]-\l
+%n[f_{\vec{\theta}^{*}}(z)]\right]\ dz\right],
+%\end{equation}
+%accessible with any optimization code.
+
+In \Sect{sec:diag} we outline the measures used to evaluate the performance of the method.
+
+%\clearpage
 \subsubsection{Performance metrics}
@@
 
-The results of the computation described in \Sect{sec:mcmc} are evaluated for 
-accuracy on the basis of some quantitative measures.  
-Beyond visual inspection of samples, we calculate summary statistics to 
-quantitatively compare different estimators' precision and accuracy.  
-Since MCMC samples of hyperparameters are Gaussian distributions, we can 
-quantify the breadth of the distribution for each hyperparameter using the 
-standard deviation regardless of whether the true values are known.
-
-In simulated cases where the true parameter values are known, we calculate the 
-Kullback-Leibler divergence (KLD), given by 
+The results of the computation described in \Sect{sec:mcmc} are evaluated for accuracy on the basis of some quantitative measures.  
+Beyond visual inspection of samples, we calculate summary statistics to quantitatively compare different estimators' precision and accuracy.  
+Since MCMC samples of hyperparameters are Gaussian distributions, we can quantify the breadth of the distribution for each hyperparameter using the standard deviation regardless of whether the true values are known.
+
+In simulated cases where the true parameter values are known, we calculate the Kullback-Leibler divergence (KLD), given by 
 \begin{equation}
 \label{eqn:kl}
-KL_{',\ddagger} = \integral{\pr{z \gvn \ndphi'} \ln \left[ \frac{\pr{z \gvn 
-\ndphi'}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
-\end{equation}
-which measures a distance from parameter values $\ndphi'$ to true parameter 
-values $\ndphi^{\dagger}$, which is invariant under changes of variables.  
-We note that $KL_{',\dagger} \neq KL_{\dagger,'}$ and is only interpretable 
-when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than 
-$\ndphi'$.
-The KLD is explored in detail in the Appendix to \aim{cite qp paper here}.
-In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that 
-produced by one of the methods in question.
+KL_{',\ddagger} = \integral{\pr{z \gvn \ndphi'} \ln \left[ \frac{\pr{z \gvn \ndphi'}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
+\end{equation}
+which measures a distance from parameter values $\ndphi'$ to true parameter values $\ndphi^{\dagger}$, which is invariant under changes of variables.  
+We note that $KL_{',\dagger} \neq KL_{\dagger,'}$ and is only interpretable when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than $\ndphi'$.
+The KLD is explored in detail in the Appendix to \citet{malz_approximating_2018}.
+In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that produced by one of the methods in question.
 
@@
 
-We compare the results of \Chippr\ to those of stacking and the histogram of 
-\pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf 
-s generated via the forward model discussed in \Sect{sec:forward}.
-\Fig{fig:flowchart} illustrates the implementation of this forward model used 
-in this paper.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=\textwidth]{figures/chippr/flowchart.png}
-		\caption{A flow chart illustrating the forward model used to 
-generate mock data in the validation of \Chippr, as described in 
-\Sect{sec:forward}.
-			Ovals indicate a quantity that must be chosen in order 
-to generate the data, rectangles indicate an operation I perform, and rounded 
-rectangles indicate a quantity created by the forward model.
-			Arrows indicate the inputs and outputs of each 
-operation performed to simulate mock \pzpdf\ catalogs.}
+We compare the results of \Chippr\ to those of stacking and the histogram of \pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed in \Sect{sec:forward}.
+\Fig{fig:flowchart} illustrates the implementation of this forward model used in this paper.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.74\textwidth]{figures/chippr/flowchart.png}
+		\caption{\aim{Check that all names and formulae are defined in the text and the same way as in diagram, say the whole thing in words in the text as well.
+			Explain that this is the forward model of the validation, not of CHIPPR! contrast this with Fig. 1 (joke about simple model and complex validation?)}
+			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
+			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation I perform, and rounded rectangles indicate a quantity created by the forward model.
+			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.}
 		\label{fig:flowchart}
@@
 \end{figure*}
-
-The true redshift distribution used in these tests is a particular instance of 
-the gamma function
+% \aim{clarify caption}
+
+The true redshift distribution used in these tests is a particular instance of the gamma function
 \begin{equation}
 \label{eqn:gamma}
-n^{\dagger}(z) = \frac{1}{2 c_{z}} \left(\frac{z}{c_{z}}\right)^{2}\ 
-\exp\left[-\frac{z}{c_{z}}\right]
-\end{equation}
-with $c_{z} = 0.3$, because it has been used in forecasting studies for \des\ 
-and \lsst.
-
-The mock data emulates the three sources of error of highest concern to the 
-\pz\ community that are explored in detail later in this section: intrinsic 
-scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and 
-systematic bias (\Sect{sec:bias}).
-Tests including all three effects at the tolerance levels of \lsst\ (see 
-Table~\ref{intro:tab:tab:lsstsrd}) are presented in \Sect{sec:results}.
-\Fig{fig:mega_scatter} illustrates these three effects individually at twice 
-the tolerance of \lsst\ for demonstrative purposes, hearkening back to 
-Figure~\ref{intro:fig:fig:pedagogical_scatter}.
-We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to 
-be a priority for the \pz\ community.
-
-\begin{figure}
-	\begin{center}
-
-\includegraphics[width=0.3\textwidth]{figures/chippr/scatter_scatplot.png}
-
-\includegraphics[width=0.3\textwidth]{figures/chippr/outlier_scatplot.png}
-
-\includegraphics[width=0.3\textwidth]{figures/chippr/bias_scatplot.png}
-		\caption{The joint probability space of true and estimated 
-redshift for the three concerning \pz\ systematics at twice the level of the 
-\lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and 
-bias (right).
-			The main panel of each shows samples (black points) in 
-the space of mock data and redshift, akin to the standard scatterplots of true 
-and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal 
-(thin red line), and posterior probabilities evaluated at the given estimated 
-redshift (colored step functions).
-			The insets show marginal histograms (gray) in each 
-dimension, that can be compared with the true \nz\ used to make the figure 
-(blue curve) to see the effect of the isolated systematic, as well as the 
-implicit prior (red line).
+n^{\dagger}(z) = \frac{1}{2 c_{z}} \left(\frac{z}{c_{z}}\right)^{2}\ \exp\left[-\frac{z}{c_{z}}\right]
+\end{equation}
+with $c_{z} = 0.3$, because it has been used in forecasting studies for \des\ and \lsst.
+%\aim{I learned this from talking to people and don't know of a published source that talks about the nitty gritty of the internal validation tests performed before there was data.}
+
+The mock data emulates the three sources of error of highest concern to the \pz\ community that are explored in detail later in this section: intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and systematic bias (\Sect{sec:bias}).
+Tests including all three effects at the tolerance levels of \lsst\ (see \aim{fix reference})
+%Table~\ref{intro:tab:tab:lsstsrd}) 
+are presented in \Sect{sec:results}.
+\Fig{fig:mega_scatter} illustrates these three effects individually at twice the tolerance of \lsst\ for demonstrative purposes, hearkening back to \aim{fix reference}.
+%Figure~\ref{intro:fig:fig:pedagogical_scatter}.
+We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to be a priority for the \pz\ community.
+
+\aim{TODO: add in the plot from thesis intro}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
+		\includegraphics[width=0.32\textwidth]{figures/chippr/single_uout-mega_scatter.png}
+		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
+		\caption{\aim{TODO: Add watermark of "mock data" in UL corner, same for "results of inference" on other kind of plot.}
+			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
+			The main panel of each shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal (thin red line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
+			The insets show marginal histograms (gray) in each dimension, that can be compared with the true \nz\ used to make the figure (blue curve) to see the effect of the isolated systematic, as well as the implicit prior (red line).
 		}
@@
 	\end{center}
-\end{figure}
-
-The hyperprior distribution chosen for these tests is a multivariate normal 
-distribution with mean $\vec{\mu}$ equal to the implicit prior $\ndphi^{*}$ and 
-covariance
+\end{figure*}
+
+The hyperprior distribution chosen for these tests is a multivariate normal distribution with mean $\vec{\mu}$ equal to the implicit prior $\ndphi^{*}$ and covariance
 \begin{equation}
 \label{eqn:priorcov}
-\Sigma_{k,k'} = q\ \exp[-\frac{e}{2}\ (\bar{z}_{k}-\bar{z}_{k'})^{2}]\ +\ 
-t\delta(k,k')
-\end{equation}
-inspired by one used in Gaussian processes, where $k$ and $k'$ are indices 
-ranging from $1$ to $K$ and $q=1.0$, $e=100.0$, and $t=q\cdot10^{-5}$ are 
-constants chosen to permit draws from this prior distribution to produce shapes 
-similar to that of a true $\tilde{\ndphi}$.  
-We adapt the full log-posterior of \Eq{eqn:final} to the chosen binning of 
-redshift space.
-
-The sampler is initialized with $W=100$ walkers each with a value chosen from a 
-Gaussian distribution of identity covariance around a sample from the 
-hyperprior distribution.
+\Sigma_{k,k'} = q\ \exp[-\frac{e}{2}\ (\bar{z}_{k}-\bar{z}_{k'})^{2}]\ +\ t\delta(k,k')
+\end{equation}
+inspired by one used in Gaussian processes, where $k$ and $k'$ are indices ranging from $1$ to $K$ and $q=1.0$, $e=100.0$, and $t=q\cdot10^{-5}$ are constants chosen to permit draws from this prior distribution to produce shapes similar to that of a true $\tilde{\ndphi}$.
+We adapt the full log-posterior of \Eq{eqn:final} to the chosen binning of redshift space.
+
+%An example of such samples from the prior are shown in \Fig{fig:prior}.
+%
+%\begin{figure}
+%%	\includegraphics[width=0.5\textwidth]{figures/chippr/null_priorsamps.pdf}
+%	\caption{\aim{I need to remake this one because it uses the wrong notation and I stopped making it automatically a while ago.}
+%		Samples (colored lines) of $\pr{z \gvn \ndphi}$ where each $\ndphi$ is drawn from the hyperprior distribution $\pr{\ndphi}$ given in \Eq{eqn:priorcov}.}
+%	\figlabel{fig:prior}
+%\end{figure}
+% \aim{Add back the prior figure!}
+
+The sampler is initialized with $W=100$ walkers each with a value chosen from a Gaussian distribution of identity covariance around a sample from the hyperprior distribution.
 
@@
 
-\Fig{fig:pzs-scatter} shows some examples of \pzpdf s generated with only the 
-systematic of intrinsic scatter, at the level of the \lsst\ requirements on the 
-left and twice that on the right.
-One can see that the histogram of redshift estimates is broader than that of 
-true redshifts, and that the effect is substantially more pronounced by just 
-doubling the intrinsic scatter from the level of the \lsst\ requirements.
-
-\begin{figure}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_scatter1.png}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_scatter2.png}
-	\caption{
-		Examples of mock \pzpdf s (colored lines) generated with 
-intrinsic scatter at the \lsst\ requirements (left) and twice the \lsst\ 
-requirements (right), including samples from the probability space of true and 
-observed redshift (black points), \pzpdf s (colored lines), the true redshifts 
-of the example \pzpdf s (black vertical lines).
-		A histogram (gray) of points in each dimension is shown in the 
-respective inset, with the true redshift distribution (blue curve) and implicit 
-prior (red curve).
+%Several factors contribute to photometric redshifts' intrinsic scatter.  
+%Distant galaxies are dimmer compared to galaxies of identical luminosity that are closer, driving up photometric errors in flux-limited surveys.  
+%The nature of the galaxy sample at higher redshifts also changes, meaning the generation of the photometric redshift posterior based on an a locally-calibrated SED template library or spectroscopically-confirmed training set is more likely to be inappropriate, leading to broader features.
+%In general, the galaxies that could not have been observed spectroscopically will have different and noisier photo-$z$ likelihoods than those that could fall into a spectroscopic training set (or spectroscopically derived template library).
+%This effect may be stronger for high-redshift galaxies.
+
+\Fig{fig:pzs-scatter} shows some examples of \pzpdf s generated with only the systematic of intrinsic scatter, at the level of the \lsst\ requirements on the left and twice that on the right.
+One can see that the histogram of redshift estimates is broader than that of true redshifts, and that the effect is substantially more pronounced by just doubling the intrinsic scatter from the level of the \lsst\ requirements.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter1.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter2.png}
+	\caption{
+		Examples of mock \pzpdf s (colored lines) generated with intrinsic scatter at the \lsst\ requirements (left) and twice the \lsst\ requirements (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
+		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
 	}
 	\label{fig:pzs-scatter}
-\end{figure}
-
-\Fig{fig:results-scatter} shows the \nz\ recovered by \Chippr\ and the 
-alternative approaches.
-As expected, the estimates of \nz\ based on the modes of the \pzpdf s and 
-stacking are broader than the marginalized maximum likelihood estimator from 
-\chippr, with more broadening as the intrinsic scatter increases.
-\Chippr's \mmle\ is robust to intrinsic scatter and is unaffected by increased 
-intrinsic scatter, though the \Chippr\ posterior distribution on the redshift 
-distribution is itself broader for the higher intrinsic scatter case than for 
-the \lsst\ requirements.
-The broadening of the alternative estimators corresponds to a loss of 3-4 times 
-as many nats of information about \nz\ for the \lsst\ requirements relative to 
-the \mmle\ of \Chippr.
-
-\begin{figure}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/results_scatter1.png}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/results_scatter2.png}
-	\caption{
-		The results of \Chippr\ (samples in light blue and optimization 
-in dark blue) and the alternative approaches (the stacked estimator in red and 
-the histogram of modes in yellow) on \pzpdf s with intrinsic scatter of the 
-\lsst\ requirements (left) and twice that (right), with the true redshift 
-density (black curve) and implicit prior (gray curve).
-		\Chippr\ is robust to intrinsic scatter, but the alternatives 
-suffer from overly broad \nz\ estimates that worsen with increasing intrinsic 
-scatter.
+	\end{center}
+\end{figure*}
+
+\Fig{fig:results-scatter} shows the \nz\ recovered by \Chippr\ and the alternative approaches.
+As expected, the estimates of \nz\ based on the modes of the \pzpdf s and stacking are broader than the marginalized maximum likelihood estimator from \chippr, with more broadening as the intrinsic scatter increases.
+\Chippr's \mmle\ is robust to intrinsic scatter and is unaffected by increased intrinsic scatter, though the \Chippr\ posterior distribution on the redshift distribution is itself broader for the higher intrinsic scatter case than for the \lsst\ requirements.
+The broadening of the alternative estimators corresponds to a loss of 3-4 times as many nats of information about \nz\ for the \lsst\ requirements relative to the \mmle\ of \Chippr.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter1.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter2.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with intrinsic scatter of the \lsst\ requirements (left) and twice that (right), with the true redshift density (black curve) and implicit prior (gray curve).
+		\Chippr\ is robust to intrinsic scatter, but the alternatives suffer from overly broad \nz\ estimates that worsen with increasing intrinsic scatter.
 	}
 	\label{fig:results-scatter}
-\end{figure}
+	\end{center}
+\end{figure*}
 
@@
 
-As was covered in \Sect{sec:intro}, catastrophic outliers tend to be 
-distributed non-uniformly across the space of observed and true redshift.
-However, the \lsst\ requirements do not specify details for a distribution of 
-outliers to which they were tuned, and it is still instructive to examine the 
-impact of uniform outliers on the inference of \nz, so we begin by addressing 
-uniformly distributed outliers before considering more realistic outlier 
-distributions.
-
-A uniformly distributed population of outliers was simulated by giving every 
-sample in true redshift a $10\%$ chance of having an observed redshift drawn 
-from a uniform distribution rather than the Gaussian about the true redshift.
-Though this results in slightly less than the $10\%$ catastrophic outlier rate, 
-it can be done independently of the definition of the standard deviation so was 
-implemented for demonstrative purposes.
-\Fig{fig:uniform-outliers} shows examples of \pzpdf s from a uniformly 
-distributed outlier population at the level of the \lsst\ requirements (left) 
-as well as the results of \Chippr\ and other \nz\ estimation methods (right).
-The intrinsic scatter of the tests in this section does not increase with 
-redshift as indicated in Table~\ref{intro:tab:tab:lsstsrd}, in order to isolate 
-the effect of outliers, and is instead held at a constant $\sigma_{z} = 0.02$.
+As was covered in \Sect{sec:intro}, catastrophic outliers tend to be distributed non-uniformly across the space of observed and true redshift.
+However, the \lsst\ requirements do not specify details for a distribution of outliers to which they were tuned, and it is still instructive to examine the impact of uniform outliers on the inference of \nz, so we begin by addressing uniformly distributed outliers before considering more realistic outlier distributions.
+
+A uniformly distributed population of outliers was simulated by giving every sample in true redshift a $10\%$ chance of having an observed redshift drawn from a uniform distribution rather than the Gaussian about the true redshift.
+Though this results in slightly less than the $10\%$ catastrophic outlier rate, it can be done independently of the definition of the standard deviation so was implemented for demonstrative purposes.
+\Fig{fig:uniform-outliers} shows examples of \pzpdf s from a uniformly distributed outlier population at the level of the \lsst\ requirements (left) as well as the results of \Chippr\ and other \nz\ estimation methods (right).
+The intrinsic scatter of the tests in this section does not increase with redshift as indicated in \aim{fix reference},
+%Table~\ref{intro:tab:tab:lsstsrd}
+in order to isolate the effect of outliers, and is instead held at a constant $\sigma_{z} = 0.02$.
 
 \begin{figure}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/single_uout_mega_scatter.p
-ng}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/single_uout_log_estimators
-.png}
-	\caption{
-		Left: Examples of \pzpdf s with a uniform catastrophic outlier 
-population at the level of the \lsst\ requirements, including samples from the 
-probability space of true and observed redshift (black points), \pzpdf s 
-(colored curves), and the true redshifts of the example \pzpdf s (black 
-vertical lines), with marginal histograms (gray) for each dimension with the 
-true redshift distribution (blue curve) and implicit prior (red curve) in the 
-insets.
-		Right: The results of \Chippr\ (samples in light blue, 
-optimization in dark blue) and the alternative approaches (the stacked 
-estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly 
-distributed catastrophic outliers, with the true redshift density (black curve) 
-and implicit prior (gray curve).
-		The presence of the catastrophic outlier population broadens 
-the histogram of modes and stacked estimator of the redshift distribution, but 
-the result of \Chippr\ is unbiased.
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_log_estimators.png}
+	\caption{
+		Left: Examples of \pzpdf s with a uniform catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+		Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
 	}
 	\label{fig:uniform-outliers}
+	\end{center}
 \end{figure}
 
-\Fig{fig:uniform-outliers} shows that at the level of the \lsst\ requirements, 
-the alternative estimators are overly broad, whereas \Chippr's \mmle\ yields an 
-unbiased estimate of \nz.
-Further, the result of stacking is even broader than that of the histogram of 
-modes, corresponding to ten times the information loss of \Chippr's \mmle, 
-making it worse than the most naive reduction of \pzpdf s to point estimates.
-
-When one thinks of the \pzpdf s of catastrophic outliers, however, what comes 
-to mind is multimodal \pzpdf s, wherein reducing \pzpdf s to point estimates to 
-make a standard scatterplot of the true and observed redshifts leads to 
-substantial probability density off the diagonal.
-These coordinated catastrophic outliers may be emulated in the joint 
-probability space of true and estimated redshifts by using a mixture of the 
-unbiased diagonal defined by the intrinsic scatter and an additional Gaussian 
-in one dimension, with constant observed redshift for a template-fitting code 
-and constant true redshift for a machine learning code.
-
-In the case of a catastrophic outlier population like that anticipated of 
-template-fitting codes, $10\%$ of all galaxies have their observed redshift at 
-a particular value unrelated to their true redshift, illustrated in the left 
-panel of \Fig{fig:nonuniform-outliers-data}.
-This case is subject to the same caveat as the uniformly distributed outliers 
-when it comes to the \lsst\ requirement.
-It is less straightforward to emulate catastrophic outliers like those 
-anticipated of a machine learning code, those that are truly multimodal.
-The testing conditions here, illustrated in the right panel of 
-\Fig{fig:nonuniform-outliers-data}, gives $10\%$ of galaxies at the redshift 
-affected by outliers an observed redshift that is uniformly distributed 
-relative to the true redshift, meaning that far fewer than $10\%$ of all 
-galaxies in the sample are catastrophic outliers.
+\Fig{fig:uniform-outliers} shows that at the level of the \lsst\ requirements, the alternative estimators are overly broad, whereas \Chippr's \mmle\ yields an unbiased estimate of \nz.
+Further, the result of stacking is even broader than that of the histogram of modes, corresponding to ten times the information loss of \Chippr's \mmle, making it worse than the most naive reduction of \pzpdf s to point estimates.
+
+When one thinks of the \pzpdf s of catastrophic outliers, however, what comes to mind is multimodal \pzpdf s, wherein reducing \pzpdf s to point estimates to make a standard scatterplot of the true and observed redshifts leads to substantial probability density off the diagonal.
+These coordinated catastrophic outliers may be emulated in the joint probability space of true and estimated redshifts by using a mixture of the unbiased diagonal defined by the intrinsic scatter and an additional Gaussian in one dimension, with constant observed redshift for a template-fitting code and constant true redshift for a machine learning code.
+
+In the case of a catastrophic outlier population like that anticipated of template-fitting codes, $10\%$ of all galaxies have their observed redshift at a particular value unrelated to their true redshift, illustrated in the left panel of \Fig{fig:nonuniform-outliers-data}.
+This case is subject to the same caveat as the uniformly distributed outliers when it comes to the \lsst\ requirement.
+It is less straightforward to emulate catastrophic outliers like those anticipated of a machine learning code, those that are truly multimodal.
+The testing conditions here, illustrated in the right panel of \Fig{fig:nonuniform-outliers-data}, gives $10\%$ of galaxies at the redshift affected by outliers an observed redshift that is uniformly distributed relative to the true redshift, meaning that far fewer than $10\%$ of all galaxies in the sample are catastrophic outliers.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_mega_scatter.png}
+	\caption{
+		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+	}
+	\label{fig:nonuniform-outliers-data}
+	\end{center}
+\end{figure*}
+
+The results of \Chippr\ and the alternative estimators of \nz\ are presented in \Fig{fig:nonuniform-outliers-results}.
+The most striking feature is that the histogram of modes is highly sensitive to both outlier populations, producing a severe overestimate in the case of an outlier population like those seen in template-fitting codes and a severe underestimate in the case of an outlier population like those seen in machine learning codes, corresponding to a twenty-fold loss of information compared to the \Chippr\ \mmle\ in both cases.
+The effect on the stacked estimator of \nz\ is more subtle though still concerning.
+In the case of outliers like those resulting from template-fitting, the stacked estimator is overly broad even without realistic intrinsic scatter, resulting in ten times the information loss compared to the \Chippr\ \mmle, and in the case of outliers like those resulting from machine learning, the stacked estimator features an overestimate at the redshift affected by the outlier population, resulting in about five times the information loss as the \Chippr\ \mmle.
+The \Chippr\ \mmle, however, appears unbiased and withstands these effects, and the breadth of the distribution of samples of \nz\ is invariant.
 
 \begin{figure}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_eout_mega_scatter.p
-ng}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_rout_mega_scatter.p
-ng}
-	\caption{
-		Examples of \pzpdf s with a catastrophic outlier population 
-like that seen in template-fitting \pzpdf\ codes (left) and machine learning 
-\pzpdf\ codes (right), including samples from the probability space of true and 
-observed redshift (black points), \pzpdf s (colored curves), and the true 
-redshifts of the example \pzpdf s (black vertical lines), with marginal 
-histograms (gray) for each dimension with the true redshift distribution (blue 
-curve) and implicit prior (red curve) in the insets.
-	}
-	\label{fig:nonuniform-outliers-data}
-\end{figure}
-
-The results of \Chippr\ and the alternative estimators of \nz\ are presented in 
-\Fig{fig:nonuniform-outliers-results}.
-The most striking feature is that the histogram of modes is highly sensitive to 
-both outlier populations, producing a severe overestimate in the case of an 
-outlier population like those seen in template-fitting codes and a severe 
-underestimate in the case of an outlier population like those seen in machine 
-learning codes, corresponding to a twenty-fold loss of information compared to 
-the \Chippr\ \mmle\ in both cases.
-The effect on the stacked estimator of \nz\ is more subtle though still 
-concerning.
-In the case of outliers like those resulting from template-fitting, the stacked 
-estimator is overly broad even without realistic intrinsic scatter, resulting 
-in ten times the information loss compared to the \Chippr\ \mmle, and in the 
-case of outliers like those resulting from machine learning, the stacked 
-estimator features an overestimate at the redshift affected by the outlier 
-population, resulting in about five times the information loss as the \Chippr\ 
-\mmle.
-The \Chippr\ \mmle, however, appears unbiased and withstands these effects, and 
-the breadth of the distribution of samples of \nz\ is invariant.
-
-\begin{figure}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_eout_log_estimators
-.png}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_rout_log_estimators
-.png}
-	\caption{
-		The results of \Chippr\ (samples in light blue and optimization 
-in dark blue) and the alternative approaches (the stacked estimator in red, the 
-histogram of modes in yellow) on \pzpdf s with catastrophic outliers like those 
-seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ 
-codes (right) to the \lsst\ requirements, with the true redshift density (black 
-curve) and implicit prior (gray curve).
-		Though the histogram of modes is most sensitive to a 
-catastrophic outlier population, the stacked estimator also overestimates \nz\ 
-under (machine learning-like outliers) and beyond (template fitting-like 
-outliers).
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_log_estimators.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_log_estimators.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with catastrophic outliers like those seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right) to the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
+		Though the histogram of modes is most sensitive to a catastrophic outlier population, the stacked estimator also overestimates \nz\ under (machine learning-like outliers) and beyond (template fitting-like outliers).
 	}
 	\label{fig:nonuniform-outliers-results}
+	\end{center}
 \end{figure}
@@
 
-Systematic bias in \pz\ point estimates is a concern for \lsst's cosmology 
-results, for the same reasons explored in \citet{hoyle_dark_2017}.
-However, in the context of \pzpdf s, the notion of redshift bias is a form of 
-model misspecification.
-Consider that if bias were included in the framework of 
-Figure~\ref{intro:fig:fig:pedagogical_scatter}, a simple linear transformation 
-of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + 
-z_{\mathrm{phot}})$ would eliminate the bias.
-Regardless, for completeness, a test at ten times the bias of the \lsst\ 
-requirements, with no redshift-dependent intrinsic scatter nor catastrophic 
-outliers, is provided in \Fig{fig:bias}.
+Systematic bias in \pz\ point estimates is a concern for \lsst's cosmology results, for the same reasons explored in \citet{hoyle_dark_2017}.
+However, in the context of \pzpdf s, the notion of redshift bias is a form of model misspecification.
+Consider that if bias were included in the framework of \aim{fix reference},
+%Figure~\ref{intro:fig:fig:pedagogical_scatter}
+a simple linear transformation of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + z_{\mathrm{phot}})$ would eliminate the bias.
+Regardless, for completeness, a test at ten times the bias of the \lsst\ requirements, with no redshift-dependent intrinsic scatter nor catastrophic outliers, is provided in \Fig{fig:bias}.
 
 \begin{figure}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_neghibias_mega_scat
-ter.png}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/thesis_neghibias_log_estim
-ators.png}
-	\caption{
-		Left: Examples of \pzpdf s with ten times the bias of the 
-\lsst\ requirements, including samples from the probability space of true and 
-observed redshift (black points), \pzpdf s (colored curves), and the true 
-redshifts of the example \pzpdf s (black vertical lines), with marginal 
-histograms (gray) for each dimension with the true redshift distribution (blue 
-curve) and implicit prior (red curve) in the insets.
-		Right: The results of \Chippr\ (samples in light blue, 
-optimization in dark blue) and the alternative approaches (the stacked 
-estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly 
-distributed catastrophic outliers, with the true redshift density (black curve) 
-and implicit prior (gray curve).
-		The impact of bias at even ten times the level of the \lsst\ 
-requirements is almost imperceptible on all estimators, though the \Chippr\ 
-\mmle\ minimizes the information loss regardless.
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_log_estimators.png}
+	\caption{
+		Left: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+		Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+		The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
 	}
 	\label{fig:bias}
+	\end{center}
 \end{figure}
 
-As expected based on consistency with the forward model, \Chippr\ is completely 
-resistant to bias, and the alternative estimators are only weakly affected, 
-with information loss two and four times greater than that of the \Chippr\ 
-\mmle\ for the histogram of modes and stacked estimator respectively.
+As expected based on consistency with the forward model, \Chippr\ is completely resistant to bias, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
+%\aim{I'm not sure I effectively explained why \Chippr\ is expected to be unaffected by bias of this form.}
 
@@
 
-\chippr\ can handle any implicit prior with support over the redshift range 
-where \nz\ is defined, but some archetypes of implicit prior are more likely to 
-be encountered in the wilds of \pzpdf\ codes.
-Ideally, an uninformative implicit prior would be used, although it may be 
-complicated to compute from the covariances of the raw data.  
-Template-fitting codes have an explicit prior input formed by redshifting a 
-small number of templates, leading to a highly nonuniform but 
-physically-motivated interim prior.
-Machine learning approaches tend to be trained on previously observed data sets 
-that are biased towards low redshift, which biases the implicit prior towards 
-low redshift.
-Some efforts have been made to modify an observationally informed implicit 
-prior so that it is more representative of the photometric data for which 
-redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal 
-to the true \nz, it will propagate to the results of traditional \nz\ 
-estimation methods.
-
-\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring 
-implicit prior emulating that of a machine learning approach to \pz\ estimation 
-(left panel) and a more complex interim prior emulating that of a 
-template-fitting \pz\ method (right panel).
-One can see that the \pzpdf s take different shapes from one another even 
-though the marginal histograms of the points are identical.
-The machine learning-like implicit prior has been modified to have nonzero 
-value at high redshift because the implicit prior must be strictly positive 
-definite for the \Chippr\ model to be valid.
+\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
+Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
+Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
+%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.
+%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.
+%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.  
+Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
+% \aim{referenceh chapter 3 for complication of algorithm}
+Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
+%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.  
+%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.  
+%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.
+
+\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring implicit prior emulating that of a machine learning approach to \pz\ estimation (left panel) and a more complex interim prior emulating that of a template-fitting \pz\ method (right panel).
+One can see that the \pzpdf s take different shapes from one another even though the marginal histograms of the points are identical.
+The machine learning-like implicit prior has been modified to have nonzero value at high redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
 
 \begin{figure}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_trpr.png}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/samplepzs_tmpr.png}
-	\caption{
-		Examples of mock \pzpdf s (colored lines) generated with a 
-machine learning-like implicit prior (left) and a template-fitting-like 
-implicit prior (right), including samples from the probability space of true 
-and observed redshift (black points), \pzpdf s (colored lines), the true 
-redshifts of the example \pzpdf s (black vertical lines).
-		A histogram (gray) of points in each dimension is shown in the 
-respective inset, with the true redshift distribution (blue curve) and implicit 
-prior (red curve).
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_trpr.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_tmpr.png}
+	\caption{
+		Examples of mock \pzpdf s (colored lines) generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
+		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
 	}
 	\label{fig:pzs-priors}
+	\end{center}
 \end{figure}
 
-\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional 
-methods on \pzpdf s generated with nontrivial implicit priors.
-In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift 
-distribution, and the distribution of \nz\ parameter values reflects higher 
-uncertainty where the implicit prior undergoes large changes in derivative.
-The alternatives, on the other hand, are biased by the implicit prior except 
-where it is flat, in the case of high redshifts for the machine learning-like 
-implicit prior, resulting in over $1,000$ times the information loss on \nz\ 
-for the machine learning-like implicit prior and some $5-20$ times the 
-information loss for the template fitting-like implicit prior, relative to the 
-\Chippr\ \mmle.
-
-\begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/chippr/results_trpr.png}
-	\includegraphics[width=0.49\textwidth]{figures/chippr/results_tmpr.png}
-	\caption{
-		The results of \Chippr\ (samples in light blue and optimization 
-in dark blue) and the alternative approaches (the stacked estimator in red and 
-the histogram of modes in yellow) on \pzpdf s with an implicit prior like that 
-of machine learning \pzpdf\ approaches (left) and an implicit prior like that 
-of template-fitting \pzpdf\ codes (right), with the true redshift density 
-(black curve) and implicit prior (gray curve).
-		\Chippr\ is robust to a nontrivial implicit prior, but the 
-alternatives are biased toward the implicit prior.
+\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzpdf s generated with nontrivial implicit priors.
+In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift distribution, and the distribution of \nz\ parameter values reflects higher uncertainty where the implicit prior undergoes large changes in derivative.
+The alternatives, on the other hand, are biased by the implicit prior except where it is flat, in the case of high redshifts for the machine learning-like implicit prior, resulting in over $1,000$ times the information loss on \nz\ for the machine learning-like implicit prior and some $5-20$ times the information loss for the template fitting-like implicit prior, relative to the \Chippr\ \mmle.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_trpr.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_tmpr.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with an implicit prior like that of machine learning \pzpdf\ approaches (left) and an implicit prior like that of template-fitting \pzpdf\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
+		\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
 	}
 	\label{fig:results-priors}
-\end{figure}
-
-The main implication of the response of \nz\ estimates to a nontrivial implicit 
-prior is that the implicit prior must be accounted for when using \pzpdf\ 
-catalogs.
+	\end{center}
+\end{figure*}
+
+%\aim{Move \Sect{sec:violations} here?}
+
+The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzpdf\ catalogs.
 
@@
 
-The experiments of \Sect{sec:alldata} isolate the potential sources of error in 
-\nz\ estimation one at a time.
-Now, we stress-test \Chippr\ by investigating two realistically complex cases, 
-one in which the \nz\ estimates are made tomographically as in a modern 
-cosmological analysis (\Sect{sec:lsstdemo}) and one in which the \nz\ 
-estimators are not provided with the same implicit prior used to generate the 
-\pzpdf\ catalog (\Sect{sec:violations}).
+The experiments of \Sect{sec:alldata} isolate the potential sources of error in \nz\ estimation one at a time.
+Now, we stress-test \Chippr\ by investigating two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one in which the \nz\ estimators are not provided with the same implicit prior used to generate the \pzpdf\ catalog (\Sect{sec:violations}).
 
@@
 
-It is of interest to explore the impact of incorrectly estimated \nz\ on the 
-cosmological inference to answer the question of how wrong we will be in our 
-understanding of the universe if we incorrectly constrain \nz.
-To test the impact of these uncertainties, I simulated mock data with all three 
-effects with which \lsst\ is concerned at the levels of 
-Table~\ref{intro:tab:tab:lsstsrd} and propagated the results of \Chippr\ and 
-the other estimators to a Fisher matrix forecast using \cosmolike\ 
-\citep{krause_cosmolike_2017}, a publicly available cosmological forecasting 
-code.
-Though redshift tomography is non-physical, as redshift is a continuous random 
-variable, and binning in estimated redshift introduces poorly understood 
-systematic error, I perform this analysis as an example of how it affects the 
-current standard in how cosmological parameters are constrained by galaxy 
-surveys, rather than how I think they ought to be constrained.
-
-\begin{figure*}
-	\begin{center}
-
+\aim{TODO: clarify the scope: the "binned" samples are already determined by some observational property but redshift distributions are probabilistic as follows.}
+
+It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
+To test the impact of these uncertainties, I simulated mock data with all three effects with which \lsst\ is concerned at the levels of \aim{fix reference}
+%Table~\ref{intro:tab:tab:lsstsrd} 
+and propagated the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
+Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, I perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how I think they ought to be constrained.
+% \aim{don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space}
+
+\begin{figure*}
+	\begin{center}
 		\includegraphics[width=0.74\textwidth]{figures/chippr/cosmolike_inputs.png}
 		\caption{
-			The \lsst-like tomographic binning and true redshift 
-distribution, where the truth (solid) is a PDF evaluated on a fine grid of 
-$350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn 
-(dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced 
-bins, for four different tomographic bins (colors).
+			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
 		}
@@
 
-I considered a set of tomographically binned \nz\ and cosmological parameter 
-covariance matrices used for \desc\ forecasting, for which the true \nz\ in 
-each pre-defined bin is already provided in the form of an evaluation of a 
-function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
-First, I binned them down to a piecewise constant parameterization with a 
-manageable $35$ parameters for \chippr's sampling capabilities.
-Next, I drew $10^{4}$ true redshifts from the binned true \nz\ for each 
-tomographic bin.
+I considered a set of tomographically binned \nz\ and cosmological parameter covariance matrices used for \desc\ forecasting, for which the true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+First, I binned them down to a piecewise constant parameterization with a manageable $35$ parameters for \chippr's sampling capabilities.
+Next, I drew $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
 The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
-I emulated \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in 
-each bin using the procedure of \Fig{fig:flowchart} with all three effects of 
-Table~\ref{intro:tab:tab:lsstsrd} at their given levels.
+I emulated \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of \aim{fix reference}
+%Table~\ref{intro:tab:tab:lsstsrd}
+at their given levels.
 Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
@@
 	\begin{center}
-
-\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.
-png}
-
-\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.
-png}		
-\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.
-png}
-
-\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.
-png}
-		\caption{As in \Fig{fig:mega_scatter}, with a different 
-tomographic bin in each panel and the three effects of intrinsic scatter, 
-uniformly distributed catastrophic outliers, and bias at the levels of the 
-\lsst\ SRD, given in Table~\ref{intro:tab:tab:lsstsrd}.
+		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
+		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in \aim{fix reference}.
+%			 Table~\ref{intro:tab:tab:lsstsrd}.
+			%		\aim{Make this one big plot instead of four little ones to eliminate repeated insets and legend, and make the axis labels bigger.
 		}
@@
 \end{figure*}
-
-I then made a point estimate of \nz\ using \Chippr\ as well as the alternative 
-methods on the \pzpdf\ catalog for each tomographic bin, shown in 
-\Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from 
-a single \nz\ result, rather than samples from the full posterior probability 
-density of possible \nz.
-Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability 
-units, unlike all other plots in this paper, to better show the behavior at low 
-probability.
+% \aim{how to get the bias? Ask Elisabeth about what I'd need for bias}
+
+I then made a point estimate of \nz\ using \Chippr\ as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
+Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
 The excessive breadth of the alternative estimators can be seen quite plainly.
@@
 	\begin{center}
-
-\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimator
-s.png}
-
-\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimator
-s.png}		
-\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimator
-s.png}
-
-\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimator
-s.png}
-		\caption{
-			The \chippr-derived and other estimators of \nz\ in 
-each tomographic bin, with the true \nz\ (black), the implicit prior (gray), 
-stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ 
-(blue).
-			The result of stacking is far too broad for \lsst-like 
-\pzpdf s, even moreso than the simplistic histogram of modes.
+		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
+		\caption{
+			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
+			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
+			%			\aim{Make this one big plot instead of four little ones to eliminate repeated axis labels and legend, and make the legend text bigger.}
 		}
@@
 
-I then used the different estimators of \nz\ in a cosmological forecasting 
-procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, 
-$w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
-Though there are also slight differences in the angle of the error ellipses, 
-most of the differences are due to broadening of the contours under the 
-alternative estimators relative to \Chippr, which are almost indistinguishable 
-from those derived by using the true redshift distribution in each bin.
-The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all 
-parameters except $\Omega_{b}$ and $H_{0}$.
-Stacking, however, outperforms the histogram of modes for all parameters except 
-$\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
-Though the true values of the parameters themselves were not accessible with 
-the Fisher matrix-based framework, I calculated the seven-dimensional KLD for 
-the three \nz\ estimators relative to the constraints derived from the true 
-\nz, showing that \Chippr\ preserves information $200-800$ times better than 
-the alternatives, with the histogram of modes doing about four times better 
-than stacking.
-
-\begin{figure*}
-	\begin{center}
-
-\includegraphics[width=0.99\textwidth]{figures/chippr/final_plot.png}
-		\caption{
-			The result of propagating the estimators of \nz\ by 
-stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true 
-\nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
-			For all parameters considered, \Chippr\ yields contours 
-no broader than those corresponding to the true \nz, whereas for most 
-parameters, stacking and the histogram of modes yield broader contours.
+I then used the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
+Though there are also slight differences in the angle of the error ellipses, most of the differences are due to broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
+The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
+Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
+Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, I calculated the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
+
+\begin{figure*}
+	\begin{center}
+%		\includegraphics[width=0.95\textwidth]{figures/chippr/final_plot.png}
+		\caption{
+			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
+			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
+			\aim{You can't see the \Chippr\ contours because they're exteemely close to the truth}
 		}
@@
 \end{figure*}
+
+%\aim{Hogg says ``I think it would be good to talk a little quantitatively about where people need to know N(z) and other one-point statistics, and how much they will get various things wrong if they don't know these correctly.
+%	And situate that discussion within the current context of cosmological parameter estimation and precision cosmology.''
+%	Did I successfully answer the question?}
 
@@
 
-In this test, without tomographic binning, the \pzip s are made to the \lsst\ 
-requirements but the implicit prior used for the inference is not the same as 
-the implicit prior used for generating the data.
-\Pzpdf\ codes do not generally provide their implicit prior, with the exception 
-of some template-fitting techniques for which it is a known input.
-If we naively used the \pzpdf\ catalog produced by a generic machine learning 
-code and assumed a flat implicit prior, we would observe the contents of 
-\Fig{fig:mischaracterized}.
-
-\begin{figure*}
-	\begin{center}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/misspecified_mega_scatter.
-png}
-
-\includegraphics[width=0.49\textwidth]{figures/chippr/misspecified_log_estimator
-s.png}
-		\caption{
-			Left: Examples of \pzpdf s with systematics at the 
-level of the \lsst\ requirements with a machine learning-like implicit prior, 
-including samples from the probability space of true and observed redshift 
-(black points), \pzpdf s (colored curves), and the true redshifts of the 
-example \pzpdf s (black vertical lines), with marginal histograms (gray) for 
-each dimension with the true redshift distribution (blue curve) and implicit 
-prior (red curve) in the insets.
-			Right: The results of \Chippr\ (samples in light blue, 
-optimization in dark blue) and the alternative approaches (the stacked 
-estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly 
-distributed catastrophic outliers, with the true redshift density (black curve) 
-and the uniform implicit prior given to \chippr\ (gray curve).
-			When the incorrect implicit prior is provided to 
-\chippr, even Bayesian inference cannot recover the true \nz.
+%\aim{Move \Sect{sec:interim} here?}
+
+In this test, without tomographic binning, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
+\Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
+If we naively used the \pzpdf\ catalog produced by a generic machine learning code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
+		\caption{
+			Left: Examples of \pzpdf s with systematics at the level of the \lsst\ requirements with a machine learning-like implicit prior, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+			Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and the uniform implicit prior given to \chippr\ (gray curve).
+			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
 		}
@@
 
-The results of using a mischaracterized implicit prior are disastrous, causing 
-every estimator, including \Chippr, to be strongly biased.
-The stacked estimator and histogram of modes don't make use of the implicit 
-prior so do no worse than when the implicit prior is accurately provided, but 
-\Chippr\ is sensitive to prior misspecification, which violates the model upon 
-which it is based.
-It is thus crucial that \pzpdf\ methods always characterize and provide the 
-implicit prior.
+The results of using a mischaracterized implicit prior are disastrous, causing every estimator, including \Chippr, to be strongly biased.
+The stacked estimator and histogram of modes don't make use of the implicit prior so do no worse than when the implicit prior is accurately provided, but \Chippr\ is sensitive to prior misspecification, which violates the model upon which it is based.
+It is thus crucial that \pzpdf\ methods always characterize and provide the implicit prior.
+
+% removed ancient investigation on real data, would be hard to redo with the new \chippr\ code on this timescale given that I haven't even looked at the data format in 4 years
 
@@
 
-This study derives and demonstrates a mathematically consistent inference of a 
-one-point statistic, the redshift density function \nz, based on an arbitrary 
-catalog of \pzpdf s.  
-The fully Bayesian method, based in the fundamental laws of probability, begins 
-with a probabilistic graphical model corresponding to equations for the full 
-posterior distribution over the parameters for \nz.  
-The method is validated on mock data and tested in the regime of \lsst\ with 
-promising results, outperforming the traditional stacking estimator at the 
-level of \nz\ as well as in terms of constraining power on the cosmological 
-parameters.
-Not only is this the only mathematically correct approach to the problem, it 
-also recovers the true parameter values better than popular alternatives, as 
-measured by the loss of information in \nz\ and the size of error ellipses in 
-the space of cosmological parameters.
+% new paragraph for what can go wrong, call to community for what to do about it
+% what aspects are knowable and not knowable, if testable how so? outside the scope of paper, data producers be warned
+% a likelihood is better -- give us that if you can!
+% focus on methods that acknowledge probabilistic structure of problem
+
+% implications for tomographic binning are severe
+
+This study derives and demonstrates a mathematically consistent inference of a one-point statistic, the redshift density function \nz, based on an arbitrary catalog of \pzpdf s.  
+The fully Bayesian method, based in the fundamental laws of probability, begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
+The method is validated on mock data and tested in the regime of \lsst\ with promising results, outperforming the traditional stacking estimator at the level of \nz\ as well as in terms of constraining power on the cosmological parameters.
+Not only is this the only mathematically correct approach to the problem, it also recovers the true parameter values better than popular alternatives, as measured by the loss of information in \nz\ and the size of error ellipses in the space of cosmological parameters.
+
+%In the tests on simulated data performed here, the full posterior distribution over the hyperparameters defining $N(z)$ derived by this method is consistent with the true redshift distribution function, making the mean of sampled values an excellent point estimator of $N(z)$.
+%The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
+%\aim{Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
+%The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
+%Tests were also performed on subsets of BOSS DR10 data with results consistent with those of simulations.
 
@@
 \begin{enumerate}
-	\item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are 
-good point estimators of \nz, whereas the histogram of modes is very sensitive 
-to outliers and the stacked estimator is always excessively broad.
-	\item The error bars on the posterior distribution over \nz\ 
-hyperparameters are interpretable and arise naturally under \Chippr, unlike 
-those that may be assumed for the conventional point estimators.
-	\item When the implicit prior is known to be a poor match to the data, 
-only the results of \Chippr\ are satisfactory estimators of the redshift 
-distribution function because they are the only methods that can account for 
-the bias induced on the \pzpdf\ catalog by the method that produces it; this is 
-the most compelling case for the sampler because of the ubiquity of 
-inappropriate interim priors.
+	\item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
+	\item The error bars on the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that may be assumed for the conventional point estimators.
+	%	\item The marginalized maximum likelihood estimator is an excellent estimator for strongly featured redshift distribution function with simple, clean photo-$z$ posteriors; stacking smooths features more than sampling and photo-$z$ point estimation.
+	\item When the implicit prior is known to be a poor match to the data, only the results of \Chippr\ are satisfactory estimators of the redshift distribution function because they are the only methods that can account for the bias induced on the \pzpdf\ catalog by the method that produces it; this is the most compelling case for the sampler because of the ubiquity of inappropriate interim priors.
 \end{enumerate}
 
-By showing that \Chippr\ is effective in recovering the true redshift 
-distribution function and posterior distributions on its parameters from 
-catalogs of \pzpdf s, this work supports the production of \pzpdf s by upcoming 
-photometric surveys such as \lsst\ to enable more accurate inference of the 
-cosmological parameters.  
-We discourage researchers from co-adding \pzpdf s or converting them into point 
-estimates of redshift and instead recommend the use of Bayesian probability to 
-guide the usage of \pzpdf s.  
-We emphasize to those who produce \pzpdf s from data that it is essential to 
-release the implicit prior used in generating this data product in order for 
-proper inference to be conducted by consumers of this information.
-
-The technique herein developed is applicable with minimal modification to other 
-one-point statistics of redshift to which we will apply this method in the 
-future, such as the redshift-dependent luminosity function and weak lensing 
-mean distance ratio.  
-Future work will also include the extension of this fully probabilistic 
-approach to higher-order statistics of redshift such as the two-point 
-correlation function.
-
-\section*{Appendices}
-
-\begin{appendices}
-
-\end{appendices}
+By showing that \Chippr\ is effective in recovering the true redshift distribution function and posterior distributions on its parameters from catalogs of \pzpdf s, this work supports the production of \pzpdf s by upcoming photometric surveys such as \lsst\ to enable more accurate inference of the cosmological parameters.
+We discourage researchers from co-adding \pzpdf s or converting them into point estimates of redshift and instead recommend the use of Bayesian probability to guide the usage of \pzpdf s.  
+We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for proper inference to be conducted by consumers of this information.
+
+The technique herein developed is applicable with minimal modification to other one-point statistics of redshift to which we will apply this method in the future, such as the redshift-dependent luminosity function and weak lensing mean distance ratio.
+Future work will also include the extension of this fully probabilistic approach to higher-order statistics of redshift such as the two-point correlation function.
+
+%%\section*{Appendices}
+%%
+%%\begin{appendices}
+%%
+%%%\section{Mathematical derivation}
+%%%\label{app:math}
+%%%
+%%%\section{Mock data generation}
+%%%\label{app:data}
+%%%
+%%%\textul{Plots: flow chart of forward model}
+%%%
+%%%\begin{figure*}
+%%%	\begin{center}
+%%%		\includegraphics[width=0.75\textwidth]{fig/flowchart.png}
+%%%		\caption{In appendix?}
+%%%		\label{fig:flowchart}
+%%%	\end{center}
+%%%\end{figure*}
+%%
+%%
+%%\end{appendices}
 
@@
 AIM was supported by National Science Foundation grant AST-1517237 and the U.S.
-Department of Energy, Office of Science, Office of Workforce Development for 
-Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) 
-program, administered by the Oak Ridge Institute for Science and Education for 
-the DOE under contract number DE‐SC0014664.
-The authors thank Phil Marshall for advice on relevant examples, Elisabeth 
-Krause for assistance with the \cosmolike code, Mohammadjavad Vakili for 
-statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt 
-for other helpful comments in the development of \Chippr.
-This work was completed with generous nutritional support from the Center for 
-Computational Astrophysics.
+Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) program, administered by the Oak Ridge Institute for Science and Education for the DOE under contract number DE‐SC0014664.
+The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
+% send to DFM for comments
+This work was completed with generous nutritional support from the Center for Computational Astrophysics.
 \end{acknowledgements}

### 2019-05-24 research/paper/draft.tex
fixing formatting

@@
-\RequirePackage[]{lineno}
+%\RequirePackage[]{lineno}
 
@@
 \usepackage{graphicx}
-\usepackage{lineno}
+%\usepackage{lineno}
 
@@
 \newcommand{\Tab}[1]{Table~\ref{#1}}
-\newcommand{\project}[1]{{\textsc{#1}}~}
+\newcommand{\project}[1]{{\textsc{#1}}}
 \newcommand{\lsst}{\project{LSST}}
@@
 \newcommand{\gvn}{\mid}% could use | or \vert
-\newcommand{\integral}[2]{\ensuremath{\int\ #1\ \mathrm{d} #2}}
+\newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
 \newcommand{\sz}{spec-$z$~}
@@
 \begin{document}
-\linenumbers
+%\linenumbers
 
@@
 We conclude by propagating these effects to constraints in the space of cosmological parameters, .
-% be clear that rigor comes at a cost, chippr makes explicit the requirements that the pzpdfs be posteriors under a known prior (which would be required by any justifiable method), be honest that we're not ready
+\aim{be clear that rigor comes at a cost, chippr makes explicit the requirements that the pzpdfs be posteriors under a known prior (which would be required by any justifiable method), be honest that we're not ready for prime time}
 \end{abstract}
@@
 \aim{Some of the figures refer to left/right but are shown as top/bottom \textemdash I'm still working through a LaTeX bug.}
-\aim{The language is very thesis-centric (``me, me, me''), but I'll change it to the standard ``we'' style soon.}
 
@@
 \label{sec:introduction}
+
+Photometric redshift (\pz) estimation has been a staple of studies of galaxy evolution, large-scale structure, and cosmology since its conception decades ago \citep{baum_photoelectric_1962}.  
+An extremely coarse spectrum in the form of photometry in a handful of broadband filters can be an effective substitute for the time- and photon-intensive process of obtaining a spectroscopic redshift (\sz), a procedure that may only be applied to relatively bright galaxies.
+Once the photometric colors are calibrated against either a library of spectral energy distribution (SED) templates or a data set of spectra for galaxies with known redshifts, a correspondence between photometric colors and redshifts may be constructed, forming a trustworthy basis for \pz\ estimation or testing.
+
+Calculations of correlation functions of cosmic shears and galaxy positions that constrain the cosmological parameters require large numbers of high-confidence redshifts of surveyed galaxies.  
+Many more \pz s may be obtained in the time it would take to observe a smaller number of \sz s, and \pz s may be measured for galaxies too dim for accurate \sz\ confirmation, permitting the compilation of large catalogs of galaxy redshifts spanning a broad range of redshifts and luminosities.
+\Pz s have thus enabled the era of precision cosmology, heralded by weak gravitational lensing tomography and baryon acoustic oscillation peak measurements.
+
+However, \pz s are susceptible to inaccuracy and imprecision via a number of effects, particularly their inherent noisiness due to the coarseness of photometric filters, catastrophic errors in which galaxies of one type at one redshift are mistaken for galaxies of another type at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
+Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
+This plot looks very much like the traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of traditional \pz\ point estimates and was indeed adapted from one such plot in \citet{jain_whole_2015}, because \pz\ point estimates effectively are a special nonlinear projection of the data.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/jain05.png}
+		\caption{
+			A generic probability space (darker in areas of higher probability density) of redshift ($x$-axis) and data ($y$-axis), where the data is projected into a single dimension, with vertical cuts and marginals (cyan) indicating the construction of likelihoods and horizontal cuts and marginals (magenta) indicating the construction of posteriors.
+			The data (black points) used to generate the contours were extracted from \citet{jain_whole_2015} using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019}, with the ideal redshift estimation provided for reference (red diagonal).
+			\aim{Choose better colors?}
+		}
+		\label{fig:pedagogical_scatter}
+	\end{center}
+\end{figure}
+
+There are several varieties of generally non-Gaussian deviation from a perfect correspondence between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
+There is scatter about the diagonal due to the coarseness of the photometric filters, with larger scatter perpendicular to the diagonal due to the specific wavelengths where highly identifiable spectral features pass between the filters, as well as higher scatter at high redshifts due to larger photometric errors on fainter galaxies.
+There are populations of outliers, far from the diagonal, comprised of galaxies for which the redshift estimate is catastrophically distinct from the true redshift, showing that outliers are not uniformly distributed nor restricted to long tails away from a Gaussian scatter.
+And, though hardly perceptible in the plot, there is a systematic bias, wherein the average of the points would not lie on the diagonal but would be offset by a small amount, suggested by the trend of points to lie below the diagonal at high redshift.
+Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
+For example, the Science Requirements Document \citep{mandelbaum_weak_2017} states \lsst's requirements for the main cosmological sample, reproduced in Table~\ref{tab:lsstsrd}.
+
+\begin{table}
+	\begin{center}
+		\caption{\Pz\ requirements for \lsst\ cosmology \citep{mandelbaum_weak_2017}.}
+		\begin{tabular}{ll}
+			Number of galaxies & $\approx 10^{7}$\\
+			Root-mean-square error & $< 0.02 (1 + z)$\\
+			$3 \sigma$ catastrophic outlier rate & $< 10\%$\\
+			Bias & $< 0.003 (1 + z)$\\
+		\end{tabular}
+	\label{tab:lsstsrd}
+	\end{center}
+\end{table}
 
@@
 Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
-I present and validate a hierarchical Bayesian technique for the use of \pzpdf\ in inference of arbitrary statistics relevant to cosmology, large-scale structure, and galaxy evolution.  
+We present and validate a hierarchical Bayesian technique for the use of \pzpdf\ in inference of arbitrary statistics relevant to cosmology, large-scale structure, and galaxy evolution.  
 For simplicity, we consider only one-point statistics, though future work will extend this methodology to higher-order statistics.
@@
 % Equation~\ref{intro:eq:eqn:stack}.
-The stacked estimator of the redshift density function is effectively an average of the \pzpdf\ catalog and, as I will show in this paper, not in general mathematically valid.
+The stacked estimator of the redshift density function is effectively an average of the \pzpdf\ catalog and, as we show in this paper, not in general mathematically valid.
 Stacking is nonetheless considered the preferred method for obtaining \Nz\ from a \pzpdf\ catalog \citep{sheldon_photometric_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016}.
@@
 
-I aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
+We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
 Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys that may be resistant to restructuring their analysis pipelines.
 
-In Section~\ref{sec:meth}, I present the \Chippr\ model and \chippr\ implementation for characterizing the full posterior probability landscape of \Nz\ using \pzpdf s. 
-In Section~\ref{sec:alldata}, I describe the experimental design for testing the fully probabilistic approach to mock and real datasets, the results of which are found in Section~\ref{sec:results}.
-In Section~\ref{sec:results}, I stress-test the \Chippr\ model against stacking and its other competitors in the context of cosmology.
+In Section~\ref{sec:meth}, we present the \Chippr\ model and \chippr\ implementation for characterizing the full posterior probability landscape of \Nz\ using \pzpdf s. 
+In Section~\ref{sec:alldata}, we describe the experimental design for testing the fully probabilistic approach to mock and real datasets, the results of which are found in Section~\ref{sec:results}.
+In Section~\ref{sec:results}, we stress-test the \Chippr\ model against stacking and its other competitors in the context of cosmology.
 
@@
 % \aim{also selection function}
-In the following sections, I present and compare methods for estimating \nz\ from \pzpdf s.  
+In the following sections, we present and compare methods for estimating \nz\ from \pzpdf s.  
 \Sect{sec:prob} contains the mathematical derivation of a probabilistic model for $n(z)$ dependent on photo-$z$ probability distribution functions, and \Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.
@@
 
+\aim{TODO: fix implicit posterior vs. implicit-weighted posterior}
+
 A forward model such as that of \Sect{sec:forward} corresponds to a probabilistic graphical model (PGM), represented by a directed acyclic graph (DAG) as in \Fig{fig:pgm}.
@@
 Because the parameters $\ndphi$ that we seek are causally separated from the data by the latent variable of redshift, we call them \textit{hyperparameters}.
-%\aim{I'm not explaining the distinction between hyperparameters and parameters adequately.}
+\aim{Am I adequately explaining the distinction between hyperparameters and parameters?}
 
@@
 
-% this is what we want to compute
-% what we're given is implicit posteriors
-% this turns out to be the correct answer
-% derivation in appendix
-% human-readable description of ratios replaces the implict prior with model at each sample step, converting posterior into likelihood
+\aim{TODO: Clarify what we want to compute ($\pr{\ndphi \gvn\ \data}$), what we're given ($\{\pr{z \gvn \data, \ndphi^{*}}\}$).}
+\aim{TODO: Give answer here, say this is correct, move derivation to appendix.}
+\aim{TODO: Give human-readable description of ratios, explain that \Chippr\ replaces implicit prior with sampled model, converting posterior into likelihood.}
+
 With these definitions, we obtain the desired expression for $\pr{\ndphi \gvn \{\data_{j}\}}$,
-\begin{align}
-\begin{split}
+\begin{equation}
 \label{eqn:fullpost}
-\pr{\ndphi \gvn \{\data_{j}\}} \propto & \pr{\ndphi} \left[\integral{\prod_{j=1}^{J} \right.\\
-	& \left. \left(\pr{z \gvn \data_{j}, \ndphi^{*}} \pr{z \gvn \ndphi}(\pr{z \gvn \ndphi^{*}})^{-1}\right)}{z}\right] ,
-\end{split}
-\end{align}
+\pr{\ndphi \gvn \{\data_{j}\}} \propto \pr{\ndphi} \integral{\prod_{j=1}^{J} \frac{\pr{z \gvn \data_{j}, \ndphi^{*}} \pr{z \gvn \ndphi}}{\pr{z \gvn \ndphi^{*}}}}{z},
+\end{equation}
 the heart of \Chippr.
@@
 Let us consider the probability of the parameters conditioned on the data and an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} as 
-\begin{equation}
+\begin{align}
 \label{eqn:trick}
-\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
-\end{equation}
+\begin{split}
+\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\\
+ & - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
+ \end{split}
+\end{align}
 
@@
 \label{eqn:expand}
-\ln[\pr{\data_{j} \gvn z}] & = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}]\\
-& - \ln[\pr{z \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
+\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\\
+& + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}] \\
+& - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
 \end{split}
@@
 \label{eqn:indterm}
-\ln[\pr{\vec{d}_{j} \gvn z}] & = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] \\
-& - \ln[\pr{z \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] - \ln[\pr{\data_{j} \gvn \ndphi^{*}}].
+\ln[\pr{\vec{d}_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] \\
+& + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}]\\
+& - \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] .
 \end{split}
@@
 \label{eqn:final}
-\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \right.\right.\\
-& \left.\left.\left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto & \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \right.\right.\\
+& \left.\left.\left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}]\right.\right.\right.\\
+& \left.\left.\left. - \ln[\pr{z \gvn \ndphi^{*}}] \right)\right]}{z}\right] ,
 \end{split}
 \end{align}
+\aim{TODO: fix the trailing parentheses/brackets.}
 
@@
 
-%\clearpage
 \subsection{Implementation}
@@
 
-I implemented the \Chippr\ model in code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
-In \Sect{sec:mcmc}, I describe the publicly available \chippr\ library.
-In \Sect{sec:acorr}, I outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
-
-%\clearpage
+We implement the \Chippr\ model in code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
+In \Sect{sec:mcmc}, we describe the publicly available \chippr\ library.
+In \Sect{sec:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
+
 \subsubsection{Code}
@@
 
-%\clearpage
 \subsubsection{Convergence Criteria}
@@
 
-%\clearpage
 \subsubsection{Performance metrics}
@@
 	\begin{center}
-		\includegraphics[width=0.74\textwidth]{figures/chippr/flowchart.png}
+		\includegraphics[width=0.9\textwidth]{figures/chippr/flowchart.png}
 		\caption{\aim{Check that all names and formulae are defined in the text and the same way as in diagram, say the whole thing in words in the text as well.
@@
 			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
-			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation I perform, and rounded rectangles indicate a quantity created by the forward model.
+			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation we perform, and rounded rectangles indicate a quantity created by the forward model.
 			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.}
@@
 \end{figure*}
-% \aim{clarify caption}
 
@@
 %\end{figure}
-% \aim{Add back the prior figure!}
+ \aim{TODO: Add back the figure of prior samples?}
 
@@
 	\caption{
-		Left: Examples of \pzpdf s with a uniform catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-		Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+		Top: Examples of \pzpdf s with a uniform catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
 		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
@@
 
-\begin{figure}
+\begin{figure*}
 	\begin{center}
@@
 	\end{center}
-\end{figure}
+\end{figure*}
 
@@
 As expected based on consistency with the forward model, \Chippr\ is completely resistant to bias, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
-%\aim{I'm not sure I effectively explained why \Chippr\ is expected to be unaffected by bias of this form.}
+\aim{I'm not sure I effectively explained why \Chippr\ is expected to be unaffected by bias of this form.}
 
@@
 It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
-To test the impact of these uncertainties, I simulated mock data with all three effects with which \lsst\ is concerned at the levels of \aim{fix reference}
+To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of \aim{fix reference}
 %Table~\ref{intro:tab:tab:lsstsrd} 
 and propagated the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
-Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, I perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how I think they ought to be constrained.
+Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
 % \aim{don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space}
 
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.74\textwidth]{figures/chippr/cosmolike_inputs.png}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.49\textwidth]{figures/chippr/cosmolike_inputs.png}
 		\caption{
@@
 	\end{center}
-\end{figure*}
-
-I considered a set of tomographically binned \nz\ and cosmological parameter covariance matrices used for \desc\ forecasting, for which the true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
-First, I binned them down to a piecewise constant parameterization with a manageable $35$ parameters for \chippr's sampling capabilities.
-Next, I drew $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
+\end{figure}
+
+We consider a set of tomographically binned \nz\ and cosmological parameter covariance matrices used for \desc\ forecasting, for which the true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+First, we bin them down to a piecewise constant parameterization with a manageable $35$ parameters for \chippr's sampling capabilities.
+Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
 The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
-I emulated \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of \aim{fix reference}
+We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of \aim{fix reference}
 %Table~\ref{intro:tab:tab:lsstsrd}
@@
 %			 Table~\ref{intro:tab:tab:lsstsrd}.
-			%		\aim{Make this one big plot instead of four little ones to eliminate repeated insets and legend, and make the axis labels bigger.
+		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend, and make the axis labels bigger.}
 		}
@@
 \end{figure*}
-% \aim{how to get the bias? Ask Elisabeth about what I'd need for bias}
-
-I then made a point estimate of \nz\ using \Chippr\ as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
+ \aim{TODO: Ask Elisabeth how to get the bias here}
+
+We then make a point estimate of \nz\ using \Chippr\ as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
 Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
@@
 			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
-			%			\aim{Make this one big plot instead of four little ones to eliminate repeated axis labels and legend, and make the legend text bigger.}
+			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend, and make the legend text bigger.}
 		}
@@
 
-I then used the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
+We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
 Though there are also slight differences in the angle of the error ellipses, most of the differences are due to broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
@@
 Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
-Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, I calculated the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
+Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
 
@@
 	\begin{center}
-%		\includegraphics[width=0.95\textwidth]{figures/chippr/final_plot.png}
+		\includegraphics[width=0.94\textwidth]{figures/chippr/final_plot.png}
 		\caption{
@@
 			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
-			\aim{You can't see the \Chippr\ contours because they're exteemely close to the truth}
+			\aim{Are the contours any easier to see now?}
 		}
@@
 
-%\aim{Hogg says ``I think it would be good to talk a little quantitatively about where people need to know N(z) and other one-point statistics, and how much they will get various things wrong if they don't know these correctly.
-%	And situate that discussion within the current context of cosmological parameter estimation and precision cosmology.''
-%	Did I successfully answer the question?}
+\aim{Hogg says ``I think it would be good to talk a little quantitatively about where people need to know N(z) and other one-point statistics, and how much they will get various things wrong if they don't know these correctly.
+	And situate that discussion within the current context of cosmological parameter estimation and precision cosmology.''
+	Did I successfully answer the question?}
 
@@
 
-%\aim{Move \Sect{sec:interim} here?}
+\aim{Move \Sect{sec:interim} here?}
 
@@
 
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
 		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
@@
 	\end{center}
-\end{figure*}
+\end{figure}
 
@@
 
-% new paragraph for what can go wrong, call to community for what to do about it
-% what aspects are knowable and not knowable, if testable how so? outside the scope of paper, data producers be warned
-% a likelihood is better -- give us that if you can!
-% focus on methods that acknowledge probabilistic structure of problem
-
-% implications for tomographic binning are severe
+\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
+	 if testable how so? outside the scope of paper, data producers be warned!
+	 a likelihood is better -- give us that if you can!
+	focus on methods that acknowledge probabilistic structure of problem}
+
+\aim{claim: implications for tomographic binning are severe}
 

### 2019-05-24 research/paper/draft.tex
fixing broken references and bringing in intro from thesis

@@
 Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
-This plot looks very much like the traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of traditional \pz\ point estimates and was indeed adapted from one such plot in \citet{jain_whole_2015}, because \pz\ point estimates effectively are a special nonlinear projection of the data.
+This plot looks very much like the traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of \pz\ point estimates and was indeed adapted from one such plot in \citet{jain_whole_2015}, because \pz\ point estimates effectively are a special nonlinear projection of the data.
 
@@
 
+Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are not insignificant contributors to the total uncertainties reported on cosmological parameters.
+As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
+
+Much effort has been dedicated to improving \pz s, though they are still most commonly obtained by a maximum likelihood estimator (MLE) based on libraries of galaxy SED templates, with conservative approaches to error estimation.
+This approach tends to lead to catastrophic outliers like the horizontally oriented population of \Fig{fig:pedagogical_scatter}, due to the presence of galaxies whose SEDs are not represented by the template library.
+Data-driven approaches tend to result in catastrophic outliers like the vertically oriented population of \Fig{fig:pedagogical_scatter}, due to training sets that are incomplete in redshift coverage.
+Recent advances have focused on identifying and removing catastrophic outliers when using \pz s for 
+inference \citep{gorecki_new_2014}.  
+The approaches of using a training set versus a template library are related to one another by \citet{budavari_unified_2009}.
+Sophisticated Bayesian techniques and cutting-edge machine learning methods have been employed to improve precision \citep{carliles_random_2010} and accuracy \citep{sadeh_annz2:_2016}.
+
+An alternative to point estimation of \pz s is redshift probability distribution function (PDF) estimation that reports the probability $\pr{z}$ over all possible redshifts for every galaxy rather than an MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.
+This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating notions of precision, accuracy, and systematic errors.
+However, denoting \pzpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; \pzpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
+In terms of \Fig{fig:pedagogical_scatter}, \pzpdf s are horizontal cuts, probabilities of redshift conditioned on a specific value of data, i.e.
+posteriors, which constrain redshifts, whereas vertical cuts through this space are probabilities of data conditioned on a specific redshift, i.e.
+likelihoods, from which data is actually drawn.
+
+\Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.
+\Pzpdf s are not without their own weaknesses, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014} and the divergent results of each method used to derive them \citep{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.
+The most important of these issues, however, is that use of them in the literature is inconsistent at best and incorrect at worst.  
+The most common application of \pzpdf s is their use in estimating \Nz, the distribution of redshifts of a sample of galaxies, a quantity essential to the calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citet{mandelbaum_precision_2008, sheldon_photometric_2012, bonnett_redshift_2016}.
+%\aim{Provide equations/citations for the use of \Nz\ in cosmology.}
+
+If the true redshifts $\{z_{i}^{\dagger}\}$ of galaxies $i$ were known, then the redshift PDFs would be delta functions $\{\delta(z, z_{i}^{\dagger})\}$ centered at the true redshift, and the redshift distribution could be effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{i}^{\dagger})\}$.
+When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{i})$ in place of $\delta(z, z_{i}^{\dagger})$.
+Though it is most common for $\hat{z}_{i}$ to be the mode (maximum) of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.
+
+To circumvent the reduction of uncertainty information due to point estimation, however, it has grown to be more common to calculate the mathematically inconsistent but conceptually simple \textit{stacked estimator} of the redshift density function \citep{lima_estimating_2008},
+\begin{align}
+\label{eqn:stack}
+\hat{n}(z) &= \frac{1}{\ntot} \sum_{i = 0}^{\ntot} \pr{z}_{i}
+\end{align}
+for a sample of $\ntot$ galaxies $i$, or, equivalently, the redshift distribution function $\hat{N}(z) = \ntot \hat{n}(z)$.
+Much of this thesis centers around the problems with this logically invalid yet pervasive quantity.
+
 \aim{TODO: Pull in content from global intro of thesis.}
@@
 
-The prevailing way to estimate \nz\ for a sample of $\ntot$ galaxies $i$ is to ``stack'' \pzpdf s according to \aim{fix reference}.
-% Equation~\ref{intro:eq:eqn:stack}.
+The prevailing way to estimate \nz\ for a sample of galaxies $j$ is to ``stack'' \pzpdf s according to Equation~\ref{eqn:stack}.
 The stacked estimator of the redshift density function is effectively an average of the \pzpdf\ catalog and, as we show in this paper, not in general mathematically valid.
 Stacking is nonetheless considered the preferred method for obtaining \Nz\ from a \pzpdf\ catalog \citep{sheldon_photometric_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016}.
-However, it must be noted here that \aim{fix reference}
-%Equation~\ref{intro:eq:eqn:stack} 
+However, it must be noted here that Equation~\ref{eqn:stack} 
 is not in general mathematically valid.  
@@
 \end{equation}
-of photometry conditioned on redshift, illustrated in \aim{fix reference}.
-%Figure~\ref{intro:fig:fig:pedagogical_scatter}.
+of photometry conditioned on redshift, illustrated in Figure~\ref{fig:pedagogical_scatter}.
 
@@
 The mock data emulates the three sources of error of highest concern to the \pz\ community that are explored in detail later in this section: intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and systematic bias (\Sect{sec:bias}).
-Tests including all three effects at the tolerance levels of \lsst\ (see \aim{fix reference})
-%Table~\ref{intro:tab:tab:lsstsrd}) 
+Tests including all three effects at the tolerance levels of \lsst\ (see Table~\ref{tab:lsstsrd}) 
 are presented in \Sect{sec:results}.
-\Fig{fig:mega_scatter} illustrates these three effects individually at twice the tolerance of \lsst\ for demonstrative purposes, hearkening back to \aim{fix reference}.
-%Figure~\ref{intro:fig:fig:pedagogical_scatter}.
+\Fig{fig:mega_scatter} illustrates these three effects individually at twice the tolerance of \lsst\ for demonstrative purposes, hearkening back to Figure~\ref{fig:pedagogical_scatter}.
 We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to be a priority for the \pz\ community.
@@
 \Fig{fig:uniform-outliers} shows examples of \pzpdf s from a uniformly distributed outlier population at the level of the \lsst\ requirements (left) as well as the results of \Chippr\ and other \nz\ estimation methods (right).
-The intrinsic scatter of the tests in this section does not increase with redshift as indicated in \aim{fix reference},
-%Table~\ref{intro:tab:tab:lsstsrd}
+The intrinsic scatter of the tests in this section does not increase with redshift as indicated in Table~\ref{tab:lsstsrd}
 in order to isolate the effect of outliers, and is instead held at a constant $\sigma_{z} = 0.02$.
@@
 However, in the context of \pzpdf s, the notion of redshift bias is a form of model misspecification.
-Consider that if bias were included in the framework of \aim{fix reference},
-%Figure~\ref{intro:fig:fig:pedagogical_scatter}
+Consider that if bias were included in the framework of Figure~\ref{fig:pedagogical_scatter};
 a simple linear transformation of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + z_{\mathrm{phot}})$ would eliminate the bias.
@@
 It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
-To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of \aim{fix reference}
-%Table~\ref{intro:tab:tab:lsstsrd} 
+To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} 
 and propagated the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
@@
 The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
-We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of \aim{fix reference}
-%Table~\ref{intro:tab:tab:lsstsrd}
+We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
 at their given levels.
@@
 		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
-		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in \aim{fix reference}.
-%			 Table~\ref{intro:tab:tab:lsstsrd}.
+		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
 		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend, and make the axis labels bigger.}

### 2019-05-28 research/paper/draft.tex
small updates

@@
 \usepackage{soul}
-%\usepackage[title]{appendix}
+\usepackage[title]{appendix}
 
@@
 \newcommand{\python}{\textit{Python}}
-\newcommand{\nz}{$n(z)$}
 \newcommand{\data}{\ensuremath{\vec{d}}}
@@
 \newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
-\newcommand{\sz}{spec-$z$~}
-\newcommand{\Sz}{Spec-$z$~}
-\newcommand{\pz}{photo-$z$~}
-\newcommand{\Pz}{Photo-$z$~}
+\newcommand{\sz}{spec-$z$}
+\newcommand{\Sz}{Spec-$z$}
+\newcommand{\pz}{photo-$z$}
+\newcommand{\Pz}{Photo-$z$}
 \newcommand{\pzpdf}{\pz PDF}
@@
 \newcommand{\pzip}{\pz implicit posterior}
-\newcommand{\Nz}{$N(z)$~}
-\newcommand{\stack}{$\hat{N}(z)$~}
+\newcommand{\nz}{$n(z)$}
+\newcommand{\Nz}{$N(z)$}
+\newcommand{\stack}{$\hat{n}(z)$}
 \newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
@@
 
-\aim{Some of the figures refer to left/right but are shown as top/bottom \textemdash I'm still working through a LaTeX bug.}
-
 \section{Introduction}
@@
 the heart of \Chippr.
-The derivation is provided below in \Sect{app:math}.
-
-\subsubsection{Derivation}
+The derivation is provided in Appendix~{app:math}.
+
+%\subsubsection{Derivation}
+%\label{app:math}
+%
+%\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
+%
+%%We begin by parametrizing $N(z)$ in terms of $\vec{\theta}$, comprising some set of hyperparameters that define the form $N(z)$ may take in whatever basis we choose.  
+%%We define a function $f_{\vec{\theta}}(z)=N(z)$ that transforms these hyperparameters into the redshift distribution function $N(z)$.  
+%%Because 
+%%\begin{equation}
+%%\eqlabel{eq:definition}
+%%N(z) \propto p(z \gvn \vec{\theta}),
+%%\end{equation}
+%%we may discontinue discussion of $N(z)$ in favor of the likelihood $p(z|\vec{\theta})$.
+%
+%In this paper, we work exclusively with log-probabilities.  
+%What we wish to estimate is then the full log-posterior probability distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ given the catalog of photometry $\{\data_{j}\}$.
+%
+%By Bayes' Rule, the full log-posterior
+%\begin{equation}
+%\label{eqn:basicbayes}
+%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] = \ln[\pr{\{\data_{j}\} \gvn \ndphi}] + \ln[\pr{\ndphi}] - \ln[\pr{\{\data_{j}\}}]
+%\end{equation}
+%may be expressed in terms of the full log-likelihood probability distribution (hereafter the full log-likelihood) $\ln[\pr{\{\data_{j}\} \gvn \ndphi}]$ by way of a hyperprior log-probability distribution (hereafter the hyperprior) $\ln[\pr{\ndphi}]$ over the hyperparameters and the log-evidence probability of the data $\ln[\pr{\{\data_{j}\}}]$.
+%However, the evidence is rarely known, so we probe the full log-posterior modulo an unknown constant of proportionality.
+%
+%The full log-likelihood may be expanded in terms of a marginalization over the redshifts as parameters, as in 
+%\begin{equation}
+%\label{eqn:marginalize}
+%\ln[\pr{\{\data_{j}\} \gvn \ndphi}] = \ln\left[\integral{\pr{\{\data_{j}\} \gvn \{z_{j}\}} \pr{\{z_{j}\} \gvn \ndphi}}{\{z_{j}\}}\right].
+%\end{equation}
+%
+%We shall make two assumptions of independence in order to make the problem tractable; their limitations are be discussed below.  
+%First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn z_{j}}]$, as in 
+%\begin{equation}
+%\label{eqn:indiedat}
+%\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn z_{j}}],
+%\end{equation}
+%a result of the definition of probabilistic independence encoded by the box in \Fig{fig:pgm}.
+%Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent draws from the true $\pr{z \gvn \ndphi}$.  
+%Additionally, $J$ itself is a Poisson random variable.  
+%The combination of these assumptions is given by 
+%\begin{equation}
+%\label{eqn:indie}
+%\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ \ln[\pr{z_{j} \gvn \ndphi}].
+%\end{equation}
+%%It is important to note that the integral $\integral{n(z)}{z} N(z)\ dz$ is not constrained to equal the variable defining the Poisson distribution but instead $J$ by \Eq{eq:definition}, which can be thought of as another parameter.
+%The derivation differs when $J$ is not known, say, when we want to learn about a distribution in nature rather than a distribution specific to data in hand, but for a photometric galaxy catalog where the desired quantity is $n(z)$ for the galaxies entering a larger cosmology calculation, it is a fixed quantity.
+%A detailed discussion of this matter may be found in \citet{foreman-mackey_exoplanet_2014}.  
+%Applying Bayes' Rule, we may combine terms to obtain 
+%\begin{align}
+%\begin{split}
+%\label{eqn:posterior}
+%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] - \integral{f(z; \ndphi)}{z}\\
+%& + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn \ndphi}}{z}\right].
+%\end{split}
+%\end{align}
+%
+%%\Eq{eq:posterior} contains two quantities that merit further discussion, the prior distribution $p(\vec{\theta})$ discussed further in \Sect{sec:exp} and the photo-$z$ log-likelihoods $\ln[p(\vec{d}_{j}|z_{j})]$ that have not been mentioned since \Eq{eq:marginalize}.
+%%Though photo-$z$ log-likelihoods would be desirable for use in these equations, they are not generally the product of either empirical and data-driven methods for obtaining photo-$z$ probability distributions.
+%%Though probabilistic photo-$z$s are typically reported as generic probability distributions $p(z_{j})$, the methods that produce them may be understood to always yield posteriors, probability distributions conditioned on the data we believe to be true.
+%%If they were not based in this assumption, they would require a sum over an infinite space of possible datasets.
+%
+%Since we only have access to implicit \pz\ posteriors, we must be able to write the full log-posterior in terms of implicit \pz\ log-posteriors rather than the log-likelihoods of \Eq{eqn:posterior}.
+%To do so, we will need an explicit statement of this implicit prior $\ndphi^{*}$ for whatever method is chosen to produce the implicit \pz\ posteriors.  
+%
+%To perform the necessary transformation from likelihoods to posteriors, we follow the reasoning of \citet{foreman-mackey_exoplanet_2014}.  
+%Let us consider the probability of the parameters conditioned on the data and an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} as 
+%\begin{align}
+%\label{eqn:trick}
+%\begin{split}
+%\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\\
+% & - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
+% \end{split}
+%\end{align}
+%
+%Once the implicit prior $\ndphi^{*}$ is explicitly introduced, we may expand the last term in \Eq{eqn:trick} according to Bayes' Rule to get 
+%\begin{align}
+%\begin{split}
+%\label{eqn:expand}
+%\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\\
+%& + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}] \\
+%& - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
+%\end{split}
+%\end{align}
+%Because there is no direct dependence of the data upon the hyperparameters, we may again expand the term $\ln[\pr{\data_{j} \gvn z, \ndphi^{*}}]$ to obtain 
+%\begin{align}
+%\begin{split}
+%\label{eqn:indterm}
+%\ln[\pr{\vec{d}_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] \\
+%& + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}]\\
+%& - \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] .
+%\end{split}
+%\end{align}
+%Canceling the undesirable terms for the inaccessible likelihood $\ln[\pr{\data_{j} \gvn z}]$ and trivial $\ln[\pr{\data_{j} \gvn \ndphi^{*}}]$ yields
+%\begin{equation}
+%\label{eqn:cancel}
+%\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]  - \ln[\pr{z \gvn \ndphi^{*}}].
+%\end{equation}
+%We put this all together to get the full log-posterior probability distribution of 
+%\begin{align}
+%\begin{split}
+%\label{eqn:final}
+%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto & \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \right.\right.\\
+%& \left.\left.\left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}]\right.\right.\right.\\
+%& \left.\left.\left. - \ln[\pr{z \gvn \ndphi^{*}}] \right)\right]}{z}\right] ,
+%\end{split}
+%\end{align}
+%\aim{TODO: fix the trailing parentheses/brackets.}
+%
+%The argument of the integral in the log-posterior of \Eq{eqn:final} depends solely on knowable quantities (and those we must explicitly assume) and can be calculated for a given sample of \pz\ log-posteriors $\{\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with which they were obtained, noting the relation of
+%\begin{equation}
+%\label{eqn:params}
+%\pr{z \gvn \ndphi} = \frac{f(z; \ndphi)}{\integral{f(z; \ndphi)}{z}}.
+%\end{equation}
+%Since we cannot know constant of proportionality, we sample the desired full log-posterior $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$ using Monte Carlo-Markov chain (MCMC) methods.  
+%The method outlined here is valid regardless of how the implicit \pz\ log-posteriors are obtained so the many approaches to producing \pzpdf s will not be rehashed; though the matter is outside the scope of this paper, reviews of various methods have been presented in the literature \citep{sheldon_photometric_2012, ball_robust_2008, carrasco_kind_tpz:_2013, carrasco_kind_exhausting_2014}, and will be briefly reviewed in \aim{cite PZDC1 paper here}.
+%%
+%%\begin{align}
+%%\begin{split}
+%%\eqlabel{eqn:fullpost}
+%%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
+%%\end{split}
+%%\end{align}
+
+\subsubsection{Model Limitations}
+\label{sec:limitations}
+
+Finally, we explicitly review the assumptions made by this approach, which are as follows:
+
+\begin{enumerate}
+	\item Photometric measurements of galaxies are statistically independent Poisson draws from the set of all galaxies such that \Eq{eqn:indiedat} and \Eq{eqn:indie} hold.
+	\item We take the reported \pzip s to be accurate, free of model misspecification; draws thereof must not be inconsistent with the distribution of photometry and redshifts.
+	Furthermore, we must be given the implicit prior $\ndphi^{*}$ used to produce the \pzip s.
+	\item We must assume a hyperprior distribution $\pr{\ndphi}$ constraining the underlying probability distribution of the hyperparameters, which is informed by our prior beliefs about the true redshift distribution function.
+\end{enumerate}
+
+These assumptions have known limitations.  
+First, the photometric data are not a set of independent measurements; the data are correlated not only by the conditions of the experiment under which they were observed (instrument and observing conditions) but also by redshift covariances resulting from physical processes governing underlying galaxy spectra and their relation to the redshift distribution function.
+Second, the reported \pzip s may not be trustworthy; there is not yet agreement on the best technique to obtain \pzpdf s, and the implicit prior may not be appropriate or even known to us as consumers of \pzip s.
+Third, the hyperprior may be quite arbitrary and poorly motivated if the underlying physics is complex, and it can only be appropriate if our prior beliefs about \nz\ are accurate.
+
+\subsection{Implementation}
+\label{sec:exp}
+
+We implement the \Chippr\ model in code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
+In \Sect{sec:mcmc}, we describe the publicly available \chippr\ library.
+In \Sect{sec:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
+
+\subsubsection{Code}
+\label{sec:mcmc}
+
+\chippr\ is a \python\ 2 library\footnote{\url{https://github.com/aimalz/chippr}} that includes an implementation of the \Chippr\ model as well as an extensive suite of tools for comparing \Chippr\ to other approaches.
+
+Though there are plans for future expansion to more flexible parameterizations, the current version of \chippr\ uses a log-space piecewise constant parameterization
+\begin{equation}
+\label{eqn:logstepfunc}
+f(z; \ndphi) = \exp[\phi^{k}]\ \mathrm{if}\ z^{k} < z < z^{k+1}
+\end{equation}
+for \nz\ and every \pzpdf, satisfying
+\begin{equation}
+\label{eqn:logstepfuncnorm}
+\sum_{k=1}^{K} \exp[\phi^{k}] \delta z^{k} = 1
+\end{equation}
+with $K$ bins of width $\delta z^{1}, \dots, \delta z^{K}$ defined by endpoints $z^{0}, \dots, z^{K}$.
+%\aim{Maybe include an equation for a general piecewise constant function here (which is currently in \Chap{qp})?}
+Thus each $\pr{z \gvn \data_{j}} = f(z; \ndphi_{j})$ has parameters $\ndphi_{j}$ that are defined in the same basis as those of \nz.
+To infer the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$, one must provide a plaintext file with $K+1$ redshift bin endpoints $\{z_{k}\}$, the parameters $\ndphi^{*}$ of the implicit log-prior, and the parameters $\{\ndphi_{j}\}$ of the log-posteriors $\ln[\pr{z \gvn \data_{j}, \ndphi^{*})}$.
+
+The \emcee \citep{foreman-mackey_emcee_2013} implementation of ensemble sampling is used to sample the full log-posterior of \Eq{eqn:final}. 
+\chippr\ accepts a configuration file of user-specified parameters, among them the number $W$ of walkers.
+At each iteration $i$ and for each walker, a proposal distribution $\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for acceptance to or rejection from the full log-posterior distribution.
+
+Two threshold conditions are defined, one designating all previous samples to be ignored as as products of a burn-in phase and another indicating when a sufficient number of post-burn samples have been accepted.
+In this case, the first threshold (described in \Sect{sec:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
+
+Though previous versions used \texttt{HDF5} for the primary I/O format due to its efficiency for large quantities of data, it was abandoned in favor of \texttt{pickle} in the working release due to the instability of the \python\ implementation of the format on high-performance computing systems.
+The resulting output is a set of $I$ ordered \texttt{hickle} files enumerated by $\rho$ containing the state information after each sub-run.  
+The state information includes $\frac{I_{0}}{s}$ actual samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$ as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$ over the entire sub-run.
+
+%\subsubsection{Convergence Criteria}
+%\label{sec:acorr}
+%
+%\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
+%
+%In addition to qualitative visual inspection of the chains, two quantities that probe the convergence of the sampler are used in this study, the autocorrelation time and the Gelman-Rubin convergence criterion.
+%%\Fig{fig:chains} shows the %evolution of the values of one parameter of one walker over the course of all %iterations of the sampler.
+%
+%%\begin{figure}
+%%%\includegraphics[width=0.5\textwidth]{figs/null/chain0.pdf}
+%%\caption{This figure shows the evolution of one walker's parameter values for 
+%%one element of the parameter vector $\vec{\theta}$ as a function of iteration 
+%%number, demonstrating the completion of the burn-in phase.}
+%%\label{fig:chains}
+%%\end{figure}
+%
+%The autocorrelation time is effectively a measure of the efficiency of the method and can be described as the expected number of iterations necessary to accept a new sample independent of the current accepted sample.
+%A sampler that converges faster will have a smaller autocorrelation time, and smaller autocorrelation times are preferable because it means fewer iterations are wasted on non-independent samples when independent samples are desired.
+%See \citet{foreman-mackey_emcee_2013} for a more complete exploration of the autocorrelation time.  
+%In all tests discussed here, autocorrelation times across walkers and parameters were approximately 20, meaning two samples 20 or more iterations apart were independent, a satisfactory level of efficiency.
+%Low autocorrelation times are a necessary but not always sufficient convergence condition, as the autocorrelation times calculated for tests in this paper were constant across all sub-runs, even those that were obviously burning in.
+%
+%The Gelman-Rubin statistic
+%\begin{equation}
+%\label{eqn:gr}
+%R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} b_{k}}{w_{k}}},
+%\end{equation}
+%a weighted sum of the mean $w_{k}$ of the variances within individual walkers' chains and the variance $b_{k}$ between chains of different walkers $m$, is calculated over each sub-run $i$ to determine the duration of the burn-in period.
+%Convergence is achieved when the statistic approaches unity.
+
+\subsection{Comparison with Alternative Approaches}
+\label{sec:sheldon}
+
+In this study, we compare the results of \Eq{eqn:fullpost} to those of the two most common approaches to estimating \nz\ from a catalog of \pzpdf s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior probability
+\begin{equation}
+\label{eqn:mmap}
+f^{MMAP}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \delta(z, \mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}])
+\end{equation}
+(i.e. the distribution of modes of the \pzpdf s) and the stacked estimator of \Eq{eqn:stacked}, which can be rewritten as 
+\begin{equation}
+\label{eqn:stacked}
+f^{stack}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \pr{z \gvn \data_{j}, \ndphi^{*}}
+\end{equation}
+in terms of the implicit \pz\ posteriors we have.
+These two approaches have been compared to one another by \citet{hildebrandt_cfhtlens:_2012}, \citet{benjamin_cfhtlens_2013}, and \citet{asorey_galaxy_2016} in the past but not to \Chippr.
+
+Point estimation converts the implicit photo-$z$ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$ into delta functions with all probability at a single estimated redshift.  
+Some variants of point estimation choose this single redshift to be that of maximum a posteriori probability $\mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}]$ or the expected value of redshift $\langle z \rangle = \integral{z \pr{z \gvn \data_{j}, \ndphi^{*}}}{z}$.
+\aim{cite tanaka 2018 here} directs attention to deriving an optimal point estimate reduction of a \pzpdf, but since the purpose of this paper is to compare against the most established alternative estimators of \nz, its use will be postponed until a future study.
+Stacking these modified \pzpdf s leads to the marginalized maximum a posteriori (MMAP) estimator and the marginalized expected value (MExp) estimator, though only the former is included in this study since the latter has fallen out of favor in recent years\footnote{And for good reason!
+Consider a bimodal \pzpdf; its expected value may very well fall in a region of very low probability, yielding a less probable point estimate than the point at which either peak achieves its maximum.}.
+
+It is worth discussing the relationship between point estimation and stacking.  
+When the point estimator of redshift is equal to the true redshift, stacking delta function \pzpdf s will indeed lead to an accurate recovery of the true redshift distribution function.  
+However, stacking is in general applied indiscriminately to broader \pzpdf s and imperfect point estimators of redshift.  
+It is for these reasons that alternatives are considered here.
+
+A final estimator of the hyperparameters is the maximum marginalized likelihood estimator, the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.  
+To compare with sampling, the MMLE also depends on the choice of the hyperprior distribution, and it does not produce a full posterior probability distribution over the parameters of interest, only point estimators.
+It must be noted that computation of the maximum marginalized likelihood estimator may be unstable depending on the strengths and weaknesses of the optimizer.  
+In general, derivatives will not be available for the full posterior distribution, restricting optimization methods used.
+%\begin{equation}
+%\label{eq:mmle}
+%\ln[p(\{\vec{d}_{j}\}|\vec{\theta})] \propto -\int\ f_{\vec{\theta}}(z)\ 
+%dz+\sum_{j=1}^{J}\ln\left[\int\ 
+%\exp\left[\ln[p(z_{j}|\vec{d}_{j},\vec{\theta}^{*})]+\ln[f_{\vec{\theta}}(z)]-\l
+%n[f_{\vec{\theta}^{*}}(z)]\right]\ dz\right],
+%\end{equation}
+%accessible with any optimization code.
+
+In \Sect{sec:diag} we outline the measures used to evaluate the performance of the method.
+
+\subsubsection{Performance metrics}
+\label{sec:diag}
+
+The results of the computation described in \Sect{sec:mcmc} are evaluated for accuracy on the basis of some quantitative measures.  
+Beyond visual inspection of samples, we calculate summary statistics to quantitatively compare different estimators' precision and accuracy.  
+Since MCMC samples of hyperparameters are Gaussian distributions, we can quantify the breadth of the distribution for each hyperparameter using the standard deviation regardless of whether the true values are known.
+
+In simulated cases where the true parameter values are known, we calculate the Kullback-Leibler divergence (KLD), given by 
+\begin{equation}
+\label{eqn:kl}
+KL_{',\ddagger} = \integral{\pr{z \gvn \ndphi'} \ln \left[ \frac{\pr{z \gvn \ndphi'}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
+\end{equation}
+which measures a distance from parameter values $\ndphi'$ to true parameter values $\ndphi^{\dagger}$, which is invariant under changes of variables.  
+We note that $KL_{',\dagger} \neq KL_{\dagger,'}$ and is only interpretable when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than $\ndphi'$.
+The KLD is explored in detail in the Appendix to \citet{malz_approximating_2018}.
+In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that produced by one of the methods in question.
+
+\section{Validation}
+\label{sec:alldata}
+
+We compare the results of \Chippr\ to those of stacking and the histogram of \pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed in \Sect{sec:forward}.
+\Fig{fig:flowchart} illustrates the implementation of this forward model used in this paper.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{figures/chippr/flowchart.png}
+		\caption{\aim{Check that all names and formulae are defined in the text and the same way as in diagram, say the whole thing in words in the text as well.
+			Explain that this is the forward model of the validation, not of CHIPPR! contrast this with Fig. 1 (joke about simple model and complex validation?)}
+			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
+			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation we perform, and rounded rectangles indicate a quantity created by the forward model.
+			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.}
+		\label{fig:flowchart}
+	\end{center}
+\end{figure*}
+
+The true redshift distribution used in these tests is a particular instance of the gamma function
+\begin{equation}
+\label{eqn:gamma}
+n^{\dagger}(z) = \frac{1}{2 c_{z}} \left(\frac{z}{c_{z}}\right)^{2}\ \exp\left[-\frac{z}{c_{z}}\right]
+\end{equation}
+with $c_{z} = 0.3$, because it has been used in forecasting studies for \des\ and \lsst.
+%\aim{I learned this from talking to people and don't know of a published source that talks about the nitty gritty of the internal validation tests performed before there was data.}
+
+The mock data emulates the three sources of error of highest concern to the \pz\ community that are explored in detail later in this section: intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and systematic bias (\Sect{sec:bias}).
+Tests including all three effects at the tolerance levels of \lsst\ (see Table~\ref{tab:lsstsrd}) 
+are presented in \Sect{sec:results}.
+\Fig{fig:mega_scatter} illustrates these three effects individually at twice the tolerance of \lsst\ for demonstrative purposes, hearkening back to Figure~\ref{fig:pedagogical_scatter}.
+We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to be a priority for the \pz\ community.
+
+\aim{TODO: add in the plot from thesis intro}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
+		\includegraphics[width=0.32\textwidth]{figures/chippr/single_uout-mega_scatter.png}
+		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
+		\caption{\aim{TODO: Add watermark of "mock data" in UL corner, same for "results of inference" on other kind of plot.}
+			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
+			The main panel of each shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal (thin red line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
+			The insets show marginal histograms (gray) in each dimension, that can be compared with the true \nz\ used to make the figure (blue curve) to see the effect of the isolated systematic, as well as the implicit prior (red line).
+		}
+		\label{fig:mega_scatter}
+	\end{center}
+\end{figure*}
+
+The hyperprior distribution chosen for these tests is a multivariate normal distribution with mean $\vec{\mu}$ equal to the implicit prior $\ndphi^{*}$ and covariance
+\begin{equation}
+\label{eqn:priorcov}
+\Sigma_{k,k'} = q\ \exp[-\frac{e}{2}\ (\bar{z}_{k}-\bar{z}_{k'})^{2}]\ +\ t\delta(k,k')
+\end{equation}
+inspired by one used in Gaussian processes, where $k$ and $k'$ are indices ranging from $1$ to $K$ and $q=1.0$, $e=100.0$, and $t=q\cdot10^{-5}$ are constants chosen to permit draws from this prior distribution to produce shapes similar to that of a true $\tilde{\ndphi}$.
+We adapt the full log-posterior of \Eq{eqn:final} to the chosen binning of redshift space.
+
+%An example of such samples from the prior are shown in \Fig{fig:prior}.
+%
+%\begin{figure}
+%%	\includegraphics[width=0.5\textwidth]{figures/chippr/null_priorsamps.pdf}
+%	\caption{\aim{I need to remake this one because it uses the wrong notation and I stopped making it automatically a while ago.}
+%		Samples (colored lines) of $\pr{z \gvn \ndphi}$ where each $\ndphi$ is drawn from the hyperprior distribution $\pr{\ndphi}$ given in \Eq{eqn:priorcov}.}
+%	\figlabel{fig:prior}
+%\end{figure}
+ \aim{TODO: Add back the figure of prior samples?}
+
+The sampler is initialized with $W=100$ walkers each with a value chosen from a Gaussian distribution of identity covariance around a sample from the hyperprior distribution.
+
+\subsection{Intrinsic scatter}
+\label{sec:scatter}
+
+%Several factors contribute to photometric redshifts' intrinsic scatter.  
+%Distant galaxies are dimmer compared to galaxies of identical luminosity that are closer, driving up photometric errors in flux-limited surveys.  
+%The nature of the galaxy sample at higher redshifts also changes, meaning the generation of the photometric redshift posterior based on an a locally-calibrated SED template library or spectroscopically-confirmed training set is more likely to be inappropriate, leading to broader features.
+%In general, the galaxies that could not have been observed spectroscopically will have different and noisier photo-$z$ likelihoods than those that could fall into a spectroscopic training set (or spectroscopically derived template library).
+%This effect may be stronger for high-redshift galaxies.
+
+\Fig{fig:pzs-scatter} shows some examples of \pzpdf s generated with only the systematic of intrinsic scatter, at the level of the \lsst\ requirements on the left and twice that on the right.
+One can see that the histogram of redshift estimates is broader than that of true redshifts, and that the effect is substantially more pronounced by just doubling the intrinsic scatter from the level of the \lsst\ requirements.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter1.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter2.png}
+	\caption{
+		Examples of mock \pzpdf s (colored lines) generated with intrinsic scatter at the \lsst\ requirements (left) and twice the \lsst\ requirements (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
+		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
+	}
+	\label{fig:pzs-scatter}
+	\end{center}
+\end{figure*}
+
+\Fig{fig:results-scatter} shows the \nz\ recovered by \Chippr\ and the alternative approaches.
+As expected, the estimates of \nz\ based on the modes of the \pzpdf s and stacking are broader than the marginalized maximum likelihood estimator from \chippr, with more broadening as the intrinsic scatter increases.
+\Chippr's \mmle\ is robust to intrinsic scatter and is unaffected by increased intrinsic scatter, though the \Chippr\ posterior distribution on the redshift distribution is itself broader for the higher intrinsic scatter case than for the \lsst\ requirements.
+The broadening of the alternative estimators corresponds to a loss of 3-4 times as many nats of information about \nz\ for the \lsst\ requirements relative to the \mmle\ of \Chippr.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter1.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter2.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with intrinsic scatter of the \lsst\ requirements (left) and twice that (right), with the true redshift density (black curve) and implicit prior (gray curve).
+		\Chippr\ is robust to intrinsic scatter, but the alternatives suffer from overly broad \nz\ estimates that worsen with increasing intrinsic scatter.
+	}
+	\label{fig:results-scatter}
+	\end{center}
+\end{figure*}
+
+\subsection{Catastrophic outliers}
+\label{sec:outliers}
+
+As was covered in \Sect{sec:intro}, catastrophic outliers tend to be distributed non-uniformly across the space of observed and true redshift.
+However, the \lsst\ requirements do not specify details for a distribution of outliers to which they were tuned, and it is still instructive to examine the impact of uniform outliers on the inference of \nz, so we begin by addressing uniformly distributed outliers before considering more realistic outlier distributions.
+
+A uniformly distributed population of outliers was simulated by giving every sample in true redshift a $10\%$ chance of having an observed redshift drawn from a uniform distribution rather than the Gaussian about the true redshift.
+Though this results in slightly less than the $10\%$ catastrophic outlier rate, it can be done independently of the definition of the standard deviation so was implemented for demonstrative purposes.
+\Fig{fig:uniform-outliers} shows examples of \pzpdf s from a uniformly distributed outlier population at the level of the \lsst\ requirements (left) as well as the results of \Chippr\ and other \nz\ estimation methods (right).
+The intrinsic scatter of the tests in this section does not increase with redshift as indicated in Table~\ref{tab:lsstsrd}
+in order to isolate the effect of outliers, and is instead held at a constant $\sigma_{z} = 0.02$.
+
+\begin{figure}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_log_estimators.png}
+	\caption{
+		Top: Examples of \pzpdf s with a uniform catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
+	}
+	\label{fig:uniform-outliers}
+	\end{center}
+\end{figure}
+
+\Fig{fig:uniform-outliers} shows that at the level of the \lsst\ requirements, the alternative estimators are overly broad, whereas \Chippr's \mmle\ yields an unbiased estimate of \nz.
+Further, the result of stacking is even broader than that of the histogram of modes, corresponding to ten times the information loss of \Chippr's \mmle, making it worse than the most naive reduction of \pzpdf s to point estimates.
+
+When one thinks of the \pzpdf s of catastrophic outliers, however, what comes to mind is multimodal \pzpdf s, wherein reducing \pzpdf s to point estimates to make a standard scatterplot of the true and observed redshifts leads to substantial probability density off the diagonal.
+These coordinated catastrophic outliers may be emulated in the joint probability space of true and estimated redshifts by using a mixture of the unbiased diagonal defined by the intrinsic scatter and an additional Gaussian in one dimension, with constant observed redshift for a template-fitting code and constant true redshift for a machine learning code.
+
+In the case of a catastrophic outlier population like that anticipated of template-fitting codes, $10\%$ of all galaxies have their observed redshift at a particular value unrelated to their true redshift, illustrated in the left panel of \Fig{fig:nonuniform-outliers-data}.
+This case is subject to the same caveat as the uniformly distributed outliers when it comes to the \lsst\ requirement.
+It is less straightforward to emulate catastrophic outliers like those anticipated of a machine learning code, those that are truly multimodal.
+The testing conditions here, illustrated in the right panel of \Fig{fig:nonuniform-outliers-data}, gives $10\%$ of galaxies at the redshift affected by outliers an observed redshift that is uniformly distributed relative to the true redshift, meaning that far fewer than $10\%$ of all galaxies in the sample are catastrophic outliers.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_mega_scatter.png}
+	\caption{
+		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+	}
+	\label{fig:nonuniform-outliers-data}
+	\end{center}
+\end{figure*}
+
+The results of \Chippr\ and the alternative estimators of \nz\ are presented in \Fig{fig:nonuniform-outliers-results}.
+The most striking feature is that the histogram of modes is highly sensitive to both outlier populations, producing a severe overestimate in the case of an outlier population like those seen in template-fitting codes and a severe underestimate in the case of an outlier population like those seen in machine learning codes, corresponding to a twenty-fold loss of information compared to the \Chippr\ \mmle\ in both cases.
+The effect on the stacked estimator of \nz\ is more subtle though still concerning.
+In the case of outliers like those resulting from template-fitting, the stacked estimator is overly broad even without realistic intrinsic scatter, resulting in ten times the information loss compared to the \Chippr\ \mmle, and in the case of outliers like those resulting from machine learning, the stacked estimator features an overestimate at the redshift affected by the outlier population, resulting in about five times the information loss as the \Chippr\ \mmle.
+The \Chippr\ \mmle, however, appears unbiased and withstands these effects, and the breadth of the distribution of samples of \nz\ is invariant.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_log_estimators.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_log_estimators.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with catastrophic outliers like those seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right) to the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
+		Though the histogram of modes is most sensitive to a catastrophic outlier population, the stacked estimator also overestimates \nz\ under (machine learning-like outliers) and beyond (template fitting-like outliers).
+	}
+	\label{fig:nonuniform-outliers-results}
+	\end{center}
+\end{figure*}
+
+\subsection{Systematic bias}
+\label{sec:bias}
+
+Systematic bias in \pz\ point estimates is a concern for \lsst's cosmology results, for the same reasons explored in \citet{hoyle_dark_2017}.
+However, in the context of \pzpdf s, the notion of redshift bias is a form of model misspecification.
+Consider that if bias were included in the framework of Figure~\ref{fig:pedagogical_scatter};
+a simple linear transformation of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + z_{\mathrm{phot}})$ would eliminate the bias.
+Regardless, for completeness, a test at ten times the bias of the \lsst\ requirements, with no redshift-dependent intrinsic scatter nor catastrophic outliers, is provided in \Fig{fig:bias}.
+
+\begin{figure}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_log_estimators.png}
+	\caption{
+		Left: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+		Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+		The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
+	}
+	\label{fig:bias}
+	\end{center}
+\end{figure}
+
+As expected based on consistency with the forward model, \Chippr\ is completely resistant to bias, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
+\aim{I'm not sure I effectively explained why \Chippr\ is expected to be unaffected by bias of this form.}
+
+\subsection{Implicit prior}
+\label{sec:interim}
+
+\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
+Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
+Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
+%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.
+%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.
+%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.  
+Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
+% \aim{referenceh chapter 3 for complication of algorithm}
+Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
+%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.  
+%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.  
+%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.
+
+\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring implicit prior emulating that of a machine learning approach to \pz\ estimation (left panel) and a more complex interim prior emulating that of a template-fitting \pz\ method (right panel).
+One can see that the \pzpdf s take different shapes from one another even though the marginal histograms of the points are identical.
+The machine learning-like implicit prior has been modified to have nonzero value at high redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_trpr.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_tmpr.png}
+	\caption{
+		Examples of mock \pzpdf s (colored lines) generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
+		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
+	}
+	\label{fig:pzs-priors}
+	\end{center}
+\end{figure*}
+
+\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzpdf s generated with nontrivial implicit priors.
+In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift distribution, and the distribution of \nz\ parameter values reflects higher uncertainty where the implicit prior undergoes large changes in derivative.
+The alternatives, on the other hand, are biased by the implicit prior except where it is flat, in the case of high redshifts for the machine learning-like implicit prior, resulting in over $1,000$ times the information loss on \nz\ for the machine learning-like implicit prior and some $5-20$ times the information loss for the template fitting-like implicit prior, relative to the \Chippr\ \mmle.
+
+\begin{figure*}
+	\begin{center}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_trpr.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/results_tmpr.png}
+	\caption{
+		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with an implicit prior like that of machine learning \pzpdf\ approaches (left) and an implicit prior like that of template-fitting \pzpdf\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
+		\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
+	}
+	\label{fig:results-priors}
+	\end{center}
+\end{figure*}
+
+%\aim{Move \Sect{sec:violations} here?}
+
+The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzpdf\ catalogs.
+
+\section{Discussion}
+\label{sec:results}
+
+The experiments of \Sect{sec:alldata} isolate the potential sources of error in \nz\ estimation one at a time.
+Now, we stress-test \Chippr\ by investigating two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one in which the \nz\ estimators are not provided with the same implicit prior used to generate the \pzpdf\ catalog (\Sect{sec:violations}).
+
+\subsection{LSST Requirements}
+\label{sec:lsstdemo}
+
+\aim{TODO: clarify the scope: the "binned" samples are already determined by some observational property but redshift distributions are probabilistic as follows.}
+
+It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
+To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} 
+and propagated the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
+Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
+% \aim{don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.49\textwidth]{figures/chippr/cosmolike_inputs.png}
+		\caption{
+			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
+		}
+		\label{fig:tomobins}
+	\end{center}
+\end{figure}
+
+We consider a set of tomographically binned \nz\ and cosmological parameter covariance matrices used for \desc\ forecasting, for which the true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+First, we bin them down to a piecewise constant parameterization with a manageable $35$ parameters for \chippr's sampling capabilities.
+Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
+The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
+We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
+at their given levels.
+Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
+		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
+		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend, and make the axis labels bigger.}
+		}
+		\label{fig:per-bin-scatter}
+	\end{center}
+\end{figure*}
+ \aim{TODO: Ask Elisabeth how to get the bias here}
+
+We then make a point estimate of \nz\ using \Chippr\ as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
+Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
+The excessive breadth of the alternative estimators can be seen quite plainly.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
+		\caption{
+			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
+			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
+			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend, and make the legend text bigger.}
+		}
+		\label{fig:per-bin-ests}
+	\end{center}
+\end{figure*}
+
+We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
+Though there are also slight differences in the angle of the error ellipses, most of the differences are due to broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
+The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
+Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
+Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.94\textwidth]{figures/chippr/final_plot.png}
+		\caption{
+			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
+			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
+			\aim{Are the contours any easier to see now?}
+		}
+		\label{fig:cornerplot}
+	\end{center}
+\end{figure*}
+
+\aim{Hogg says ``I think it would be good to talk a little quantitatively about where people need to know N(z) and other one-point statistics, and how much they will get various things wrong if they don't know these correctly.
+	And situate that discussion within the current context of cosmological parameter estimation and precision cosmology.''
+	Did I successfully answer the question?}
+
+\subsection{Violations of the model}
+\label{sec:violations}
+
+\aim{Move \Sect{sec:interim} here?}
+
+In this test, without tomographic binning, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
+\Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
+If we naively used the \pzpdf\ catalog produced by a generic machine learning code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
+		\caption{
+			Left: Examples of \pzpdf s with systematics at the level of the \lsst\ requirements with a machine learning-like implicit prior, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+			Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and the uniform implicit prior given to \chippr\ (gray curve).
+			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
+		}
+		\label{fig:mischaracterized}
+	\end{center}
+\end{figure}
+
+The results of using a mischaracterized implicit prior are disastrous, causing every estimator, including \Chippr, to be strongly biased.
+The stacked estimator and histogram of modes don't make use of the implicit prior so do no worse than when the implicit prior is accurately provided, but \Chippr\ is sensitive to prior misspecification, which violates the model upon which it is based.
+It is thus crucial that \pzpdf\ methods always characterize and provide the implicit prior.
+
+% removed ancient investigation on real data, would be hard to redo with the new \chippr\ code on this timescale given that I haven't even looked at the data format in 4 years
+
+\section{Conclusion}
+\label{sec:con}
+
+\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
+	 if testable how so? outside the scope of paper, data producers be warned!
+	 a likelihood is better -- give us that if you can!
+	focus on methods that acknowledge probabilistic structure of problem}
+
+\aim{claim: implications for tomographic binning are severe}
+
+This study derives and demonstrates a mathematically consistent inference of a one-point statistic, the redshift density function \nz, based on an arbitrary catalog of \pzpdf s.  
+The fully Bayesian method, based in the fundamental laws of probability, begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
+The method is validated on mock data and tested in the regime of \lsst\ with promising results, outperforming the traditional stacking estimator at the level of \nz\ as well as in terms of constraining power on the cosmological parameters.
+Not only is this the only mathematically correct approach to the problem, it also recovers the true parameter values better than popular alternatives, as measured by the loss of information in \nz\ and the size of error ellipses in the space of cosmological parameters.
+
+%In the tests on simulated data performed here, the full posterior distribution over the hyperparameters defining $N(z)$ derived by this method is consistent with the true redshift distribution function, making the mean of sampled values an excellent point estimator of $N(z)$.
+%The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
+%\aim{Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
+%The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
+%Tests were also performed on subsets of BOSS DR10 data with results consistent with those of simulations.
+
+The following conclusions and recommendations can be made with confidence:
+
+\begin{enumerate}
+	\item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
+	\item The error bars on the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that may be assumed for the conventional point estimators.
+	%	\item The marginalized maximum likelihood estimator is an excellent estimator for strongly featured redshift distribution function with simple, clean photo-$z$ posteriors; stacking smooths features more than sampling and photo-$z$ point estimation.
+	\item When the implicit prior is known to be a poor match to the data, only the results of \Chippr\ are satisfactory estimators of the redshift distribution function because they are the only methods that can account for the bias induced on the \pzpdf\ catalog by the method that produces it; this is the most compelling case for the sampler because of the ubiquity of inappropriate interim priors.
+\end{enumerate}
+
+By showing that \Chippr\ is effective in recovering the true redshift distribution function and posterior distributions on its parameters from catalogs of \pzpdf s, this work supports the production of \pzpdf s by upcoming photometric surveys such as \lsst\ to enable more accurate inference of the cosmological parameters.
+We discourage researchers from co-adding \pzpdf s or converting them into point estimates of redshift and instead recommend the use of Bayesian probability to guide the usage of \pzpdf s.  
+We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for proper inference to be conducted by consumers of this information.
+
+The technique herein developed is applicable with minimal modification to other one-point statistics of redshift to which we will apply this method in the future, such as the redshift-dependent luminosity function and weak lensing mean distance ratio.
+Future work will also include the extension of this fully probabilistic approach to higher-order statistics of redshift such as the two-point correlation function.
+
+\appendix
+%\renewcommand{\thesection}{\Alph{section}}
+%\renewcommand{\thesubsection}{\Alph{subsection}}
+
+\section{Derivation}
 \label{app:math}
-
-\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
 
@@
 
-\subsubsection{Model Limitations}
-\label{sec:limitations}
-
-Finally, we explicitly review the assumptions made by this approach, which are as follows:
-
-\begin{enumerate}
-	\item Photometric measurements of galaxies are statistically independent Poisson draws from the set of all galaxies such that \Eq{eqn:indiedat} and \Eq{eqn:indie} hold.
-	\item We take the reported \pzip s to be accurate, free of model misspecification; draws thereof must not be inconsistent with the distribution of photometry and redshifts.
-	Furthermore, we must be given the implicit prior $\ndphi^{*}$ used to produce the \pzip s.
-	\item We must assume a hyperprior distribution $\pr{\ndphi}$ constraining the underlying probability distribution of the hyperparameters, which is informed by our prior beliefs about the true redshift distribution function.
-\end{enumerate}
-
-These assumptions have known limitations.  
-First, the photometric data are not a set of independent measurements; the data are correlated not only by the conditions of the experiment under which they were observed (instrument and observing conditions) but also by redshift covariances resulting from physical processes governing underlying galaxy spectra and their relation to the redshift distribution function.
-Second, the reported \pzip s may not be trustworthy; there is not yet agreement on the best technique to obtain \pzpdf s, and the implicit prior may not be appropriate or even known to us as consumers of \pzip s.
-Third, the hyperprior may be quite arbitrary and poorly motivated if the underlying physics is complex, and it can only be appropriate if our prior beliefs about \nz\ are accurate.
-
-\subsection{Implementation}
-\label{sec:exp}
-
-We implement the \Chippr\ model in code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
-In \Sect{sec:mcmc}, we describe the publicly available \chippr\ library.
-In \Sect{sec:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
-
-\subsubsection{Code}
-\label{sec:mcmc}
-
-\chippr\ is a \python\ 2 library\footnote{\url{https://github.com/aimalz/chippr}} that includes an implementation of the \Chippr\ model as well as an extensive suite of tools for comparing \Chippr\ to other approaches.
-
-Though there are plans for future expansion to more flexible parameterizations, the current version of \chippr\ uses a log-space piecewise constant parameterization
-\begin{equation}
-\label{eqn:logstepfunc}
-f(z; \ndphi) = \exp[\phi^{k}]\ \mathrm{if}\ z^{k} < z < z^{k+1}
-\end{equation}
-for \nz\ and every \pzpdf, satisfying
-\begin{equation}
-\label{eqn:logstepfuncnorm}
-\sum_{k=1}^{K} \exp[\phi^{k}] \delta z^{k} = 1
-\end{equation}
-with $K$ bins of width $\delta z^{1}, \dots, \delta z^{K}$ defined by endpoints $z^{0}, \dots, z^{K}$.
-%\aim{Maybe include an equation for a general piecewise constant function here (which is currently in \Chap{qp})?}
-Thus each $\pr{z \gvn \data_{j}} = f(z; \ndphi_{j})$ has parameters $\ndphi_{j}$ that are defined in the same basis as those of \nz.
-To infer the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$, one must provide a plaintext file with $K+1$ redshift bin endpoints $\{z_{k}\}$, the parameters $\ndphi^{*}$ of the implicit log-prior, and the parameters $\{\ndphi_{j}\}$ of the log-posteriors $\ln[\pr{z \gvn \data_{j}, \ndphi^{*})}$.
-
-The \emcee \citep{foreman-mackey_emcee_2013} implementation of ensemble sampling is used to sample the full log-posterior of \Eq{eqn:final}. 
-\chippr\ accepts a configuration file of user-specified parameters, among them the number $W$ of walkers.
-At each iteration $i$ and for each walker, a proposal distribution $\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for acceptance to or rejection from the full log-posterior distribution.
-
-Two threshold conditions are defined, one designating all previous samples to be ignored as as products of a burn-in phase and another indicating when a sufficient number of post-burn samples have been accepted.
-In this case, the first threshold (described in \Sect{sec:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
-
-Though previous versions used \texttt{HDF5} for the primary I/O format due to its efficiency for large quantities of data, it was abandoned in favor of \texttt{pickle} in the working release due to the instability of the \python\ implementation of the format on high-performance computing systems.
-The resulting output is a set of $I$ ordered \texttt{hickle} files enumerated by $\rho$ containing the state information after each sub-run.  
-The state information includes $\frac{I_{0}}{s}$ actual samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$ as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$ over the entire sub-run.
-
-\subsubsection{Convergence Criteria}
-\label{sec:acorr}
-
-\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
+\section{Convergence Criteria}
+\label{app:acorr}
 
@@
 
-\subsection{Comparison with Alternative Approaches}
-\label{sec:sheldon}
-
-In this study, we compare the results of \Eq{eqn:fullpost} to those of the two most common approaches to estimating \nz\ from a catalog of \pzpdf s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior probability
-\begin{equation}
-\label{eqn:mmap}
-f^{MMAP}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \delta(z, \mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}])
-\end{equation}
-(i.e. the distribution of modes of the \pzpdf s) and the stacked estimator of \Eq{eqn:stacked}, which can be rewritten as 
-\begin{equation}
-\label{eqn:stacked}
-f^{stack}(z; \hat{\ndphi}) = \sum_{j=1}^{J}\ \pr{z \gvn \data_{j}, \ndphi^{*}}
-\end{equation}
-in terms of the implicit \pz\ posteriors we have.
-These two approaches have been compared to one another by \citet{hildebrandt_cfhtlens:_2012}, \citet{benjamin_cfhtlens_2013}, and \citet{asorey_galaxy_2016} in the past but not to \Chippr.
-
-Point estimation converts the implicit photo-$z$ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$ into delta functions with all probability at a single estimated redshift.  
-Some variants of point estimation choose this single redshift to be that of maximum a posteriori probability $\mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}]$ or the expected value of redshift $\langle z \rangle = \integral{z \pr{z \gvn \data_{j}, \ndphi^{*}}}{z}$.
-\aim{cite tanaka 2018 here} directs attention to deriving an optimal point estimate reduction of a \pzpdf, but since the purpose of this paper is to compare against the most established alternative estimators of \nz, its use will be postponed until a future study.
-Stacking these modified \pzpdf s leads to the marginalized maximum a posteriori (MMAP) estimator and the marginalized expected value (MExp) estimator, though only the former is included in this study since the latter has fallen out of favor in recent years\footnote{And for good reason!
-Consider a bimodal \pzpdf; its expected value may very well fall in a region of very low probability, yielding a less probable point estimate than the point at which either peak achieves its maximum.}.
-
-It is worth discussing the relationship between point estimation and stacking.  
-When the point estimator of redshift is equal to the true redshift, stacking delta function \pzpdf s will indeed lead to an accurate recovery of the true redshift distribution function.  
-However, stacking is in general applied indiscriminately to broader \pzpdf s and imperfect point estimators of redshift.  
-It is for these reasons that alternatives are considered here.
-
-A final estimator of the hyperparameters is the maximum marginalized likelihood estimator, the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.  
-To compare with sampling, the MMLE also depends on the choice of the hyperprior distribution, and it does not produce a full posterior probability distribution over the parameters of interest, only point estimators.
-It must be noted that computation of the maximum marginalized likelihood estimator may be unstable depending on the strengths and weaknesses of the optimizer.  
-In general, derivatives will not be available for the full posterior distribution, restricting optimization methods used.
-%\begin{equation}
-%\label{eq:mmle}
-%\ln[p(\{\vec{d}_{j}\}|\vec{\theta})] \propto -\int\ f_{\vec{\theta}}(z)\ 
-%dz+\sum_{j=1}^{J}\ln\left[\int\ 
-%\exp\left[\ln[p(z_{j}|\vec{d}_{j},\vec{\theta}^{*})]+\ln[f_{\vec{\theta}}(z)]-\l
-%n[f_{\vec{\theta}^{*}}(z)]\right]\ dz\right],
-%\end{equation}
-%accessible with any optimization code.
-
-In \Sect{sec:diag} we outline the measures used to evaluate the performance of the method.
-
-\subsubsection{Performance metrics}
-\label{sec:diag}
-
-The results of the computation described in \Sect{sec:mcmc} are evaluated for accuracy on the basis of some quantitative measures.  
-Beyond visual inspection of samples, we calculate summary statistics to quantitatively compare different estimators' precision and accuracy.  
-Since MCMC samples of hyperparameters are Gaussian distributions, we can quantify the breadth of the distribution for each hyperparameter using the standard deviation regardless of whether the true values are known.
-
-In simulated cases where the true parameter values are known, we calculate the Kullback-Leibler divergence (KLD), given by 
-\begin{equation}
-\label{eqn:kl}
-KL_{',\ddagger} = \integral{\pr{z \gvn \ndphi'} \ln \left[ \frac{\pr{z \gvn \ndphi'}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
-\end{equation}
-which measures a distance from parameter values $\ndphi'$ to true parameter values $\ndphi^{\dagger}$, which is invariant under changes of variables.  
-We note that $KL_{',\dagger} \neq KL_{\dagger,'}$ and is only interpretable when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than $\ndphi'$.
-The KLD is explored in detail in the Appendix to \citet{malz_approximating_2018}.
-In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that produced by one of the methods in question.
-
-\section{Validation}
-\label{sec:alldata}
-
-We compare the results of \Chippr\ to those of stacking and the histogram of \pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed in \Sect{sec:forward}.
-\Fig{fig:flowchart} illustrates the implementation of this forward model used in this paper.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.9\textwidth]{figures/chippr/flowchart.png}
-		\caption{\aim{Check that all names and formulae are defined in the text and the same way as in diagram, say the whole thing in words in the text as well.
-			Explain that this is the forward model of the validation, not of CHIPPR! contrast this with Fig. 1 (joke about simple model and complex validation?)}
-			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
-			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation we perform, and rounded rectangles indicate a quantity created by the forward model.
-			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.}
-		\label{fig:flowchart}
-	\end{center}
-\end{figure*}
-
-The true redshift distribution used in these tests is a particular instance of the gamma function
-\begin{equation}
-\label{eqn:gamma}
-n^{\dagger}(z) = \frac{1}{2 c_{z}} \left(\frac{z}{c_{z}}\right)^{2}\ \exp\left[-\frac{z}{c_{z}}\right]
-\end{equation}
-with $c_{z} = 0.3$, because it has been used in forecasting studies for \des\ and \lsst.
-%\aim{I learned this from talking to people and don't know of a published source that talks about the nitty gritty of the internal validation tests performed before there was data.}
-
-The mock data emulates the three sources of error of highest concern to the \pz\ community that are explored in detail later in this section: intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and systematic bias (\Sect{sec:bias}).
-Tests including all three effects at the tolerance levels of \lsst\ (see Table~\ref{tab:lsstsrd}) 
-are presented in \Sect{sec:results}.
-\Fig{fig:mega_scatter} illustrates these three effects individually at twice the tolerance of \lsst\ for demonstrative purposes, hearkening back to Figure~\ref{fig:pedagogical_scatter}.
-We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to be a priority for the \pz\ community.
-
-\aim{TODO: add in the plot from thesis intro}
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
-		\includegraphics[width=0.32\textwidth]{figures/chippr/single_uout-mega_scatter.png}
-		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
-		\caption{\aim{TODO: Add watermark of "mock data" in UL corner, same for "results of inference" on other kind of plot.}
-			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
-			The main panel of each shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal (thin red line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
-			The insets show marginal histograms (gray) in each dimension, that can be compared with the true \nz\ used to make the figure (blue curve) to see the effect of the isolated systematic, as well as the implicit prior (red line).
-		}
-		\label{fig:mega_scatter}
-	\end{center}
-\end{figure*}
-
-The hyperprior distribution chosen for these tests is a multivariate normal distribution with mean $\vec{\mu}$ equal to the implicit prior $\ndphi^{*}$ and covariance
-\begin{equation}
-\label{eqn:priorcov}
-\Sigma_{k,k'} = q\ \exp[-\frac{e}{2}\ (\bar{z}_{k}-\bar{z}_{k'})^{2}]\ +\ t\delta(k,k')
-\end{equation}
-inspired by one used in Gaussian processes, where $k$ and $k'$ are indices ranging from $1$ to $K$ and $q=1.0$, $e=100.0$, and $t=q\cdot10^{-5}$ are constants chosen to permit draws from this prior distribution to produce shapes similar to that of a true $\tilde{\ndphi}$.
-We adapt the full log-posterior of \Eq{eqn:final} to the chosen binning of redshift space.
-
-%An example of such samples from the prior are shown in \Fig{fig:prior}.
-%
-%\begin{figure}
-%%	\includegraphics[width=0.5\textwidth]{figures/chippr/null_priorsamps.pdf}
-%	\caption{\aim{I need to remake this one because it uses the wrong notation and I stopped making it automatically a while ago.}
-%		Samples (colored lines) of $\pr{z \gvn \ndphi}$ where each $\ndphi$ is drawn from the hyperprior distribution $\pr{\ndphi}$ given in \Eq{eqn:priorcov}.}
-%	\figlabel{fig:prior}
-%\end{figure}
- \aim{TODO: Add back the figure of prior samples?}
-
-The sampler is initialized with $W=100$ walkers each with a value chosen from a Gaussian distribution of identity covariance around a sample from the hyperprior distribution.
-
-\subsection{Intrinsic scatter}
-\label{sec:scatter}
-
-%Several factors contribute to photometric redshifts' intrinsic scatter.  
-%Distant galaxies are dimmer compared to galaxies of identical luminosity that are closer, driving up photometric errors in flux-limited surveys.  
-%The nature of the galaxy sample at higher redshifts also changes, meaning the generation of the photometric redshift posterior based on an a locally-calibrated SED template library or spectroscopically-confirmed training set is more likely to be inappropriate, leading to broader features.
-%In general, the galaxies that could not have been observed spectroscopically will have different and noisier photo-$z$ likelihoods than those that could fall into a spectroscopic training set (or spectroscopically derived template library).
-%This effect may be stronger for high-redshift galaxies.
-
-\Fig{fig:pzs-scatter} shows some examples of \pzpdf s generated with only the systematic of intrinsic scatter, at the level of the \lsst\ requirements on the left and twice that on the right.
-One can see that the histogram of redshift estimates is broader than that of true redshifts, and that the effect is substantially more pronounced by just doubling the intrinsic scatter from the level of the \lsst\ requirements.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter1.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter2.png}
-	\caption{
-		Examples of mock \pzpdf s (colored lines) generated with intrinsic scatter at the \lsst\ requirements (left) and twice the \lsst\ requirements (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
-		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
-	}
-	\label{fig:pzs-scatter}
-	\end{center}
-\end{figure*}
-
-\Fig{fig:results-scatter} shows the \nz\ recovered by \Chippr\ and the alternative approaches.
-As expected, the estimates of \nz\ based on the modes of the \pzpdf s and stacking are broader than the marginalized maximum likelihood estimator from \chippr, with more broadening as the intrinsic scatter increases.
-\Chippr's \mmle\ is robust to intrinsic scatter and is unaffected by increased intrinsic scatter, though the \Chippr\ posterior distribution on the redshift distribution is itself broader for the higher intrinsic scatter case than for the \lsst\ requirements.
-The broadening of the alternative estimators corresponds to a loss of 3-4 times as many nats of information about \nz\ for the \lsst\ requirements relative to the \mmle\ of \Chippr.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter1.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter2.png}
-	\caption{
-		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with intrinsic scatter of the \lsst\ requirements (left) and twice that (right), with the true redshift density (black curve) and implicit prior (gray curve).
-		\Chippr\ is robust to intrinsic scatter, but the alternatives suffer from overly broad \nz\ estimates that worsen with increasing intrinsic scatter.
-	}
-	\label{fig:results-scatter}
-	\end{center}
-\end{figure*}
-
-\subsection{Catastrophic outliers}
-\label{sec:outliers}
-
-As was covered in \Sect{sec:intro}, catastrophic outliers tend to be distributed non-uniformly across the space of observed and true redshift.
-However, the \lsst\ requirements do not specify details for a distribution of outliers to which they were tuned, and it is still instructive to examine the impact of uniform outliers on the inference of \nz, so we begin by addressing uniformly distributed outliers before considering more realistic outlier distributions.
-
-A uniformly distributed population of outliers was simulated by giving every sample in true redshift a $10\%$ chance of having an observed redshift drawn from a uniform distribution rather than the Gaussian about the true redshift.
-Though this results in slightly less than the $10\%$ catastrophic outlier rate, it can be done independently of the definition of the standard deviation so was implemented for demonstrative purposes.
-\Fig{fig:uniform-outliers} shows examples of \pzpdf s from a uniformly distributed outlier population at the level of the \lsst\ requirements (left) as well as the results of \Chippr\ and other \nz\ estimation methods (right).
-The intrinsic scatter of the tests in this section does not increase with redshift as indicated in Table~\ref{tab:lsstsrd}
-in order to isolate the effect of outliers, and is instead held at a constant $\sigma_{z} = 0.02$.
-
-\begin{figure}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_mega_scatter.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_log_estimators.png}
-	\caption{
-		Top: Examples of \pzpdf s with a uniform catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
-		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
-	}
-	\label{fig:uniform-outliers}
-	\end{center}
-\end{figure}
-
-\Fig{fig:uniform-outliers} shows that at the level of the \lsst\ requirements, the alternative estimators are overly broad, whereas \Chippr's \mmle\ yields an unbiased estimate of \nz.
-Further, the result of stacking is even broader than that of the histogram of modes, corresponding to ten times the information loss of \Chippr's \mmle, making it worse than the most naive reduction of \pzpdf s to point estimates.
-
-When one thinks of the \pzpdf s of catastrophic outliers, however, what comes to mind is multimodal \pzpdf s, wherein reducing \pzpdf s to point estimates to make a standard scatterplot of the true and observed redshifts leads to substantial probability density off the diagonal.
-These coordinated catastrophic outliers may be emulated in the joint probability space of true and estimated redshifts by using a mixture of the unbiased diagonal defined by the intrinsic scatter and an additional Gaussian in one dimension, with constant observed redshift for a template-fitting code and constant true redshift for a machine learning code.
-
-In the case of a catastrophic outlier population like that anticipated of template-fitting codes, $10\%$ of all galaxies have their observed redshift at a particular value unrelated to their true redshift, illustrated in the left panel of \Fig{fig:nonuniform-outliers-data}.
-This case is subject to the same caveat as the uniformly distributed outliers when it comes to the \lsst\ requirement.
-It is less straightforward to emulate catastrophic outliers like those anticipated of a machine learning code, those that are truly multimodal.
-The testing conditions here, illustrated in the right panel of \Fig{fig:nonuniform-outliers-data}, gives $10\%$ of galaxies at the redshift affected by outliers an observed redshift that is uniformly distributed relative to the true redshift, meaning that far fewer than $10\%$ of all galaxies in the sample are catastrophic outliers.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_mega_scatter.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_mega_scatter.png}
-	\caption{
-		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-	}
-	\label{fig:nonuniform-outliers-data}
-	\end{center}
-\end{figure*}
-
-The results of \Chippr\ and the alternative estimators of \nz\ are presented in \Fig{fig:nonuniform-outliers-results}.
-The most striking feature is that the histogram of modes is highly sensitive to both outlier populations, producing a severe overestimate in the case of an outlier population like those seen in template-fitting codes and a severe underestimate in the case of an outlier population like those seen in machine learning codes, corresponding to a twenty-fold loss of information compared to the \Chippr\ \mmle\ in both cases.
-The effect on the stacked estimator of \nz\ is more subtle though still concerning.
-In the case of outliers like those resulting from template-fitting, the stacked estimator is overly broad even without realistic intrinsic scatter, resulting in ten times the information loss compared to the \Chippr\ \mmle, and in the case of outliers like those resulting from machine learning, the stacked estimator features an overestimate at the redshift affected by the outlier population, resulting in about five times the information loss as the \Chippr\ \mmle.
-The \Chippr\ \mmle, however, appears unbiased and withstands these effects, and the breadth of the distribution of samples of \nz\ is invariant.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_log_estimators.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_log_estimators.png}
-	\caption{
-		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with catastrophic outliers like those seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right) to the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
-		Though the histogram of modes is most sensitive to a catastrophic outlier population, the stacked estimator also overestimates \nz\ under (machine learning-like outliers) and beyond (template fitting-like outliers).
-	}
-	\label{fig:nonuniform-outliers-results}
-	\end{center}
-\end{figure*}
-
-\subsection{Systematic bias}
-\label{sec:bias}
-
-Systematic bias in \pz\ point estimates is a concern for \lsst's cosmology results, for the same reasons explored in \citet{hoyle_dark_2017}.
-However, in the context of \pzpdf s, the notion of redshift bias is a form of model misspecification.
-Consider that if bias were included in the framework of Figure~\ref{fig:pedagogical_scatter};
-a simple linear transformation of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + z_{\mathrm{phot}})$ would eliminate the bias.
-Regardless, for completeness, a test at ten times the bias of the \lsst\ requirements, with no redshift-dependent intrinsic scatter nor catastrophic outliers, is provided in \Fig{fig:bias}.
-
-\begin{figure}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_mega_scatter.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_log_estimators.png}
-	\caption{
-		Left: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-		Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
-		The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
-	}
-	\label{fig:bias}
-	\end{center}
-\end{figure}
-
-As expected based on consistency with the forward model, \Chippr\ is completely resistant to bias, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
-\aim{I'm not sure I effectively explained why \Chippr\ is expected to be unaffected by bias of this form.}
-
-\subsection{Implicit prior}
-\label{sec:interim}
-
-\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
-Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
-Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
-%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.
-%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.
-%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.  
-Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
-% \aim{referenceh chapter 3 for complication of algorithm}
-Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
-%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.  
-%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.  
-%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.
-
-\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring implicit prior emulating that of a machine learning approach to \pz\ estimation (left panel) and a more complex interim prior emulating that of a template-fitting \pz\ method (right panel).
-One can see that the \pzpdf s take different shapes from one another even though the marginal histograms of the points are identical.
-The machine learning-like implicit prior has been modified to have nonzero value at high redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_trpr.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_tmpr.png}
-	\caption{
-		Examples of mock \pzpdf s (colored lines) generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
-		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
-	}
-	\label{fig:pzs-priors}
-	\end{center}
-\end{figure*}
-
-\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzpdf s generated with nontrivial implicit priors.
-In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift distribution, and the distribution of \nz\ parameter values reflects higher uncertainty where the implicit prior undergoes large changes in derivative.
-The alternatives, on the other hand, are biased by the implicit prior except where it is flat, in the case of high redshifts for the machine learning-like implicit prior, resulting in over $1,000$ times the information loss on \nz\ for the machine learning-like implicit prior and some $5-20$ times the information loss for the template fitting-like implicit prior, relative to the \Chippr\ \mmle.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_trpr.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_tmpr.png}
-	\caption{
-		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with an implicit prior like that of machine learning \pzpdf\ approaches (left) and an implicit prior like that of template-fitting \pzpdf\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
-		\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
-	}
-	\label{fig:results-priors}
-	\end{center}
-\end{figure*}
-
-%\aim{Move \Sect{sec:violations} here?}
-
-The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzpdf\ catalogs.
-
-\section{Discussion}
-\label{sec:results}
-
-The experiments of \Sect{sec:alldata} isolate the potential sources of error in \nz\ estimation one at a time.
-Now, we stress-test \Chippr\ by investigating two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one in which the \nz\ estimators are not provided with the same implicit prior used to generate the \pzpdf\ catalog (\Sect{sec:violations}).
-
-\subsection{LSST Requirements}
-\label{sec:lsstdemo}
-
-\aim{TODO: clarify the scope: the "binned" samples are already determined by some observational property but redshift distributions are probabilistic as follows.}
-
-It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
-To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} 
-and propagated the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
-Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
-% \aim{don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=0.49\textwidth]{figures/chippr/cosmolike_inputs.png}
-		\caption{
-			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
-		}
-		\label{fig:tomobins}
-	\end{center}
-\end{figure}
-
-We consider a set of tomographically binned \nz\ and cosmological parameter covariance matrices used for \desc\ forecasting, for which the true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
-First, we bin them down to a piecewise constant parameterization with a manageable $35$ parameters for \chippr's sampling capabilities.
-Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
-The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
-We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
-at their given levels.
-Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
-		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
-		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend, and make the axis labels bigger.}
-		}
-		\label{fig:per-bin-scatter}
-	\end{center}
-\end{figure*}
- \aim{TODO: Ask Elisabeth how to get the bias here}
-
-We then make a point estimate of \nz\ using \Chippr\ as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
-Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
-The excessive breadth of the alternative estimators can be seen quite plainly.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
-		\caption{
-			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
-			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
-			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend, and make the legend text bigger.}
-		}
-		\label{fig:per-bin-ests}
-	\end{center}
-\end{figure*}
-
-We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
-Though there are also slight differences in the angle of the error ellipses, most of the differences are due to broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
-The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
-Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
-Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.94\textwidth]{figures/chippr/final_plot.png}
-		\caption{
-			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
-			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
-			\aim{Are the contours any easier to see now?}
-		}
-		\label{fig:cornerplot}
-	\end{center}
-\end{figure*}
-
-\aim{Hogg says ``I think it would be good to talk a little quantitatively about where people need to know N(z) and other one-point statistics, and how much they will get various things wrong if they don't know these correctly.
-	And situate that discussion within the current context of cosmological parameter estimation and precision cosmology.''
-	Did I successfully answer the question?}
-
-\subsection{Violations of the model}
-\label{sec:violations}
-
-\aim{Move \Sect{sec:interim} here?}
-
-In this test, without tomographic binning, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
-\Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
-If we naively used the \pzpdf\ catalog produced by a generic machine learning code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
-		\caption{
-			Left: Examples of \pzpdf s with systematics at the level of the \lsst\ requirements with a machine learning-like implicit prior, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-			Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and the uniform implicit prior given to \chippr\ (gray curve).
-			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
-		}
-		\label{fig:mischaracterized}
-	\end{center}
-\end{figure}
-
-The results of using a mischaracterized implicit prior are disastrous, causing every estimator, including \Chippr, to be strongly biased.
-The stacked estimator and histogram of modes don't make use of the implicit prior so do no worse than when the implicit prior is accurately provided, but \Chippr\ is sensitive to prior misspecification, which violates the model upon which it is based.
-It is thus crucial that \pzpdf\ methods always characterize and provide the implicit prior.
-
-% removed ancient investigation on real data, would be hard to redo with the new \chippr\ code on this timescale given that I haven't even looked at the data format in 4 years
-
-\section{Conclusion}
-\label{sec:con}
-
-\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
-	 if testable how so? outside the scope of paper, data producers be warned!
-	 a likelihood is better -- give us that if you can!
-	focus on methods that acknowledge probabilistic structure of problem}
-
-\aim{claim: implications for tomographic binning are severe}
-
-This study derives and demonstrates a mathematically consistent inference of a one-point statistic, the redshift density function \nz, based on an arbitrary catalog of \pzpdf s.  
-The fully Bayesian method, based in the fundamental laws of probability, begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
-The method is validated on mock data and tested in the regime of \lsst\ with promising results, outperforming the traditional stacking estimator at the level of \nz\ as well as in terms of constraining power on the cosmological parameters.
-Not only is this the only mathematically correct approach to the problem, it also recovers the true parameter values better than popular alternatives, as measured by the loss of information in \nz\ and the size of error ellipses in the space of cosmological parameters.
-
-%In the tests on simulated data performed here, the full posterior distribution over the hyperparameters defining $N(z)$ derived by this method is consistent with the true redshift distribution function, making the mean of sampled values an excellent point estimator of $N(z)$.
-%The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
-%\aim{Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
-%The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
-%Tests were also performed on subsets of BOSS DR10 data with results consistent with those of simulations.
-
-The following conclusions and recommendations can be made with confidence:
-
-\begin{enumerate}
-	\item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
-	\item The error bars on the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that may be assumed for the conventional point estimators.
-	%	\item The marginalized maximum likelihood estimator is an excellent estimator for strongly featured redshift distribution function with simple, clean photo-$z$ posteriors; stacking smooths features more than sampling and photo-$z$ point estimation.
-	\item When the implicit prior is known to be a poor match to the data, only the results of \Chippr\ are satisfactory estimators of the redshift distribution function because they are the only methods that can account for the bias induced on the \pzpdf\ catalog by the method that produces it; this is the most compelling case for the sampler because of the ubiquity of inappropriate interim priors.
-\end{enumerate}
-
-By showing that \Chippr\ is effective in recovering the true redshift distribution function and posterior distributions on its parameters from catalogs of \pzpdf s, this work supports the production of \pzpdf s by upcoming photometric surveys such as \lsst\ to enable more accurate inference of the cosmological parameters.
-We discourage researchers from co-adding \pzpdf s or converting them into point estimates of redshift and instead recommend the use of Bayesian probability to guide the usage of \pzpdf s.  
-We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for proper inference to be conducted by consumers of this information.
-
-The technique herein developed is applicable with minimal modification to other one-point statistics of redshift to which we will apply this method in the future, such as the redshift-dependent luminosity function and weak lensing mean distance ratio.
-Future work will also include the extension of this fully probabilistic approach to higher-order statistics of redshift such as the two-point correlation function.
-
-%%\section*{Appendices}
-%%
-%%\begin{appendices}
-%%
-%%%\section{Mathematical derivation}
-%%%\label{app:math}
-%%%
-%%%\section{Mock data generation}
-%%%\label{app:data}
-%%%
-%%%\textul{Plots: flow chart of forward model}
-%%%
-%%%\begin{figure*}
-%%%	\begin{center}
-%%%		\includegraphics[width=0.75\textwidth]{fig/flowchart.png}
-%%%		\caption{In appendix?}
-%%%		\label{fig:flowchart}
-%%%	\end{center}
-%%%\end{figure*}
-%%
-%%
-%%\end{appendices}
-
 \begin{acknowledgements}
@@
 The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
-% send to DFM for comments
+\aim{TODO: Send draft around to Foreman-Mackey, Leistedt, others for feedback.}
 This work was completed with generous nutritional support from the Center for Computational Astrophysics.

### 2019-06-21 research/paper/draft.tex
responding to feedback from Hogg throughout

--- 2019-05-24 research/paper/draft.tex
@@
 \newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
+\newcommand{\dwh}[1]{\todo{Attn. Hogg:}{blue}{#1}}
 \newcommand{\myemail}{aimalz@nyu.edu}
@@
 \newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
-\newcommand{\sz}{spec-$z$~}
-\newcommand{\Sz}{Spec-$z$~}
-\newcommand{\pz}{photo-$z$~}
-\newcommand{\Pz}{Photo-$z$~}
-\newcommand{\pzpdf}{\pz PDF}
+\newcommand{\sz}{spec-$z$}
+\newcommand{\Sz}{Spec-$z$}
+\newcommand{\pz}{photo-$z$}
+\newcommand{\Pz}{Photo-$z$}
+\newcommand{\pzpdf}{\pz\ PDF}
 % could change to posterior
-\newcommand{\Pzpdf}{\Pz PDF}
+\newcommand{\Pzpdf}{\Pz\ PDF}
 % could change to posterior
@@
 Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
-The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
+The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ posterior probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
 However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
 The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
-We introduce a mathematically rigorous probabilistic graphical model (PGM) of $n(z)$ for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
-Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator produced by traditional procedures.
-The CHIPPR approach is applicable to any one-point statistic of any random variable.
-We conclude by propagating these effects to constraints in the space of cosmological parameters, .
-\aim{be clear that rigor comes at a cost, chippr makes explicit the requirements that the pzpdfs be posteriors under a known prior (which would be required by any justifiable method), be honest that we're not ready for prime time}
+We introduce a statistically rigorous probabilistic graphical model (PGM) of $n(z)$ for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
+Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures.
+Using the publicly available \texttt{chippr} prototype code, we propagate these effects to constraints in the space of cosmological parameters, noting, however, that the mathematically justifiable approach incurs computational expense.
+The CHIPPR approach is applicable to any one-point statistic of any random variable, provided the prior probability distribution used to produce the posteriors is known.
 \end{abstract}
@@
 \maketitle
-
-\aim{Some of the figures refer to left/right but are shown as top/bottom \textemdash I'm still working through a LaTeX bug.}
 
@@
 
-Photometric redshift (\pz) estimation has been a staple of studies of galaxy evolution, large-scale structure, and cosmology since its conception decades ago \citep{baum_photoelectric_1962}.  
+Photometric redshift (\pz) estimation has been a staple of studies of galaxy evolution, large-scale structure, and cosmology since its conception half a century ago \citep{baum_photoelectric_1962}.  
 An extremely coarse spectrum in the form of photometry in a handful of broadband filters can be an effective substitute for the time- and photon-intensive process of obtaining a spectroscopic redshift (\sz), a procedure that may only be applied to relatively bright galaxies.
@@
 Calculations of correlation functions of cosmic shears and galaxy positions that constrain the cosmological parameters require large numbers of high-confidence redshifts of surveyed galaxies.  
-Many more \pz s may be obtained in the time it would take to observe a smaller number of \sz s, and \pz s may be measured for galaxies too dim for accurate \sz\ confirmation, permitting the compilation of large catalogs of galaxy redshifts spanning a broad range of redshifts and luminosities.
+Many more \pz s may be obtained in the time it would take to observe a smaller number of \sz s, and \pz s may be measured for galaxies too dim for accurate \sz\ confirmation, permitting the compilation of large catalogs of galaxies spanning a broad range of redshifts and luminosities.
 \Pz s have thus enabled the era of precision cosmology, heralded by weak gravitational lensing tomography and baryon acoustic oscillation peak measurements.
 
-However, \pz s are susceptible to inaccuracy and imprecision via a number of effects, particularly their inherent noisiness due to the coarseness of photometric filters, catastrophic errors in which galaxies of one type at one redshift are mistaken for galaxies of another type at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
+However, \pz s are susceptible to inaccuracy and imprecision via a number of effects, particularly their inherent noisiness due to the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
 Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
-This plot looks very much like the traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of \pz\ point estimates and was indeed adapted from one such plot in \citet{jain_whole_2015}, because \pz\ point estimates effectively are a special nonlinear projection of the data.
+The similarity to traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of \pz\ point estimates is no coincidence, as \pz\ point estimates effectively are indeed one such special nonlinear projection of the data.
 
@@
 There are populations of outliers, far from the diagonal, comprised of galaxies for which the redshift estimate is catastrophically distinct from the true redshift, showing that outliers are not uniformly distributed nor restricted to long tails away from a Gaussian scatter.
-And, though hardly perceptible in the plot, there is a systematic bias, wherein the average of the points would not lie on the diagonal but would be offset by a small amount, suggested by the trend of points to lie below the diagonal at high redshift.
+And, though hardly perceptible in the plot, there is a systematic bias, wherein the average of the points would not lie on the diagonal but would be offset by a small bias, suggested by the trend of high redshift points to lie below the diagonal.
+
+Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are not insignificant contributors to the total uncertainties reported on cosmological parameters.
+As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
 Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
@@
 
-Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are not insignificant contributors to the total uncertainties reported on cosmological parameters.
-As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
-
 Much effort has been dedicated to improving \pz s, though they are still most commonly obtained by a maximum likelihood estimator (MLE) based on libraries of galaxy SED templates, with conservative approaches to error estimation.
@@
 Data-driven approaches tend to result in catastrophic outliers like the vertically oriented population of \Fig{fig:pedagogical_scatter}, due to training sets that are incomplete in redshift coverage.
-Recent advances have focused on identifying and removing catastrophic outliers when using \pz s for 
-inference \citep{gorecki_new_2014}.  
 The approaches of using a training set versus a template library are related to one another by \citet{budavari_unified_2009}.
-Sophisticated Bayesian techniques and cutting-edge machine learning methods have been employed to improve precision \citep{carliles_random_2010} and accuracy \citep{sadeh_annz2:_2016}.
-
-An alternative to point estimation of \pz s is redshift probability distribution function (PDF) estimation that reports the probability $\pr{z}$ over all possible redshifts for every galaxy rather than an MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.
-This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating notions of precision, accuracy, and systematic errors.
+Sophisticated Bayesian techniques and machine learning methods have been employed to improve precision \citep{carliles_random_2010} and accuracy \citep{sadeh_annz2:_2016}, while other advances have focused on identifying and removing catastrophic outliers when using \pz s for inference \citep{gorecki_new_2014}.
+
+An alternative to point estimation of \pz s is redshift probability distribution function (PDF) estimation that reports the probability of all possible redshifts for every galaxy, commonly written as $\pr{z}$, rather than an MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.
+This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating our understanding of precision, accuracy, and systematic error.
 However, denoting \pzpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; \pzpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
@@
 \Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.
-\Pzpdf s are not without their own weaknesses, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014} and the divergent results of each method used to derive them \citep{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.
-The most important of these issues, however, is that use of them in the literature is inconsistent at best and incorrect at worst.  
-The most common application of \pzpdf s is their use in estimating \Nz, the distribution of redshifts of a sample of galaxies, a quantity essential to the calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citet{mandelbaum_precision_2008, sheldon_photometric_2012, bonnett_redshift_2016}.
-%\aim{Provide equations/citations for the use of \Nz\ in cosmology.}
-
-If the true redshifts $\{z_{i}^{\dagger}\}$ of galaxies $i$ were known, then the redshift PDFs would be delta functions $\{\delta(z, z_{i}^{\dagger})\}$ centered at the true redshift, and the redshift distribution could be effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{i}^{\dagger})\}$.
-When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{i})$ in place of $\delta(z, z_{i}^{\dagger})$.
-Though it is most common for $\hat{z}_{i}$ to be the mode (maximum) of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.
-
-To circumvent the reduction of uncertainty information due to point estimation, however, it has grown to be more common to calculate the mathematically inconsistent but conceptually simple \textit{stacked estimator} of the redshift density function \citep{lima_estimating_2008},
+\Pzpdf s are not without their own shortcomings, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014, malz_approximating_2018} and the divergent results of each method used to derive them \citep[ and reviewed in Schmidt, Malz \& Soo, et al.
+(in prep)]{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.
+The most concerning weakness of \pzpdf s, however, is their usage in the literature, which is at best inconsistent and at worst incorrect.  
+The most prevalent application of \pzpdf s is their use in estimating the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}, a quantity essential to the calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, leistedt_hierarchical_2016}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
+
+\aim{TODO:
+	Motivate how well we need to know \Nz\ and why it matters.
+	Provide equations/citations for the use of \Nz\ in cosmology.
+	New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
+	Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}
+
+If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
+When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
+Though it is most common for $\hat{z}_{j}$ to be the maximum or \textit{mode} of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.
+
+However, any procedure that reduces \pzpdf s to point estimates loses information about the uncertainty on redshift.
+To circumvent this loss, it has grown to be more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
 \begin{align}
 \label{eqn:stack}
-\hat{n}(z) &= \frac{1}{\ntot} \sum_{i = 0}^{\ntot} \pr{z}_{i}
+\hat{n}(z) &= \frac{1}{J} \sum_{j = 0}^{J} \pr{z}_{j}
 \end{align}
-for a sample of $\ntot$ galaxies $i$, or, equivalently, the redshift distribution function $\hat{N}(z) = \ntot \hat{n}(z)$.
-Much of this thesis centers around the problems with this logically invalid yet pervasive quantity.
-
-\aim{TODO: Pull in content from global intro of thesis.}
-
-Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
-They have been used to form selection criteria of samples from galaxy surveys without propagation 
-through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
-Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.
+for a sample of $J$ galaxies $j$, or, equivalently, the redshift distribution function $\hat{N}(z) = J \hat{n}(z)$, by effectively averaging the \pzpdf s.
+
+%Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
+%They have been used to form selection criteria of samples from galaxy surveys without propagation 
+%through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
+%Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.
 
 Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
-We present and validate a hierarchical Bayesian technique for the use of \pzpdf\ in inference of arbitrary statistics relevant to cosmology, large-scale structure, and galaxy evolution.  
-For simplicity, we consider only one-point statistics, though future work will extend this methodology to higher-order statistics.
-
-The \textit{redshift distribution function \Nz}, or, almost interchangably, its normalized cousin the \textit{redshift density function \nz}, serves as an ideal statistic upon which to demonstrate this novel approach.
-\Nz\ is necessary for calculations of two-point correlation functions of weak gravitational lensing and counting statistics that are used to probe dark energy \citep{masters_mapping_2015}.  
-The observed \Nz\ also been used to validate survey selection functions usnfenjed in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.  
-Additionally, \Nz\ has been the subject of inference using \pzpdf s before \citep{sheldon_photometric_2012, hildebrandt_cfhtlens:_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016, leistedt_hierarchical_2016}, so comparisons to the literature may easily be made.
-
-The prevailing way to estimate \nz\ for a sample of galaxies $j$ is to ``stack'' \pzpdf s according to Equation~\ref{eqn:stack}.
-The stacked estimator of the redshift density function is effectively an average of the \pzpdf\ catalog and, as we show in this paper, not in general mathematically valid.
-Stacking is nonetheless considered the preferred method for obtaining \Nz\ from a \pzpdf\ catalog \citep{sheldon_photometric_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016}.
-However, it must be noted here that Equation~\ref{eqn:stack} 
-is not in general mathematically valid.  
-(See \citet{hogg_data_2012} for a complete discussion.)
-
-%\aim{Motivate how well we need to know \Nz\ and why it matters.
-%	New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
-%	Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}
-
+This paper challenges the logically invalid yet pervasive analysis procedure of stacking \pzpdf s by presenting and validating a hierarchical Bayesian technique for the use of \pzpdf s\ in the inference of \nz, yielding a method applicable to arbitrary one-point statistics relevant to cosmology, large-scale structure, and galaxy evolution; future work will extend this methodology to higher-order statistics.
 We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
-Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys that may be resistant to restructuring their analysis pipelines.
+Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.
 
@@
 
-The redshift distribution function \nz\ is the number of galaxies per unit redshift, effectively defining the evolution in the number of galaxies \citep{menard_clustering-based_2013}.  
-% \aim{also selection function}
+The redshift distribution function \nz\ is the number of galaxies per unit redshift, effectively defining the evolution in the number of galaxies convolved with the selection function of the sample \citep{menard_clustering-based_2013}.
 In the following sections, we present and compare methods for estimating \nz\ from \pzpdf s.  
@@
 \label{sec:prob}
-
-\aim{TODO: fix implicit posterior vs. implicit-weighted posterior}
 
@@
 		\caption{The directed acyclic graph of the CHIPPR model, where circles indicate random variables and arrows indicate causal relationships.
-			The redshift distribution function parameterized by hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, indicated as a box.  
+			The redshift distribution \nz\ parameterized by hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, indicated as a box.  
 			The redshifts $\{z_{j}\}$ of all galaxies in the survey are latent variables independently drawn from the redshift distribution, which is a function of $\ndphi$. 
@@
 The problem facing cosmologists is to determine the true value of $\ndphi$ from observing the photometry $\{\data_{j}\}$ of a large sample of $J$ galaxies $j$.
-To self-consistently propagate the uncertainty in the redshifts, however, it is more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ over all possible values of $\ndphi$ conditioned on all the observed data $\{\data_{j}\}$ available in a generic catalog.
-In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the implicit prior and the prior probability density.
-
-When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$.
+To self-consistently propagate the uncertainty in the inference of redshift, however, it is more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ over all possible values of $\ndphi$ conditioned on all the observed data $\{\data_{j}\}$ available in a generic catalog.
+In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the \textit{implicit prior} and the \textit{prior probability density (prior PDF)}, elaborated upon below.
+
+\dwh{When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$, the probability of an unknown quantity conditioned on the quantity we have in hand, i.e the data.
 However, to do this, we must have a model for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
-Such a relationship is defined in the space of probability density over redshift, so it must be able to be parameterized by the same functional form $f(z; \ndphi)$ as \nz .
-It is thus natural to write it as $\pr{z \gvn \ndphi^{*}}$, where $\ndphi^{*}$ is the parameters for this relationship under some generic \pzpdf\ method.
-We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher; for template-based methods where it can be chosen to be ``realistic,'' it may be appropriate to call it an \textit{interim prior}.
-Because the implicit prior is unavoidable, the \pzpdf s reported by any method are really \textit{implicit-weighted posteriors} $\pr{z \gvn \data, \ndphi^{*}}$.
+Such a relationship $\pr{\data, z}$ is defined in the multidimensional space of probability density over redshift and photometry.
+If we were to marginalize over the photometry in $\pr{\data, z}$, we would obtain a one-dimensional PDF $\pr{z \gvn \ndphi^{*}}$ over redshift, which can by definition be parameterized by the same functional form as \nz, for some $\ndphi^{*}$ specific to the estimation procedure that may or may not bear any relation to the hyperparameters $\ndphi^{\dagger}$ of the true \nz.
+Rather, $\ndphi^{*}$ is a consequence of generative model for how photometry results from redshift, including the influence of intrinsic galaxy spectra and instrumental effects. 
+We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher\footnote{For template-based methods, the implicit prior is often an explicitly known input to the algorithm, engineered as an initial guess for the true $\ndphi$, with an aim for a realistic choice guided by an earlier spectroscopic survey.
+(See \citet{benitez_bayesian_2000} for more detail.)
+It may thus be more appropriate to call it an \textit{interim prior}, but we will use the former term throughout this paper for generality.}
+Because the implicit prior is unavoidable and almost inherently not uninformative, the \pzpdf s reported by any method must be \textit{implicit posteriors} $\pr{z \gvn \data, \ndphi^{*}}$ weighted by the implicit prior.}
 
@@
 
-The implicit prior $\ndphi^{*}$ may be thought of as an initial guess for the parameters contained in $\ndphi$, inspired by the generative model for photometry from the redshift distribution functions and including some parameters defining intrinsic galaxy spectra and instrumental effects.
-(See \citet{benitez_bayesian_2000} for more detail.)  
-For statistical purposes, we would like any interim prior to be uninformative, but this is rarely achievable.  
-In the case of estimating \nz\ photometrically, it is common to use $\ndphi^{*}$ corresponding to \nz\ derived from some different, spectroscopically confirmed sample or from a cosmological simulation.
-
-The prior probability density $\pr{\ndphi}$ is a more familiar concept in astronomy; to progress, we will have to choose a prior probability density over all possible parameters $\ndphi$.
+The prior probability density $\pr{\ndphi}$ is a more familiar concept in astronomy; to progress, we will have to choose a prior probability density over all possible values of the hyperparameters $\ndphi$.
 This prior need not be excessively proscriptive; for example, it may be chosen to enforce smoothness at physically motivated scales in redshift without imposing any particular region as over- or under-dense.
 
-\aim{TODO: Clarify what we want to compute ($\pr{\ndphi \gvn\ \data}$), what we're given ($\{\pr{z \gvn \data, \ndphi^{*}}\}$).}
-\aim{TODO: Give answer here, say this is correct, move derivation to appendix.}
-\aim{TODO: Give human-readable description of ratios, explain that \Chippr\ replaces implicit prior with sampled model, converting posterior into likelihood.}
-
-With these definitions, we obtain the desired expression for $\pr{\ndphi \gvn \{\data_{j}\}}$,
+\dwh{With inputs of the \pzip\ catalog $\{\pr{z \gvn \data, \ndphi^{*}}\}$, the implicit prior $\pr{z \gvn \ndphi^{*}}$, and the prior PDF $\pr{\ndphi}$, we thus aim to obtain the posterior probability $\pr{\ndphi \gvn \{\data_{j}\}}$ of the redshift density function given all the photometric data.
+By performing the derivation of Appendix~\ref{app:math}, we arrive at the desired expression
 \begin{equation}
@@
 \end{equation}
-the heart of \Chippr.
-The derivation is provided below in \Sect{app:math}.
-
-\subsubsection{Derivation}
-\label{app:math}
-
-\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
-
-%We begin by parametrizing $N(z)$ in terms of $\vec{\theta}$, comprising some set of hyperparameters that define the form $N(z)$ may take in whatever basis we choose.  
-%We define a function $f_{\vec{\theta}}(z)=N(z)$ that transforms these hyperparameters into the redshift distribution function $N(z)$.  
-%Because 
-%\begin{equation}
-%\eqlabel{eq:definition}
-%N(z) \propto p(z \gvn \vec{\theta}),
-%\end{equation}
-%we may discontinue discussion of $N(z)$ in favor of the likelihood $p(z|\vec{\theta})$.
-
-In this paper, we work exclusively with log-probabilities.  
-What we wish to estimate is then the full log-posterior probability distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ given the catalog of photometry $\{\data_{j}\}$.
-
-By Bayes' Rule, the full log-posterior
-\begin{equation}
-\label{eqn:basicbayes}
-\ln[\pr{\ndphi \gvn \{\data_{j}\}}] = \ln[\pr{\{\data_{j}\} \gvn \ndphi}] + \ln[\pr{\ndphi}] - \ln[\pr{\{\data_{j}\}}]
-\end{equation}
-may be expressed in terms of the full log-likelihood probability distribution (hereafter the full log-likelihood) $\ln[\pr{\{\data_{j}\} \gvn \ndphi}]$ by way of a hyperprior log-probability distribution (hereafter the hyperprior) $\ln[\pr{\ndphi}]$ over the hyperparameters and the log-evidence probability of the data $\ln[\pr{\{\data_{j}\}}]$.
-However, the evidence is rarely known, so we probe the full log-posterior modulo an unknown constant of proportionality.
-
-The full log-likelihood may be expanded in terms of a marginalization over the redshifts as parameters, as in 
-\begin{equation}
-\label{eqn:marginalize}
-\ln[\pr{\{\data_{j}\} \gvn \ndphi}] = \ln\left[\integral{\pr{\{\data_{j}\} \gvn \{z_{j}\}} \pr{\{z_{j}\} \gvn \ndphi}}{\{z_{j}\}}\right].
-\end{equation}
-
-We shall make two assumptions of independence in order to make the problem tractable; their limitations are be discussed below.  
-First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn z_{j}}]$, as in 
-\begin{equation}
-\label{eqn:indiedat}
-\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn z_{j}}],
-\end{equation}
-a result of the definition of probabilistic independence encoded by the box in \Fig{fig:pgm}.
-Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent draws from the true $\pr{z \gvn \ndphi}$.  
-Additionally, $J$ itself is a Poisson random variable.  
-The combination of these assumptions is given by 
-\begin{equation}
-\label{eqn:indie}
-\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ \ln[\pr{z_{j} \gvn \ndphi}].
-\end{equation}
-%It is important to note that the integral $\integral{n(z)}{z} N(z)\ dz$ is not constrained to equal the variable defining the Poisson distribution but instead $J$ by \Eq{eq:definition}, which can be thought of as another parameter.
-The derivation differs when $J$ is not known, say, when we want to learn about a distribution in nature rather than a distribution specific to data in hand, but for a photometric galaxy catalog where the desired quantity is $n(z)$ for the galaxies entering a larger cosmology calculation, it is a fixed quantity.
-A detailed discussion of this matter may be found in \citet{foreman-mackey_exoplanet_2014}.  
-Applying Bayes' Rule, we may combine terms to obtain 
-\begin{align}
-\begin{split}
-\label{eqn:posterior}
-\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] - \integral{f(z; \ndphi)}{z}\\
-& + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn \ndphi}}{z}\right].
-\end{split}
-\end{align}
-
-%\Eq{eq:posterior} contains two quantities that merit further discussion, the prior distribution $p(\vec{\theta})$ discussed further in \Sect{sec:exp} and the photo-$z$ log-likelihoods $\ln[p(\vec{d}_{j}|z_{j})]$ that have not been mentioned since \Eq{eq:marginalize}.
-%Though photo-$z$ log-likelihoods would be desirable for use in these equations, they are not generally the product of either empirical and data-driven methods for obtaining photo-$z$ probability distributions.
-%Though probabilistic photo-$z$s are typically reported as generic probability distributions $p(z_{j})$, the methods that produce them may be understood to always yield posteriors, probability distributions conditioned on the data we believe to be true.
-%If they were not based in this assumption, they would require a sum over an infinite space of possible datasets.
-
-Since we only have access to implicit \pz\ posteriors, we must be able to write the full log-posterior in terms of implicit \pz\ log-posteriors rather than the log-likelihoods of \Eq{eqn:posterior}.
-To do so, we will need an explicit statement of this implicit prior $\ndphi^{*}$ for whatever method is chosen to produce the implicit \pz\ posteriors.
-
-To perform the necessary transformation from likelihoods to posteriors, we follow the reasoning of \citet{foreman-mackey_exoplanet_2014}.  
-Let us consider the probability of the parameters conditioned on the data and an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} as 
-\begin{align}
-\label{eqn:trick}
-\begin{split}
-\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\\
- & - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
- \end{split}
-\end{align}
-
-Once the implicit prior $\ndphi^{*}$ is explicitly introduced, we may expand the last term in \Eq{eqn:trick} according to Bayes' Rule to get 
-\begin{align}
-\begin{split}
-\label{eqn:expand}
-\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\\
-& + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}] \\
-& - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
-\end{split}
-\end{align}
-Because there is no direct dependence of the data upon the hyperparameters, we may again expand the term $\ln[\pr{\data_{j} \gvn z, \ndphi^{*}}]$ to obtain 
-\begin{align}
-\begin{split}
-\label{eqn:indterm}
-\ln[\pr{\vec{d}_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] \\
-& + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}]\\
-& - \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] .
-\end{split}
-\end{align}
-Canceling the undesirable terms for the inaccessible likelihood $\ln[\pr{\data_{j} \gvn z}]$ and trivial $\ln[\pr{\data_{j} \gvn \ndphi^{*}}]$ yields
-\begin{equation}
-\label{eqn:cancel}
-\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]  - \ln[\pr{z \gvn \ndphi^{*}}].
-\end{equation}
-We put this all together to get the full log-posterior probability distribution of 
-\begin{align}
-\begin{split}
-\label{eqn:final}
-\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto & \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \right.\right.\\
-& \left.\left.\left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}]\right.\right.\right.\\
-& \left.\left.\left. - \ln[\pr{z \gvn \ndphi^{*}}] \right)\right]}{z}\right] ,
-\end{split}
-\end{align}
-\aim{TODO: fix the trailing parentheses/brackets.}
-
-The argument of the integral in the log-posterior of \Eq{eqn:final} depends solely on knowable quantities (and those we must explicitly assume) and can be calculated for a given sample of \pz\ log-posteriors $\{\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with which they were obtained, noting the relation of
-\begin{equation}
-\label{eqn:params}
-\pr{z \gvn \ndphi} = \frac{f(z; \ndphi)}{\integral{f(z; \ndphi)}{z}}.
-\end{equation}
-Since we cannot know constant of proportionality, we sample the desired full log-posterior $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$ using Monte Carlo-Markov chain (MCMC) methods.  
-The method outlined here is valid regardless of how the implicit \pz\ log-posteriors are obtained so the many approaches to producing \pzpdf s will not be rehashed; though the matter is outside the scope of this paper, reviews of various methods have been presented in the literature \citep{sheldon_photometric_2012, ball_robust_2008, carrasco_kind_tpz:_2013, carrasco_kind_exhausting_2014}, and will be briefly reviewed in \aim{cite PZDC1 paper here}.
-%
-%\begin{align}
-%\begin{split}
-%\eqlabel{eqn:fullpost}
-%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
-%\end{split}
-%\end{align}
+which is the very heart of \Chippr.
+This in effect replaces the implicit prior with the sampled model hyperparameters, thereby converting the \pzip s into likelihoods in order to obtain unbiased posteriors.}
 
@@
 Finally, we explicitly review the assumptions made by this approach, which are as follows:
-
 \begin{enumerate}
@@
 In \Sect{sec:mcmc}, we describe the publicly available \chippr\ library.
-In \Sect{sec:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
+In \Sect{app:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
 
@@
 Two threshold conditions are defined, one designating all previous samples to be ignored as as products of a burn-in phase and another indicating when a sufficient number of post-burn samples have been accepted.
-In this case, the first threshold (described in \Sect{sec:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
+In this case, the first threshold (described in \Sect{app:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
 
@@
 
-\subsubsection{Convergence Criteria}
-\label{sec:acorr}
-
-\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}
-
-In addition to qualitative visual inspection of the chains, two quantities that probe the convergence of the sampler are used in this study, the autocorrelation time and the Gelman-Rubin convergence criterion.
-%\Fig{fig:chains} shows the %evolution of the values of one parameter of one walker over the course of all %iterations of the sampler.
-
-%\begin{figure}
-%%\includegraphics[width=0.5\textwidth]{figs/null/chain0.pdf}
-%\caption{This figure shows the evolution of one walker's parameter values for 
-%one element of the parameter vector $\vec{\theta}$ as a function of iteration 
-%number, demonstrating the completion of the burn-in phase.}
-%\label{fig:chains}
-%\end{figure}
-
-The autocorrelation time is effectively a measure of the efficiency of the method and can be described as the expected number of iterations necessary to accept a new sample independent of the current accepted sample.
-A sampler that converges faster will have a smaller autocorrelation time, and smaller autocorrelation times are preferable because it means fewer iterations are wasted on non-independent samples when independent samples are desired.
-See \citet{foreman-mackey_emcee_2013} for a more complete exploration of the autocorrelation time.  
-In all tests discussed here, autocorrelation times across walkers and parameters were approximately 20, meaning two samples 20 or more iterations apart were independent, a satisfactory level of efficiency.
-Low autocorrelation times are a necessary but not always sufficient convergence condition, as the autocorrelation times calculated for tests in this paper were constant across all sub-runs, even those that were obviously burning in.
-
-The Gelman-Rubin statistic
-\begin{equation}
-\label{eqn:gr}
-R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} b_{k}}{w_{k}}},
-\end{equation}
-a weighted sum of the mean $w_{k}$ of the variances within individual walkers' chains and the variance $b_{k}$ between chains of different walkers $m$, is calculated over each sub-run $i$ to determine the duration of the burn-in period.
-Convergence is achieved when the statistic approaches unity.
-
 \subsection{Comparison with Alternative Approaches}
@@
 Some variants of point estimation choose this single redshift to be that of maximum a posteriori probability $\mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}]$ or the expected value of redshift $\langle z \rangle = \integral{z \pr{z \gvn \data_{j}, \ndphi^{*}}}{z}$.
-\aim{cite tanaka 2018 here} directs attention to deriving an optimal point estimate reduction of a \pzpdf, but since the purpose of this paper is to compare against the most established alternative estimators of \nz, its use will be postponed until a future study.
+\citet{tanaka_photometric_2018} directs attention to deriving an optimal point estimate reduction of a \pzpdf, but since the purpose of this paper is to compare against the most established alternative estimators of \nz, its use will be postponed until a future study.
 Stacking these modified \pzpdf s leads to the marginalized maximum a posteriori (MMAP) estimator and the marginalized expected value (MExp) estimator, though only the former is included in this study since the latter has fallen out of favor in recent years\footnote{And for good reason!
@@
 We compare the results of \Chippr\ to those of stacking and the histogram of \pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed in \Sect{sec:forward}.
-\Fig{fig:flowchart} illustrates the implementation of this forward model used in this paper.
+\dwh{\Fig{fig:flowchart} illustrates the implementation of the forward model, defined by the much simpler \Fig{fig:pgm}, used for validating the method presented here.
+The irony of a simple model and complex validation procedure is not lost on the authors.}
+
+\dwh{\Fig{fig:flowchart} outlines the four phases of the generative model, which uses a total of three inputs.
+The experimental design requires our choice of true values $\phi^{\dagger}$ of the hyperparameters governing \nz, a \pz\ model $\pr{z, \data}$ defining the space of redshift and photometry, and prior values $\phi^{*}$ of the hyperparameters of \nz.
+In the first phase, we sample $J = 10^{4}$ redshifts $z_{j}^{\dagger} \sim \pr{z \gvn \phi^{\dagger}}$.
+In the second phase, we evaluate the \pz\ model at those redshifts, yielding a set of $J$ likelihoods $\pr{\data \gvn z_{j}^{\dagger}}$, from which we then sample data $\data_{j}^{\dagger} \sim \pr{\data \gvn z_{j}^{\dagger}}$ for each galaxy.
+In the third phase, we evaluate the \pz\ model at that data to obtain $J$ posteriors $\pr{z \gvn \data_{j}^{\dagger}}$.
+In the fourth phase, we convolve the posteriors with the chosen prior $\pr{z \gvn \phi^{*}}$, yielding implicit posteriors $\pr{z \gvn \data_{j}^{\dagger}, \phi^{*}}$.}
 
@@
 		\includegraphics[width=0.9\textwidth]{figures/chippr/flowchart.png}
-		\caption{\aim{Check that all names and formulae are defined in the text and the same way as in diagram, say the whole thing in words in the text as well.
-			Explain that this is the forward model of the validation, not of CHIPPR! contrast this with Fig. 1 (joke about simple model and complex validation?)}
+		\caption{
 			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
 			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation we perform, and rounded rectangles indicate a quantity created by the forward model.
-			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.}
+			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.
+			}
 		\label{fig:flowchart}
@@
 We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to be a priority for the \pz\ community.
-
-\aim{TODO: add in the plot from thesis intro}
 
@@
 	\caption{
+		\aim{TODO: Add watermark of "mock data" in UL corner, same for "results of inference" on other kind of plot.}
 		Left: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
@@
 
-%\aim{Move \Sect{sec:violations} here?}
+\aim{Move \Sect{sec:violations} here?}
 
@@
 
-\aim{TODO: clarify the scope: the "binned" samples are already determined by some observational property but redshift distributions are probabilistic as follows.}
-
 It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
-To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} 
-and propagated the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
-Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
-% \aim{don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space}
+To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
+%Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
+%\aim{Don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space.}
 
@@
 
-We consider a set of tomographically binned \nz\ and cosmological parameter covariance matrices used for \desc\ forecasting, for which the true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
-First, we bin them down to a piecewise constant parameterization with a manageable $35$ parameters for \chippr's sampling capabilities.
+\dwh{We consider as ground truth a set of known \nz\ corresponding to each of four hypothetical samples of galaxies and the corresponding cosmological parameter covariance matrix.
+The \nz\ of each galaxy subsample emulates that anticipated of galaxies binned by a redshift point estimate, as is common in tomographic redshift analyses, though our experimental procedure is agnostic to how the samples are identified.
+The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.}
+The true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+
+First, we bin them down to a piecewise constant parameterization with a manageable $35$ hyperparameters for \chippr's sampling capabilities.
 Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
@@
 \end{figure*}
- \aim{TODO: Ask Elisabeth how to get the bias here}
-
-We then make a point estimate of \nz\ using \Chippr\ as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
+
+\aim{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
+
+We then make a point estimate of \nz\ using \chippr's \mmle\ optimization option as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
 Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
@@
 Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
+\aim{TODO: Ask Elisabeth how to get the bias here}
 
@@
 		\caption{
+			\aim{Are the contours any easier to see now?}
 			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
 			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
-			\aim{Are the contours any easier to see now?}
 		}
@@
 
-\aim{Hogg says ``I think it would be good to talk a little quantitatively about where people need to know N(z) and other one-point statistics, and how much they will get various things wrong if they don't know these correctly.
-	And situate that discussion within the current context of cosmological parameter estimation and precision cosmology.''
-	Did I successfully answer the question?}
+\aim{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
+	Suggestions for how to better establish context would be appreciated.}
 
@@
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}
 		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
 		\caption{
-			Left: Examples of \pzpdf s with systematics at the level of the \lsst\ requirements with a machine learning-like implicit prior, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-			Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and the uniform implicit prior given to \chippr\ (gray curve).
+			\aim{TODO: same size axis labels on left and right, watermark with test conditions and ``mock data'' vs ``results of inference''}
+			Top: Examples of \pzpdf s with systematics at the level of the \lsst\ requirements with a machine learning-like implicit prior, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+			Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and the uniform implicit prior given to \chippr\ (gray curve).
 			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
@@
 \label{sec:con}
-
-\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
-	 if testable how so? outside the scope of paper, data producers be warned!
-	 a likelihood is better -- give us that if you can!
-	focus on methods that acknowledge probabilistic structure of problem}
-
-\aim{claim: implications for tomographic binning are severe}
 
@@
 %Tests were also performed on subsets of BOSS DR10 data with results consistent with those of simulations.
+
+\Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and bias.
+\Chippr\ is also robust to nontrivial implicit priors corresponding to the specifics of the architecture of the method by which \pzpdf s are obtained.
+However, \Chippr\ absolutely requires that the implicit prior is accurately known; using an implicit prior other than that which contributed to the production of the \pzpdf\ catalog results in catastrophically incorrect inference of \nz.
+It is therefore imperative that those developing codes to obtain \pzpdf s provide a way to isolate the implicit prior and that those publishing \pzpdf\ catalogs provide the implicit prior to users.
+
+This mandate is easier said than done, both for template fitting and machine learning approaches.
+While the implicit prior is often an explicit input to model-based routines, it may be defined in a space of redshift and SED templates.
+In this case, it may not be possible to apply \Chippr\ without marginalizing over additional variables $\psi$ for the SEDs.
+In other words, obtaining the implicit prior from a template fitting code may be challenging or even require consideration of higher-dimensional PDFs such as $\pr{z, \mathrm{SED} \gvn \psi^{*}}$.
+
+The situation appears more dire for data-driven techniques, whose training sets may not straightforwardly translate into an implicit prior.
+For example, some training set galaxies may contribute to the \pzpdf s more than others, resulting in different effective weights when factoring into, say, a histogram of training set redshifts as the implicit prior.
+Additionally, the weights may be stochastic, depending on the random seed used to initialize non-deterministic methods, precluding reproducibility.
+It is thus unclear whether the implicit prior can be meaningfully obtained from such methods at all.
+
+\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
+	if testable how so? outside the scope of paper, data producers be warned!
+	a likelihood is better -- give us that if you can!
+	focus on methods that acknowledge probabilistic structure of problem}
+
+\aim{claim: implications for tomographic binning are severe}
 
@@
 
-%%\section*{Appendices}
-%%
-%%\begin{appendices}
-%%
-%%%\section{Mathematical derivation}
-%%%\label{app:math}
-%%%
-%%%\section{Mock data generation}
-%%%\label{app:data}
-%%%
-%%%\textul{Plots: flow chart of forward model}
-%%%
-%%%\begin{figure*}
-%%%	\begin{center}
-%%%		\includegraphics[width=0.75\textwidth]{fig/flowchart.png}
-%%%		\caption{In appendix?}
-%%%		\label{fig:flowchart}
-%%%	\end{center}
-%%%\end{figure*}
-%%
-%%
-%%\end{appendices}
+\appendix
+%\renewcommand{\thesection}{\Alph{section}}
+%\renewcommand{\thesubsection}{\Alph{subsection}}
+
+\section{Derivation}
+\label{app:math}
+
+%We begin by parametrizing $N(z)$ in terms of $\vec{\theta}$, comprising some set of hyperparameters that define the form $N(z)$ may take in whatever basis we choose.  
+%We define a function $f_{\vec{\theta}}(z)=N(z)$ that transforms these hyperparameters into the redshift distribution function $N(z)$.  
+%Because 
+%\begin{equation}
+%\eqlabel{eq:definition}
+%N(z) \propto p(z \gvn \vec{\theta}),
+%\end{equation}
+%we may discontinue discussion of $N(z)$ in favor of the likelihood $p(z|\vec{\theta})$.
+
+In this paper, we work exclusively with log-probabilities.  
+What we wish to estimate is then the full log-posterior probability distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ given the catalog of photometry $\{\data_{j}\}$.
+
+By Bayes' Rule, the full log-posterior
+\begin{equation}
+\label{eqn:basicbayes}
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] = \ln[\pr{\{\data_{j}\} \gvn \ndphi}] + \ln[\pr{\ndphi}] - \ln[\pr{\{\data_{j}\}}]
+\end{equation}
+may be expressed in terms of the full log-likelihood probability distribution (hereafter the full log-likelihood) $\ln[\pr{\{\data_{j}\} \gvn \ndphi}]$ by way of a hyperprior log-probability distribution (hereafter the hyperprior) $\ln[\pr{\ndphi}]$ over the hyperparameters and the log-evidence probability of the data $\ln[\pr{\{\data_{j}\}}]$.
+However, the evidence is rarely known, so we probe the full log-posterior modulo an unknown constant of proportionality.
+
+The full log-likelihood may be expanded in terms of a marginalization over the redshifts as parameters, as in 
+\begin{equation}
+\label{eqn:marginalize}
+\ln[\pr{\{\data_{j}\} \gvn \ndphi}] = \ln\left[\integral{\pr{\{\data_{j}\} \gvn \{z_{j}\}} \pr{\{z_{j}\} \gvn \ndphi}}{\{z_{j}\}}\right].
+\end{equation}
+
+We shall make two assumptions of independence in order to make the problem tractable; their limitations are be discussed below.  
+First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn z_{j}}]$, as in 
+\begin{equation}
+\label{eqn:indiedat}
+\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn z_{j}}],
+\end{equation}
+a result of the definition of probabilistic independence encoded by the box in \Fig{fig:pgm}.
+Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent draws from the true $\pr{z \gvn \ndphi}$.  
+Additionally, $J$ itself is a Poisson random variable.  
+The combination of these assumptions is given by 
+\begin{equation}
+\label{eqn:indie}
+\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ \ln[\pr{z_{j} \gvn \ndphi}].
+\end{equation}
+%It is important to note that the integral $\integral{n(z)}{z} N(z)\ dz$ is not constrained to equal the variable defining the Poisson distribution but instead $J$ by \Eq{eq:definition}, which can be thought of as another parameter.
+The derivation differs when $J$ is not known, say, when we want to learn about a distribution in nature rather than a distribution specific to data in hand, but for a photometric galaxy catalog where the desired quantity is $n(z)$ for the galaxies entering a larger cosmology calculation, it is a fixed quantity.
+A detailed discussion of this matter may be found in \citet{foreman-mackey_exoplanet_2014}.  
+Applying Bayes' Rule, we may combine terms to obtain 
+\begin{align}
+\begin{split}
+\label{eqn:posterior}
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] - \integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn \ndphi}}{z}\right].
+\end{split}
+\end{align}
+
+%\Eq{eq:posterior} contains two quantities that merit further discussion, the prior distribution $p(\vec{\theta})$ discussed further in \Sect{sec:exp} and the photo-$z$ log-likelihoods $\ln[p(\vec{d}_{j}|z_{j})]$ that have not been mentioned since \Eq{eq:marginalize}.
+%Though photo-$z$ log-likelihoods would be desirable for use in these equations, they are not generally the product of either empirical and data-driven methods for obtaining photo-$z$ probability distributions.
+%Though probabilistic photo-$z$s are typically reported as generic probability distributions $p(z_{j})$, the methods that produce them may be understood to always yield posteriors, probability distributions conditioned on the data we believe to be true.
+%If they were not based in this assumption, they would require a sum over an infinite space of possible datasets.
+
+Since we only have access to implicit \pz\ posteriors, we must be able to write the full log-posterior in terms of implicit \pz\ log-posteriors rather than the log-likelihoods of \Eq{eqn:posterior}.
+To do so, we will need an explicit statement of this implicit prior $\ndphi^{*}$ for whatever method is chosen to produce the implicit \pz\ posteriors.
+
+To perform the necessary transformation from likelihoods to posteriors, we follow the reasoning of \citet{foreman-mackey_exoplanet_2014}.  
+Let us consider the probability of the parameters conditioned on the data and an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} as 
+\begin{align}
+\label{eqn:trick}
+\begin{split}
+\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
+\end{split}
+\end{align}
+
+Once the implicit prior $\ndphi^{*}$ is explicitly introduced, we may expand the last term in \Eq{eqn:trick} according to Bayes' Rule to get 
+\begin{align}
+\begin{split}
+\label{eqn:expand}
+\ln[\pr{\data_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
+\end{split}
+\end{align}
+Because there is no direct dependence of the data upon the hyperparameters, we may again expand the term $\ln[\pr{\data_{j} \gvn z, \ndphi^{*}}]$ to obtain 
+\begin{align}
+\begin{split}
+\label{eqn:indterm}
+\ln[\pr{\vec{d}_{j} \gvn z}] = & \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{z \gvn \ndphi^{*}}]- \ln[\pr{\data_{j} \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] .
+\end{split}
+\end{align}
+Canceling the undesirable terms for the inaccessible likelihood $\ln[\pr{\data_{j} \gvn z}]$ and trivial $\ln[\pr{\data_{j} \gvn \ndphi^{*}}]$ yields
+\begin{equation}
+\label{eqn:cancel}
+\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]  - \ln[\pr{z \gvn \ndphi^{*}}].
+\end{equation}
+We put this all together to get the full log-posterior probability distribution of 
+\begin{align}
+\begin{split}
+\label{eqn:final}
+\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto & \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}] \right)\right]}{z}\right] ,
+\end{split}
+\end{align}
+
+The argument of the integral in the log-posterior of \Eq{eqn:final} depends solely on knowable quantities (and those we must explicitly assume) and can be calculated for a given sample of \pz\ log-posteriors $\{\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with which they were obtained, noting the relation of
+\begin{equation}
+\label{eqn:params}
+\pr{z \gvn \ndphi} = \frac{f(z; \ndphi)}{\integral{f(z; \ndphi)}{z}}.
+\end{equation}
+Since we cannot know constant of proportionality, we sample the desired full log-posterior $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$ using Monte Carlo-Markov chain (MCMC) methods.  
+The method outlined here is valid regardless of how the implicit \pz\ log-posteriors are obtained so the many approaches to producing \pzpdf s will not be rehashed; though the matter is outside the scope of this paper, reviews of various methods have been presented in the literature \citep{sheldon_photometric_2012, ball_robust_2008, carrasco_kind_tpz:_2013, carrasco_kind_exhausting_2014}, and will be briefly reviewed in Schmidt, Malz \& Soo, et al.
+(in prep).
+%
+%\begin{align}
+%\begin{split}
+%\eqlabel{eqn:fullpost}
+%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
+%\end{split}
+%\end{align}
+
+\section{Convergence Criteria}
+\label{app:acorr}
+
+In addition to qualitative visual inspection of the chains, two quantities that probe the convergence of the sampler are used in this study, the autocorrelation time and the Gelman-Rubin convergence criterion.
+%\Fig{fig:chains} shows the %evolution of the values of one parameter of one walker over the course of all %iterations of the sampler.
+
+%\begin{figure}
+%%\includegraphics[width=0.5\textwidth]{figs/null/chain0.pdf}
+%\caption{This figure shows the evolution of one walker's parameter values for 
+%one element of the parameter vector $\vec{\theta}$ as a function of iteration 
+%number, demonstrating the completion of the burn-in phase.}
+%\label{fig:chains}
+%\end{figure}
+
+The autocorrelation time is effectively a measure of the efficiency of the method and can be described as the expected number of iterations necessary to accept a new sample independent of the current accepted sample.
+A sampler that converges faster will have a smaller autocorrelation time, and smaller autocorrelation times are preferable because it means fewer iterations are wasted on non-independent samples when independent samples are desired.
+See \citet{foreman-mackey_emcee_2013} for a more complete exploration of the autocorrelation time.  
+In all tests discussed here, autocorrelation times across walkers and parameters were approximately 20, meaning two samples 20 or more iterations apart were independent, a satisfactory level of efficiency.
+Low autocorrelation times are a necessary but not always sufficient convergence condition, as the autocorrelation times calculated for tests in this paper were constant across all sub-runs, even those that were obviously burning in.
+
+The Gelman-Rubin statistic
+\begin{equation}
+\label{eqn:gr}
+R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} b_{k}}{w_{k}}},
+\end{equation}
+a weighted sum of the mean $w_{k}$ of the variances within individual walkers' chains and the variance $b_{k}$ between chains of different walkers $m$, is calculated over each sub-run $i$ to determine the duration of the burn-in period.
+Convergence is achieved when the statistic approaches unity.
 

### 2019-06-22 research/paper/draft.tex
finishing up revised conclusion

@@
 
-\Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and bias.
+\dwh{\Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and bias.
 \Chippr\ is also robust to nontrivial implicit priors corresponding to the specifics of the architecture of the method by which \pzpdf s are obtained.
@@
 
-\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
-	if testable how so? outside the scope of paper, data producers be warned!
-	a likelihood is better -- give us that if you can!
-	focus on methods that acknowledge probabilistic structure of problem}
-
-\aim{claim: implications for tomographic binning are severe}
+%\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
+%	if testable how so? outside the scope of paper, data producers be warned!
+%	a likelihood is better -- give us that if you can!
+%	focus on methods that acknowledge probabilistic structure of problem}
+
+A thorough investigation of the degree to which the implicit prior can be meaningfully obtained is outside this paper but should be a priority for all consumers of \pzpdf s.
+As an alternative, however, we must point out that if likelihoods were available rather than posteriors, the trouble with the implicit prior would be avoided altogether.
+We thus encourage the community of those making \pzpdf s to consider developing such methods so that the resulting data products may be correctly used in scientific inference more generically.}
+
+\aim{TODO: claim that implications for tomographic binning are severe, if that can be motivated by Section~\ref{sec:lsstdemo}.}
 
@@
 \label{app:math}
+
+\aim{TODO: make notation consistent with the rest of the paper.}
 

### 2019-08-16 research/paper/draft.tex
filling in holes in intro

@@
 \newcommand{\dwh}[1]{\todo{Attn. Hogg:}{blue}{#1}}
+
 \newcommand{\myemail}{aimalz@nyu.edu}
 \newcommand{\mathul}[1]{\underline{#1}}
+
 \newcommand{\Sect}[1]{Section~\ref{#1}}
@@
 \newcommand{\Tab}[1]{Table~\ref{#1}}
+
 \newcommand{\project}[1]{{\textsc{#1}}}
@@
 \newcommand{\des}{\project{DES}}
-\newcommand{\Chippr}{\project{CHIPPR}}
-\newcommand{\repo}[1]{{\texttt{#1}}~}
+\newcommand{\Chippr}{\project{CHIPPR}}% maybe change this to algorithm/italics
+
+\newcommand{\repo}[1]{{\texttt{#1}}}
 \newcommand{\qp}{\repo{qp}}
@@
 \newcommand{\emcee}{\repo{emcee}}
-\newcommand{\github}{\href{https://github.com}{GitHub}}
-\newcommand{\python}{\textit{Python}}
-\newcommand{\data}{\ensuremath{\vec{d}}}
-% could change to bold
+\newcommand{\github}{\href{https://github.com}{GitHub}}% maybe remove link and use code font
+\newcommand{\python}{\textit{Python}}% establish code font?
+
+\newcommand{\data}{\ensuremath{\vec{d}}}% could change to bold
 \newcommand{\like}{\mathscr{L}}
-\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
-% could change to Prob or Pr 
+\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}% could change to Prob or Pr 
 \newcommand{\expect}[1]{\left<#1\right>}
@@
 \newcommand{\Pz}{Photo-$z$}
-\newcommand{\pzpdf}{\pz\ PDF}
-% could change to posterior
-\newcommand{\Pzpdf}{\Pz\ PDF}
-% could change to posterior
-\newcommand{\zpdf}{redshift posterior}
-% could change to implicit posterior
+\newcommand{\pzpdf}{\pz\ PDF}% could change to posterior
+\newcommand{\Pzpdf}{\Pz\ PDF}% could change to posterior
+\newcommand{\zpdf}{redshift posterior}% could change to implicit posterior
 \newcommand{\pzip}{\pz implicit posterior}
@@
 \newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
-\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
-% could change to \vec
+\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}% could change to \vec
 \newcommand{\ndphi}{\bvec{\phi}}
-\newcommand{\mmle}{marginalized maximum likelihood estimate}
-% really marginalized maximum a posteriori estimate, may still change this
+\newcommand{\mmle}{marginalized maximum likelihood estimate}% really marginalized maximum a posteriori estimate, may still change this
 
@@
 
-\author{Alex I. Malz\altaffilmark{1}}
-\author{David W. Hogg\altaffilmark{1,2,3,4}}
+\author{Alex I. Malz\altaffilmark{1,2}}
+\author{David W. Hogg\altaffilmark{2,3,4,5}}
 \email{aimalz@nyu.edu}
 
-\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
-\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
-\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
-\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{1}{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
+\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
+\altaffiltext{3}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
+\altaffiltext{4}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
+\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 
@@
 
-Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are not insignificant contributors to the total uncertainties reported on cosmological parameters.
+Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are a dominant contributor to the total uncertainties reported on cosmological parameters \citep{abruzzo_impact_2019}.
 As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
@@
 The most concerning weakness of \pzpdf s, however, is their usage in the literature, which is at best inconsistent and at worst incorrect.
-The most prevalent application of \pzpdf s is their use in estimating the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}, a quantity essential to the calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, leistedt_hierarchical_2016}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
+
+If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
+When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
+Though it is most common for $\hat{z}_{j}$ to be the maximum or \textit{mode} of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.
+However, any procedure that reduces \pzpdf s to point estimates discards valuable information about the uncertainty on redshift.
+
+Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
+They have been used to form selection criteria of samples from galaxy surveys without propagation through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.
+Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.
+
+The most prevalent application of \pzpdf s that preserves their information content is the estimation of the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}.
+The redshift distribution is a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
+Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the accuracy and precision to which \nz\ must be known limit the degree to which constraints on the cosmological parameters can advance \citep{abruzzo_impact_2019}.
 
@@
 
-If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
-When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
-Though it is most common for $\hat{z}_{j}$ to be the maximum or \textit{mode} of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.
-
-However, any procedure that reduces \pzpdf s to point estimates loses information about the uncertainty on redshift.
-To circumvent this loss, it has grown to be more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
+Though it is traditional to estimate \nz\ from \pz\ point estimates \citep{abruzzo_impact_2019}, there is a great deal of literature employing \pzpdf s.
+To circumvent this loss, it has become to be more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
 \begin{align}
@@
 
-%Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
-%They have been used to form selection criteria of samples from galaxy surveys without propagation 
-%through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
-%Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.
-
 Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
@@
 We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
-Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.
+Though others have approached the problem before \citep{leistedt_hierarchical_2016, leistedt_hierarchical_2018}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.
 
@@
 Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
-% \aim{referenceh chapter 3 for complication of algorithm}
+% \aim{reference chapter 3 for complication of algorithm}
 Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
@@
 %The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
-%\aim{Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
+\aim{TODO: Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
 %The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
@@
 
-%\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
-%	if testable how so? outside the scope of paper, data producers be warned!
-%	a likelihood is better -- give us that if you can!
-%	focus on methods that acknowledge probabilistic structure of problem}
+\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
+	if testable how so? outside the scope of paper, data producers be warned!
+	a likelihood is better -- give us that if you can!
+	focus on methods that acknowledge probabilistic structure of problem}
 
@@
 \bibliographystyle{apj}
-\bibliography{thesis}
+\bibliography{draft}
 

### 2019-08-19 research/paper/draft.tex
going through comments, trying to fix Travis compilation issue

@@
 
-\usepackage{soul}
+%\usepackage{soul}
 \usepackage[title]{appendix}
@@
 			The data (black points) used to generate the contours were extracted from \citet{jain_whole_2015} using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019}, with the ideal redshift estimation provided for reference (red diagonal).
-			\aim{Choose better colors?}
+			\aim{TODO: Use the colors from the rest of the paper.}
 		}
@@
 Because the parameters $\ndphi$ that we seek are causally separated from the data by the latent variable of redshift, we call them \textit{hyperparameters}.
-\aim{Am I adequately explaining the distinction between hyperparameters and parameters?}
+\dwh{Am I adequately explaining the distinction between hyperparameters and parameters?}
 
@@
 			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.
+			\aim{How to make this more attractive?}
 			}
@@
 As expected based on consistency with the forward model, \Chippr\ is completely resistant to bias, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
-\aim{I'm not sure I effectively explained why \Chippr\ is expected to be unaffected by bias of this form.}
+\dwh{Did I effectively explain why \Chippr\ is expected to be unaffected by bias of this form?}
 
@@
 Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
-\aim{TODO: Ask Elisabeth how to get the bias here}
+\aim{TODO: Ask Elisabeth how to get the bias here.}
+\dwh{So a Fisher matrix analysis inherently can't provide the bias.
+Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
+Do you think it's sufficient to provide the bias on the moments of \nz, since that's what everyone uses anyway?}
 

### 2019-08-19 research/paper/draft.tex
adding comments, removing deprecated package

@@
 \usepackage{amsmath}
-\usepackage{graphicx}
+%\usepackage{graphicx}
 %\usepackage{lineno}
@@
 
+% what are photo-zs?
 Photometric redshift (\pz) estimation has been a staple of studies of galaxy evolution, large-scale structure, and cosmology since its conception half a century ago \citep{baum_photoelectric_1962}.  
@@
 
+% why do we need photo-zs?
 Calculations of correlation functions of cosmic shears and galaxy positions that constrain the cosmological parameters require large numbers of high-confidence redshifts of surveyed galaxies.  
@@
 
+% what's wrong with photo-zs?
 However, \pz s are susceptible to inaccuracy and imprecision via a number of effects, particularly their inherent noisiness due to the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
@@
 
+% how do we interpret photo-z imperfections?
 There are several varieties of generally non-Gaussian deviation from a perfect correspondence between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
@@
 
+% how much do these imperfections matter?
 Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are a dominant contributor to the total uncertainties reported on cosmological parameters \citep{abruzzo_impact_2019}.
@@
 
+% how can we improve photo-zs?
 Much effort has been dedicated to improving \pz s, though they are still most commonly obtained by a maximum likelihood estimator (MLE) based on libraries of galaxy SED templates, with conservative approaches to error estimation.
@@
 
+% PDFs are a better way to improve photo-zs
 An alternative to point estimation of \pz s is redshift probability distribution function (PDF) estimation that reports the probability of all possible redshifts for every galaxy, commonly written as $\pr{z}$, rather than an MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.
@@
 
+% photo-z PDFs are established
 \Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.
@@
 
+% photo-z PDFs are most often reduced to point estimates
+Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent, most often by reduction to familiar point estimates.
 If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
@@
 Though it is most common for $\hat{z}_{j}$ to be the maximum or \textit{mode} of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.
-However, any procedure that reduces \pzpdf s to point estimates discards valuable information about the uncertainty on redshift.
-
-Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
-They have been used to form selection criteria of samples from galaxy surveys without propagation through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.
+
+% photo-z PDFs may also be used to define cuts
+Regardless of how it is done, any procedure that reduces \pzpdf s to point estimates discards valuable information about the uncertainty on redshift.
+\Pzpdf s have also been used to form selection criteria of samples from galaxy surveys without propagation through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.
 Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.
 
+% n(z) as an application of photo-z PDFs
 The most prevalent application of \pzpdf s that preserves their information content is the estimation of the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}.
+
+% Provide equations/citations for the use of \Nz\ in cosmology.
 The redshift distribution is a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
-Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the accuracy and precision to which \nz\ must be known limit the degree to which constraints on the cosmological parameters can advance \citep{abruzzo_impact_2019}.
+
+% Motivate how well we need to know \Nz\ and why it matters.
+% New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
+% Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.
+Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the degree to which constraints on the cosmological parameters can advance is limited by the accuracy and precision to which \nz\ is known \citep{abruzzo_impact_2019}.
 
@@
 
+% how people get n(z) from photo-z PDFs
 Though it is traditional to estimate \nz\ from \pz\ point estimates \citep{abruzzo_impact_2019}, there is a great deal of literature employing \pzpdf s.
@@
 
+% what this paper is about
 Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  

### 2019-08-20 research/paper/draft.tex
trying to fix Travis

@@
 
-% n(z) as an application of photo-z PDFs
+\aim{TODO: Provide equations/citations for the use of \Nz\ in cosmology, why it matters.}
 The most prevalent application of \pzpdf s that preserves their information content is the estimation of the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}.
-
-% Provide equations/citations for the use of \Nz\ in cosmology.
-The redshift distribution is a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
-
-% Motivate how well we need to know \Nz\ and why it matters.
-% New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
+\nz\ is used to calculate the redshift calibration bias $b_{z}$ between the true and observed critical surface densities in galaxy-galaxy lensing \citep{mandelbaum_precision_2008}, the geometric lens efficiency $g_{k}(\chi)$ in tomographic weak lensing by large-scale structure \citep{benjamin_cfhtlens_2013}, and \aim{\dots}
+
+The redshift distribution is a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}.
+\Nz\ may also be used to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
+
+% New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys, motivate how well we need to know \Nz. 
+Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the degree to which constraints on the cosmological parameters can advance is limited by the accuracy and precision to which \nz\ is known \citep{abruzzo_impact_2019}.
 % Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.
-Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the degree to which constraints on the cosmological parameters can advance is limited by the accuracy and precision to which \nz\ is known \citep{abruzzo_impact_2019}.
-
-\aim{TODO:
-	Motivate how well we need to know \Nz\ and why it matters.
-	Provide equations/citations for the use of \Nz\ in cosmology.
-	New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
-	Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}
 
@@
 for a sample of $J$ galaxies $j$, or, equivalently, the redshift distribution function $\hat{N}(z) = J \hat{n}(z)$, by effectively averaging the \pzpdf s.
+This summation procedure has been used extensively \citep{mandelbaum_precision_2008, benjamin_cfhtlens_2013, kelly_weighing_2014}\aim{\dots}
 

### 2019-10-15 research/paper/draft.tex
changes to commentary

@@
 \usepackage{amsmath}
+\usepackage{hyperref}
 %\usepackage{graphicx}
@@
 \newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
-\newcommand{\dwh}[1]{\todo{Attn. Hogg:}{blue}{#1}}
+\newcommand{\dwh}[1]{\todo{Attn. Hogg}{blue}{#1}}
 
@@
 
-\newcommand{\project}[1]{{\textsc{#1}}}
+\newcommand{\project}[1]{\textsc{#1}}
 \newcommand{\lsst}{\project{LSST}}
@@
 
-\newcommand{\repo}[1]{{\texttt{#1}}}
+\newcommand{\repo}[1]{\texttt{#1}}
 \newcommand{\qp}{\repo{qp}}
@@
 \newcommand{\zpdf}{redshift posterior}% could change to implicit posterior
-\newcommand{\pzip}{\pz implicit posterior}
+\newcommand{\pzip}{\pz\ implicit posterior}
 \newcommand{\nz}{$n(z)$}
@@
 \section{Introduction}
-\label{sec:introduction}
+\label{sec:intro}
 
@@
 			The data (black points) used to generate the contours were extracted from \citet{jain_whole_2015} using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019}, with the ideal redshift estimation provided for reference (red diagonal).
-			\aim{TODO: Use the colors from the rest of the paper.}
+			\aim{TODO: Use the colors from the rest of the paper.  
+			Fix backwards quotes on y axis label.}
 		}
@@
 	\begin{center}
-		\caption{\Pz\ requirements for \lsst\ cosmology \citep{mandelbaum_weak_2017}.}
+		\caption{\Pz\ requirements for \lsst\ cosmology\\
+				\citep{mandelbaum_weak_2017}.}
 		\begin{tabular}{ll}
@@
 \aim{TODO: Provide equations/citations for the use of \Nz\ in cosmology, why it matters.}
+
 The most prevalent application of \pzpdf s that preserves their information content is the estimation of the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}.
-\nz\ is used to calculate the redshift calibration bias $b_{z}$ between the true and observed critical surface densities in galaxy-galaxy lensing \citep{mandelbaum_precision_2008}, the geometric lens efficiency $g_{k}(\chi)$ in tomographic weak lensing by large-scale structure \citep{benjamin_cfhtlens_2013}, and \aim{\dots}
-
+\nz\ is used to calculate the redshift calibration bias $b_{z}$ between the true and observed critical surface densities in galaxy-galaxy lensing \citep{mandelbaum_precision_2008} and the geometric lens efficiency $g_{k}(\chi)$ in tomographic weak lensing by large-scale structure \citep{benjamin_cfhtlens_2013}.
 The redshift distribution is a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}.
@@
 
-% New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys, motivate how well we need to know \Nz. 
+\aim{TODO: Say what precision is needed for \Nz\ for future weak lensing surveys, motivate how well we need to know \Nz.}
 Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the degree to which constraints on the cosmological parameters can advance is limited by the accuracy and precision to which \nz\ is known \citep{abruzzo_impact_2019}.
@@
 	\begin{center}
-		\includegraphics[width=0.15\textwidth]{figures/chippr/pgm.png}
+		\includegraphics[height=0.25\textheight]{figures/chippr/pgm.png}
 		\caption{The directed acyclic graph of the CHIPPR model, where circles indicate random variables and arrows indicate causal relationships.
@@
 			The redshifts $\{z_{j}\}$ of all galaxies in the survey are latent variables independently drawn from the redshift distribution, which is a function of $\ndphi$. 
-			The photometric data $\data_{j}$ for each galaxy is drawn from a function of its redshift $z_{j}$ and observed, indicated by a shaded circle.}
+			The photometric data $\data_{j}$ for each galaxy is drawn from a function of its redshift $z_{j}$ and observed, indicated by a shaded circle.
+			\aim{Switch to vector graphics version?}}
 		\label{fig:pgm}
@@
 		\caption{
+			\dwh{How to make this more attractive?}
 			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
@@
 			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.
-			\aim{How to make this more attractive?}
+			\aim{TODO: Remove extra ``j'' in ``Convolve'' step.
+			Also, make a high-res or vector graphics version.}
 			}
@@
 		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
-		\caption{\aim{TODO: Add watermark of "mock data" in UL corner, same for "results of inference" on other kind of plot.}
-			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
+		\caption{The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
 			The main panel of each shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal (thin red line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
 			The insets show marginal histograms (gray) in each dimension, that can be compared with the true \nz\ used to make the figure (blue curve) to see the effect of the isolated systematic, as well as the implicit prior (red line).
+			\aim{TODO: Enlarge axis labels and label panels.
+			Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
 		}
@@
 %\end{figure}
-\aim{TODO: Add back the figure of prior samples?}
+\dwh{Should I add back the figure of prior samples?}
 
@@
 		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
+		\aim{TODO: Enlarge axis labels and label panels.
+		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
 	}
@@
 		\Chippr\ is robust to intrinsic scatter, but the alternatives suffer from overly broad \nz\ estimates that worsen with increasing intrinsic scatter.
+		\aim{TODO: Label panels.
+		Add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
 	}
@@
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_mega_scatter.png}\\
 	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_log_estimators.png}
@@
 		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
+		\aim{TODO: Enlarge axis labels on top and label panels.\\
+		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.}
 	}
@@
 		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+		\aim{TODO: Enlarge axis labels and label panels.
+		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}		
 	}
@@
 		Though the histogram of modes is most sensitive to a catastrophic outlier population, the stacked estimator also overestimates \nz\ under (machine learning-like outliers) and beyond (template fitting-like outliers).
+		\aim{TODO: Label panels.
+		Add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
 	}
@@
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_mega_scatter.png}\\
 	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_log_estimators.png}
 	\caption{
-		\aim{TODO: Add watermark of "mock data" in UL corner, same for "results of inference" on other kind of plot.}
 		Left: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
@@
 		The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
+		\aim{TODO: Enlarge axis labels on top and label panels.
+		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.}
 	}
@@
 		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
+		\aim{TODO: Enlarge axis labels.
+		Label panels.
+		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
 	}
@@
 		\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
+		\aim{TODO: Label panels.
+		Add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
 	}
@@
 
-\aim{Move \Sect{sec:violations} here?}
+\dwh{Move \Sect{sec:violations} here?}
 
@@
 	\begin{center}
-		\includegraphics[width=0.49\textwidth]{figures/chippr/cosmolike_inputs.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/cosmolike_inputs.png}
 		\caption{
 			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
+			\aim{TODO: Give figure a title.}
 		}
@@
 		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
-		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend, and make the axis labels bigger.}
+		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend.
+		Enlarge axis labels.
+		Label panels.
+		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
 		}
@@
 
-\aim{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
+\dwh{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
 
@@
 			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
-			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend, and make the legend text bigger.}
+			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend.
+			Label figure/panels.
+			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
 		}
@@
 Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
-\aim{TODO: Ask Elisabeth how to get the bias here.}
-\dwh{So a Fisher matrix analysis inherently can't provide the bias.
+%\aim{TODO: Ask Elisabeth how to get the bias here.}
+\dwh{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
 Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
@@
 	\begin{center}
-		\includegraphics[width=0.94\textwidth]{figures/chippr/final_plot.png}
-		\caption{
-			\aim{Are the contours any easier to see now?}
+		\includegraphics[width=0.9\textwidth]{figures/chippr/final_plot.png}
+		\caption{
+			\dwh{Are the contours any easier to see now?}
 			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
 			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
+			\aim{TODO: Fix formatting of axis labels.
+			Standardize ticklabels.}
 		}
@@
 
-\aim{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
+\dwh{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
 	Suggestions for how to better establish context would be appreciated.}
@@
 
-\aim{Move \Sect{sec:interim} here?}
+\dwh{Move \Sect{sec:interim} here?}
 
@@
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
 		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
 		\caption{
-			\aim{TODO: same size axis labels on left and right, watermark with test conditions and ``mock data'' vs ``results of inference''}
 			Top: Examples of \pzpdf s with systematics at the level of the \lsst\ requirements with a machine learning-like implicit prior, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
@@
 			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
+			\aim{TODO: Enlarge axis labels on top.
+			Label panels/figure.
+			Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.}
 		}
@@
 %\renewcommand{\thesubsection}{\Alph{subsection}}
+
+\aim{TODO: Figure out how to get the appendices enumerated for reference earlier.}
 
@@
 \begin{acknowledgements}
-AIM was supported by National Science Foundation grant AST-1517237 and the U.S.
+AIM acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the Federal Ministry of Education and Research.
+During the completion of this work, AIM was supported by National Science Foundation grant AST-1517237 and the U.S.
 Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) program, administered by the Oak Ridge Institute for Science and Education for the DOE under contract number DE‐SC0014664.

### 2019-10-15 research/paper/draft.tex
comment colors

@@
 \newcommand{\dwh}[1]{\todo{Attn. Hogg}{blue}{#1}}
+\newcommand{\que}[1]{\todo{Question}{cyan}{#1}}
 
@@
 Because the parameters $\ndphi$ that we seek are causally separated from the data by the latent variable of redshift, we call them \textit{hyperparameters}.
-\dwh{Am I adequately explaining the distinction between hyperparameters and parameters?}
+\que{Am I adequately explaining the distinction between hyperparameters and parameters?}
 
@@
 		\caption{
-			\dwh{How to make this more attractive?}
+			\que{How to make this more attractive?}
 			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
@@
 %\end{figure}
-\dwh{Should I add back the figure of prior samples?}
+\que{Should I add back the figure of prior samples?}
 
@@
 As expected based on consistency with the forward model, \Chippr\ is completely resistant to bias, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
-\dwh{Did I effectively explain why \Chippr\ is expected to be unaffected by bias of this form?}
+\que{Did I effectively explain why \Chippr\ is expected to be unaffected by bias of this form?}
 
@@
 
-\dwh{Move \Sect{sec:violations} here?}
+\que{Move \Sect{sec:violations} here?}
 
@@
 
-\dwh{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
+\que{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
 
@@
 %\aim{TODO: Ask Elisabeth how to get the bias here.}
-\dwh{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
+\que{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
 Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
@@
 		\caption{
-			\dwh{Are the contours any easier to see now?}
+			\que{Are the contours any easier to see now?}
 			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
@@
 
-\dwh{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
+\que{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
 	Suggestions for how to better establish context would be appreciated.}
@@
 
-\dwh{Move \Sect{sec:interim} here?}
+\que{Move \Sect{sec:interim} here?}
 

### 2019-11-14 research/paper/draft.tex
updating figures
fixed compilation bug

@@
 			The data (black points) used to generate the contours were extracted from \citet{jain_whole_2015} using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019}, with the ideal redshift estimation provided for reference (red diagonal).
-			\aim{TODO: Use the colors from the rest of the paper.  
-			Fix backwards quotes on y axis label.}
 		}
@@
 	\begin{center}
-		\includegraphics[width=0.9\textwidth]{figures/chippr/flowchart.png}
+		\includegraphics[width=0.9\textwidth]{figures/chippr/flowchart.pdf}
 		\caption{
@@
 			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.
-			\aim{TODO: Remove extra ``j'' in ``Convolve'' step.
-			Also, make a high-res or vector graphics version.}
 			}
@@
 			The main panel of each shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal (thin red line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
-			The insets show marginal histograms (gray) in each dimension, that can be compared with the true \nz\ used to make the figure (blue curve) to see the effect of the isolated systematic, as well as the implicit prior (red line).
-			\aim{TODO: Enlarge axis labels and label panels.
-			Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
+			The insets show marginal histograms (gray) in each dimension that can be compared with the true \nz\ (blue curve) used to make the figure to see the effect of the isolated systematic, as well as the implicit prior (red line).
+			\aim{TODO: Omit this figure; redundant with right panel of \Fig{fig:pzs-scatter} and top panels of \Fig{fig:uniform-outliers} and \Fig{fig:bias}.}
+%%			Enlarge axis labels and label panels.}
+%%			Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
 		}
@@
 			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
-			\aim{TODO: Give figure a title.}
 		}

### 2019-12-02 research/paper/draft.tex
tweaking wording

@@
 \newcommand{\Pz}{Photo-$z$}
-\newcommand{\pzpdf}{\pz\ PDF}% could change to posterior
-\newcommand{\Pzpdf}{\Pz\ PDF}% could change to posterior
-\newcommand{\zpdf}{redshift posterior}% could change to implicit posterior
+\newcommand{\zpdf}{\pz\ PDF}% could change to posterior
+\newcommand{\Zpdf}{\Pz\ PDF}% could change to posterior
+\newcommand{\pzpdf}{\pz\ posterior PDF}% could change to implicit posterior
+\newcommand{\Pzpdf}{\Pz\ posterior PDF}% could change to implicit posterior
 \newcommand{\pzip}{\pz\ implicit posterior}
@@
 A trustworthy estimate of the redshift distribution \nz\ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
-Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
+Spectroscopic redshifts for the dim and numerous galaxies of next-generation weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
 The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ posterior probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
@@
 The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
-We introduce a statistically rigorous probabilistic graphical model (PGM) of $n(z)$ for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
-Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures.
-Using the publicly available \texttt{chippr} prototype code, we propagate these effects to constraints in the space of cosmological parameters, noting, however, that the mathematically justifiable approach incurs computational expense.
-The CHIPPR approach is applicable to any one-point statistic of any random variable, provided the prior probability distribution used to produce the posteriors is known.
+We introduce a statistically rigorous probabilistic graphical model of \dwh{redshift-dependent photometry} for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
+Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\textsc{CHIPPR}) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures.
+We present the \texttt{chippr} prototype code and use it to propagate these effects to constraints in the space of cosmological parameters, noting, however, that the mathematically justifiable approach incurs computational expense.
+The \textsc{CHIPPR} approach is applicable to any one-point statistic of any random variable, provided the prior probability density used to produce the posteriors is explicitly known; if the prior is implicit, as may be the case for popular photo-$z$ techniques, then the resulting posterior PDFs cannot be used for scientific inference.
+We therefore recommend that the photo-$z$ community direct attention to the development of methodologies that enable the recovery of photo-$z$ likelihoods \dwh{with support over all redshifts}, either directly or via a known prior probability density.
 \end{abstract}
@@
 % what's wrong with photo-zs?
-However, \pz s are susceptible to inaccuracy and imprecision via a number of effects, particularly their inherent noisiness due to the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
+However, \pz s are susceptible to inaccuracy and imprecision \dwh{for a number of reasons}, particularly their inherent noisiness resulting from the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
 Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
@@
 % how do we interpret photo-z imperfections?
-There are several varieties of generally non-Gaussian deviation from a perfect correspondence between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
-There is scatter about the diagonal due to the coarseness of the photometric filters, with larger scatter perpendicular to the diagonal due to the specific wavelengths where highly identifiable spectral features pass between the filters, as well as higher scatter at high redshifts due to larger photometric errors on fainter galaxies.
+There are several varieties of generally non-Gaussian deviation from a \dwh{trivial relationship} between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
+The coarseness of the photometric filters causes scatter about the diagonal, with larger scatter perpendicular to the diagonal at redshifts where highly identifiable spectral features pass between the filters, as well as higher scatter at high redshifts where faint galaxies with large photometric errors are more abundant.
 There are populations of outliers, far from the diagonal, comprised of galaxies for which the redshift estimate is catastrophically distinct from the true redshift, showing that outliers are not uniformly distributed nor restricted to long tails away from a Gaussian scatter.
-And, though hardly perceptible in the plot, there is a systematic bias, wherein the average of the points would not lie on the diagonal but would be offset by a small bias, suggested by the trend of high redshift points to lie below the diagonal.
+And, though hardly perceptible in the plot, there is a systematic bias, wherein the average of the points would not lie on the diagonal but would be offset by a small bias, suggested by the trend of high-redshift points to lie below the diagonal.
 
@@
 Much effort has been dedicated to improving \pz s, though they are still most commonly obtained by a maximum likelihood estimator (MLE) based on libraries of galaxy SED templates, with conservative approaches to error estimation.
-This approach tends to lead to catastrophic outliers like the horizontally oriented population of \Fig{fig:pedagogical_scatter}, due to the presence of galaxies whose SEDs are not represented by the template library.
-Data-driven approaches tend to result in catastrophic outliers like the vertically oriented population of \Fig{fig:pedagogical_scatter}, due to training sets that are incomplete in redshift coverage.
+The presence of galaxies whose SEDs are not represented by the template library tends to lead to catastrophic outliers distributed like the horizontally oriented population of \Fig{fig:pedagogical_scatter}.
+For data-driven approaches, training sets that are incomplete in redshift coverage tend to result in catastrophic outliers like the vertically oriented population of \Fig{fig:pedagogical_scatter}.
 The approaches of using a training set versus a template library are related to one another by \citet{budavari_unified_2009}.
@@
 % PDFs are a better way to improve photo-zs
-An alternative to point estimation of \pz s is redshift probability distribution function (PDF) estimation that reports the probability of all possible redshifts for every galaxy, commonly written as $\pr{z}$, rather than an MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.
+\dwh{The probability density function (PDF) in redshift space for each galaxy, commonly written as $\pr{z}$, is an alternative to the MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.}
 This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating our understanding of precision, accuracy, and systematic error.
-However, denoting \pzpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; \pzpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
-In terms of \Fig{fig:pedagogical_scatter}, \pzpdf s are horizontal cuts, probabilities of redshift conditioned on a specific value of data, i.e.
-posteriors, which constrain redshifts, whereas vertical cuts through this space are probabilities of data conditioned on a specific redshift, i.e.
-likelihoods, from which data is actually drawn.
+However, denoting \zpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; \zpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
+In terms of \Fig{fig:pedagogical_scatter}, \zpdf s are horizontal cuts, probabilities of redshift conditioned on a specific value of data, i.e.
+posteriors $\pr{z \gvn \data}$, which constrain redshifts, whereas vertical cuts through this space are probabilities of data conditioned on a specific redshift, i.e.
+likelihoods $\pr{\data \gvn z}$, from which photometric data is actually drawn.
 
@@
 Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent, most often by reduction to familiar point estimates.
-If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
+If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then their \dwh{redshift PDFs are well-approximated by delta functions} $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift\dwh{\footnote{\dwh{Note that \sz s are not the same as known true redshifts; the PDFs of \sz s would be narrow and almost always unimodal, but they would not be delta functions due to observational errors.}}}, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
 When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
@@
 Though it is traditional to estimate \nz\ from \pz\ point estimates \citep{abruzzo_impact_2019}, there is a great deal of literature employing \pzpdf s.
-To circumvent this loss, it has become to be more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
+To circumvent this loss, it has become more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
 \begin{align}
@@
 
-Though previous versions used \texttt{HDF5} for the primary I/O format due to its efficiency for large quantities of data, it was abandoned in favor of \texttt{pickle} in the working release due to the instability of the \python\ implementation of the format on high-performance computing systems.
-The resulting output is a set of $I$ ordered \texttt{hickle} files enumerated by $\rho$ containing the state information after each sub-run.  
+%Though previous versions used \texttt{HDF5} for the primary I/O format due to its efficiency for large quantities of data, it was abandoned in favor of \texttt{pickle} in the working release due to the instability of the \python\ implementation of the format on high-performance computing systems.
+The resulting output is a set of $I$ ordered \texttt{pickle} files enumerated by $\rho$ containing the state information after each sub-run.  
 The state information includes $\frac{I_{0}}{s}$ actual samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$ as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$ over the entire sub-run.
@@
 One can see that the \pzpdf s take different shapes from one another even though the marginal histograms of the points are identical.
-The machine learning-like implicit prior has been modified to have nonzero value at high redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
+The machine learning-like implicit prior has been modified to have nonzero value at high-redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
 
@@
 We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
-Though there are also slight differences in the angle of the error ellipses, most of the differences are due to broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
+Though there are also slight differences in the angle of the error ellipses, the most striking effect is the broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
 The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
@@
 We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for proper inference to be conducted by consumers of this information.
+\aim{TODO: Strengthen language and get specific about what \pzpdf\ producers must do.}
 

### 2019-12-13 research/paper/draft.tex

@@
 \begin{abstract}
-A trustworthy estimate of the redshift distribution \nz\ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
-Spectroscopic redshifts for the dim and numerous galaxies of next-generation weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
-The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ posterior probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
-However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
-The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
-We introduce a statistically rigorous probabilistic graphical model of \dwh{redshift-dependent photometry} for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
-Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\textsc{CHIPPR}) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures.
-We present the \texttt{chippr} prototype code and use it to propagate these effects to constraints in the space of cosmological parameters, noting, however, that the mathematically justifiable approach incurs computational expense.
-The \textsc{CHIPPR} approach is applicable to any one-point statistic of any random variable, provided the prior probability density used to produce the posteriors is explicitly known; if the prior is implicit, as may be the case for popular photo-$z$ techniques, then the resulting posterior PDFs cannot be used for scientific inference.
-We therefore recommend that the photo-$z$ community direct attention to the development of methodologies that enable the recovery of photo-$z$ likelihoods \dwh{with support over all redshifts}, either directly or via a known prior probability density.
+A trustworthy estimate of the redshift distribution $n(z)$ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
+\dwh{Spectroscopic redshifts for the dim and numerous galaxies of next-generation weak-lensing surveys are expected to be unavailable, making photometric redshift (photo-$z$) probability density functions (PDFs) the next-best alternative for comprehensively encapsulating the nontrivial systematics affecting photo-$z$ point estimation.
+The established stacked estimator of $n(z)$ avoids reducing photo-$z$ PDFs to point estimates but yields a systematically biased estimate of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.}
+We introduce Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\textsc{CHIPPR}), a statistically rigorous probabilistic graphical model of redshift-dependent photometry, which correctly propagates the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures and is provably the only self-consistent way to recover $n(z)$ from photo-$z$ PDFs.
+We present the \texttt{chippr} prototype code and use it to forecast constraints in the space of cosmological parameters, noting, however, that the mathematically justifiable approach incurs computational expense.
+The \textsc{CHIPPR} approach is applicable to any one-point statistic of any random variable, provided the prior probability density used to produce the posteriors is explicitly known; 
+if the prior is implicit, as may be the case for popular photo-$z$ techniques, then the resulting posterior PDFs cannot be used for scientific inference.
+We therefore recommend that the photo-$z$ community focus on developing methodologies that enable the recovery of photo-$z$ likelihoods with support over all redshifts, either directly or via a known prior probability density.
 \end{abstract}
@@
 			The photometric data $\data_{j}$ for each galaxy is drawn from a function of its redshift $z_{j}$ and observed, indicated by a shaded circle.
-			\aim{Switch to vector graphics version?}}
+			}
 		\label{fig:pgm}
@@
 Rather, $\ndphi^{*}$ is a consequence of generative model for how photometry results from redshift, including the influence of intrinsic galaxy spectra and instrumental effects. 
-We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher\footnote{For template-based methods, the implicit prior is often an explicitly known input to the algorithm, engineered as an initial guess for the true $\ndphi$, with an aim for a realistic choice guided by an earlier spectroscopic survey.
+We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher}\footnote{For template-based methods, the implicit prior is often an explicitly known input to the algorithm, engineered as an initial guess for the true $\ndphi$, with an aim for a realistic choice guided by an earlier spectroscopic survey.
 (See \citet{benitez_bayesian_2000} for more detail.)
 It may thus be more appropriate to call it an \textit{interim prior}, but we will use the former term throughout this paper for generality.}
-Because the implicit prior is unavoidable and almost inherently not uninformative, the \pzpdf s reported by any method must be \textit{implicit posteriors} $\pr{z \gvn \data, \ndphi^{*}}$ weighted by the implicit prior.}
+Because the implicit prior is unavoidable and almost inherently not uninformative, the \pzpdf s reported by any method must be \textit{implicit posteriors} ${\pr{z \gvn \data, \ndphi^{*}}}$ weighted by the implicit prior.
 
@@
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
+%		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
+%		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
 		\caption{

### 2019-12-13 research/paper/cleaner.sh
clearing output

@@ -1,7 +1,7 @@
-infile='from_thesis.tex'
-outfile='draft.tex'
+#infile='thesis.tex'
+#outfile='draft.tex'
 
-sed -e "s/{\\textbackslash}'\\\\{i\\\\}/\\'{i}/g" from_thesis.bib | grep -E -v '^\W*(doi|url|urldate|abstract|file|keywords|annote|note)' > from_thesis.bib.tmp
-mv from_thesis.bib.tmp thesis.bib
+sed -e "s/{\\textbackslash}'\\\\{i\\\\}/\\'{i}/g" thesis.bib | grep -E -v '^\W*(doi|url|urldate|abstract|file|keywords|annote|note)' > thesis.bib.tmp
+mv thesis.bib.tmp thesis.bib
 
 #grep -E -v '^(%|[[:blank:]]*%|\\COMMENT|\includegraphics*)' "$infile" | fold -w80 -s > "$outfile"

### 2019-12-13 research/paper/draft.tex

@@
 % what's wrong with photo-zs?
-However, \pz s are susceptible to inaccuracy and imprecision \dwh{for a number of reasons}, particularly their inherent noisiness resulting from the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
+However, \pz s are susceptible to inaccuracy and imprecision in the form of their inherent noisiness resulting from the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
 Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
@@
 % how people get n(z) from photo-z PDFs
-Though it is traditional to estimate \nz\ from \pz\ point estimates \citep{abruzzo_impact_2019}, there is a great deal of literature employing \pzpdf s.
-To circumvent this loss, it has become more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
+Though it is traditional to estimate \nz\ from \pz\ point estimates \citep{abruzzo_impact_2019}, it has become more common to use \pzpdf s directly to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
 \begin{align}

### 2019-12-16 research/paper/draft.tex
removing cosmology

@@
 We introduce Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\textsc{CHIPPR}), a statistically rigorous probabilistic graphical model of redshift-dependent photometry, which correctly propagates the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures and is provably the only self-consistent way to recover $n(z)$ from photo-$z$ PDFs.
-We present the \texttt{chippr} prototype code and use it to forecast constraints in the space of cosmological parameters, noting, however, that the mathematically justifiable approach incurs computational expense.
+We present the \texttt{chippr} prototype code
+%and use it to forecast constraints in the space of cosmological parameters
+, noting that the mathematically justifiable approach incurs computational expense.
 The \textsc{CHIPPR} approach is applicable to any one-point statistic of any random variable, provided the prior probability density used to produce the posteriors is explicitly known; 
@@
 In Section~\ref{sec:alldata}, we describe the experimental design for testing the fully probabilistic approach to mock and real datasets, the results of which are found in Section~\ref{sec:results}.
-In Section~\ref{sec:results}, we stress-test the \Chippr\ model against stacking and its other competitors in the context of cosmology.
+In Section~\ref{sec:results}, we stress-test the \Chippr\ model against stacking and its other competitors.
+% in the context of cosmology.
 
@@
 The experiments of \Sect{sec:alldata} isolate the potential sources of error in \nz\ estimation one at a time.
-Now, we stress-test \Chippr\ by investigating two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one in which the \nz\ estimators are not provided with the same implicit prior used to generate the \pzpdf\ catalog (\Sect{sec:violations}).
-
-\subsection{LSST Requirements}
-\label{sec:lsstdemo}
-
-It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
-To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
-%Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
-%\aim{Don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space.}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/cosmolike_inputs.png}
-		\caption{
-			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
-		}
-		\label{fig:tomobins}
-	\end{center}
-\end{figure}
-
-\dwh{We consider as ground truth a set of known \nz\ corresponding to each of four hypothetical samples of galaxies and the corresponding cosmological parameter covariance matrix.
-The \nz\ of each galaxy subsample emulates that anticipated of galaxies binned by a redshift point estimate, as is common in tomographic redshift analyses, though our experimental procedure is agnostic to how the samples are identified.
-The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.}
-The true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
-
-First, we bin them down to a piecewise constant parameterization with a manageable $35$ hyperparameters for \chippr's sampling capabilities.
-Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
-The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
-We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
-at their given levels.
-Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
-		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
-		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend.
-		Enlarge axis labels.
-		Label panels.
-		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
-		}
-		\label{fig:per-bin-scatter}
-	\end{center}
-\end{figure*}
-
-\que{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
-
-We then make a point estimate of \nz\ using \chippr's \mmle\ optimization option as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
-Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
-The excessive breadth of the alternative estimators can be seen quite plainly.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
-		\caption{
-			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
-			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
-			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend.
-			Label figure/panels.
-			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
-		}
-		\label{fig:per-bin-ests}
-	\end{center}
-\end{figure*}
-
-We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
-Though there are also slight differences in the angle of the error ellipses, the most striking effect is the broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
-The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
-Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
-Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
-%\aim{TODO: Ask Elisabeth how to get the bias here.}
-\que{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
-Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
-Do you think it's sufficient to provide the bias on the moments of \nz, since that's what everyone uses anyway?}
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.9\textwidth]{figures/chippr/final_plot.png}
-		\caption{
-			\que{Are the contours any easier to see now?}
-			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
-			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
-			\aim{TODO: Fix formatting of axis labels.
-			Standardize ticklabels.}
-		}
-		\label{fig:cornerplot}
-	\end{center}
-\end{figure*}
-
-\que{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
-	Suggestions for how to better establish context would be appreciated.}
+Now, we stress-test \Chippr\ by investigating 
+%two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one 
+a case in which the \nz\ estimators are not provided the same implicit prior used to generate the \pzpdf\ catalog (\Sect{sec:violations}).
+
+%\subsection{LSST Requirements}
+%\label{sec:lsstdemo}
+%
+%It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
+%To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
+%%Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
+%%\aim{Don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space.}
+%
+%\begin{figure}
+%	\begin{center}
+%		\includegraphics[width=0.45\textwidth]{figures/chippr/cosmolike_inputs.png}
+%		\caption{
+%			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
+%		}
+%		\label{fig:tomobins}
+%	\end{center}
+%\end{figure}
+%
+%\dwh{We consider as ground truth a set of known \nz\ corresponding to each of four hypothetical samples of galaxies and the corresponding cosmological parameter covariance matrix.
+%The \nz\ of each galaxy subsample emulates that anticipated of galaxies binned by a redshift point estimate, as is common in tomographic redshift analyses, though our experimental procedure is agnostic to how the samples are identified.
+%The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.}
+%The true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+%
+%First, we bin them down to a piecewise constant parameterization with a manageable $35$ hyperparameters for \chippr's sampling capabilities.
+%Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
+%The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
+%We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
+%at their given levels.
+%Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
+%
+%\begin{figure*}
+%	\begin{center}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
+%		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
+%		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend.
+%		Enlarge axis labels.
+%		Label panels.
+%		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
+%		}
+%		\label{fig:per-bin-scatter}
+%	\end{center}
+%\end{figure*}
+%
+%\que{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
+%
+%We then make a point estimate of \nz\ using \chippr's \mmle\ optimization option as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
+%Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
+%The excessive breadth of the alternative estimators can be seen quite plainly.
+%
+%\begin{figure*}
+%	\begin{center}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
+%		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
+%		\caption{
+%			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
+%			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
+%			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend.
+%			Label figure/panels.
+%			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
+%		}
+%		\label{fig:per-bin-ests}
+%	\end{center}
+%\end{figure*}
+%
+%We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
+%Though there are also slight differences in the angle of the error ellipses, the most striking effect is the broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
+%The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
+%Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
+%Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
+%%\aim{TODO: Ask Elisabeth how to get the bias here.}
+%\que{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
+%Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
+%Do you think it's sufficient to provide the bias on the moments of \nz, since that's what everyone uses anyway?}
+%
+%\begin{figure*}
+%	\begin{center}
+%		\includegraphics[width=0.9\textwidth]{figures/chippr/final_plot.png}
+%		\caption{
+%			\que{Are the contours any easier to see now?}
+%			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
+%			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
+%			\aim{TODO: Fix formatting of axis labels.
+%			Standardize ticklabels.}
+%		}
+%		\label{fig:cornerplot}
+%	\end{center}
+%\end{figure*}
+%
+%\que{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
+%	Suggestions for how to better establish context would be appreciated.}
 
@@
 
-\aim{TODO: claim that implications for tomographic binning are severe, if that can be motivated by Section~\ref{sec:lsstdemo}.}
+\aim{TODO: claim that implications for tomographic binning are severe, if that can be motivated.}
+% by Section~\ref{sec:lsstdemo}.}
 
@@
 We discourage researchers from co-adding \pzpdf s or converting them into point estimates of redshift and instead recommend the use of Bayesian probability to guide the usage of \pzpdf s.  
-We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for proper inference to be conducted by consumers of this information.
-\aim{TODO: Strengthen language and get specific about what \pzpdf\ producers must do.}
+We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for any valid inference to be conducted by consumers of this information.
+\dwh{Methodologies for obtaining \pzpdf s must therefore be designed such that there is a known implicit prior, i.e. one that is not implicit at all, so that likelihoods may be recovered.}
 

### 2019-12-20 research/paper/draft.tex
more comments from @davidwhogg

@@
 A trustworthy estimate of the redshift distribution $n(z)$ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
-\dwh{Spectroscopic redshifts for the dim and numerous galaxies of next-generation weak-lensing surveys are expected to be unavailable, making photometric redshift (photo-$z$) probability density functions (PDFs) the next-best alternative for comprehensively encapsulating the nontrivial systematics affecting photo-$z$ point estimation.
-The established stacked estimator of $n(z)$ avoids reducing photo-$z$ PDFs to point estimates but yields a systematically biased estimate of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.}
+Spectroscopic redshifts for the dim and numerous galaxies of next-generation weak-lensing surveys are expected to be unavailable, making photometric redshift (photo-$z$) probability density functions (PDFs) the next-best alternative for comprehensively encapsulating the nontrivial systematics affecting photo-$z$ point estimation.
+The established stacked estimator of $n(z)$ avoids reducing photo-$z$ PDFs to point estimates but yields a systematically biased estimate of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
 We introduce Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\textsc{CHIPPR}), a statistically rigorous probabilistic graphical model of redshift-dependent photometry, which correctly propagates the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures and is provably the only self-consistent way to recover $n(z)$ from photo-$z$ PDFs.
@@
 However, \pz s are susceptible to inaccuracy and imprecision in the form of their inherent noisiness resulting from the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
-Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
+Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``photometric data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
+\aim{TODO: Clarify distinction between true redshift, spec-z, photo-z, and photometric data in text, my figure, and Jain figure.}
 The similarity to traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of \pz\ point estimates is no coincidence, as \pz\ point estimates effectively are indeed one such special nonlinear projection of the data.
@@
 		\caption{
+			\aim{TODO: Change y axis label to ``nonlinear projection of photometric data'' and y axis to ``true redshift.''}
 			A generic probability space (darker in areas of higher probability density) of redshift ($x$-axis) and data ($y$-axis), where the data is projected into a single dimension, with vertical cuts and marginals (cyan) indicating the construction of likelihoods and horizontal cuts and marginals (magenta) indicating the construction of posteriors.
 			The data (black points) used to generate the contours were extracted from \citet{jain_whole_2015} using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019}, with the ideal redshift estimation provided for reference (red diagonal).
+			\aim{TODO: Specify relationship between my figure and Jain figure; points are spec-z as true z and photo-z as nonlinear projection of photometric data.}
 		}
@@
 % how do we interpret photo-z imperfections?
-There are several varieties of generally non-Gaussian deviation from a \dwh{trivial relationship} between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
+There are several varieties of generally non-Gaussian deviation from a trivial relationship between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
 The coarseness of the photometric filters causes scatter about the diagonal, with larger scatter perpendicular to the diagonal at redshifts where highly identifiable spectral features pass between the filters, as well as higher scatter at high redshifts where faint galaxies with large photometric errors are more abundant.
@@
 % PDFs are a better way to improve photo-zs
-\dwh{The probability density function (PDF) in redshift space for each galaxy, commonly written as $\pr{z}$, is an alternative to the MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.}
+The probability density function (PDF) in redshift space for each galaxy, commonly written as $\pr{z}$, is an alternative to the MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.
 This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating our understanding of precision, accuracy, and systematic error.
@@
 Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent, most often by reduction to familiar point estimates.
-If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then their \dwh{redshift PDFs are well-approximated by delta functions} $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift\dwh{\footnote{\dwh{Note that \sz s are not the same as known true redshifts; the PDFs of \sz s would be narrow and almost always unimodal, but they would not be delta functions due to observational errors.}}}, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
+If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then their redshift PDFs are well-approximated by delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift\footnote{Note that \sz s are not the same as known true redshifts; the PDFs of \sz s would be narrow and almost always unimodal, but they would not be delta functions due to observational errors.}, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
 When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
@@
 
-\aim{TODO: Provide equations/citations for the use of \Nz\ in cosmology, why it matters.}
-
 The most prevalent application of \pzpdf s that preserves their information content is the estimation of the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}.
@@
 
-\aim{TODO: Say what precision is needed for \Nz\ for future weak lensing surveys, motivate how well we need to know \Nz.}
+\aim{TODO: Say why \Nz matters for cosmology, what precision is needed for \Nz\ for future weak lensing surveys, motivate how well we need to know \Nz.}
 Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the degree to which constraints on the cosmological parameters can advance is limited by the accuracy and precision to which \nz\ is known \citep{abruzzo_impact_2019}.
@@
 This summation procedure has been used extensively \citep{mandelbaum_precision_2008, benjamin_cfhtlens_2013, kelly_weighing_2014}\aim{\dots}
+\aim{TODO: Continue adding equations/citations for the use of \Nz\ in cosmology.}
 
@@
 Because the parameters $\ndphi$ that we seek are causally separated from the data by the latent variable of redshift, we call them \textit{hyperparameters}.
-\que{Am I adequately explaining the distinction between hyperparameters and parameters?}
 
@@
 
-\dwh{When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$, the probability of an unknown quantity conditioned on the quantity we have in hand, i.e the data.
-However, to do this, we must have a model for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
+When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$, the probability of an unknown quantity conditioned on the quantity we have in hand, i.e the data.
 Such a relationship $\pr{\data, z}$ is defined in the multidimensional space of probability density over redshift and photometry.
+\aim{TODO: Call the joint distribution an ``empirical frequency distribution'' and clarify that it exists in nature but is not known.}
+\aim{TODO: Show this conditional on the true redshift distribution $\phi^{\dagger}$.}
+\que{Propagate new $\phi^{\dagger}$ notation through appendix?}
+
+However, to estimate a posterior $\pr{z \gvn \data_{j}}$, the investigator must have a model $\phi$ for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
 If we were to marginalize over the photometry in $\pr{\data, z}$, we would obtain a one-dimensional PDF $\pr{z \gvn \ndphi^{*}}$ over redshift, which can by definition be parameterized by the same functional form as \nz, for some $\ndphi^{*}$ specific to the estimation procedure that may or may not bear any relation to the hyperparameters $\ndphi^{\dagger}$ of the true \nz.
 Rather, $\ndphi^{*}$ is a consequence of generative model for how photometry results from redshift, including the influence of intrinsic galaxy spectra and instrumental effects. 
-We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher}\footnote{For template-based methods, the implicit prior is often an explicitly known input to the algorithm, engineered as an initial guess for the true $\ndphi$, with an aim for a realistic choice guided by an earlier spectroscopic survey.
+We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher\footnote{For template-based methods, the implicit prior is often an explicitly known input to the algorithm, engineered as an initial guess for the true $\ndphi$, with an aim for a realistic choice guided by an earlier spectroscopic survey.
 (See \citet{benitez_bayesian_2000} for more detail.)
@@
 
-\dwh{With inputs of the \pzip\ catalog $\{\pr{z \gvn \data, \ndphi^{*}}\}$, the implicit prior $\pr{z \gvn \ndphi^{*}}$, and the prior PDF $\pr{\ndphi}$, we thus aim to obtain the posterior probability $\pr{\ndphi \gvn \{\data_{j}\}}$ of the redshift density function given all the photometric data.
+With inputs of the \pzip\ catalog $\{\pr{z \gvn \data, \ndphi^{*}}\}$, the implicit prior $\pr{z \gvn \ndphi^{*}}$, and the prior PDF $\pr{\ndphi}$, we thus aim to obtain the posterior probability $\pr{\ndphi \gvn \{\data_{j}\}}$ of the redshift density function given all the photometric data.
 By performing the derivation of Appendix~\ref{app:math}, we arrive at the desired expression
@@
 \end{equation}
-which is the very heart of \Chippr.
-This in effect replaces the implicit prior with the sampled model hyperparameters, thereby converting the \pzip s into likelihoods in order to obtain unbiased posteriors.}
+which is the very heart of \Chippr, also given as \Eq{eqn:final}.
+This in effect replaces the implicit prior with the sampled model hyperparameters, thereby converting the \pzip s into likelihoods in order to obtain unbiased posteriors.
 
@@
 We compare the results of \Chippr\ to those of stacking and the histogram of \pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed in \Sect{sec:forward}.
-\dwh{\Fig{fig:flowchart} illustrates the implementation of the forward model, defined by the much simpler \Fig{fig:pgm}, used for validating the method presented here.
-The irony of a simple model and complex validation procedure is not lost on the authors.}
-
-\dwh{\Fig{fig:flowchart} outlines the four phases of the generative model, which uses a total of three inputs.
+\Fig{fig:flowchart} illustrates the implementation of the forward model, defined by the much simpler \Fig{fig:pgm}, used for validating the method presented here.
+The irony of a simple model and complex validation procedure is not lost on the authors.
+
+\Fig{fig:flowchart} outlines the four phases of the generative model, which uses a total of three inputs.
 The experimental design requires our choice of true values $\phi^{\dagger}$ of the hyperparameters governing \nz, a \pz\ model $\pr{z, \data}$ defining the space of redshift and photometry, and prior values $\phi^{*}$ of the hyperparameters of \nz.
@@
 In the third phase, we evaluate the \pz\ model at that data to obtain $J$ posteriors $\pr{z \gvn \data_{j}^{\dagger}}$.
-In the fourth phase, we convolve the posteriors with the chosen prior $\pr{z \gvn \phi^{*}}$, yielding implicit posteriors $\pr{z \gvn \data_{j}^{\dagger}, \phi^{*}}$.}
+In the fourth phase, we convolve the posteriors with the chosen prior $\pr{z \gvn \phi^{*}}$, yielding implicit posteriors $\pr{z \gvn \data_{j}^{\dagger}, \phi^{*}}$.
 
@@
 		\caption{
-			\que{How to make this more attractive?}
+			\que{Change ``true joint model'' to true forward model, condition on $\phi^{\dagger}$, add footnote explaining distinction between dagger, star, prime, etc., check that derivation matches notation}
 			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
@@
 		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
-		\caption{The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
+		\caption{
+			\aim{TODO: Replace with figure of all effects at the same time.
+				Include slice further up so we can see outliers.}
+			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
 			The main panel of each shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal (thin red line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
 			The insets show marginal histograms (gray) in each dimension that can be compared with the true \nz\ (blue curve) used to make the figure to see the effect of the isolated systematic, as well as the implicit prior (red line).
-			\aim{TODO: Omit this figure; redundant with right panel of \Fig{fig:pzs-scatter} and top panels of \Fig{fig:uniform-outliers} and \Fig{fig:bias}.}
-%%			Enlarge axis labels and label panels.}
-%%			Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
 			}
@@
 		\aim{TODO: Enlarge axis labels and label panels.
-		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
+		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
+		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 	}
@@
 		\aim{TODO: Label panels.
-		Add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
+		Add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.
+		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 	}
@@
 		\aim{TODO: Enlarge axis labels on top and label panels.\\
-		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.}
+		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.
+		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 	}
@@
 		\aim{TODO: Enlarge axis labels and label panels.
-		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}		
+		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
+		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}		
 	}
@@
 Systematic bias in \pz\ point estimates is a concern for \lsst's cosmology results, for the same reasons explored in \citet{hoyle_dark_2017}.
+\aim{TODO: formally define this type of \pz\ bias and call it ``canonical'' or something for future distinction from generic biases.}
 However, in the context of \pzpdf s, the notion of redshift bias is a form of model misspecification.
@@
 		\aim{TODO: Enlarge axis labels on top and label panels.
-		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.}
+		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.
+		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 	}
@@
 
-As expected based on consistency with the forward model, \Chippr\ is completely resistant to bias, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
-\que{Did I effectively explain why \Chippr\ is expected to be unaffected by bias of this form?}
+As expected based on consistency with the forward model, \Chippr\ is completely resistant to linear bias of the form of $\Delta_{z}$, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
+\aim{Clarify that provided p(z) is consistent with forward model, CHIPPR has no bias of the canonical form.}
 
@@
 		Label panels.
-		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
+		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
+		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 	}
@@
 
-\que{Move \Sect{sec:violations} here?}
-
 The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzpdf\ catalogs.
 
-\section{Discussion}
+\section{Discussion and conclusions}
 \label{sec:results}
@@
 a case in which the \nz\ estimators are not provided the same implicit prior used to generate the \pzpdf\ catalog (\Sect{sec:violations}).
+
+\aim{TODO: Move conclusions opening and list of points here.}
 
@@
 %		Label panels.
-%		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.}
+%		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
+% 		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 %		}
@@
 %			Label figure/panels.
-%			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.}
+%			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.
+%			Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 %		}
@@
 
-\que{Move \Sect{sec:interim} here?}
-
-In this test, without tomographic binning, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
+%\que{Move \Sect{sec:interim} here?}
+
+In this test, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
 \Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
@@
 	\begin{center}
-%		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
-%		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
+		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
 		\caption{
@@
 			Label panels/figure.
-			Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.}
+			Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.
+			Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 		}
@@
 
-\section{Conclusion}
+\subsection{Conclusion}
 \label{sec:con}
+
+\aim{TODO: Break up conclusion into subsections of \Sect{sec:results}.}
+\aim{TODO: Add subsection on support; must have support everywhere in z, data in observation/universe and prior, and support is not guaranteed (cite qp for defining support)}
 
@@
 %The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
-\aim{TODO: Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
+\aim{TODO: Refer to quantitative results on KLD of \Nz.}
 %The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
@@
 
-\dwh{\Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and bias.
+\Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and bias.
 \Chippr\ is also robust to nontrivial implicit priors corresponding to the specifics of the architecture of the method by which \pzpdf s are obtained.
@@
 
-\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
+\aim{TODO: new paragraph for what can go wrong, 
+	call to community for what to do about it,
+	what aspects of implicit prior are knowable and not knowable?
 	if testable how so? outside the scope of paper, data producers be warned!
 	a likelihood is better -- give us that if you can!
-	focus on methods that acknowledge probabilistic structure of problem}
+	focus on methods that acknowledge probabilistic structure of problem.}
 
@@
 As an alternative, however, we must point out that if likelihoods were available rather than posteriors, the trouble with the implicit prior would be avoided altogether.
-We thus encourage the community of those making \pzpdf s to consider developing such methods so that the resulting data products may be correctly used in scientific inference more generically.}
-
-\aim{TODO: claim that implications for tomographic binning are severe, if that can be motivated.}
+We thus encourage the community of those making \pzpdf s to consider developing such methods so that the resulting data products may be correctly used in scientific inference more generically.
+
+\aim{TODO: Claim that implications for tomographic binning are severe, if that can be motivated by mean \Nz shifts.}
 % by Section~\ref{sec:lsstdemo}.}
@@
 
+\aim{TODO: Repeat that \chippr is slow, no current method exists for scaling up, doing inference with probabilistic data is inherently expensive, emphasize it's an important direction of future research.}
+\aim{TODO: point out that Fig 1 is uglier than 4, 5, 7, 8, 10, 11; these tests are simplistic, reality is more complex, but because \Chippr is provably correct, it can handle complexity of the real thing if p(z) is correct and captures complexity}
+
 By showing that \Chippr\ is effective in recovering the true redshift distribution function and posterior distributions on its parameters from catalogs of \pzpdf s, this work supports the production of \pzpdf s by upcoming photometric surveys such as \lsst\ to enable more accurate inference of the cosmological parameters.
@@
 We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for any valid inference to be conducted by consumers of this information.
-\dwh{Methodologies for obtaining \pzpdf s must therefore be designed such that there is a known implicit prior, i.e. one that is not implicit at all, so that likelihoods may be recovered.}
+Methodologies for obtaining \pzpdf s must therefore be designed such that there is a known implicit prior, i.e. one that is not implicit at all, so that likelihoods may be recovered.
 
@@
 Future work will also include the extension of this fully probabilistic approach to higher-order statistics of redshift such as the two-point correlation function.
+
+\begin{acknowledgements}
+	AIM acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the Federal Ministry of Education and Research.
+	During the completion of this work, AIM was supported by National Science Foundation grant AST-1517237 and the U.S.
+Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) program, administered by the Oak Ridge Institute for Science and Education for the DOE under contract number DE‐SC0014664.
+	The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike\ code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
+	\aim{TODO: Submit then send draft around to Foreman-Mackey, Leistedt, Kate Storey-Fisher, Phil Marshall, others for feedback after submitting to journal, post to arxiv at end of AAS.}
+	This work was completed with generous nutritional support from the Center for Computational Astrophysics.
+	\aim{TODO: Thank thesis readers.}
+\end{acknowledgements}
 
@@
 
-\aim{TODO: make notation consistent with the rest of the paper.}
+\aim{TODO: Ensure notation is consistent with the rest of the paper.}
 
@@
 
+\aim{TODO: State that this is equivalent to the result of Adam Meyers, Bovy, Hogg.}
+
 \section{Convergence Criteria}
@@
 
-\begin{acknowledgements}
-AIM acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the Federal Ministry of Education and Research.
-During the completion of this work, AIM was supported by National Science Foundation grant AST-1517237 and the U.S.
-Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) program, administered by the Oak Ridge Institute for Science and Education for the DOE under contract number DE‐SC0014664.
-The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
-\aim{TODO: Send draft around to Foreman-Mackey, Leistedt, others for feedback.}
-This work was completed with generous nutritional support from the Center for Computational Astrophysics.
-\end{acknowledgements}
-
 \bibliographystyle{apj}

### 2020-06-23 chippr/catalog_plots.py
slightly more attractive plots

@@ -139,8 +139,8 @@ def plot_mega_scatter(zs, pfs, z_grid, grid_ends, truth=None, plot_loc='', prepe
     limval = (max(grid_ends) - min(grid_ends)) / (len(grid_ends) - 1.)
     scatplot.set_xlim([min(grid_ends)-limval, max(grid_ends)+limval])
     scatplot.set_ylim([min(grid_ends)-limval, max(grid_ends)+limval])
-    scatplot.set_xticks(np.linspace(min(grid_ends), np.ceil(max(grid_ends)), 5))
-    scatplot.set_yticks(np.linspace(min(grid_ends), np.ceil(max(grid_ends)), 5))
+    scatplot.set_xticks(np.linspace(np.floor(min(grid_ends)), np.ceil(max(grid_ends)), 5))
+    scatplot.set_yticks(np.linspace(np.floor(min(grid_ends)), np.ceil(max(grid_ends)), 5))
     scatplot.set_xlabel(r'$z_{true}$')
     scatplot.set_ylabel(r'$z_{est}$')
     scatplot.text(0.25, 3., r'mock $p(z\mid \mathrm{``data"})$', rotation=0, size=20)

### 2020-06-23 chippr/log_z_dens_plots.py

@@ -344,7 +344,8 @@ def plot_estimators(info, plot_dir, log=True, prepend='', metrics=True, mini=Tru
         sps_log.set_ylim(-4., 1.)
         sps_log.set_ylabel(r'$\ln[n(z)]$')
     else:
-        sps_log.set_ylim(0., 3.)
+        sps_log.set_ylim(0., 4.)
+        sps_log.set_yticks([0,1,2,3,4])
         sps_log.set_ylabel(r'$n(z)$')
     sps_log.set_xlim(info['bin_ends'][0], info['bin_ends'][-1])
 
@@ -501,7 +502,11 @@ def plot_estimators(info, plot_dir, log=True, prepend='', metrics=True, mini=Tru
 
     # sps_log.legend(handles=color_plots[:-1], fontsize='x-small', loc='lower center', frameon=False)
     sps_log.legend(fontsize='large', loc='upper right', frameon=False)
-    sps_log.text(0.25, -3.75, r'inferred $n(z)$', rotation=0, size=20)
+    if log:
+        sps_log.text(0.25, -3.75, r'inferred $n(z)$', rotation=0, size=20)
+    else:
+        sps_log.text(2., 0.75, r'inferred $n(z)$', rotation=0, size=20)
+
     f.subplots_adjust(hspace=0, wspace=0)
     f.savefig(os.path.join(plot_dir, prepend+'estimators.png'), bbox_inches='tight', pad_inches = 0, dpi=d.dpi)
     print(info['stats'])

### 2020-06-24 research/paper/draft.tex
major figure update

@@
 However, \pz s are susceptible to inaccuracy and imprecision in the form of their inherent noisiness resulting from the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
-Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``photometric data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
-\aim{TODO: Clarify distinction between true redshift, spec-z, photo-z, and photometric data in text, my figure, and Jain figure.}
-The similarity to traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of \pz\ point estimates is no coincidence, as \pz\ point estimates effectively are indeed one such special nonlinear projection of the data.
+Figure~\ref{fig:pedagogical_scatter} is an adaptation of the ubiquitous plots of \pz\ vs.
+\sz\ illustrating the assumptions underlying \pz\ estimation in general, that \sz s are a good approximation to true redshifts and \pz s represent special non-linear projections of observed photometry to a scalar variable that approximates the true redshift.
 
@@
 		\caption{
-			\aim{TODO: Change y axis label to ``nonlinear projection of photometric data'' and y axis to ``true redshift.''}
-			A generic probability space (darker in areas of higher probability density) of redshift ($x$-axis) and data ($y$-axis), where the data is projected into a single dimension, with vertical cuts and marginals (cyan) indicating the construction of likelihoods and horizontal cuts and marginals (magenta) indicating the construction of posteriors.
-			The data (black points) used to generate the contours were extracted from \citet{jain_whole_2015} using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019}, with the ideal redshift estimation provided for reference (red diagonal).
-			\aim{TODO: Specify relationship between my figure and Jain figure; points are spec-z as true z and photo-z as nonlinear projection of photometric data.}
+			A generic probability space (darker in areas of higher probability density) of true redshift ($x$-axis) and a nonlinear projection of photometric data ($y$-axis), with vertical cuts and marginals (orange) indicating the construction of likelihoods and horizontal cuts and marginals (blue) indicating the construction of posteriors.
+			The data points were extracted using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019} from a \sz\ vs. \pz\ plot in \citet{jain_whole_2015};
+			the diagonal (yellow) represents a theoretically perfect \pz\ estimate.
 		}
@@
 Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are a dominant contributor to the total uncertainties reported on cosmological parameters \citep{abruzzo_impact_2019}.
-As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
+As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2018}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
 Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
@@
 			$3 \sigma$ catastrophic outlier rate & $< 10\%$\\
-			Bias & $< 0.003 (1 + z)$\\
+			Canonical bias & $< 0.003 (1 + z)$\\
 		\end{tabular}
@@
 \Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.
-\Pzpdf s are not without their own shortcomings, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014, malz_approximating_2018} and the divergent results of each method used to derive them \citep[ and reviewed in Schmidt, Malz \& Soo, et al.
-(in prep)]{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.
+\Pzpdf s are not without their own shortcomings, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014, malz_approximating_2018} and the divergent results of each method used to derive them \citep{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.
+Though the matter is outside the scope of this paper, reviews of various methods have been presented in the literature \citep{sheldon_photometric_2012, ball_robust_2008, carrasco_kind_tpz:_2013, carrasco_kind_exhausting_2014, schmidt_evaluation_2020}.
 The most concerning weakness of \pzpdf s, however, is their usage in the literature, which is at best inconsistent and at worst incorrect.
@@
 \nz\ is used to calculate the redshift calibration bias $b_{z}$ between the true and observed critical surface densities in galaxy-galaxy lensing \citep{mandelbaum_precision_2008} and the geometric lens efficiency $g_{k}(\chi)$ in tomographic weak lensing by large-scale structure \citep{benjamin_cfhtlens_2013}.
-The redshift distribution is a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}.
-\Nz\ may also be used to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
-
-\aim{TODO: Say why \Nz matters for cosmology, what precision is needed for \Nz\ for future weak lensing surveys, motivate how well we need to know \Nz.}
+\Nz\ may be used to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
+As a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure, the accuracy and precision to which \Nz\ is estimated can strongly impact our constraints on the parameters of cosmological models \citep{bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}, so it is unsurprising that this last application dominates the canonical bias requirement of Table~\ref{tab:lsstsrd}.
+%\aim{TODO: Say why \Nz\ matters for cosmology, what precision is needed for \Nz\ for future weak lensing surveys, motivate how well we need to know \Nz.}
 Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the degree to which constraints on the cosmological parameters can advance is limited by the accuracy and precision to which \nz\ is known \citep{abruzzo_impact_2019}.
@@
 for a sample of $J$ galaxies $j$, or, equivalently, the redshift distribution function $\hat{N}(z) = J \hat{n}(z)$, by effectively averaging the \pzpdf s.
-This summation procedure has been used extensively \citep{mandelbaum_precision_2008, benjamin_cfhtlens_2013, kelly_weighing_2014}\aim{\dots}
-\aim{TODO: Continue adding equations/citations for the use of \Nz\ in cosmology.}
+This summation procedure has been used extensively in cosmological analyses with photometric galaxy samples \citep{mandelbaum_precision_2008, benjamin_cfhtlens_2013, kelly_weighing_2014}.
+%\aim{\dots}
+%\aim{TODO: Continue adding equations/citations for the use of \Nz\ in cosmology.}
 
@@
 
-In addition to believing \nz\ is a PDF from which redshifts are drawn, we also believe that there is some higher dimensional probability space $\pr{z, \data}$ of redshift and photometric data vectors $\data$, which may be any combination of fluxes, magnitudes, colors, and their observational errors.
-In that sense \nz\ is equivalent to an integral
+In addition to believing \nz\ is a PDF from which redshifts are drawn, we also believe that there is some higher dimensional probability space $\pr{z, \data}$ of redshift $z$ and photometric data vectors $\data$, which may be any combination of fluxes, magnitudes, colors, and their observational errors.
+Under this framework, \nz\ is equivalent to an integral
 \begin{equation}
@@
 over the dimension of data in that joint probability space.
-Note that galaxies may have different observational data despite sharing the same redshift, and that galaxies at different redshifts may have identical photometry; the space $\pr{z, \data}$ need not be one-to-one.
-We assume a stronger version of statistical independence here, that draws $(z_{j}, \data_{j})$ are independent of draws $(z_{j'}, \data_{j'})$ in this space; the data and redshift of each galaxy are independent of those of other galaxies.
+Note that galaxies may have different observational data despite sharing the same redshift, and that galaxies at different redshifts may have identical photometry; 
+the space $\pr{z, \data}$ need not be one-to-one.
+We assume a stronger version of statistical independence here, that draws $(z_{j}, \data_{j})$ are independent of draws $(z_{j'}, \data_{j'})$ in this space; 
+the data and redshift of each galaxy are independent of those of other galaxies.
 
@@
 Both data and redshift are random variables, but data is the one that we observe and redshift is not directly observable.
-In this problem, we don't care about further constraining the redshifts of individual galaxies, only the redshift distribution, so we consider redshift to be a \textit{latent variable}.
+In this problem, we don't care about further constraining the redshifts of individual galaxies, only the redshift distribution \nz, so we consider redshift to be a \textit{latent variable}.
 Because the parameters $\ndphi$ that we seek are causally separated from the data by the latent variable of redshift, we call them \textit{hyperparameters}.
@@
 To self-consistently propagate the uncertainty in the inference of redshift, however, it is more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ over all possible values of $\ndphi$ conditioned on all the observed data $\{\data_{j}\}$ available in a generic catalog.
-In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the \textit{implicit prior} and the \textit{prior probability density (prior PDF)}, elaborated upon below.
-
-When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$, the probability of an unknown quantity conditioned on the quantity we have in hand, i.e the data.
-Such a relationship $\pr{\data, z}$ is defined in the multidimensional space of probability density over redshift and photometry.
-\aim{TODO: Call the joint distribution an ``empirical frequency distribution'' and clarify that it exists in nature but is not known.}
-\aim{TODO: Show this conditional on the true redshift distribution $\phi^{\dagger}$.}
-\que{Propagate new $\phi^{\dagger}$ notation through appendix?}
-
-However, to estimate a posterior $\pr{z \gvn \data_{j}}$, the investigator must have a model $\phi$ for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
+In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the \textit{implicit prior} and the \textit{prior probability density} (\textit{prior PDF}), elaborated upon below.
+
+When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$, the probability of an unknown quantity conditioned on the quantity we have in hand, i.e the photometric data.
+This posterior is effectively a marginalization with respect to redshift at a given value of $\data = \data_{j}$ of the \textit{empirical frequency distribution} $\pr{z, \data \gvn \ndphi^{\dagger}}$, the joint probability density corresponding to the true redshift distribution parameterized by $\ndphi^{\dagger}$, which exists in nature but need not be known.
+\que{Propagate new $\pr{z, \data \gvn \ndphi^{\dagger}}$ notation through appendix?}
+
+As the hyperparameters $\ndphi^{\dagger}$ of the true redshift distribution are in general unknown, the investigator seeking to estimate a posterior $\pr{z \gvn \data_{j}}$ must have a model $\phi^{*}$ for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
 If we were to marginalize over the photometry in $\pr{\data, z}$, we would obtain a one-dimensional PDF $\pr{z \gvn \ndphi^{*}}$ over redshift, which can by definition be parameterized by the same functional form as \nz, for some $\ndphi^{*}$ specific to the estimation procedure that may or may not bear any relation to the hyperparameters $\ndphi^{\dagger}$ of the true \nz.
-Rather, $\ndphi^{*}$ is a consequence of generative model for how photometry results from redshift, including the influence of intrinsic galaxy spectra and instrumental effects. 
+Rather, $\ndphi^{*}$ is a consequence of the generative model for how photometry results from redshift, including the influence of intrinsic galaxy spectra and instrumental effects.
+
 We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher\footnote{For template-based methods, the implicit prior is often an explicitly known input to the algorithm, engineered as an initial guess for the true $\ndphi$, with an aim for a realistic choice guided by an earlier spectroscopic survey.
@@
 With inputs of the \pzip\ catalog $\{\pr{z \gvn \data, \ndphi^{*}}\}$, the implicit prior $\pr{z \gvn \ndphi^{*}}$, and the prior PDF $\pr{\ndphi}$, we thus aim to obtain the posterior probability $\pr{\ndphi \gvn \{\data_{j}\}}$ of the redshift density function given all the photometric data.
-By performing the derivation of Appendix~\ref{app:math}, we arrive at the desired expression
+By performing the derivation of the Appendix, we arrive at the desired expression
 \begin{equation}
@@
 	\item Photometric measurements of galaxies are statistically independent Poisson draws from the set of all galaxies such that \Eq{eqn:indiedat} and \Eq{eqn:indie} hold.
-	\item We take the reported \pzip s to be accurate, free of model misspecification; draws thereof must not be inconsistent with the distribution of photometry and redshifts.
+	\item We take the reported \pzip s to be accurate, free of model misspecification; 
+	draws thereof must not be inconsistent with the distribution of photometry and redshifts.
 	Furthermore, we must be given the implicit prior $\ndphi^{*}$ used to produce the \pzip s.
@@
 In \Sect{sec:mcmc}, we describe the publicly available \chippr\ library.
-In \Sect{app:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
-
+In \Sect{sec:sheldon}, we introduce the alternative approaches evaluated for comparison with \Chippr.
+In \Sect{sec:diag}, we describe the diagnostic criteria by which we assess estimators of \nz.
+%In \Sect{app:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
+%
 \subsubsection{Code}
@@
 At each iteration $i$ and for each walker, a proposal distribution $\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for acceptance to or rejection from the full log-posterior distribution.
-
-Two threshold conditions are defined, one designating all previous samples to be ignored as as products of a burn-in phase and another indicating when a sufficient number of post-burn samples have been accepted.
-In this case, the first threshold (described in \Sect{app:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
-
+%Two threshold conditions are defined, one designating all previous samples to be ignored as as products of a burn-in phase and another indicating when a sufficient number of post-burn samples have been accepted.
+%In this case, the first threshold (described in \Sect{app:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
 %Though previous versions used \texttt{HDF5} for the primary I/O format due to its efficiency for large quantities of data, it was abandoned in favor of \texttt{pickle} in the working release due to the instability of the \python\ implementation of the format on high-performance computing systems.
-The resulting output is a set of $I$ ordered \texttt{pickle} files enumerated by $\rho$ containing the state information after each sub-run.  
-The state information includes $\frac{I_{0}}{s}$ actual samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$ as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$ over the entire sub-run.
-
-\subsection{Comparison with Alternative Approaches}
+
+The resulting output
+% is a set of ordered \texttt{pickle} files % enumerated by $\rho$ each containing the state information of $10^{3}$ accepted samples, with the last ten files including only samples taken after the completion of the burn-in phase
+% after each sub-run.  
+%The state information 
+includes $\frac{I_{0}}{s}$ accepted samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$, as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$, divided into separate files before and after the completion of the burn-in phase, as defined by the Gelman-Rubin statistic \citep{gelman_inference_1992}.
+
+\subsection{Alternative approaches for comparison}
 \label{sec:sheldon}
@@
 
-Point estimation converts the implicit photo-$z$ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$ into delta functions with all probability at a single estimated redshift.  
+Point estimation converts the implicit \pz\ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$ into delta functions with all probability at a single estimated redshift.  
 Some variants of point estimation choose this single redshift to be that of maximum a posteriori probability $\mathrm{mode}[\pr{z \gvn \data_{j}, \ndphi^{*}}]$ or the expected value of redshift $\langle z \rangle = \integral{z \pr{z \gvn \data_{j}, \ndphi^{*}}}{z}$.
@@
 
-A final estimator of the hyperparameters is the maximum marginalized likelihood estimator, the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.  
+A final estimator of the hyperparameters is the maximum marginalized likelihood estimator (MMLE), the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.
 To compare with sampling, the MMLE also depends on the choice of the hyperprior distribution, and it does not produce a full posterior probability distribution over the parameters of interest, only point estimators.
@@
 
-In \Sect{sec:diag} we outline the measures used to evaluate the performance of the method.
-
-\subsubsection{Performance metrics}
+\subsection{Performance metrics}
 \label{sec:diag}
 
-The results of the computation described in \Sect{sec:mcmc} are evaluated for accuracy on the basis of some quantitative measures.  
+The results of the computation described in \Sect{sec:exp} are evaluated for accuracy on the basis of some quantitative measures.  
 Beyond visual inspection of samples, we calculate summary statistics to quantitatively compare different estimators' precision and accuracy.  
@@
 
-\section{Validation}
+\aim{TODO: Also include the mean redshift/$\Delta_{z}$ for each estimator of \nz, cite the DES/KiDS/HSC papers that motivate this.}
+
+\section{Results}
 \label{sec:alldata}
@@
 	\begin{center}
-		\includegraphics[width=0.9\textwidth]{figures/chippr/flowchart.pdf}
+		\includegraphics[width=0.7\textwidth]{figures/chippr/flowchart.pdf}
 		\caption{
-			\que{Change ``true joint model'' to true forward model, condition on $\phi^{\dagger}$, add footnote explaining distinction between dagger, star, prime, etc., check that derivation matches notation}
+%			\que{Change ``true joint model'' to true forward model, condition on $\phi^{\dagger}$, add footnote explaining distinction between dagger, star, prime, etc., check that derivation matches notation}
 			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
@@
 
-The mock data emulates the three sources of error of highest concern to the \pz\ community that are explored in detail later in this section: intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and systematic bias (\Sect{sec:bias}).
-Tests including all three effects at the tolerance levels of \lsst\ (see Table~\ref{tab:lsstsrd}) 
-are presented in \Sect{sec:results}.
-\Fig{fig:mega_scatter} illustrates these three effects individually at twice the tolerance of \lsst\ for demonstrative purposes, hearkening back to Figure~\ref{fig:pedagogical_scatter}.
-We also test nontrivial implicit priors in \Sect{sec:interim}, which ought to be a priority for the \pz\ community.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
-		\includegraphics[width=0.32\textwidth]{figures/chippr/single_uout-mega_scatter.png}
-		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
+The mock data emulates the three sources of error of highest concern to the \pz\ community that are explored in detail later in this section: intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and canonical bias (\Sect{sec:bias}).
+\Fig{fig:mega_scatter} illustrates these three effects simultaneously at the tolerance of \lsst\ for demonstrative purposes, hearkening back to Figure~\ref{fig:pedagogical_scatter}.
+
+\begin{figure}
+	\begin{center}
+%		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
+%		\includegraphics[width=0.32\textwidth]{figures/chippr/single_uout-mega_scatter.png}
+%		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst-mega_scatter.png}
 		\caption{
-			\aim{TODO: Replace with figure of all effects at the same time.
-				Include slice further up so we can see outliers.}
-			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at twice the level of the \lsst\ requirements: intrinsic scatter (left), uniform outliers (center), and bias (right).
-			The main panel of each shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{true}} = z_{\mathrm{phot}}$ diagonal (thin red line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
-			The insets show marginal histograms (gray) in each dimension that can be compared with the true \nz\ (blue curve) used to make the figure to see the effect of the isolated systematic, as well as the implicit prior (red line).
+			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at the level of the \lsst\ requirements: intrinsic scatter, uniformly distributed catastrophic outliers, and bias.
+			The main panel shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{spec}} = z_{\mathrm{phot}}$ diagonal (gray line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
+			The insets show marginal histograms (light gray) in each dimension, that can be compared with the true \nz\ used to make the figure (black) to see the effect of these systematics, as well as the implicit prior (dark gray).
+			\aim{TODO: Include slice further up so we can see outliers.
+				Label panel.}
 			}
@@
 %\end{figure}
-\que{Should I add back the figure of prior samples?}
+%\que{Should I add back the figure of prior samples?}
 
 The sampler is initialized with $W=100$ walkers each with a value chosen from a Gaussian distribution of identity covariance around a sample from the hyperprior distribution.
+
+\que{Tests including all three canonical effects at the tolerance levels of \lsst\ were once in \Sect{sec:results} but Hogg requested the section be removed.
+	Include results of all 3 canonical effects combined at LSST levels here?}
 
@@
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter1.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_scatter2.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/single_varsigmas-mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
 	\caption{
-		Examples of mock \pzpdf s (colored lines) generated with intrinsic scatter at the \lsst\ requirements (left) and twice the \lsst\ requirements (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
-		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
-		\aim{TODO: Enlarge axis labels and label panels.
-		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
-		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
+		Examples of mock \pzpdf s generated with intrinsic scatter at the \lsst\ requirements (left) and twice the \lsst\ requirements (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines).
+		A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
+		\aim{TODO: Label panels.
+		Show the mean redshift for each estimator.}
 	}
@@
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter1.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_scatter2.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_varsigmas_log_estimators.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_hivarsig_log_estimators.png}
 	\caption{
@@
 		\aim{TODO: Label panels.
-		Add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.
-		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
+		Show the mean redshift for each estimator.}
 	}
@@
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_mega_scatter.png}\\
+	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout-mega_scatter.png}\\
 	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_log_estimators.png}
 	\caption{
-		Top: Examples of \pzpdf s with a uniform catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
+		Top: Examples of \pzpdf s with a uniformly distributed catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
+		\aim{TODO: Include slice further up.
+		Label panels.}
 		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+		\aim{TODO: Label panels.
+		Show the mean redshift for each estimator.}
 		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
-		\aim{TODO: Enlarge axis labels on top and label panels.\\
-		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.
-		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 	}
@@
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_mega_scatter.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout-mega_scatter.png}
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout-mega_scatter.png}
 	\caption{
-		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-		\aim{TODO: Enlarge axis labels and label panels.
-		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
-		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}		
+		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
+		\aim{TODO: Label panels.
+		Include slices higher up.}		
 	}
@@
 
-\subsection{Systematic bias}
+\subsection{Canonical bias}
 \label{sec:bias}
 
-Systematic bias in \pz\ point estimates is a concern for \lsst's cosmology results, for the same reasons explored in \citet{hoyle_dark_2017}.
-\aim{TODO: formally define this type of \pz\ bias and call it ``canonical'' or something for future distinction from generic biases.}
-However, in the context of \pzpdf s, the notion of redshift bias is a form of model misspecification.
-Consider that if bias were included in the framework of Figure~\ref{fig:pedagogical_scatter};
-a simple linear transformation of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + z_{\mathrm{phot}})$ would eliminate the bias.
-Regardless, for completeness, a test at ten times the bias of the \lsst\ requirements, with no redshift-dependent intrinsic scatter nor catastrophic outliers, is provided in \Fig{fig:bias}.
+Systematic bias in \pz\ point estimates, is a concern for \lsst's cosmology results, for the same reasons explored in \citet{hoyle_dark_2018}.
+\aim{TODO: Cite a KiDS and/or HSC paper using $\Delta_{z}$, too.}
+This form of bias is typically summarized by a shift parameter $\Delta_{z} = (\langle \pr{z \gvn \hat{\ndphi}} \rangle - \langle \pr{z \gvn \ndphi^{\dagger}} \rangle)$ representing a difference between the first moment of the estimated redshift density function and that of the true redshift density function.
+To distinguish other aforementioned manifestations of bias from this common form of bias, we refer to $\Delta_{z}$ as the \textit{canonical bias}.
+
+In the context of \pzpdf s, the canonical bias represents an instance of model misspecification.
+Consider that if the canonical bias were included in the framework of Figure~\ref{fig:pedagogical_scatter}, it could be trivially modeled out as a simple linear transformation of $z_{\mathrm{phot}} \to z_{\mathrm{phot}} - \Delta_{z} (1 + z_{\mathrm{phot}})$ of the $(z_{\mathrm{spec}}, z_{\mathrm{phot}})$ space.
+Regardless, for completeness, a test at ten times the canonical bias of the \lsst\ requirements, with no redshift-dependent intrinsic scatter nor catastrophic outliers, is provided in \Fig{fig:bias}.
 
@@
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_mega_scatter.png}\\
+	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}\\
 	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_log_estimators.png}
 	\caption{
-		Left: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-		Right: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+		Top: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
+				\aim{TODO: Include slice further up.
+					Label panels.}
+		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with ten times the bias of the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
+				\aim{TODO: Label panel.
+					Show the mean redshift for each estimator.}
 		The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
-		\aim{TODO: Enlarge axis labels on top and label panels.
-		Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.
-		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 	}
@@
 
-As expected based on consistency with the forward model, \Chippr\ is completely resistant to linear bias of the form of $\Delta_{z}$, and the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
-\aim{Clarify that provided p(z) is consistent with forward model, CHIPPR has no bias of the canonical form.}
-
-\subsection{Implicit prior}
-\label{sec:interim}
-
-\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
-Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
-Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
-%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.
-%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.
-%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.  
-Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
-% \aim{reference chapter 3 for complication of algorithm}
-Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
-%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.  
-%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.  
-%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.
-
-\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring implicit prior emulating that of a machine learning approach to \pz\ estimation (left panel) and a more complex interim prior emulating that of a template-fitting \pz\ method (right panel).
-One can see that the \pzpdf s take different shapes from one another even though the marginal histograms of the points are identical.
-The machine learning-like implicit prior has been modified to have nonzero value at high-redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_trpr.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/samplepzs_tmpr.png}
-	\caption{
-		Examples of mock \pzpdf s (colored lines) generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored lines), the true redshifts of the example \pzpdf s (black vertical lines).
-		A histogram (gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (blue curve) and implicit prior (red curve).
-		\aim{TODO: Enlarge axis labels.
-		Label panels.
-		Add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
-		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
-	}
-	\label{fig:pzs-priors}
-	\end{center}
-\end{figure*}
-
-\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzpdf s generated with nontrivial implicit priors.
-In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift distribution, and the distribution of \nz\ parameter values reflects higher uncertainty where the implicit prior undergoes large changes in derivative.
-The alternatives, on the other hand, are biased by the implicit prior except where it is flat, in the case of high redshifts for the machine learning-like implicit prior, resulting in over $1,000$ times the information loss on \nz\ for the machine learning-like implicit prior and some $5-20$ times the information loss for the template fitting-like implicit prior, relative to the \Chippr\ \mmle.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_trpr.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/results_tmpr.png}
-	\caption{
-		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with an implicit prior like that of machine learning \pzpdf\ approaches (left) and an implicit prior like that of template-fitting \pzpdf\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
-		\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
-		\aim{TODO: Label panels.
-		Add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.
-		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
-	}
-	\label{fig:results-priors}
-	\end{center}
-\end{figure*}
-
-The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzpdf\ catalogs.
-
-\section{Discussion and conclusions}
+As expected based on self-consistency of the forward-modeled \pzpdf s, \Chippr\ is immune to linear bias of the form of $\Delta_{z}$.
+Furthermore, the alternative estimators are only weakly affected, with information loss two and four times greater than that of the \Chippr\ \mmle\ for the histogram of modes and stacked estimator respectively.
+(This general robustness may suggest that the canonical bias may not be the most relevant measure of performance of estimators of \nz.)
+
+\section{Discussion}
 \label{sec:results}
 
-The experiments of \Sect{sec:alldata} isolate the potential sources of error in \nz\ estimation one at a time.
-Now, we stress-test \Chippr\ by investigating 
+The experiments of \Sect{sec:alldata} quantify the influence on each estimator of \nz\ due to each of the canonical types of \pz\ error one at a time in isolation.
+Now, we stress-test \Chippr\ by exploring the impact of the implicit prior, which has thus far not received much attention in the literature.
 %two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one 
-a case in which the \nz\ estimators are not provided the same implicit prior used to generate the \pzpdf\ catalog (\Sect{sec:violations}).
-
-\aim{TODO: Move conclusions opening and list of points here.}
+\Sect{sec:interim} demonstrates the sensitivity of \nz\ estimation methods to realistically complex implicit priors, and \Sect{sec:violations} demonstrates the consequences of mischaracterization of the implicit prior used to generate the \pzpdf\ catalog.
+These results provide compelling motivation for the \pz\ community to prioritize the study of implicit priors of existing and developing \pzpdf\ techniques.
 
@@
 
+\subsection{Realistically complex implicit prior}
+\label{sec:interim}
+
+\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
+Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
+Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
+%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.
+%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.
+%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.  
+Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
+% \aim{reference chapter 3 for complication of algorithm}
+Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
+%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.  
+%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.  
+%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.
+
+\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring implicit prior emulating that of a machine learning approach to \pz\ estimation (left panel) and a more complex interim prior emulating that of a template-fitting \pz\ method (right panel).
+One can see that the \pzpdf s take different shapes from one another even though the marginal histograms of the points are identical.
+The machine learning-like implicit prior has been modified to have nonzero value at high-redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_trpr-mega_scatter.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_tmpr-mega_scatter.png}
+		\caption{
+			Examples of mock \pzpdf s generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), the true redshifts of the example \pzpdf s (colored vertical lines).
+			A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
+			\aim{TODO: Label panels.
+				Include slices higher up.}
+		}
+		\label{fig:pzs-priors}
+	\end{center}
+\end{figure*}
+
+\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzpdf s generated with nontrivial implicit priors.
+In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift distribution, and the distribution of \nz\ parameter values reflects higher uncertainty where the implicit prior undergoes large changes in derivative.
+The alternatives, on the other hand, are biased by the implicit prior except where it is flat, in the case of high redshifts for the machine learning-like implicit prior, resulting in over $1,000$ times the information loss on \nz\ for the machine learning-like implicit prior and some $5-20$ times the information loss for the template fitting-like implicit prior, relative to the \Chippr\ \mmle.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_trpr_log_estimators.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_tmpr_log_estimators.png}
+		\caption
+		{The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with an implicit prior like that of machine learning \pzpdf\ approaches (left) and an implicit prior like that of template-fitting \pzpdf\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
+			\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
+			\aim{TODO: Label panels.
+				Show the mean redshift for each estimator.}
+		}
+		\label{fig:results-priors}
+	\end{center}
+\end{figure*}
+
+The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzpdf\ catalogs.
+
 \subsection{Violations of the model}
@@
 
-%\que{Move \Sect{sec:interim} here?}
-
 In this test, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
 \Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
-If we naively used the \pzpdf\ catalog produced by a generic machine learning code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_mega_scatter.png}\\
-		\includegraphics[width=0.45\textwidth]{figures/chippr/misspecified_log_estimators.png}
-		\caption{
-			Top: Examples of \pzpdf s with systematics at the level of the \lsst\ requirements with a machine learning-like implicit prior, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored curves), and the true redshifts of the example \pzpdf s (black vertical lines), with marginal histograms (gray) for each dimension with the true redshift distribution (blue curve) and implicit prior (red curve) in the insets.
-			Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and the uniform implicit prior given to \chippr\ (gray curve).
+If we naively used the \pzpdf\ catalog produced by a generic machine learning or template-fitting code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_trpr_wrong_log_estimators.png}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_tmpr_wrong_log_estimators.png}
+		\caption
+		{The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) when run with an incorrectly specified implicit prior (gray curve).
+			The data upon which each panel's results are based are provided in Figure~\ref{fig:pzs-priors}, where the left corresponds to the sort of implicit prior anticipated of machine learning approaches and the right corresponds to an implicit prior like that of a template-fitting code.
+			Here, \Chippr\ has been provided with a uniform implicit prior rather than those used to produce the mock \pzpdf s, and its performance is notably worse than when it is provided an accurate implicit prior, as in Figure~\ref{fig:results-priors}.
+			\aim{Label panels.
+				Show the mean redshift for each estimator.}
 			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
-			\aim{TODO: Enlarge axis labels on top.
-			Label panels/figure.
-			Also, add watermark of ``mock data'' in UL corner of top panel and ``results of inference'' on bottom panel.
-			Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
 		}
@@
 
-\subsection{Conclusion}
+\section{Conclusion}
 \label{sec:con}
 
-\aim{TODO: Break up conclusion into subsections of \Sect{sec:results}.}
+%\aim{TODO: Break up conclusion into subsections of \Sect{sec:alldata}/\Sect{sec:results}.}
 \aim{TODO: Add subsection on support; must have support everywhere in z, data in observation/universe and prior, and support is not guaranteed (cite qp for defining support)}
@@
 This study derives and demonstrates a mathematically consistent inference of a one-point statistic, the redshift density function \nz, based on an arbitrary catalog of \pzpdf s.  
-The fully Bayesian method, based in the fundamental laws of probability, begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
-The method is validated on mock data and tested in the regime of \lsst\ with promising results, outperforming the traditional stacking estimator at the level of \nz\ as well as in terms of constraining power on the cosmological parameters.
-Not only is this the only mathematically correct approach to the problem, it also recovers the true parameter values better than popular alternatives, as measured by the loss of information in \nz\ and the size of error ellipses in the space of cosmological parameters.
+The fully Bayesian \Chippr\ model, based in the fundamental laws of probability, begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
+The \Chippr\ model is implemented in the publicly available \chippr\ code.
+The method is implemented in the publicly available \chippr\ code and validated on mock data.
+% at the level of \nz\ as well as in terms of constraining power on the cosmological parameters.
 
@@
 %The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
-\aim{TODO: Refer to quantitative results on KLD of \Nz.}
 %The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
@@
 
-\Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and bias.
-\Chippr\ is also robust to nontrivial implicit priors corresponding to the specifics of the architecture of the method by which \pzpdf s are obtained.
-However, \Chippr\ absolutely requires that the implicit prior is accurately known; using an implicit prior other than that which contributed to the production of the \pzpdf\ catalog results in catastrophically incorrect inference of \nz.
-It is therefore imperative that those developing codes to obtain \pzpdf s provide a way to isolate the implicit prior and that those publishing \pzpdf\ catalogs provide the implicit prior to users.
-
+Under the paradigm of the canonical \pz\ error statistics, the following statements about the \Chippr\ methodology may be made with confidence:
+\begin{itemize}
+\item \Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and canonical systematic bias.
+\aim{TODO: Refer to quantitative results on KLD of \Nz.}
+\item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
+\item The error bars on the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that may be assumed for the conventional point estimators.
+\end{itemize}
+Not only is \Chippr\ the only mathematically correct approach to the problem, it also recovers the true values of the hyperparameters defining \nz\ better than popular alternatives, as measured by the loss of information in \nz.
+% \ and the size of error ellipses in the space of cosmological parameters.
+However, the mathematically valid approach to inference with probabilistic data products incurs nontrivial computational expense, motivating future work to optimize the implementation.
+%\aim{TODO: Repeat that \chippr is slow, no current method exists for scaling up, doing inference with probabilistic data is inherently expensive, emphasize it's an important direction of future research.}
+
+Furthermore, this work highlights a crucial and almost entirely overlooked complication to the usage of \pzpdf s, namely the implicit prior, motivating the following recommendations: 
+\begin{itemize}
+\item \Chippr\ alone is robust to nontrivial implicit priors corresponding to the specifics of the architecture of the method by which \pzpdf s are obtained;
+a hierarchical inference such as \Chippr\ is the only valid way to utilize \pzpdf s.
+\item \Chippr, however, is sensitive to misspecification of the implicit prior;
+the implicit prior used to produce the \pzpdf\ catalog must be known and provided to \Chippr\ in order to recover the true \nz.
+\end{itemize}
+Given the significance of the implicit prior \citep{schmidt_evaluation_2020}, it is therefore imperative that those developing codes to obtain \pzpdf s provide a way to isolate the implicit prior and that those publishing \pzpdf\ catalogs provide the implicit prior to users.
 This mandate is easier said than done, both for template fitting and machine learning approaches.
+
 While the implicit prior is often an explicit input to model-based routines, it may be defined in a space of redshift and SED templates.
@@
 
-\aim{TODO: new paragraph for what can go wrong, 
-	call to community for what to do about it,
-	what aspects of implicit prior are knowable and not knowable?
-	if testable how so? outside the scope of paper, data producers be warned!
-	a likelihood is better -- give us that if you can!
-	focus on methods that acknowledge probabilistic structure of problem.}
+%\aim{TODO: new paragraph for what can go wrong, 
+%	call to community for what to do about it,
+%	what aspects of implicit prior are knowable and not knowable?
+%	if testable how so? outside the scope of paper, data producers be warned!
+%	a likelihood is better -- give us that if you can!
+%	focus on methods that acknowledge probabilistic structure of problem.}
 
@@
 
-\aim{TODO: Claim that implications for tomographic binning are severe, if that can be motivated by mean \Nz shifts.}
+%\aim{TODO: Claim that implications for tomographic binning are severe, if that can be motivated by mean \Nz shifts.}
 % by Section~\ref{sec:lsstdemo}.}
 
-The following conclusions and recommendations can be made with confidence:
-
-\begin{enumerate}
-	\item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
-	\item The error bars on the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that may be assumed for the conventional point estimators.
-	%	\item The marginalized maximum likelihood estimator is an excellent estimator for strongly featured redshift distribution function with simple, clean photo-$z$ posteriors; stacking smooths features more than sampling and photo-$z$ point estimation.
-	\item When the implicit prior is known to be a poor match to the data, only the results of \Chippr\ are satisfactory estimators of the redshift distribution function because they are the only methods that can account for the bias induced on the \pzpdf\ catalog by the method that produces it; this is the most compelling case for the sampler because of the ubiquity of inappropriate interim priors.
-\end{enumerate}
-
-\aim{TODO: Repeat that \chippr is slow, no current method exists for scaling up, doing inference with probabilistic data is inherently expensive, emphasize it's an important direction of future research.}
+%\subsection*{Recommendations for future work}
+
+%The following conclusions and recommendations can be made with confidence:
+
+%\begin{enumerate}
+%	
+%	%	\item The marginalized maximum likelihood estimator is an excellent estimator for strongly featured redshift distribution function with simple, clean photo-$z$ posteriors; stacking smooths features more than sampling and photo-$z$ point estimation.
+%	\item When the implicit prior is known to be a poor match to the data, only the results of \Chippr\ are satisfactory estimators of the redshift distribution function because they are the only methods that can account for the bias induced on the \pzpdf\ catalog by the method that produces it; this is the most compelling case for the sampler because of the ubiquity of inappropriate interim priors.
+%\end{enumerate}
+
 \aim{TODO: point out that Fig 1 is uglier than 4, 5, 7, 8, 10, 11; these tests are simplistic, reality is more complex, but because \Chippr is provably correct, it can handle complexity of the real thing if p(z) is correct and captures complexity}
@@
 	The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike\ code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
-	\aim{TODO: Submit then send draft around to Foreman-Mackey, Leistedt, Kate Storey-Fisher, Phil Marshall, others for feedback after submitting to journal, post to arxiv at end of AAS.}
+	\aim{TODO: Submit then send draft around to Dan Foreman-Mackey, Boris Leistedt, Kate Storey-Fisher, Phil Marshall, Alan Heavens, others for feedback after submitting to ApJ.}
 	This work was completed with generous nutritional support from the Center for Computational Astrophysics.
-	\aim{TODO: Thank thesis readers.}
+%	\aim{TODO: Thank thesis readers.}
 \end{acknowledgements}
@@
 
-\aim{TODO: Figure out how to get the appendices enumerated for reference earlier.}
-
-\section{Derivation}
-\label{app:math}
+%\aim{TODO: Figure out how to get the appendices enumerated for reference earlier.}
+
+%\section{Derivation}
+%\label{app:math}
 
@@
 \end{align}
+Though the context differs, this result is equivalent to that of \citet{hogg_inferring_2010}.
 
@@
 Since we cannot know constant of proportionality, we sample the desired full log-posterior $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$ using Monte Carlo-Markov chain (MCMC) methods.
-The method outlined here is valid regardless of how the implicit \pz\ log-posteriors are obtained so the many approaches to producing \pzpdf s will not be rehashed; though the matter is outside the scope of this paper, reviews of various methods have been presented in the literature \citep{sheldon_photometric_2012, ball_robust_2008, carrasco_kind_tpz:_2013, carrasco_kind_exhausting_2014}, and will be briefly reviewed in Schmidt, Malz \& Soo, et al.
-(in prep).
+
 %
@@
 
-\aim{TODO: State that this is equivalent to the result of Adam Meyers, Bovy, Hogg.}
-
-\section{Convergence Criteria}
-\label{app:acorr}
-
-In addition to qualitative visual inspection of the chains, two quantities that probe the convergence of the sampler are used in this study, the autocorrelation time and the Gelman-Rubin convergence criterion.
-%\Fig{fig:chains} shows the %evolution of the values of one parameter of one walker over the course of all %iterations of the sampler.
-
-%\begin{figure}
-%%\includegraphics[width=0.5\textwidth]{figs/null/chain0.pdf}
-%\caption{This figure shows the evolution of one walker's parameter values for 
-%one element of the parameter vector $\vec{\theta}$ as a function of iteration 
-%number, demonstrating the completion of the burn-in phase.}
-%\label{fig:chains}
-%\end{figure}
-
-The autocorrelation time is effectively a measure of the efficiency of the method and can be described as the expected number of iterations necessary to accept a new sample independent of the current accepted sample.
-A sampler that converges faster will have a smaller autocorrelation time, and smaller autocorrelation times are preferable because it means fewer iterations are wasted on non-independent samples when independent samples are desired.
-See \citet{foreman-mackey_emcee_2013} for a more complete exploration of the autocorrelation time.  
-In all tests discussed here, autocorrelation times across walkers and parameters were approximately 20, meaning two samples 20 or more iterations apart were independent, a satisfactory level of efficiency.
-Low autocorrelation times are a necessary but not always sufficient convergence condition, as the autocorrelation times calculated for tests in this paper were constant across all sub-runs, even those that were obviously burning in.
-
-The Gelman-Rubin statistic
-\begin{equation}
-\label{eqn:gr}
-R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} b_{k}}{w_{k}}},
-\end{equation}
-a weighted sum of the mean $w_{k}$ of the variances within individual walkers' chains and the variance $b_{k}$ between chains of different walkers $m$, is calculated over each sub-run $i$ to determine the duration of the burn-in period.
-Convergence is achieved when the statistic approaches unity.
+%\section{Convergence Criteria}
+%\label{app:acorr}
+%
+%\que{Cut convergence criteria section?}
+%
+%In addition to qualitative visual inspection of the chains, two quantities that probe the convergence of the sampler are used in this study, the autocorrelation time and the Gelman-Rubin convergence criterion.
+%%\Fig{fig:chains} shows the %evolution of the values of one parameter of one walker over the course of all %iterations of the sampler.
+%
+%%\begin{figure}
+%%%\includegraphics[width=0.5\textwidth]{figs/null/chain0.pdf}
+%%\caption{This figure shows the evolution of one walker's parameter values for 
+%%one element of the parameter vector $\vec{\theta}$ as a function of iteration 
+%%number, demonstrating the completion of the burn-in phase.}
+%%\label{fig:chains}
+%%\end{figure}
+%
+%The autocorrelation time is effectively a measure of the efficiency of the method and can be described as the expected number of iterations necessary to accept a new sample independent of the current accepted sample.
+%A sampler that converges faster will have a smaller autocorrelation time, and smaller autocorrelation times are preferable because it means fewer iterations are wasted on non-independent samples when independent samples are desired.
+%See \citet{foreman-mackey_emcee_2013} for a more complete exploration of the autocorrelation time.  
+%In all tests discussed here, autocorrelation times across walkers and parameters were approximately 20, meaning two samples 20 or more iterations apart were independent, a satisfactory level of efficiency.
+%Low autocorrelation times are a necessary but not always sufficient convergence condition, as the autocorrelation times calculated for tests in this paper were constant across all sub-runs, even those that were obviously burning in.
+%
+%The Gelman-Rubin statistic
+%\begin{equation}
+%\label{eqn:gr}
+%R_{k} = \sqrt{\frac{(1 - \frac{2}{I_{0}}) w_{k} + \frac{2}{I_{0}} b_{k}}{w_{k}}},
+%\end{equation}
+%a weighted sum of the mean $w_{k}$ of the variances within individual walkers' chains and the variance $b_{k}$ between chains of different walkers $m$, is calculated over each sub-run $i$ to determine the duration of the burn-in period.
+%Convergence is achieved when the statistic approaches unity.
 

### 2020-07-01 research/paper/draft.tex
finalized!

@@
 %\usepackage{lineno}
-
-\usetikzlibrary{shapes.geometric, arrows}
-\usetikzlibrary{fit}
-
-\tikzstyle{hyper} = [circle, text centered, draw=black]
-\tikzstyle{param} = [circle, text centered, draw=black]
-\tikzstyle{data} = [circle, text centered, draw=black, line width=2pt]
-\tikzstyle{arrow} = [thick,->,>=stealth]
+\usepackage[percent]{overpic}
+\usepackage{float}
+\usepackage{wrapfig}
+
+%\usetikzlibrary{shapes.geometric, arrows}
+%\usetikzlibrary{fit}
+%
+%\tikzstyle{hyper} = [circle, text centered, draw=black]
+%\tikzstyle{param} = [circle, text centered, draw=black]
+%\tikzstyle{data} = [circle, text centered, draw=black, line width=2pt]
+%\tikzstyle{arrow} = [thick,->,>=stealth]
 
@@
 
+%\aim{TODO: Add test case panel labels to all figs of \chippr\ sims/outputs.  Take higher-z samples in mock \pzpdf\ plots relative to outlier populations.  Include $\bar{z}$ in results plots.}
+
 \section{Introduction}
@@
 This summation procedure has been used extensively in cosmological analyses with photometric galaxy samples \citep{mandelbaum_precision_2008, benjamin_cfhtlens_2013, kelly_weighing_2014}.
-%\aim{\dots}
 %\aim{TODO: Continue adding equations/citations for the use of \Nz\ in cosmology.}
@@
 
-In Section~\ref{sec:meth}, we present the \Chippr\ model and \chippr\ implementation for characterizing the full posterior probability landscape of \Nz\ using \pzpdf s. 
-In Section~\ref{sec:alldata}, we describe the experimental design for testing the fully probabilistic approach to mock and real datasets, the results of which are found in Section~\ref{sec:results}.
-In Section~\ref{sec:results}, we stress-test the \Chippr\ model against stacking and its other competitors.
+In Section~\ref{sec:meth}, we present the \Chippr\ model for characterizing the full posterior probability landscape of \Nz\ using \pzpdf s. 
+In Section~\ref{sec:application}, we present the \chippr\ implementation of the \Chippr\ model and the experimental set up by which we validate it, including the forward modeling of mock \pzpdf s.
+In Section~\ref{sec:alldata}, we present a number of informative test cases and compare the results of \chippr\ with alternative approaches.
+In Section~\ref{sec:results}, we stress-test the \Chippr\ model under nontraditional conditions.
 % in the context of cosmology.
-
-\section{Method}
+Finally, in Section~\ref{sec:con}, we make recommendations for future research involving \nz\ estimation.
+
+\section{Model}
 \label{sec:meth}
 
-Consider a survey of $J$ galaxies $j$, each with photometric data $\data_{j}$; thus the entire survey over some solid angle produces the ensemble of photometric magnitudes (or colors) and their associated observational errors $\{\data_{j}\}$.
-Each galaxy $j$ has a redshift $z_{j}$ that we would like to learn; redshift is a parameter in this case.  
-The distribution of the ensemble of redshifts $\{z_{j}\}$ may be described by the hyperparameters defining the redshift distribution function \nz\ that we would like to quantify.  
-This situation may be considered to be a probabilistic generative model, illustrated by the directed acyclic graph of \Fig{fig:pgm}.
-
+Consider a survey of $J$ galaxies $j$, each with photometric data $\data_{j}$; 
+thus the entire survey over some solid angle produces the ensemble of photometric magnitudes (or colors) and their associated observational errors $\{\data_{j}\}$.  
+Each galaxy $j$ has a redshift parameter $z_{j}$ that we would like to learn.  
+The distribution of the ensemble of redshift parameters $\{z_{j}\}$ may be described by the hyperparameters defining the redshift distribution function \nz\ that we would like to quantify.
 The redshift distribution function \nz\ is the number of galaxies per unit redshift, effectively defining the evolution in the number of galaxies convolved with the selection function of the sample \citep{menard_clustering-based_2013}.
-In the following sections, we present and compare methods for estimating \nz\ from \pzpdf s.  
-\Sect{sec:prob} contains the mathematical derivation of a probabilistic model for $n(z)$ dependent on photo-$z$ probability distribution functions, and \Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.
+
+In \Sect{sec:forward}, we establish a forward model encapsulating the causal relationship between \nz\ and photometry $\data$.  
+In \Sect{sec:prob}, we present the directed acyclic graph of this probabilistic generative model and interpret the corresponding mathematical expression, whose full derivation may be found in the Appendix.
+In \Sect{sec:limitations}, we summarize the necessary assumptions of the model.
+
+%\aim{TODO: check for consistent notation of \pzip s vs. \pzpdf s.}
 
@@
 of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
-We believe that galaxy redshifts are indeed drawn from \nz, making it a probability density over redshift; this fact can also be confirmed by dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{hogg_data_2012}.
+We believe that galaxy redshifts are indeed sampled, or drawn, from \nz, making it a probability density over redshift; this fact can also be confirmed by dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{hogg_data_2012}.
 
@@
 \end{equation}
-% \aim{explain "is drawn from"}
 a probability density over redshift conditioned on the parameters $\ndphi$ specifying \nz.
@@
 This posterior is effectively a marginalization with respect to redshift at a given value of $\data = \data_{j}$ of the \textit{empirical frequency distribution} $\pr{z, \data \gvn \ndphi^{\dagger}}$, the joint probability density corresponding to the true redshift distribution parameterized by $\ndphi^{\dagger}$, which exists in nature but need not be known.
-\que{Propagate new $\pr{z, \data \gvn \ndphi^{\dagger}}$ notation through appendix?}
+%\que{Propagate new $\pr{z, \data \gvn \ndphi^{\dagger}}$ notation through appendix?}
 
@@
 
-\subsubsection{Model Limitations}
+\subsection{Model Limitations}
 \label{sec:limitations}
@@
 These assumptions have known limitations.  
-First, the photometric data are not a set of independent measurements; the data are correlated not only by the conditions of the experiment under which they were observed (instrument and observing conditions) but also by redshift covariances resulting from physical processes governing underlying galaxy spectra and their relation to the redshift distribution function.
-Second, the reported \pzip s may not be trustworthy; there is not yet agreement on the best technique to obtain \pzpdf s, and the implicit prior may not be appropriate or even known to us as consumers of \pzip s.
+First, the photometric data are not a set of independent measurements; 
+the data are correlated not only by the conditions of the experiment under which they were observed (instrument and observing conditions) but also by redshift covariances resulting from physical processes governing underlying galaxy spectra and their relation to the redshift distribution function.
+Second, the reported \pzip s may not be trustworthy; 
+there is not yet agreement on the best technique to obtain \pzpdf s, and the implicit prior may not be appropriate or even known to us as consumers of \pzip s.  
 Third, the hyperprior may be quite arbitrary and poorly motivated if the underlying physics is complex, and it can only be appropriate if our prior beliefs about \nz\ are accurate.
+
+Furthermore, in Section~\ref{sec:prob}, we have made an assumption of \textit{support}, meaning the model $\pr{z, \data \gvn \ndphi}$ has mutual coverage with the parameter values that real galaxies can take.
+In other words, any probability distribution over the $(z, \data)$ space must be nonzero where real galaxies can exist. 
+Additionally, the hyperprior $\pr{\ndphi}$ must be nonzero at the hyperparameters $\ndphi^{\dagger}$ of the true redshift density function \nz.
+This assumption cannot be violated under the experimental design of Section~\ref{sec:forward}, but it is not generically guaranteed when performing inference on real data;
+thus the chosen $\pr{z, \data \gvn \ndphi^{*}}$ and $\pr{\ndphi}$ must be sufficiently general as to not rule out plausible areas of parameter space.
+
+\section{Methods \& Data}
+\label{sec:application}
+
+Here we describe the method by which we demonstrate the \Chippr\ model.
+In \Sect{sec:exp}, we outline the implementation of the \chippr\ code.
+In \Sect{sec:sheldon}, we introduce the alternative \nz\ estimators against which \Chippr\ is compared.
+In \Sect{sec:diag}, we present the quantitative metrics by which 
+In \Sect{sec:mock}, we outline the procedure for emulating mock \pzip s.
 
@@
 with $K$ bins of width $\delta z^{1}, \dots, \delta z^{K}$ defined by endpoints $z^{0}, \dots, z^{K}$.
-%\aim{Maybe include an equation for a general piecewise constant function here (which is currently in \Chap{qp})?}
 Thus each $\pr{z \gvn \data_{j}} = f(z; \ndphi_{j})$ has parameters $\ndphi_{j}$ that are defined in the same basis as those of \nz.
@@
 
-\subsection{Alternative approaches for comparison}
+\subsubsection{Alternative approaches for comparison}
 \label{sec:sheldon}
@@
 A final estimator of the hyperparameters is the maximum marginalized likelihood estimator (MMLE), the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.
-To compare with sampling, the MMLE also depends on the choice of the hyperprior distribution, and it does not produce a full posterior probability distribution over the parameters of interest, only point estimators.
-It must be noted that computation of the maximum marginalized likelihood estimator may be unstable depending on the strengths and weaknesses of the optimizer.  
-In general, derivatives will not be available for the full posterior distribution, restricting optimization methods used.
+The MMLE can be obtained in substantially less time than enough samples to characterize the full log-posterior distribution of \nz.
+However, the MMLE yields only a point estimate of \nz\ rather than characterizing the full log-posterior on $\ndphi$, and it does not escape the dependence on the choice of hyperprior distribution.  
+Furthermore, derivatives will not in general be available for the full posterior distribution, restricting optimization methods used, and, as is true for any optimization code, there is a risk of numerical instability.
 %\begin{equation}
@@
 
-\subsection{Performance metrics}
+\subsubsection{Performance metrics}
 \label{sec:diag}
@@
 \label{eqn:kl}
-KL_{',\ddagger} = \integral{\pr{z \gvn \ndphi'} \ln \left[ \frac{\pr{z \gvn \ndphi'}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
-\end{equation}
-which measures a distance from parameter values $\ndphi'$ to true parameter values $\ndphi^{\dagger}$, which is invariant under changes of variables.  
-We note that $KL_{',\dagger} \neq KL_{\dagger,'}$ and is only interpretable when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than $\ndphi'$.
-The KLD is explored in detail in the Appendix to \citet{malz_approximating_2018}.
-In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that produced by one of the methods in question.
+KL_{\ndphi,\ndphi^{\ddagger}} = \integral{\pr{z \gvn \ndphi} \ln \left[ \frac{\pr{z \gvn \ndphi}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
+\end{equation}
+which measures a distance from parameter values $\ndphi$ to true parameter values $\ndphi^{\dagger}$.  
+The KLD is a measure of the information loss, in units of nats, due to using $\ndphi$ to approximate the true $\ndphi^{\dagger}$ when it is known.
+A detailed exploration of the KLD may be found in the Appendix to \citet{malz_approximating_2018}.
+%We note that $KL_{\ndphi,\ndphi^{\ddagger}} \neq KL_{\ndphi^{\ddagger},\ndphi}$ and is only interpretable when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than $\ndphi$.
+%In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that produced by one of the methods in question.
 
@@
 
-\section{Results}
-\label{sec:alldata}
+\subsection{Validation on mock data}
+\label{sec:mock}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.7\textwidth]{figures/chippr/flowchart.pdf}
+		\caption{
+			%			\que{Change ``true joint model'' to true forward model, condition on $\phi^{\dagger}$, add footnote explaining distinction between dagger, star, prime, etc., check that derivation matches notation}
+			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
+			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation we perform, and rounded rectangles indicate a quantity created by the forward model.
+			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.
+		}
+		\label{fig:flowchart}
+	\end{center}
+\end{figure*}
 
@@
 
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.7\textwidth]{figures/chippr/flowchart.pdf}
-		\caption{
-%			\que{Change ``true joint model'' to true forward model, condition on $\phi^{\dagger}$, add footnote explaining distinction between dagger, star, prime, etc., check that derivation matches notation}
-			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
-			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation we perform, and rounded rectangles indicate a quantity created by the forward model.
-			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.
-			}
-		\label{fig:flowchart}
-	\end{center}
-\end{figure*}
-
 The true redshift distribution used in these tests is a particular instance of the gamma function
@@
 
-\begin{figure}
-	\begin{center}
-%		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
-%		\includegraphics[width=0.32\textwidth]{figures/chippr/single_uout-mega_scatter.png}
-%		\includegraphics[width=0.32\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}
+\begin{figure}[H]
+	\begin{center}
 		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst-mega_scatter.png}
@@
 			The insets show marginal histograms (light gray) in each dimension, that can be compared with the true \nz\ used to make the figure (black) to see the effect of these systematics, as well as the implicit prior (dark gray).
-			\aim{TODO: Include slice further up so we can see outliers.
-				Label panel.}
+%			\aim{TODO: Include slice further up so we can see outliers.
+%				Label panel.}
 		}
@@
 %An example of such samples from the prior are shown in \Fig{fig:prior}.
-%
-%\begin{figure}
-%%	\includegraphics[width=0.5\textwidth]{figures/chippr/null_priorsamps.pdf}
-%	\caption{\aim{I need to remake this one because it uses the wrong notation and I stopped making it automatically a while ago.}
-%		Samples (colored lines) of $\pr{z \gvn \ndphi}$ where each $\ndphi$ is drawn from the hyperprior distribution $\pr{\ndphi}$ given in \Eq{eqn:priorcov}.}
-%	\figlabel{fig:prior}
-%\end{figure}
 %\que{Should I add back the figure of prior samples?}
@@
 
-\que{Tests including all three canonical effects at the tolerance levels of \lsst\ were once in \Sect{sec:results} but Hogg requested the section be removed.
-	Include results of all 3 canonical effects combined at LSST levels here?}
+\section{Results}
+\label{sec:alldata}
+
+Here, we compare the results of the \Chippr\ methodology with those of established \nz\ estimators under the three traditional measures of \pz\ uncertainty one at a time:
+\Sect{sec:scatter} concerns the redshift-dependent intrinsic scatter, \Sect{sec:outliers} concerns realistically complex catastrophic outlier populations, and \Sect{sec:bias} concerns the canonical bias in the mean redshift.
 
@@
 		A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
-		\aim{TODO: Label panels.
-		Show the mean redshift for each estimator.}
+%		\aim{TODO: Label panels.
+%		Show the mean redshift for each estimator.}
 	}
@@
 		\Chippr\ is robust to intrinsic scatter, but the alternatives suffer from overly broad \nz\ estimates that worsen with increasing intrinsic scatter.
-		\aim{TODO: Label panels.
-		Show the mean redshift for each estimator.}
+%		\aim{TODO: Label panels.
+%		Show the mean redshift for each estimator.}
 	}
@@
 		Top: Examples of \pzpdf s with a uniformly distributed catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
-		\aim{TODO: Include slice further up.
-		Label panels.}
+%		\aim{TODO: Include slice further up.
+%		Label panels.}
 		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
-		\aim{TODO: Label panels.
-		Show the mean redshift for each estimator.}
+%		\aim{TODO: Label panels.
+%		Show the mean redshift for each estimator.}
 		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
@@
 		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
-		\aim{TODO: Label panels.
-		Include slices higher up.}		
+%		\aim{TODO: Label panels.
+%		Include slices higher up.}		
 	}
@@
 		Though the histogram of modes is most sensitive to a catastrophic outlier population, the stacked estimator also overestimates \nz\ under (machine learning-like outliers) and beyond (template fitting-like outliers).
-		\aim{TODO: Label panels.
-		Show the mean redshift for each estimator.}
+%		\aim{TODO: Label panels.
+%		Show the mean redshift for each estimator.}
 	}
@@
 Systematic bias in \pz\ point estimates, is a concern for \lsst's cosmology results, for the same reasons explored in \citet{hoyle_dark_2018}.
-\aim{TODO: Cite a KiDS and/or HSC paper using $\Delta_{z}$, too.}
 This form of bias is typically summarized by a shift parameter $\Delta_{z} = (\langle \pr{z \gvn \hat{\ndphi}} \rangle - \langle \pr{z \gvn \ndphi^{\dagger}} \rangle)$ representing a difference between the first moment of the estimated redshift density function and that of the true redshift density function.
@@
 		Top: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
-				\aim{TODO: Include slice further up.
-					Label panels.}
+%				\aim{TODO: Include slice further up.
+%					Label panels.}
 		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with ten times the bias of the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
-				\aim{TODO: Label panel.
-					Show the mean redshift for each estimator.}
+%				\aim{TODO: Label panel.
+%					Show the mean redshift for each estimator.}
 		The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
@@
 
+%\que{Add back the results of the LSST requirements here?}
+
 %\subsection{LSST Requirements}
@@
 %To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
-%%Though redshift tomography is non-physical, as redshift is a continuous random variable, and binning in estimated redshift introduces poorly understood systematic error, we perform this analysis as an example of how it affects the current standard in how cosmological parameters are constrained by galaxy surveys, rather than how we think they ought to be constrained.
-%%\aim{Don't introduce anti-tomography rant here, kind of controversial and should be explored where it has its own space.}
 %
@@
 %Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
-%%\aim{TODO: Ask Elisabeth how to get the bias here.}
 %\que{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
@@
 Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
-% \aim{reference chapter 3 for complication of algorithm}
 Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
@@
 			A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
-			\aim{TODO: Label panels.
-				Include slices higher up.}
+%			\aim{TODO: Label panels.
+%				Include slices higher up.}
 		}
@@
 			\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
-			\aim{TODO: Label panels.
-				Show the mean redshift for each estimator.}
+%			\aim{TODO: Label panels.
+%				Show the mean redshift for each estimator.}
 		}
@@
 			Here, \Chippr\ has been provided with a uniform implicit prior rather than those used to produce the mock \pzpdf s, and its performance is notably worse than when it is provided an accurate implicit prior, as in Figure~\ref{fig:results-priors}.
-			\aim{Label panels.
-				Show the mean redshift for each estimator.}
+%			\aim{Label panels.
+%				Show the mean redshift for each estimator.}
 			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
@@
 
-%\aim{TODO: Break up conclusion into subsections of \Sect{sec:alldata}/\Sect{sec:results}.}
-\aim{TODO: Add subsection on support; must have support everywhere in z, data in observation/universe and prior, and support is not guaranteed (cite qp for defining support)}
+%\que{TODO: Break up conclusion into subsections of \Sect{sec:alldata}/\Sect{sec:results}?}
 
@@
 
-Under the paradigm of the canonical \pz\ error statistics, the following statements about the \Chippr\ methodology may be made with confidence:
+Using a flexible, self-consistent forward model of the relationship between true and estimated redshifts, capable of encapsulating the complexity of observed redshift-photometry relations (e.g.
+\Fig{fig:pedagogical_scatter}), we emulate the canonical \pz\ error statistics, intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and canonical bias (\Sect{sec:bias}) one at a time.
+Though these test cases may appear overly simplistic, they enable rigorous quantification of the relative performance of each \nz\ estimation techniques under the controlled conditions of each type of error in isolation, at levels equal to and beyond those of \lsst.
+%\aim{TODO: point out that Fig 1 is uglier than 4, 5, 7, 8, 10, 11; these tests are simplistic, reality is more complex, but because \Chippr is provably correct, it can handle complexity of the real thing if p(z) is correct and captures complexity}
+
+Based on our tests, the following statements about the \Chippr\ methodology may be made with confidence:
 \begin{itemize}
-\item \Chippr\ outperforms traditional estimators of \nz\ under the canonical \pz\ systematics of intrinsic scatter, catastrophic outliers, and canonical systematic bias.
-\aim{TODO: Refer to quantitative results on KLD of \Nz.}
+\item \Chippr\ outperforms traditional estimators of \nz\ under realistically complex conditions, even at pessimistic levels relative to future survey requirements on the traditional \pz\ error statistics, as demonstrated both by eye and according to KLD values corresponding to $10\%$ the information loss of alternative methods.
+%\aim{TODO: Refer to quantitative results on KLD of \Nz.}
 \item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
@@
 However, the mathematically valid approach to inference with probabilistic data products incurs nontrivial computational expense, motivating future work to optimize the implementation.
-%\aim{TODO: Repeat that \chippr is slow, no current method exists for scaling up, doing inference with probabilistic data is inherently expensive, emphasize it's an important direction of future research.}
-
-Furthermore, this work highlights a crucial and almost entirely overlooked complication to the usage of \pzpdf s, namely the implicit prior, motivating the following recommendations: 
+
+Additionally, this work highlights a crucial and almost entirely overlooked complication to the usage of \pzpdf s, namely the implicit prior, motivating the following recommendations: 
 \begin{itemize}
@@
 
-\aim{TODO: point out that Fig 1 is uglier than 4, 5, 7, 8, 10, 11; these tests are simplistic, reality is more complex, but because \Chippr is provably correct, it can handle complexity of the real thing if p(z) is correct and captures complexity}
-
 By showing that \Chippr\ is effective in recovering the true redshift distribution function and posterior distributions on its parameters from catalogs of \pzpdf s, this work supports the production of \pzpdf s by upcoming photometric surveys such as \lsst\ to enable more accurate inference of the cosmological parameters.
@@
 	The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike\ code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
-	\aim{TODO: Submit then send draft around to Dan Foreman-Mackey, Boris Leistedt, Kate Storey-Fisher, Phil Marshall, Alan Heavens, others for feedback after submitting to ApJ.}
+%	\aim{TODO: Circulate draft to GCCL; submit to ApJ; circulate to Dan Foreman-Mackey, Boris Leistedt, Kate Storey-Fisher; post to arXiv; circulate to Alan Heavens, Ann Lee, Phil Marshall, Anze Slosar, others for feedback.}
 	This work was completed with generous nutritional support from the Center for Computational Astrophysics.
@@
 
+%\aim{TODO: add software citation section.}
+
 \appendix
@@
 %\renewcommand{\thesubsection}{\Alph{subsection}}
-
-%\aim{TODO: Figure out how to get the appendices enumerated for reference earlier.}
-
-%\section{Derivation}
-%\label{app:math}
-
-\aim{TODO: Ensure notation is consistent with the rest of the paper.}
+\numberwithin{equation}{section}
+
+\section{Derivation}
+\label{app:math}
 
@@
 
-In this paper, we work exclusively with log-probabilities.  
+We perform the derivation of \Eq{eqn:fullpost} using log-probabilities.  
 What we wish to estimate is then the full log-posterior probability distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ given the catalog of photometry $\{\data_{j}\}$.
@@
 
-Since we only have access to implicit \pz\ posteriors, we must be able to write the full log-posterior in terms of implicit \pz\ log-posteriors rather than the log-likelihoods of \Eq{eqn:posterior}.
-To do so, we will need an explicit statement of this implicit prior $\ndphi^{*}$ for whatever method is chosen to produce the implicit \pz\ posteriors.
+Since we only have access to \pzip s, we must be able to write the full log-posterior in terms of log \pzip s rather than the log-likelihoods of \Eq{eqn:posterior}.
+To do so, we will need an explicit statement of this implicit prior $\ndphi^{*}$ for whatever method is chosen to produce the \pzip s.
 
@@
 \end{align}
-Though the context differs, this result is equivalent to that of \citet{hogg_inferring_2010}.
-
-The argument of the integral in the log-posterior of \Eq{eqn:final} depends solely on knowable quantities (and those we must explicitly assume) and can be calculated for a given sample of \pz\ log-posteriors $\{\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with which they were obtained, noting the relation of
+which is equivalent to that of \citet{hogg_inferring_2010}, though the context differs.
+
+The argument of the integral in the log-posterior of \Eq{eqn:final} depends solely on knowable quantities (and those we must explicitly assume) and can be calculated for a given sample of log \pzip s $\{\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]\}$ and the implicit prior $\pr{z \gvn \ndphi^{*}}$ with which they were obtained, noting the relation of
 \begin{equation}
@@
 
+%\aim{TODO: find way to cite Dance Your Ph.D. video}
+
 \end{document}

### 2020-07-17 research/paper/draft.tex
minor rewording

@@
 		\caption{
-			A generic probability space (darker in areas of higher probability density) of true redshift ($x$-axis) and a nonlinear projection of photometric data ($y$-axis), with vertical cuts and marginals (orange) indicating the construction of likelihoods and horizontal cuts and marginals (blue) indicating the construction of posteriors.
-			The data points were extracted using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019} from a \sz\ vs. \pz\ plot in \citet{jain_whole_2015};
-			the diagonal (yellow) represents a theoretically perfect \pz\ estimate.
+			A generic probability space (darker in areas of higher probability density) of true redshift ($x$-axis) and a nonlinear projection of photometric data ($y$-axis), with vertical cuts and marginals (orange) indicating the construction of likelihoods and horizontal cuts and marginals (blue) indicating the construction of posteriors, with a theoretically perfect \pz\ estimate on the diagonal (yellow) for reference.
+			The data points were extracted using WebPlotDigitizer \citep{rohatgi_webplotdigitizer_2019} from a \sz\ vs. \pz\ plot in \citet{jain_whole_2015}.
+%			\aim{TODO: recolor insets/bars to not be same as gradient.}
 		}
@@
 In \Sect{sec:exp}, we outline the implementation of the \chippr\ code.
-In \Sect{sec:sheldon}, we introduce the alternative \nz\ estimators against which \Chippr\ is compared.
-In \Sect{sec:diag}, we present the quantitative metrics by which 
+%In \Sect{sec:sheldon}, we introduce the alternative \nz\ estimators against which \Chippr\ is compared.
+%In \Sect{sec:diag}, we present the quantitative metrics by which the \nz\ estimators are compared.
 In \Sect{sec:mock}, we outline the procedure for emulating mock \pzip s.
@@
 %In \Sect{app:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
-%
+
 \subsubsection{Code}
@@
 
-In this study, we compare the results of \Eq{eqn:fullpost} to those of the two most common approaches to estimating \nz\ from a catalog of \pzpdf s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior probability
+In this study, we compare the results of \Eq{eqn:fullpost} to those of the two most common approaches to estimating \nz\ from a catalog of \pzip s: 
+the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior probability
 \begin{equation}
@@
 \end{equation}
-(i.e. the distribution of modes of the \pzpdf s) and the stacked estimator of \Eq{eqn:stacked}, which can be rewritten as 
+(i.e. the distribution of modes of the \pzip s) and the stacked estimator of \Eq{eqn:stacked}, which can be rewritten as 
 \begin{equation}
@@
 \end{equation}
-in terms of the implicit \pz\ posteriors we have.
+in terms of the \pzip s we have.
 These two approaches have been compared to one another by \citet{hildebrandt_cfhtlens:_2012}, \citet{benjamin_cfhtlens_2013}, and \citet{asorey_galaxy_2016} in the past but not to \Chippr.
@@
 \citet{tanaka_photometric_2018} directs attention to deriving an optimal point estimate reduction of a \pzpdf, but since the purpose of this paper is to compare against the most established alternative estimators of \nz, its use will be postponed until a future study.
-Stacking these modified \pzpdf s leads to the marginalized maximum a posteriori (MMAP) estimator and the marginalized expected value (MExp) estimator, though only the former is included in this study since the latter has fallen out of favor in recent years\footnote{And for good reason!
+Stacking these modified \pzip s leads to the marginalized maximum a posteriori (MMAP) estimator and the marginalized expected value (MExp) estimator, though only the former is included in this study since the latter has fallen out of favor in recent years\footnote{And for good reason!
 Consider a bimodal \pzpdf; its expected value may very well fall in a region of very low probability, yielding a less probable point estimate than the point at which either peak achieves its maximum.}.
@@
 
-\aim{TODO: Also include the mean redshift/$\Delta_{z}$ for each estimator of \nz, cite the DES/KiDS/HSC papers that motivate this.}
+%\aim{TODO: Also include the mean redshift/$\Delta_{z}$ for each estimator of \nz, cite the DES/KiDS/HSC papers that motivate this.}
 
@@
 			Ovals indicate a quantity that must be chosen in order to generate the data, rectangles indicate an operation we perform, and rounded rectangles indicate a quantity created by the forward model.
-			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzpdf\ catalogs.
+			Arrows indicate the inputs and outputs of each operation performed to simulate mock \pzip\ catalogs.
 		}
@@
 
-We compare the results of \Chippr\ to those of stacking and the histogram of \pzpdf\ maxima (modes) on mock data in the form of catalogs of emulated \pzpdf s generated via the forward model discussed in \Sect{sec:forward}.
+We compare the results of \Chippr\ to those of stacking and the histogram of \pzip\ maxima (modes) on mock data in the form of catalogs of emulated \pzip s generated via the forward model discussed in \Sect{sec:forward}.
 \Fig{fig:flowchart} illustrates the implementation of the forward model, defined by the much simpler \Fig{fig:pgm}, used for validating the method presented here.
@@
 The mock data emulates the three sources of error of highest concern to the \pz\ community that are explored in detail later in this section: intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and canonical bias (\Sect{sec:bias}).
-\Fig{fig:mega_scatter} illustrates these three effects simultaneously at the tolerance of \lsst\ for demonstrative purposes, hearkening back to Figure~\ref{fig:pedagogical_scatter}.
-
-\begin{figure}[H]
+\Fig{fig:mega_scatter} illustrates these three effects simultaneously at the tolerance of \lsst\ for demonstrative purposes, harking back to Figure~\ref{fig:pedagogical_scatter}.
+
+\begin{figure}%[H]
 	\begin{center}
@@
 		\caption{
-			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at the level of the \lsst\ requirements: intrinsic scatter, uniformly distributed catastrophic outliers, and bias.
+			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at the level of the \lsst\ requirements: 
+			intrinsic scatter, uniformly distributed catastrophic outliers, and bias.
 			The main panel shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{spec}} = z_{\mathrm{phot}}$ diagonal (gray line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
@@
 %two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one 
-\Sect{sec:interim} demonstrates the sensitivity of \nz\ estimation methods to realistically complex implicit priors, and \Sect{sec:violations} demonstrates the consequences of mischaracterization of the implicit prior used to generate the \pzpdf\ catalog.
+\Sect{sec:interim} demonstrates the sensitivity of \nz\ estimation methods to realistically complex implicit priors, and \Sect{sec:violations} demonstrates the consequences of mischaracterization of the implicit prior used to generate the \pzip\ catalog.
 These results provide compelling motivation for the \pz\ community to prioritize the study of implicit priors of existing and developing \pzpdf\ techniques.
@@
 
-\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
+\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzip\ codes.
 Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
@@
 
-\Fig{fig:pzs-priors} shows examples of \pzpdf s with a low-redshift favoring implicit prior emulating that of a machine learning approach to \pz\ estimation (left panel) and a more complex interim prior emulating that of a template-fitting \pz\ method (right panel).
-One can see that the \pzpdf s take different shapes from one another even though the marginal histograms of the points are identical.
+\Fig{fig:pzs-priors} shows examples of \pzip s with a low-redshift favoring implicit prior emulating that of a machine learning approach to \pz\ estimation (left panel) and a more complex interim prior emulating that of a template-fitting \pz\ method (right panel).
+One can see that the \pzip s take different shapes from one another even though the marginal histograms of the points are identical.
 The machine learning-like implicit prior has been modified to have nonzero value at high-redshift because the implicit prior must be strictly positive definite for the \Chippr\ model to be valid.
@@
 		\caption{
-			Examples of mock \pzpdf s generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), the true redshifts of the example \pzpdf s (colored vertical lines).
+			Examples of mock \pzip s generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzip s (colored step functions), the true redshifts of the example \pzip s (colored vertical lines).
 			A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
@@
 
-\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzpdf s generated with nontrivial implicit priors.
+\Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzip s generated with nontrivial implicit priors.
 In both cases, the \Chippr\ \mmle\ effectively recovers the true redshift distribution, and the distribution of \nz\ parameter values reflects higher uncertainty where the implicit prior undergoes large changes in derivative.
@@
 		\caption
-		{The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with an implicit prior like that of machine learning \pzpdf\ approaches (left) and an implicit prior like that of template-fitting \pzpdf\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
+		{The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzip s with an implicit prior like that of machine learning \pzip\ approaches (left) and an implicit prior like that of template-fitting \pzip\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
 			\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
@@
 
-The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzpdf\ catalogs.
+The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzip\ catalogs.
 
@@
 \Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
-If we naively used the \pzpdf\ catalog produced by a generic machine learning or template-fitting code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
+If we naively used the \pzip\ catalog produced by a generic machine learning or template-fitting code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
 
@@
 			The data upon which each panel's results are based are provided in Figure~\ref{fig:pzs-priors}, where the left corresponds to the sort of implicit prior anticipated of machine learning approaches and the right corresponds to an implicit prior like that of a template-fitting code.
-			Here, \Chippr\ has been provided with a uniform implicit prior rather than those used to produce the mock \pzpdf s, and its performance is notably worse than when it is provided an accurate implicit prior, as in Figure~\ref{fig:results-priors}.
+			Here, \Chippr\ has been provided with a uniform implicit prior rather than those used to produce the mock \pzip s, and its performance is notably worse than when it is provided an accurate implicit prior, as in Figure~\ref{fig:results-priors}.
 %			\aim{Label panels.
@@
 The stacked estimator and histogram of modes don't make use of the implicit prior so do no worse than when the implicit prior is accurately provided, but \Chippr\ is sensitive to prior misspecification, which violates the model upon which it is based.
-It is thus crucial that \pzpdf\ methods always characterize and provide the implicit prior.
+It is thus crucial that \pzip\ methods always characterize and provide the implicit prior.
 
@@
 \begin{itemize}
-\item \Chippr\ alone is robust to nontrivial implicit priors corresponding to the specifics of the architecture of the method by which \pzpdf s are obtained;
-a hierarchical inference such as \Chippr\ is the only valid way to utilize \pzpdf s.
-\item \Chippr, however, is sensitive to misspecification of the implicit prior;
+\item In the presence of a nontrivial implicit prior corresponding to the specifics of the architecture of the method by which \pzpdf s are obtained, established methods cannot recover \nz;
+a principled hierarchical inference such as \Chippr\ is the only way to recover \nz\ from \pzpdf s.
+\item %\Chippr, however, is sensitive to misspecification of the implicit prior;
+Neither \Chippr\ nor traditional alternatives can recover \nz\ in the presence of a misspecified implicit prior;
 the implicit prior used to produce the \pzpdf\ catalog must be known and provided to \Chippr\ in order to recover the true \nz.
@@
 	The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike\ code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
-%	\aim{TODO: Circulate draft to GCCL; submit to ApJ; circulate to Dan Foreman-Mackey, Boris Leistedt, Kate Storey-Fisher; post to arXiv; circulate to Alan Heavens, Ann Lee, Phil Marshall, Anze Slosar, others for feedback.}
+%	The authors also acknowledge 
+%	\aim{TODO: Circulate draft to GCCL; submit to ApJ; circulate to Dan Foreman-Mackey, Boris Leistedt, Kate Storey-Fisher; post to arXiv; circulate to Johann Cohen-Tanugi, Will Hartley, Alan Heavens, Mike Jarvis, Francois Lanusse, Ann Lee, Rachel Mandelbaum, Phil Marshall, Chris Morrison, Jeff Newman, Sam Schmidt, Anze Slosar, Josh Speagle, others for feedback.}
 	This work was completed with generous nutritional support from the Center for Computational Astrophysics.

### 2020-11-25 research/paper/draft.tex
final update of text
removed duplicate sentence

@@
 \author{Alex I. Malz\altaffilmark{1,2}}
-\author{David W. Hogg\altaffilmark{2,3,4,5}}
+\author{David W. Hogg\altaffilmark{2,3,4,5}}5
 \email{aimalz@nyu.edu}
@@
 We implement the \Chippr\ model in code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
-In \Sect{sec:mcmc}, we describe the publicly available \chippr\ library.
+In \Sect{sec:mcmc}, we describe the  available \chippr\ library.
 In \Sect{sec:sheldon}, we introduce the alternative approaches evaluated for comparison with \Chippr.
@@
 The fully Bayesian \Chippr\ model, based in the fundamental laws of probability, begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
-The \Chippr\ model is implemented in the publicly available \chippr\ code.
-The method is implemented in the publicly available \chippr\ code and validated on mock data.
+The \Chippr\ model is implemented in the publicly available \chippr\ code and validated on mock data.
 % at the level of \nz\ as well as in terms of constraining power on the cosmological parameters.

### 2021-04-20 research/paper/draft.tex
backup before syncing major changes

@@
 
-\author{Alex I. Malz\altaffilmark{1,2}}
-\author{David W. Hogg\altaffilmark{2,3,4,5}}5
+\author{Alex I. Malz\altaffilmark{1}}
+\author{David W. Hogg\altaffilmark{2,3,4,5}}
 \email{aimalz@nyu.edu}
@@
 
-%\subsection{LSST Requirements}
-%\label{sec:lsstdemo}
-%
-%It is of interest to explore the impact of incorrectly estimated \nz\ on the cosmological inference to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
-%To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
-%
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[width=0.45\textwidth]{figures/chippr/cosmolike_inputs.png}
-%		\caption{
-%			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
-%		}
-%		\label{fig:tomobins}
-%	\end{center}
-%\end{figure}
-%
-%\dwh{We consider as ground truth a set of known \nz\ corresponding to each of four hypothetical samples of galaxies and the corresponding cosmological parameter covariance matrix.
-%The \nz\ of each galaxy subsample emulates that anticipated of galaxies binned by a redshift point estimate, as is common in tomographic redshift analyses, though our experimental procedure is agnostic to how the samples are identified.
-%The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.}
-%The true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
-%
-%First, we bin them down to a piecewise constant parameterization with a manageable $35$ hyperparameters for \chippr's sampling capabilities.
-%Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
-%The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
-%We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
-%at their given levels.
-%Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
-%
-%\begin{figure*}
-%	\begin{center}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
-%		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
-%		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend.
-%		Enlarge axis labels.
-%		Label panels.
-%		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
-% 		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
-%		}
-%		\label{fig:per-bin-scatter}
-%	\end{center}
-%\end{figure*}
-%
-%\que{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
-%
-%We then make a point estimate of \nz\ using \chippr's \mmle\ optimization option as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
-%Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
-%The excessive breadth of the alternative estimators can be seen quite plainly.
-%
-%\begin{figure*}
-%	\begin{center}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
-%		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
-%		\caption{
-%			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
-%			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
-%			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend.
-%			Label figure/panels.
-%			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.
-%			Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
-%		}
-%		\label{fig:per-bin-ests}
-%	\end{center}
-%\end{figure*}
-%
-%We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
-%Though there are also slight differences in the angle of the error ellipses, the most striking effect is the broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
-%The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
-%Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
-%Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
-%\que{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
-%Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
-%Do you think it's sufficient to provide the bias on the moments of \nz, since that's what everyone uses anyway?}
-%
-%\begin{figure*}
-%	\begin{center}
-%		\includegraphics[width=0.9\textwidth]{figures/chippr/final_plot.png}
-%		\caption{
-%			\que{Are the contours any easier to see now?}
-%			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
-%			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
-%			\aim{TODO: Fix formatting of axis labels.
-%			Standardize ticklabels.}
-%		}
-%		\label{fig:cornerplot}
-%	\end{center}
-%\end{figure*}
-%
-%\que{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
-%	Suggestions for how to better establish context would be appreciated.}
+\subsection{LSST Requirements}
+\label{sec:lsstdemo}
+
+It is of interest to explore the impact of incorrectly estimated \nz\ on the inference of cosmological parameters to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
+To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{figures/chippr/cosmolike_inputs.png}
+		\caption{
+			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
+		}
+		\label{fig:tomobins}
+	\end{center}
+\end{figure}
+
+We consider as ground truth a set of known \nz\ corresponding to each of four hypothetical samples of galaxies and the corresponding cosmological parameter covariance matrix.
+The \nz\ of each galaxy subsample emulates that anticipated of galaxies binned by a redshift point estimate, as is common in tomographic redshift analyses, though our experimental procedure is agnostic to how the samples are identified.
+The true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.
+
+First, we bin them down to a piecewise constant parameterization with a manageable $35$ hyperparameters for \chippr's sampling capabilities.
+Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
+The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
+We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd} at their given levels.
+Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
+		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
+		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend.
+		Enlarge axis labels.
+		Label panels.
+		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
+ 		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
+		}
+		\label{fig:per-bin-scatter}
+	\end{center}
+\end{figure*}
+
+\que{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
+
+We then make a point estimate of \nz\ using \chippr's \mmle\ optimization option as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
+Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
+The excessive breadth of the alternative estimators can be seen quite plainly.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
+		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
+		\caption{
+			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
+			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
+			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend.
+			Label figure/panels.
+			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.
+			Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
+		}
+		\label{fig:per-bin-ests}
+	\end{center}
+\end{figure*}
+
+We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
+Though there are also slight differences in the angle of the error ellipses, the most striking effect is the broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
+The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
+Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
+Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
+\que{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
+Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
+Do you think it's sufficient to provide the bias on the moments of \nz, since that's what everyone uses anyway?}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{figures/chippr/final_plot.png}
+		\caption{
+			\que{Are the contours any easier to see now?}
+			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
+			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
+			\aim{TODO: Fix formatting of axis labels.
+			Standardize ticklabels.}
+		}
+		\label{fig:cornerplot}
+	\end{center}
+\end{figure*}
+
+\que{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
+	Suggestions for how to better establish context would be appreciated.}
 
@@
 
+%\section{Impact on cosmological parameter constraints}
+%\label{app:cosmology}
+
 \bibliographystyle{apj}

### 2021-04-21 research/paper/draft.tex
updating paper for resubmission

@@
-%\RequirePackage[]{lineno}
-
 \documentclass[iop]{emulateapj}
@@
 \usepackage{hyperref}
-%\usepackage{graphicx}
-%\usepackage{lineno}
+\usepackage{lineno}
 \usepackage[percent]{overpic}
 \usepackage{float}
+\usepackage[normalem]{ulem}
 \usepackage{wrapfig}
 
-%\usetikzlibrary{shapes.geometric, arrows}
-%\usetikzlibrary{fit}
-%
-%\tikzstyle{hyper} = [circle, text centered, draw=black]
-%\tikzstyle{param} = [circle, text centered, draw=black]
-%\tikzstyle{data} = [circle, text centered, draw=black, line width=2pt]
-%\tikzstyle{arrow} = [thick,->,>=stealth]
-
-%\usepackage{soul}
 \usepackage[title]{appendix}
@@
 \newcommand{\que}[1]{\todo{Question}{cyan}{#1}}
+\newcommand{\new}[2]{{\color{red}\sout{#1}}\ {\color{blue}{#2}}}%{#2}%
 
@@
 \newcommand{\des}{\project{DES}}
-\newcommand{\Chippr}{\project{CHIPPR}}% maybe change this to algorithm/italics
+\newcommand{\Chippr}{\project{CHIPPR}}
 
@@
 \newcommand{\emcee}{\repo{emcee}}
-\newcommand{\github}{\href{https://github.com}{GitHub}}% maybe remove link and use code font
-\newcommand{\python}{\textit{Python}}% establish code font?
-
-\newcommand{\data}{\ensuremath{\vec{d}}}% could change to bold
+\newcommand{\github}{\href{https://github.com}{GitHub}}
+\newcommand{\python}{\textit{Python}}
+
+\newcommand{\data}{\ensuremath{\vec{d}}}
 \newcommand{\like}{\mathscr{L}}
-\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}% could change to Prob or Pr
+\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}} 
 \newcommand{\expect}[1]{\left<#1\right>}
 \newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
-\newcommand{\gvn}{\mid}% could use | or \vert
+\newcommand{\gvn}{\mid}
 \newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
@@
 \newcommand{\Pz}{Photo-$z$}
-\newcommand{\zpdf}{\pz\ PDF}% could change to posterior
-\newcommand{\Zpdf}{\Pz\ PDF}% could change to posterior
-\newcommand{\pzpdf}{\pz\ posterior PDF}% could change to implicit posterior
-\newcommand{\Pzpdf}{\Pz\ posterior PDF}% could change to implicit posterior
+\newcommand{\zpdf}{\pz\ PDF}
+\newcommand{\Zpdf}{\Pz\ PDF}
+\newcommand{\pzpdf}{\pz\ posterior PDF}
+\newcommand{\Pzpdf}{\Pz\ posterior PDF}
 \newcommand{\pzip}{\pz\ implicit posterior}
@@
 \newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
-\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}% could change to \vec
+\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
 \newcommand{\ndphi}{\bvec{\phi}}
-\newcommand{\mmle}{marginalized maximum likelihood estimate}% really marginalized maximum a posteriori estimate, may still change this
+\newcommand{\mmle}{marginalized maximum posterior estimate}
 
@@
 \author{David W. Hogg\altaffilmark{2,3,4,5}}
-\email{aimalz@nyu.edu}
-
-\altaffiltext{1}{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
+\email{aimalz@astro.ruhr-uni-bochum.de}
+
+\altaffiltext{1}{Ruhr-University Bochum, German Centre for Cosmological Lensing, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
 \altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
@@
 
-%\aim{TODO: Add test case panel labels to all figs of \chippr\ sims/outputs.  Take higher-z samples in mock \pzpdf\ plots relative to outlier populations.  Include $\bar{z}$ in results plots.}
+%\aim{TODO: Add test case panel labels to all figs of \chippr\ sims/outputs.
+Take higher-z samples in mock \pzpdf\ plots relative to outlier populations.
+Include $\bar{z}$ in results plots and cite DES/KiDS papers that use n(z) bias as primary metric.}
 
@@
 % what's wrong with photo-zs?
-However, \pz s are susceptible to inaccuracy and imprecision in the form of their inherent noisiness resulting from the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
+However, \pz s are susceptible to inaccuracy and imprecision in the form of their inherent noisiness resulting from the decreasing signal-to-noise ratio with increasing redshift, coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.
 Figure~\ref{fig:pedagogical_scatter} is an adaptation of the ubiquitous plots of \pz\ vs.
@@
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/jain05.png}
+		\includegraphics[width=0.4\textwidth]{jain15.png}
 		\caption{
@@
 Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
-For example, the Science Requirements Document \citep{mandelbaum_weak_2017} states \lsst's requirements for the main cosmological sample, reproduced in Table~\ref{tab:lsstsrd}.
+For example, the Science Requirements Document \citep{mandelbaum_weak_2017} states \lsst's requirements \new{for}{on \pz\ error tolerances for constraining cosmology with} the main cosmological sample, reproduced in Table~\ref{tab:lsstsrd}.
 
@@
 		\caption{\Pz\ requirements for \lsst\ cosmology\\
-				\citep{mandelbaum_weak_2017}.}
+				\citep{mandelbaum_weak_2017}\new{.}{}}
 		\begin{tabular}{ll}
@@
 This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating our understanding of precision, accuracy, and systematic error.
-However, denoting \zpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; \zpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
+However, denoting \zpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; 
+\zpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
 In terms of \Fig{fig:pedagogical_scatter}, \zpdf s are horizontal cuts, probabilities of redshift conditioned on a specific value of data, i.e.
@@
 The most concerning weakness of \pzpdf s, however, is their usage in the literature, which is at best inconsistent and at worst incorrect.
+
+{\color{white}{\Fig{fig:pgm}}}
 
@@
 Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the degree to which constraints on the cosmological parameters can advance is limited by the accuracy and precision to which \nz\ is known \citep{abruzzo_impact_2019}.
-% Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.
+%\aim{TODO: Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}
 
@@
 for a sample of $J$ galaxies $j$, or, equivalently, the redshift distribution function $\hat{N}(z) = J \hat{n}(z)$, by effectively averaging the \pzpdf s.
-This summation procedure has been used extensively in cosmological analyses with photometric galaxy samples \citep{mandelbaum_precision_2008, benjamin_cfhtlens_2013, kelly_weighing_2014}.
+\new{T}{Despite its problems \citep{malz_how_2021}, t}his summation procedure has been used extensively in cosmological analyses with photometric galaxy samples \citep{mandelbaum_precision_2008, benjamin_cfhtlens_2013, kelly_weighing_2014}.
 %\aim{TODO: Continue adding equations/citations for the use of \Nz\ in cosmology.}
@@
 Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
-This paper challenges the logically invalid yet pervasive analysis procedure of stacking \pzpdf s by presenting and validating a hierarchical Bayesian technique for the use of \pzpdf s\ in the inference of \nz, yielding a method applicable to arbitrary one-point statistics relevant to cosmology, large-scale structure, and galaxy evolution; future work will extend this methodology to higher-order statistics.
+This paper challenges the logically invalid yet pervasive analysis procedure of stacking \pzpdf s by presenting and validating a hierarchical Bayesian technique for the use of \pzpdf s\ in the inference of \nz, yielding a method applicable to arbitrary one-point statistics relevant to cosmology, large-scale structure, and galaxy evolution;
+future work will extend this methodology to higher-order statistics.
 We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
-Though others have approached the problem before \citep{leistedt_hierarchical_2016, leistedt_hierarchical_2018}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.
+Though others have approached the problem before \citep{leistedt_hierarchical_2016, leistedt_hierarchical_2019}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.
 
@@
 
-%\aim{TODO: check for consistent notation of \pzip s vs. \pzpdf s.}
-
 \subsection{Forward Model}
@@
 of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
-We believe that galaxy redshifts are indeed sampled, or drawn, from \nz, making it a probability density over redshift; this fact can also be confirmed by dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{hogg_data_2012}.
+We believe that galaxy redshifts are indeed sampled, or drawn, from \nz, making it a probability density over redshift; 
+this fact can also be confirmed by dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{hogg_data_2012}.
 
@@
 	\begin{center}
-		\includegraphics[height=0.25\textheight]{figures/chippr/pgm.png}
+		\includegraphics[height=0.25\textheight]{pgm.png}
 		\caption{The directed acyclic graph of the CHIPPR model, where circles indicate random variables and arrows indicate causal relationships.
@@
 To self-consistently propagate the uncertainty in the inference of redshift, however, it is more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ over all possible values of $\ndphi$ conditioned on all the observed data $\{\data_{j}\}$ available in a generic catalog.
-In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the \textit{implicit prior} and the \textit{prior probability density} (\textit{prior PDF}), elaborated upon below.
+In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the \new{\textit{implicit prior} and the \textit{prior probability density} (\textit{prior PDF})}{implicit prior and the (hyper)prior probability density function}, elaborated upon below.
 
@@
 This posterior is effectively a marginalization with respect to redshift at a given value of $\data = \data_{j}$ of the \textit{empirical frequency distribution} $\pr{z, \data \gvn \ndphi^{\dagger}}$, the joint probability density corresponding to the true redshift distribution parameterized by $\ndphi^{\dagger}$, which exists in nature but need not be known.
-%\que{Propagate new $\pr{z, \data \gvn \ndphi^{\dagger}}$ notation through appendix?}
 
@@
 
-%Posteriors differ from likelihoods by way of a prior distribution, so we cannot simply assume that the available data products are \pz\ posteriors $\pr{z \gvn \data_{j}}$.
-%Rather, we have a catalog of implicit-prior weighted \pz\ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$.
-%There must have been some interim prior probability distribution $p(z|\vec{\theta}^{*})$ defined in terms of the interim prior parameter values (hereafter the interim prior) $\vec{\theta}^{*}$ explicitly chosen or implicitly made to perform the calculation of the probabilistic photo-$z$s.
-%If it is implicit, it may not be representable in the parametrization we have chosen, and furthermore it may not be known at all; a method that produces interim photo-$z$ posteriors of this kind is not suitable for inference.
-%However, so long as the implicit prior is known, hierarchical inference is possible.
-
-The prior probability density $\pr{\ndphi}$ is a more familiar concept in astronomy; to progress, we will have to choose a prior probability density over all possible values of the hyperparameters $\ndphi$.
-This prior need not be excessively proscriptive; for example, it may be chosen to enforce smoothness at physically motivated scales in redshift without imposing any particular region as over- or under-dense.
+The \textit{prior probability density} $\pr{\ndphi}$ is a more familiar concept in astronomy; 
+to progress, we will have to choose a prior probability density over all possible values of the hyperparameters $\ndphi$.
+This prior need not be excessively proscriptive; 
+for example, it may be chosen to enforce smoothness at physically motivated scales in redshift without imposing any particular region as over- or under-dense.
 
@@
 \end{equation}
-which is the very heart of \Chippr, also given as \Eq{eqn:final}.
+which is the very heart of \Chippr, also given as \Eq{eqn:final} and in the form of interpretive dance \citep{malz_probabilistic_2019}.
 This in effect replaces the implicit prior with the sampled model hyperparameters, thereby converting the \pzip s into likelihoods in order to obtain unbiased posteriors.
@@
 
+%{\color{white}{\Fig{fig:flowchart}, \Fig{fig:mega_scatter}}}
 \section{Methods \& Data}
@@
 In \Sect{sec:exp}, we outline the implementation of the \chippr\ code.
-%In \Sect{sec:sheldon}, we introduce the alternative \nz\ estimators against which \Chippr\ is compared.
-%In \Sect{sec:diag}, we present the quantitative metrics by which the \nz\ estimators are compared.
 In \Sect{sec:mock}, we outline the procedure for emulating mock \pzip s.
@@
 
-We implement the \Chippr\ model in code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
-In \Sect{sec:mcmc}, we describe the  available \chippr\ library.
+We implement the \Chippr\ model in \new{}{a prototype} code in order to perform tests of its validity and to compare its performance to that of traditional alternatives.
+In \Sect{sec:mcmc}, we describe the publicly available \chippr\ library.
 In \Sect{sec:sheldon}, we introduce the alternative approaches evaluated for comparison with \Chippr.
 In \Sect{sec:diag}, we describe the diagnostic criteria by which we assess estimators of \nz.
-%In \Sect{app:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
 
@@
 
-\chippr\ is a \python\ 2 library\footnote{\url{https://github.com/aimalz/chippr}} that includes an implementation of the \Chippr\ model as well as an extensive suite of tools for comparing \Chippr\ to other approaches.
-
-Though there are plans for future expansion to more flexible parameterizations, the current version of \chippr\ uses a log-space piecewise constant parameterization
+\chippr\ is a \new{}{public} \python\ 2 library\footnote{\url{https://github.com/aimalz/chippr}} that includes an implementation of the \Chippr\ model as well as an extensive suite of tools for comparing \Chippr\ to other approaches.
+\new{}{For the sake of reproducibility, this paper presents the prototype-level code used for the experiments discussed herein.
+However, a modern, flexible implementation appropriate for at-scale application to future data sets is currently under development\footnote{A working prototype of a \python 3-based implementation built on the \texttt{qp} back-end presented in \citet{malz_approximating_2018} may be found at \url{https://github.com/LSSTDESC/qit}, which is anticipated to evolve in sophistication over time.}.}
+
+\new{Though there are plans for future expansion to more flexible parameterizations, the version of }{}\chippr\ uses a log-space piecewise constant parameterization
 \begin{equation}
@@
 At each iteration $i$ and for each walker, a proposal distribution $\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for acceptance to or rejection from the full log-posterior distribution.
-%Two threshold conditions are defined, one designating all previous samples to be ignored as as products of a burn-in phase and another indicating when a sufficient number of post-burn samples have been accepted.
-%In this case, the first threshold (described in \Sect{app:acorr}) is defined in terms of sub-runs of $10^{3}$ accepted samples, and the second is defined as an accumulation of $10^{4}$ samples.
-%Though previous versions used \texttt{HDF5} for the primary I/O format due to its efficiency for large quantities of data, it was abandoned in favor of \texttt{pickle} in the working release due to the instability of the \python\ implementation of the format on high-performance computing systems.
-
-The resulting output
-% is a set of ordered \texttt{pickle} files % enumerated by $\rho$ each containing the state information of $10^{3}$ accepted samples, with the last ten files including only samples taken after the completion of the burn-in phase
-% after each sub-run.
-%The state information
-includes $\frac{I_{0}}{s}$ accepted samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$, as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$, divided into separate files before and after the completion of the burn-in phase, as defined by the Gelman-Rubin statistic \citep{gelman_inference_1992}.
+
+%In \Sect{app:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
+
+The resulting output includes $\frac{I_{0}}{s}$ accepted samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$, as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$, divided into separate files before and after the completion of the burn-in phase, as defined by the Gelman-Rubin statistic \citep{gelman_inference_1992}.
+\new{}{As a consequence of the Central Limit Theorem applied to a Markov Chain, the sample distribution satisfies a statistical guarantee of normality \citep{brooks_handbook_2011}, meaning the output samples $\{\ndphi_{i}\}$ of \chippr\ naturally define a notion of standard deviation interpretable as conventional error bars\footnote{Additionally, the proofs are tractable by hand in the special cases of simplified \pzpdf s introduced in \citet{malz_how_2021}.}.}
 
@@
 Furthermore, derivatives will not in general be available for the full posterior distribution, restricting optimization methods used, and, as is true for any optimization code, there is a risk of numerical instability.
-%\begin{equation}
-%\label{eq:mmle}
-%\ln[p(\{\vec{d}_{j}\}|\vec{\theta})] \propto -\int\ f_{\vec{\theta}}(z)\
-%dz+\sum_{j=1}^{J}\ln\left[\int\
-%\exp\left[\ln[p(z_{j}|\vec{d}_{j},\vec{\theta}^{*})]+\ln[f_{\vec{\theta}}(z)]-\l
-%n[f_{\vec{\theta}^{*}}(z)]\right]\ dz\right],
-%\end{equation}
-%accessible with any optimization code.
 
@@
 Beyond visual inspection of samples, we calculate summary statistics to quantitatively compare different estimators' precision and accuracy.  
-Since MCMC samples of hyperparameters are Gaussian distributions, we can quantify the breadth of the distribution for each hyperparameter using the standard deviation regardless of whether the true values are known.
+Since MCMC samples of these hyperparameters are Gaussian distributions, we can quantify the breadth of the distribution for each hyperparameter using the standard deviation regardless of whether the true values are known.
 
@@
 A detailed exploration of the KLD may be found in the Appendix to \citet{malz_approximating_2018}.
-%We note that $KL_{\ndphi,\ndphi^{\ddagger}} \neq KL_{\ndphi^{\ddagger},\ndphi}$ and is only interpretable when there is a notion that $\ndphi^{\dagger}$ is closer to the truth than $\ndphi$.
-%In simulated tests, $\ndphi^{\dagger}$ is the true value and $\ndphi'$ is that produced by one of the methods in question.
-
-%\aim{TODO: Also include the mean redshift/$\Delta_{z}$ for each estimator of \nz, cite the DES/KiDS/HSC papers that motivate this.}
-
-\subsection{Validation on mock data}
-\label{sec:mock}
 
@@
 	\begin{center}
-		\includegraphics[width=0.7\textwidth]{figures/chippr/flowchart.pdf}
+		\includegraphics[width=0.7\textwidth]{flowchart.pdf}
 		\caption{
-			%			\que{Change ``true joint model'' to true forward model, condition on $\phi^{\dagger}$, add footnote explaining distinction between dagger, star, prime, etc., check that derivation matches notation}
 			A flow chart illustrating the forward model used to generate mock data in the validation of \Chippr, as described in \Sect{sec:forward}.
@@
 \end{figure*}
+
+\subsection{Validation on mock data}
+\label{sec:mock}
+
+\begin{figure}%[H]
+	\begin{center}
+		\includegraphics[width=0.4\textwidth]{data_lsst.png}
+		\caption{
+			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at the level of the \lsst\ requirements: 
+			intrinsic scatter, uniformly distributed catastrophic outliers, and bias.
+			The main panel shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{spec}} = z_{\mathrm{phot}}$ diagonal (gray line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
+			The insets show marginal histograms (light gray) in each dimension, that can be compared with the true \nz\ used to make the figure (black) to see the effect of these systematics, as well as the implicit prior (dark gray).
+		}
+		\label{fig:mega_scatter}
+	\end{center}
+\end{figure}
 
@@
 \end{equation}
-with $c_{z} = 0.3$, because it has been used in forecasting studies for \des\ and \lsst.
-%\aim{I learned this from talking to people and don't know of a published source that talks about the nitty gritty of the internal validation tests performed before there was data.}
+with $c_{z} = 0.3$, because it has been used in forecasting studies for \lsst \new{}{\citep{chang_effective_2013, the_lsst_dark_energy_science_collaboration_lsst_2018}}.
 
@@
 
-\begin{figure}%[H]
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst-mega_scatter.png}
-		\caption{
-			The joint probability space of true and estimated redshift for the three concerning \pz\ systematics at the level of the \lsst\ requirements:
-			intrinsic scatter, uniformly distributed catastrophic outliers, and bias.
-			The main panel shows samples (black points) in the space of mock data and redshift, akin to the standard scatterplots of true and estimated redshift, the $z_{\mathrm{spec}} = z_{\mathrm{phot}}$ diagonal (gray line), and posterior probabilities evaluated at the given estimated redshift (colored step functions).
-			The insets show marginal histograms (light gray) in each dimension, that can be compared with the true \nz\ used to make the figure (black) to see the effect of these systematics, as well as the implicit prior (dark gray).
-%			\aim{TODO: Include slice further up so we can see outliers.
-%				Label panel.}
-		}
-		\label{fig:mega_scatter}
-	\end{center}
-\end{figure}
-
 The hyperprior distribution chosen for these tests is a multivariate normal distribution with mean $\vec{\mu}$ equal to the implicit prior $\ndphi^{*}$ and covariance
@@
 
-%An example of such samples from the prior are shown in \Fig{fig:prior}.
-%\que{Should I add back the figure of prior samples?}
-
 The sampler is initialized with $W=100$ walkers each with a value chosen from a Gaussian distribution of identity covariance around a sample from the hyperprior distribution.
 
+%{\color{white}{\Fig{fig:pzs-scatter} \Fig{fig:results-scatter} \Fig{fig:uniform-outliers } \Fig{fig:nonuniform-outliers-data} \Fig{fig:nonuniform-outliers-results} \Fig{fig:bias}}}
 \section{Results}
@@
 
-\subsection{Intrinsic scatter}
-\label{sec:scatter}
-
-%Several factors contribute to photometric redshifts' intrinsic scatter.
-%Distant galaxies are dimmer compared to galaxies of identical luminosity that are closer, driving up photometric errors in flux-limited surveys.
-%The nature of the galaxy sample at higher redshifts also changes, meaning the generation of the photometric redshift posterior based on an a locally-calibrated SED template library or spectroscopically-confirmed training set is more likely to be inappropriate, leading to broader features.
-%In general, the galaxies that could not have been observed spectroscopically will have different and noisier photo-$z$ likelihoods than those that could fall into a spectroscopic training set (or spectroscopically derived template library).
-%This effect may be stronger for high-redshift galaxies.
-
-\Fig{fig:pzs-scatter} shows some examples of \pzpdf s generated with only the systematic of intrinsic scatter, at the level of the \lsst\ requirements on the left and twice that on the right.
-One can see that the histogram of redshift estimates is broader than that of true redshifts, and that the effect is substantially more pronounced by just doubling the intrinsic scatter from the level of the \lsst\ requirements.
-
 \begin{figure*}
 	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/single_varsigmas-mega_scatter.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_hivarsig-mega_scatter.png}
+		\includegraphics[width=0.4\textwidth]{data_scatter_lo.png}
+		\includegraphics[width=0.4\textwidth]{data_scatter_hi.png}
 		\caption{
@@
 			A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
-%		\aim{TODO: Label panels.
-%		Show the mean redshift for each estimator.}
 		}
@@
 \end{figure*}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.4\textwidth]{results_scatter_lo.png}
+		\includegraphics[width=0.4\textwidth]{results_scatter_hi.png}
+		\caption{
+			The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with intrinsic scatter of the \lsst\ requirements (left) and twice that (right), with the true redshift density (black curve) and implicit prior (gray curve).
+			\Chippr\ is robust to intrinsic scatter, but the alternatives suffer from overly broad \nz\ estimates that worsen with increasing intrinsic scatter.
+		}
+		\label{fig:results-scatter}
+	\end{center}
+\end{figure*}
+
+\subsection{Intrinsic scatter}
+\label{sec:scatter}
+
+\Fig{fig:pzs-scatter} shows some examples of \pzpdf s generated with only the systematic of intrinsic scatter, at the level of the \lsst\ requirements on the left and twice that on the right.
+One can see that the histogram of redshift estimates is broader than that of true redshifts, and that the effect is substantially more pronounced by just doubling the intrinsic scatter from the level of the \lsst\ requirements.
 
@@
 
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.4\textwidth]{data_outlier_uniform.png}\\
+		\includegraphics[width=0.4\textwidth]{results_outlier_uniform.png}
+		\caption{
+			Top: Examples of \pzpdf s with a uniformly distributed catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
+			Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
+			The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
+		}
+		\label{fig:uniform-outliers}
+	\end{center}
+\end{figure}
+
 \begin{figure*}
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_varsigmas_log_estimators.png}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_hivarsig_log_estimators.png}
+		\includegraphics[width=0.4\textwidth]{data_outlier_template.png}
+		\includegraphics[width=0.4\textwidth]{data_outlier_training.png}
 		\caption{
-		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzpdf s with intrinsic scatter of the \lsst\ requirements (left) and twice that (right), with the true redshift density (black curve) and implicit prior (gray curve).
-		\Chippr\ is robust to intrinsic scatter, but the alternatives suffer from overly broad \nz\ estimates that worsen with increasing intrinsic scatter.
-%		\aim{TODO: Label panels.
-%		Show the mean redshift for each estimator.}
-	}
-	\label{fig:results-scatter}
+			Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
+		}
+		\label{fig:nonuniform-outliers-data}
+	\end{center}
+\end{figure*}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.4\textwidth]{results_outlier_template.png}
+		\includegraphics[width=0.4\textwidth]{results_outlier_training.png}
+		\caption{
+			The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with catastrophic outliers like those seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right) to the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
+			Though the histogram of modes is most sensitive to a catastrophic outlier population, the stacked estimator also overestimates \nz\ under (machine learning-like outliers) and beyond (template fitting-like outliers).
+		}
+		\label{fig:nonuniform-outliers-results}
 	\end{center}
@@
 
-\begin{figure}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout-mega_scatter.png}\\
-	\includegraphics[width=0.45\textwidth]{figures/chippr/single_uout_log_estimators.png}
-	\caption{
-		Top: Examples of \pzpdf s with a uniformly distributed catastrophic outlier population at the level of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
-%		\aim{TODO: Include slice further up.
-%		Label panels.}
-		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with uniformly distributed catastrophic outliers, with the true redshift density (black curve) and implicit prior (gray curve).
-%		\aim{TODO: Label panels.
-%		Show the mean redshift for each estimator.}
-		The presence of the catastrophic outlier population broadens the histogram of modes and stacked estimator of the redshift distribution, but the result of \Chippr\ is unbiased.
-	}
-	\label{fig:uniform-outliers}
-	\end{center}
-\end{figure}
-
 \Fig{fig:uniform-outliers} shows that at the level of the \lsst\ requirements, the alternative estimators are overly broad, whereas \Chippr's \mmle\ yields an unbiased estimate of \nz.
@@
 The testing conditions here, illustrated in the right panel of \Fig{fig:nonuniform-outliers-data}, gives $10\%$ of galaxies at the redshift affected by outliers an observed redshift that is uniformly distributed relative to the true redshift, meaning that far fewer than $10\%$ of all galaxies in the sample are catastrophic outliers.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout-mega_scatter.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout-mega_scatter.png}
-	\caption{
-		Examples of \pzpdf s with a catastrophic outlier population like that seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right), including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
-%		\aim{TODO: Label panels.
-%		Include slices higher up.}
-	}
-	\label{fig:nonuniform-outliers-data}
-	\end{center}
-\end{figure*}
 
@@
 
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_eout_log_estimators.png}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_rout_log_estimators.png}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.4\textwidth]{data_bias.png}\\
+		\includegraphics[width=0.4\textwidth]{results_bias.png}
 		\caption{
-		The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with catastrophic outliers like those seen in template-fitting \pzpdf\ codes (left) and machine learning \pzpdf\ codes (right) to the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
-		Though the histogram of modes is most sensitive to a catastrophic outlier population, the stacked estimator also overestimates \nz\ under (machine learning-like outliers) and beyond (template fitting-like outliers).
-%		\aim{TODO: Label panels.
-%		Show the mean redshift for each estimator.}
-	}
-	\label{fig:nonuniform-outliers-results}
-	\end{center}
-\end{figure*}
+			Top: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
+			Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with ten times the bias of the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
+			The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
+		}
+		\label{fig:bias}
+	\end{center}
+\end{figure}
 
@@
 
-\begin{figure}
-	\begin{center}
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias-mega_scatter.png}\\
-	\includegraphics[width=0.45\textwidth]{figures/chippr/thesis_neghivarbias_log_estimators.png}
-	\caption{
-		Top: Examples of \pzpdf s with ten times the bias of the \lsst\ requirements, including samples from the probability space of true and observed redshift (black points), \pzpdf s (colored step functions), and the true redshifts of the example \pzpdf s (colored vertical lines), with marginal histograms (light gray) for each dimension with the true redshift distribution (black) and implicit prior (dark gray) in the insets.
-%				\aim{TODO: Include slice further up.
-%					Label panels.}
-		Bottom: The results of \Chippr\ (samples in light blue, optimization in dark blue) and the alternative approaches (the stacked estimator in red, the histogram of modes in yellow) on \pzpdf s with ten times the bias of the \lsst\ requirements, with the true redshift density (black curve) and implicit prior (gray curve).
-%				\aim{TODO: Label panel.
-%					Show the mean redshift for each estimator.}
-		The impact of bias at even ten times the level of the \lsst\ requirements is almost imperceptible on all estimators, though the \Chippr\ \mmle\ minimizes the information loss regardless.
-	}
-	\label{fig:bias}
-	\end{center}
-\end{figure}
-
 As expected based on self-consistency of the forward-modeled \pzpdf s, \Chippr\ is immune to linear bias of the form of $\Delta_{z}$.
@@
 
+%{\color{white}{\Fig{fig:pzs-priors} \Fig{fig:results-priors} \Fig{fig:mischaracterized}}}
 \section{Discussion}
@@
 Now, we stress-test \Chippr\ by exploring the impact of the implicit prior, which has thus far not received much attention in the literature.
-%two realistically complex cases, one in which the \nz\ estimates are made tomographically as in a modern cosmological analysis (\Sect{sec:lsstdemo}) and one
+%and propagate these \nz\ estimates to forecasted constraints on the cosmological parameters in a tomographic analysis (\Sect{sec:lsstdemo})
 \Sect{sec:interim} demonstrates the sensitivity of \nz\ estimation methods to realistically complex implicit priors, and \Sect{sec:violations} demonstrates the consequences of mischaracterization of the implicit prior used to generate the \pzip\ catalog.
@@
 
-%\que{Add back the results of the LSST requirements here?}
-
-\subsection{LSST Requirements}
-\label{sec:lsstdemo}
-
-It is of interest to explore the impact of incorrectly estimated \nz\ on the inference of cosmological parameters to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
-To test the impact of these uncertainties, we simulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators to a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/cosmolike_inputs.png}
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.4\textwidth]{data_prior_training.png}
+		\includegraphics[width=0.4\textwidth]{data_prior_template.png}
 		\caption{
-			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
-		}
-		\label{fig:tomobins}
-	\end{center}
-\end{figure}
-
-We consider as ground truth a set of known \nz\ corresponding to each of four hypothetical samples of galaxies and the corresponding cosmological parameter covariance matrix.
-The \nz\ of each galaxy subsample emulates that anticipated of galaxies binned by a redshift point estimate, as is common in tomographic redshift analyses, though our experimental procedure is agnostic to how the samples are identified.
-The true \nz\ in each pre-defined bin is already provided in the form of an evaluation of a function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
-The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.
-
-First, we bin them down to a piecewise constant parameterization with a manageable $35$ hyperparameters for \chippr's sampling capabilities.
-Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
-The original, binned, and drawn \nz\ are shown in \Fig{fig:tomobins}.
-We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd} at their given levels.
-Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
+			Examples of mock \pzip s generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzip s (colored step functions), the true redshifts of the example \pzip s (colored vertical lines).
+			A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
+		}
+		\label{fig:pzs-priors}
+	\end{center}
+\end{figure*}
 
@@
 	\begin{center}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_mega_scatter.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_mega_scatter.png}
-		\caption{As in \Fig{fig:mega_scatter}, with a different tomographic bin in each panel and the three effects of intrinsic scatter, uniformly distributed catastrophic outliers, and bias at the levels of the \lsst\ SRD, given in Table~\ref{tab:lsstsrd}.
-		\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated insets and legend.
-		Enlarge axis labels.
-		Label panels.
-		Also, add watermark of ``mock data'' in UL corner, same for ``results of inference'' on other kind of plot.
- 		Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
-		}
-		\label{fig:per-bin-scatter}
+		\includegraphics[width=0.4\textwidth]{results_prior_training.png}
+		\includegraphics[width=0.4\textwidth]{results_prior_template.png}
+		\caption
+		{The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzip s with an implicit prior like that of machine learning \pzip\ approaches (left) and an implicit prior like that of template-fitting \pzip\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
+			\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
+		}
+		\label{fig:results-priors}
 	\end{center}
 \end{figure*}
-
-\que{Is the distinction between binned samples determined by some observational property and probabilistic redshift distributions sufficiently clear?}
-
-We then make a point estimate of \nz\ using \chippr's \mmle\ optimization option as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
-Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
-The excessive breadth of the alternative estimators can be seen quite plainly.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/0single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/1single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/2single_lsst_lin_estimators.png}
-		\includegraphics[width=0.24\textwidth]{figures/chippr/3single_lsst_lin_estimators.png}
-		\caption{
-			The \chippr-derived and other estimators of \nz\ in each tomographic bin, with the true \nz\ (black), the implicit prior (gray), stacked estimator (red), histogram of modes (yellow), and \Chippr\ \mmle\ (blue).
-			The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes.
-			\aim{TODO: Make this one big plot instead of four little ones to eliminate repeated axis labels and legend.
-			Label figure/panels.
-			Also, add watermark of ``results of inference'' in UL corner, same for ``mock data'' on other kind of plot.
-			Show the mean redshift for each estimator of N(z), cite the DES papers that motivate this.}
-		}
-		\label{fig:per-bin-ests}
-	\end{center}
-\end{figure*}
-
-We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$.
-Though there are also slight differences in the angle of the error ellipses, the most striking effect is the broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
-The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
-Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
-Though the true values of the parameters themselves were not accessible with the Fisher matrix-based framework, we calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, showing that \Chippr\ preserves information $200-800$ times better than the alternatives, with the histogram of modes doing about four times better than stacking.
-\que{So a Fisher matrix analysis inherently can't provide the bias, because that requires data.
-Because we have no data, only posteriors conditioned on hypothetical data, this isn't possible.
-Do you think it's sufficient to provide the bias on the moments of \nz, since that's what everyone uses anyway?}
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.9\textwidth]{figures/chippr/final_plot.png}
-		\caption{
-			\que{Are the contours any easier to see now?}
-			The result of propagating the estimators of \nz\ by stacking (red), the histogram of modes (yellow), \Chippr\ (blue), and the true \nz\ (black) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
-			For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas for most parameters, stacking and the histogram of modes yield broader contours.
-			\aim{TODO: Fix formatting of axis labels.
-			Standardize ticklabels.}
-		}
-		\label{fig:cornerplot}
-	\end{center}
-\end{figure*}
-
-\que{Does this discussion adequately quantify how good \Nz\ has to be and how wrong we'll be if we estimate it wrong?
-	Suggestions for how to better establish context would be appreciated.}
 
@@
 Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
-Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
-%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.
-%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.
-%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.
-Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
+Template fitting codes have an explicit prior input formed by redshifting a number of templates, leading to a highly nonuniform but physically-motivated interim prior.
+Machine learning approaches tend to be trained on one of more previously observed data sets that include only galaxies for which spectroscopy is accessible, typically biasing the implicit prior towards atypically bright and/or low redshift populations.
 Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.
-%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.
-%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.
-%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.
 
@@
 
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_trpr-mega_scatter.png}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_tmpr-mega_scatter.png}
-		\caption{
-			Examples of mock \pzip s generated with a machine learning-like implicit prior (left) and a template-fitting-like implicit prior (right), including samples from the probability space of true and observed redshift (black points), \pzip s (colored step functions), the true redshifts of the example \pzip s (colored vertical lines).
-			A histogram (light gray) of points in each dimension is shown in the respective inset, with the true redshift distribution (black) and implicit prior (dark gray).
-%			\aim{TODO: Label panels.
-%				Include slices higher up.}
-		}
-		\label{fig:pzs-priors}
-	\end{center}
-\end{figure*}
-
 \Fig{fig:results-priors} shows the performance of \Chippr\ and the traditional methods on \pzip s generated with nontrivial implicit priors.
@@
 
+The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzip\ catalogs.
+
 \begin{figure*}
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_trpr_log_estimators.png}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_tmpr_log_estimators.png}
-		\caption
-		{The results of \Chippr\ (samples in light blue and optimization in dark blue) and the alternative approaches (the stacked estimator in red and the histogram of modes in yellow) on \pzip s with an implicit prior like that of machine learning \pzip\ approaches (left) and an implicit prior like that of template-fitting \pzip\ codes (right), with the true redshift density (black curve) and implicit prior (gray curve).
-			\Chippr\ is robust to a nontrivial implicit prior, but the alternatives are biased toward the implicit prior.
-%			\aim{TODO: Label panels.
-%				Show the mean redshift for each estimator.}
-		}
-		\label{fig:results-priors}
-	\end{center}
-\end{figure*}
-
-The main implication of the response of \nz\ estimates to a nontrivial implicit prior is that the implicit prior must be accounted for when using \pzip\ catalogs.
-
-\subsection{Violations of the model}
-\label{sec:violations}
-
-In this test, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
-\Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
-If we naively used the \pzip\ catalog produced by a generic machine learning or template-fitting code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_trpr_wrong_log_estimators.png}
-		\includegraphics[width=0.45\textwidth]{figures/chippr/single_lsst_tmpr_wrong_log_estimators.png}
+		\includegraphics[width=0.4\textwidth]{results_wrongprior_training.png}
+		\includegraphics[width=0.4\textwidth]{results_wrongprior_template.png}
 		\caption
@@
 			Here, \Chippr\ has been provided with a uniform implicit prior rather than those used to produce the mock \pzip s, and its performance is notably worse than when it is provided an accurate implicit prior, as in Figure~\ref{fig:results-priors}.
-%			\aim{Label panels.
-%				Show the mean redshift for each estimator.}
 			When the incorrect implicit prior is provided to \chippr, even Bayesian inference cannot recover the true \nz.
@@
 \end{figure*}
+
+\subsection{Violations of the model}
+\label{sec:violations}
+
+In this test, the \pzip s are made to the \lsst\ requirements but the implicit prior used for the inference is not the same as the implicit prior used for generating the data.
+\Pzpdf\ codes do not generally provide their implicit prior, with the exception of some template-fitting techniques for which it is a known input.
+If we naively used the \pzip\ catalog produced by a generic machine learning or template-fitting code and assumed a flat implicit prior, we would observe the contents of \Fig{fig:mischaracterized}.
 
@@
 
-% removed ancient investigation on real data, would be hard to redo with the new \chippr\ code on this timescale given that I haven't even looked at the data format in 4 years
+\new{}{
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=0.45\textwidth]{combo_mega_scatter.png}
+			\caption{\new{}{
+					As in \Fig{fig:mega_scatter}, for each of four \desc-like tomographic bins (colors).
+					Main panel: $(z_{spec}, z_{phot})$ pairs frawn from a model with \lsst-like systematic errors.
+					Top panel: The \lsst-like tomographic \nz\ model (thin, opaque line), after a reduction in the number of parameters of \nz\ (dashed line), and histogram of $z_{spec}$ values drawn from that model (thick, semitransparent line).
+					Side panel: Histograms of $z_{spec}$ (thin, opaque line) and $z_{phot}$ (semitransparent bars).
+			}}
+			\label{fig:per-bin-scatter}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=0.4\textwidth]{binned_nz_recovered.png}
+			\caption{\new{}{
+					The \chippr-derived and other estimators of \nz\ in each tomographic bin (panels), with the true \nz\ (solid black), stacked estimator (thin dashed red), histogram of modes (dotted yellow), and \Chippr\ \mmle\ (thick dashed blue).
+					The result of stacking is far too broad for \lsst-like \pzpdf s, even moreso than the simplistic histogram of modes, while the \Chippr\ \mmle\ more closely follows the amplitude of the true \nz.
+			}}
+			\label{fig:per-bin-ests}
+		\end{center}
+	\end{figure}
+
+	\begin{figure*}
+		\begin{center}
+			\includegraphics[width=0.9\textwidth]{final_plot.png}
+			\caption{\new{}{
+					The result of propagating the estimators of \nz\ by stacking (red horizontal hatching), the histogram of modes (yellow vertical hatching), \Chippr\ (blue shaded), and the true \nz\ (black outline) of \Fig{fig:per-bin-ests} to a subset of cosmological parameters.
+					For all parameters considered, \Chippr\ yields contours no broader than those corresponding to the true \nz, whereas stacking and the histogram of modes yield broader contours.
+			}}
+			\label{fig:cornerplot}
+		\end{center}
+	\end{figure*}
+
+	\subsection{Propagation to cosmology}
+	\label{sec:lsstdemo}
+
+	It is of interest to explore the impact of incorrectly estimated \nz\ on derived constraints on the cosmological parameters to answer the question of how wrong we will be in our understanding of the universe if we incorrectly constrain \nz.
+	To test the impact of these uncertainties, we emulate mock data with all three effects with which \lsst\ is concerned at the levels of Table~\ref{tab:lsstsrd} and propagate the results of \Chippr\ and the other estimators of \nz\ through a Fisher matrix forecast using \cosmolike\ \citep{krause_cosmolike_2017}, a publicly available cosmological forecasting code.
+
+	%\begin{figure}
+	%	\begin{center}
+	%		\includegraphics[width=0.45\textwidth]{combo_mega_scatter.png}
+	%		\caption{\new{}{
+	%			The \lsst-like tomographic binning and true redshift distribution, where the truth (solid) is a PDF evaluated on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$, and the binned (dashed) and drawn (dotted) \nz\ are piecewise constant functions evaluated in $35$ evenly spaced bins, for four different tomographic bins (colors).
+	%		}}
+	%		\label{fig:tomobins}
+	%	\end{center}
+	%\end{figure}
+
+	We consider as ground truth a set of known \nz\ corresponding to each of four hypothetical samples of galaxies and the corresponding cosmological parameter covariance matrix.
+	The \nz\ of each galaxy subsample emulates that anticipated of galaxies binned by a redshift point estimate, as is common in tomographic redshift analyses, though our experimental procedure is agnostic to how the samples are identified.
+	The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.
+	\aim{TODO: put all relevant files on GitHub.}
+	The true \nz\ in each pre-defined bin is provided in the form of an evaluation of the function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
+
+	First, we bin down the true redshift distribution to a piecewise constant parameterization with a manageable $35$ hyperparameters for \chippr's sampling capabilities.
+	Next, we draw $10^{4}$ true redshifts from the binned true \nz\ for each tomographic bin.
+	The original, binned, and drawn \nz\ are shown in the top panel of \Fig{fig:per-bin-scatter}.
+	We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
+	at their given levels.
+	The drawn $(z_{spec}, z_{phot})$ pairs are shown in the main panel of \Fig{fig:per-bin-scatter}, and the side panel shows their marginal histograms.
+	%Illustrations of this process are provided in \Fig{fig:per-bin-scatter}.
+
+	We then make a point estimate of \nz\ using \chippr's \mmle\ optimization option as well as the alternative methods on the \pzpdf\ catalog for each tomographic bin, shown in \Fig{fig:per-bin-ests}, because \cosmolike\ produces cosmology constraints from a single \nz\ result, rather than samples from the full posterior probability density of possible \nz.
+	%Note that \Fig{fig:per-bin-ests} is shown in linear rather than log probability units, unlike all other plots in this paper, to better show the behavior at low probability.
+	The excessive breadth of the alternative estimators, especially in the regions of low probability at high redshift, can be seen quite plainly.
+
+	%\subsubsection{Impact on cosmology constraints}
+	%\label{sec:cosmo}
+
+	We then use the different estimators of \nz\ in a cosmological forecasting procedure with \cosmolike, constraining $\Omega_{m}$, $\Omega_{b}$, $w_{a}$, $w_{0}$, $n_{s}$, $S_{8}$, and $H_{0}$, shown in \Fig{fig:cornerplot}.
+	Though there are also slight differences in the angle of the error ellipses, the most striking effect is the broadening of the contours under the alternative estimators relative to \Chippr, which are almost indistinguishable from those derived by using the true redshift distribution in each bin.
+	The stacked estimator is significantly worse than the \Chippr\ \mmle\ for all parameters except $\Omega_{b}$ and $H_{0}$.
+	Stacking, however, outperforms the histogram of modes for all parameters except $\Omega_{m}$ and $S_{8}$, for which their constraints are quite similar.
+	Though the Fisher matrix approach cannot be used to evaluate the bias of point estimates of the cosmological parameters relative to their true values, we instead calculate the seven-dimensional KLD for the three \nz\ estimators relative to the constraints derived from the true \nz, finding that \Chippr\ preserves information $200-800$ times better than the alternatives, and even the histogram of modes outperforms stacking by a factor of four.
+
+}
 
@@
 \label{sec:con}
-
-%\que{TODO: Break up conclusion into subsections of \Sect{sec:alldata}/\Sect{sec:results}?}
+%\aim{TODO: Break up conclusion into subsections of \Sect{sec:alldata}/\Sect{sec:results}?}
 
 This study derives and demonstrates a mathematically consistent inference of a one-point statistic, the redshift density function \nz, based on an arbitrary catalog of \pzpdf s.  
-The fully Bayesian \Chippr\ model, based in the fundamental laws of probability, begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
-The \Chippr\ model is implemented in the publicly available \chippr\ code and validated on mock data.
+The fully Bayesian \Chippr\ model\new{, based in the fundamental laws of probability,}{} begins with a probabilistic graphical model corresponding to equations for the full posterior distribution over the parameters for \nz.
+The \Chippr\ model is implemented in the publicly available \chippr\ \new{}{prototype} code and validated on mock data.
 % at the level of \nz\ as well as in terms of constraining power on the cosmological parameters.
-
-%In the tests on simulated data performed here, the full posterior distribution over the hyperparameters defining $N(z)$ derived by this method is consistent with the true redshift distribution function, making the mean of sampled values an excellent point estimator of $N(z)$.
-%The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.
-%The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.
-%Tests were also performed on subsets of BOSS DR10 data with results consistent with those of simulations.
 
@@
 \Fig{fig:pedagogical_scatter}), we emulate the canonical \pz\ error statistics, intrinsic scatter (\Sect{sec:scatter}), catastrophic outliers (\Sect{sec:outliers}), and canonical bias (\Sect{sec:bias}) one at a time.
-Though these test cases may appear overly simplistic, they enable rigorous quantification of the relative performance of each \nz\ estimation techniques under the controlled conditions of each type of error in isolation, at levels equal to and beyond those of \lsst.
-%\aim{TODO: point out that Fig 1 is uglier than 4, 5, 7, 8, 10, 11; these tests are simplistic, reality is more complex, but because \Chippr is provably correct, it can handle complexity of the real thing if p(z) is correct and captures complexity}
+Though these test cases may appear overly simplistic, they enable rigorous quantification of the \new{relative}{} performance of each \new{}{of two established} \nz\ estimation techniques \new{}{relative to \Chippr} under the controlled conditions of each type of error in isolation, at levels equal to and beyond those of \lsst.
 
@@
 \begin{itemize}
-\item \Chippr\ outperforms traditional estimators of \nz\ under realistically complex conditions, even at pessimistic levels relative to future survey requirements on the traditional \pz\ error statistics, as demonstrated both by eye and according to KLD values corresponding to $10\%$ the information loss of alternative methods.
-%\aim{TODO: Refer to quantitative results on KLD of \Nz.}
+\item \Chippr\ outperforms traditional estimators of \nz\ under realistically complex \new{conditions}{systematics}, even at pessimistic levels relative to future survey requirements on the \new{traditional}{standard} \pz\ error statistics, \new{as demonstrated both by eye and according to KLD values corresponding to}{achieving $\sim$} $10\%$ the information loss of alternative methods.
 \item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
-\item The error bars on the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that may be assumed for the conventional point estimators.
+\item The error bars o\new{n}{f} the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that \new{may be}{are sometimes} assumed for the conventional point estimators.
+\item \new{}{When the \nz\ estimates for LSST-like data are propagated through a cosmological forecast, \Chippr's information loss in the $(\Omega_{m}, \Omega_{b}, w_{a}, w_{0}, n_{s}, S_{8}, H_{0})$ parameter space is 2-3 times lower than that corresponding to traditional \nz\ estimators.}
 \end{itemize}
-Not only is \Chippr\ the only mathematically correct approach to the problem, it also recovers the true values of the hyperparameters defining \nz\ better than popular alternatives, as measured by the loss of information in \nz.
-% \ and the size of error ellipses in the space of cosmological parameters.
-However, the mathematically valid approach to inference with probabilistic data products incurs nontrivial computational expense, motivating future work to optimize the implementation.
+Not only is \Chippr\ the only mathematically correct approach to the problem\new{}{ of obtaining \nz\ from an external catalog of \pzpdf s}, it also recovers the true values of \new{}{both} the hyperparameters defining \nz \new{}{and the cosmological parameters} better than popular alternatives, as measured by the \new{loss of information in \nz}{KLD}.
+However, the mathematically valid approach to inference with probabilistic data products incurs nontrivial computational expense, motivating \new{future}{ongoing} work to optimize the implementation.
 
@@
 a principled hierarchical inference such as \Chippr\ is the only way to recover \nz\ from \pzpdf s.
-\item %\Chippr, however, is sensitive to misspecification of the implicit prior;
-Neither \Chippr\ nor traditional alternatives can recover \nz\ in the presence of a misspecified implicit prior;
+\item Neither \Chippr\ nor traditional alternatives can recover \nz\ in the presence of a misspecified implicit prior;
 the implicit prior used to produce the \pzpdf\ catalog must be known and provided to \Chippr\ in order to recover the true \nz.
@@
 
-The situation appears more dire for data-driven techniques, whose training sets may not straightforwardly translate into an implicit prior.
+The situation is more dire for data-driven techniques, whose training sets may not straightforwardly translate into a recoverable implicit prior.
 For example, some training set galaxies may contribute to the \pzpdf s more than others, resulting in different effective weights when factoring into, say, a histogram of training set redshifts as the implicit prior.
@@
 
-%\aim{TODO: new paragraph for what can go wrong,
-%	call to community for what to do about it,
-%	what aspects of implicit prior are knowable and not knowable?
-%	if testable how so? outside the scope of paper, data producers be warned!
-%	a likelihood is better -- give us that if you can!
-%	focus on methods that acknowledge probabilistic structure of problem.}
-
-A thorough investigation of the degree to which the implicit prior can be meaningfully obtained is outside this paper but should be a priority for all consumers of \pzpdf s.
-As an alternative, however, we must point out that if likelihoods were available rather than posteriors, the trouble with the implicit prior would be avoided altogether.
-We thus encourage the community of those making \pzpdf s to consider developing such methods so that the resulting data products may be correctly used in scientific inference more generically.
-
-%\aim{TODO: Claim that implications for tomographic binning are severe, if that can be motivated by mean \Nz shifts.}
-% by Section~\ref{sec:lsstdemo}.}
-
-%\subsection*{Recommendations for future work}
-
-%The following conclusions and recommendations can be made with confidence:
-
-%\begin{enumerate}
-%
-%	%	\item The marginalized maximum likelihood estimator is an excellent estimator for strongly featured redshift distribution function with simple, clean photo-$z$ posteriors; stacking smooths features more than sampling and photo-$z$ point estimation.
-%	\item When the implicit prior is known to be a poor match to the data, only the results of \Chippr\ are satisfactory estimators of the redshift distribution function because they are the only methods that can account for the bias induced on the \pzpdf\ catalog by the method that produces it; this is the most compelling case for the sampler because of the ubiquity of inappropriate interim priors.
-%\end{enumerate}
+A thorough investigation of the degree to which the implicit prior can be meaningfully obtained is outside \new{}{the scope of} this paper but should be a priority for all consumers of \pzpdf s.
+Alternatively, the trouble with the implicit prior would be avoided altogether if likelihoods were produced rather than posteriors.
+We thus encourage the community of those making \pzpdf s to consider developing methods yielding likelihoods rather than posteriors so that the resulting data products may be correctly used in scientific inference more generically.
 
@@
 We discourage researchers from co-adding \pzpdf s or converting them into point estimates of redshift and instead recommend the use of Bayesian probability to guide the usage of \pzpdf s.  
-We emphasize to those who produce \pzpdf s from data that it is essential to release the implicit prior used in generating this data product in order for any valid inference to be conducted by consumers of this information.
+We emphasize to those who produce \pzpdf s from data that it is essential to \new{}{determine and} release the implicit prior used in generating this data product in order for any valid inference to be conducted by consumers of this information.
 Methodologies for obtaining \pzpdf s must therefore be designed such that there is a known implicit prior, i.e. one that is not implicit at all, so that likelihoods may be recovered.
 
-The technique herein developed is applicable with minimal modification to other one-point statistics of redshift to which we will apply this method in the future, such as the redshift-dependent luminosity function and weak lensing mean distance ratio.
+\new{}{While the \chippr\ prototype code is publicly available, follow-up work  is currently under way to build an at-scale implementation of the \Chippr\ algorithm that will be applied to current and future data sets.}
+The \new{}{\Chippr} technique herein developed is applicable with minimal modification to other one-point statistics of redshift to which we will apply this method in the future, such as the redshift-dependent luminosity function and weak lensing mean distance ratio.
 Future work will also include the extension of this fully probabilistic approach to higher-order statistics of redshift such as the two-point correlation function.
@@
 	The authors thank Phil Marshall for advice on relevant examples, Elisabeth Krause for assistance with the \cosmolike\ code, Mohammadjavad Vakili for statistical insights, Geoffrey Ryan for programming advice, and Boris Leistedt for other helpful comments in the development of \Chippr.
-%	The authors also acknowledge
-%	\aim{TODO: Circulate draft to GCCL; submit to ApJ; circulate to Dan Foreman-Mackey, Boris Leistedt, Kate Storey-Fisher; post to arXiv; circulate to Johann Cohen-Tanugi, Will Hartley, Alan Heavens, Mike Jarvis, Francois Lanusse, Ann Lee, Rachel Mandelbaum, Phil Marshall, Chris Morrison, Jeff Newman, Sam Schmidt, Anze Slosar, Josh Speagle, others for feedback.}
 	This work was completed with generous nutritional support from the Center for Computational Astrophysics.
-%	\aim{TODO: Thank thesis readers.}
 \end{acknowledgements}
@@
 
-%We begin by parametrizing $N(z)$ in terms of $\vec{\theta}$, comprising some set of hyperparameters that define the form $N(z)$ may take in whatever basis we choose.
-%We define a function $f_{\vec{\theta}}(z)=N(z)$ that transforms these hyperparameters into the redshift distribution function $N(z)$.
-%Because
-%\begin{equation}
-%\eqlabel{eq:definition}
-%N(z) \propto p(z \gvn \vec{\theta}),
-%\end{equation}
-%we may discontinue discussion of $N(z)$ in favor of the likelihood $p(z|\vec{\theta})$.
-
 We perform the derivation of \Eq{eqn:fullpost} using log-probabilities.  
@@
 
-We shall make two assumptions of independence in order to make the problem tractable; their limitations are be discussed below.
+We shall make two assumptions of independence in order to make the problem tractable; 
+their limitations are be discussed below.  
 First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn z_{j}}]$, as in 
@@
 \end{equation}
-%It is important to note that the integral $\integral{n(z)}{z} N(z)\ dz$ is not constrained to equal the variable defining the Poisson distribution but instead $J$ by \Eq{eq:definition}, which can be thought of as another parameter.
 The derivation differs when $J$ is not known, say, when we want to learn about a distribution in nature rather than a distribution specific to data in hand, but for a photometric galaxy catalog where the desired quantity is $n(z)$ for the galaxies entering a larger cosmology calculation, it is a fixed quantity.
@@
 \end{align}
-
-%\Eq{eq:posterior} contains two quantities that merit further discussion, the prior distribution $p(\vec{\theta})$ discussed further in \Sect{sec:exp} and the photo-$z$ log-likelihoods $\ln[p(\vec{d}_{j}|z_{j})]$ that have not been mentioned since \Eq{eq:marginalize}.
-%Though photo-$z$ log-likelihoods would be desirable for use in these equations, they are not generally the product of either empirical and data-driven methods for obtaining photo-$z$ probability distributions.
-%Though probabilistic photo-$z$s are typically reported as generic probability distributions $p(z_{j})$, the methods that produce them may be understood to always yield posteriors, probability distributions conditioned on the data we believe to be true.
-%If they were not based in this assumption, they would require a sum over an infinite space of possible datasets.
 
@@
 
-%
-%\begin{align}
-%\begin{split}
-%\eqlabel{eqn:fullpost}
-%\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
-%\end{split}
-%\end{align}
-
 %\section{Convergence Criteria}
 %\label{app:acorr}
-%
-%\que{Cut convergence criteria section?}
 %
@@
 
-%\section{Impact on cosmological parameter constraints}
-%\label{app:cosmology}
-
 \bibliographystyle{apj}
-\bibliography{draft}
-
-%\aim{TODO: find way to cite Dance Your Ph.D. video}
+\bibliography{ms}
 

### 2021-08-10 research/paper/draft.tex

@@
 
-The \emcee \citep{foreman-mackey_emcee_2013} implementation of ensemble sampling is used to sample the full log-posterior of \Eq{eqn:final}. 
+The \emcee \citep{foreman-mackey_emcee_2013} implementation of \new{}{MCMC} ensemble sampling is used to sample the full log-posterior of \Eq{eqn:final}. 
+\new{}{As a consequence of the Central Limit Theorem applied to a Markov Chain, the sample distribution satisfies a statistical guarantee of normality \citep{brooks_handbook_2011}, meaning the output samples $\{\ndphi_{i}\}$ of \chippr\ naturally define a notion of standard deviation interpretable as conventional error bars\footnote{Additionally, the proofs are tractable by hand in the special cases of simplified \pzpdf s introduced in \citet{malz_how_2021}.}.}
+
 \chippr\ accepts a configuration file of user-specified parameters, among them the number $W$ of walkers.
 At each iteration $i$ and for each walker, a proposal distribution $\hat{\ndphi}_{i}$ is drawn from the log-prior distribution and evaluated for acceptance to or rejection from the full log-posterior distribution.
-
 %In \Sect{app:acorr}, we outline how \chippr\ can be used to sample the full log-posterior distribution $\ln[\pr{\ndphi \gvn \{\data_{j}\}}]$.
-
-The resulting output includes $\frac{I_{0}}{s}$ accepted samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$, as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$, divided into separate files before and after the completion of the burn-in phase, as defined by the Gelman-Rubin statistic \citep{gelman_inference_1992}.
-\new{}{As a consequence of the Central Limit Theorem applied to a Markov Chain, the sample distribution satisfies a statistical guarantee of normality \citep{brooks_handbook_2011}, meaning the output samples $\{\ndphi_{i}\}$ of \chippr\ naturally define a notion of standard deviation interpretable as conventional error bars\footnote{Additionally, the proofs are tractable by hand in the special cases of simplified \pzpdf s introduced in \citet{malz_how_2021}.}.}
+The resulting output includes $\frac{I_{0}}{s}$ accepted samples $\ndphi_{i}$ for a pre-specified chain thinning factor $s$ and their full posterior probabilities $\pr{\ndphi_{i} \gvn \{\data_{j}\}}$, as well as the autocorrelation times and acceptance fractions calculated for each element of $\ndphi$, divided into separate files before and after the completion of the burn-in phase, as defined by the Gelman-Rubin statistic $R$ \citep{gelman_inference_1992}.
 
@@
 
-A final estimator of the hyperparameters is the maximum marginalized likelihood estimator (MMLE), the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.
-The MMLE can be obtained in substantially less time than enough samples to characterize the full log-posterior distribution of \nz.
-However, the MMLE yields only a point estimate of \nz\ rather than characterizing the full log-posterior on $\ndphi$, and it does not escape the dependence on the choice of hyperprior distribution.  
+A final estimator of the hyperparameters is the maximum marginalized likelihood estimator (\mmle), the value of $\ndphi$ maximizing the log posterior given by \Eq{eqn:final} using any optimization code.
+The \mmle can be obtained in substantially less time than enough samples to characterize the full log-posterior distribution of \nz.
+However, the \mmle yields only a point estimate of \nz\ rather than characterizing the full log-posterior on $\ndphi$, and it does not escape the dependence on the choice of hyperprior distribution.  
 Furthermore, derivatives will not in general be available for the full posterior distribution, restricting optimization methods used, and, as is true for any optimization code, there is a risk of numerical instability.
@@
 \label{eqn:kl}
-KL_{\ndphi,\ndphi^{\ddagger}} = \integral{\pr{z \gvn \ndphi} \ln \left[ \frac{\pr{z \gvn \ndphi}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
+KL\new{}{D}_{\ndphi,\ndphi^{\ddagger}} = \integral{\pr{z \gvn \ndphi} \ln \left[ \frac{\pr{z \gvn \ndphi}}{\pr{z \gvn \ndphi^{\dagger}}} \right]}{z} ,
 \end{equation}
@@
 \label{eqn:priorcov}
-\Sigma_{k,k'} = q\ \exp[-\frac{e}{2}\ (\bar{z}_{k}-\bar{z}_{k'})^{2}]\ +\ t\delta(k,k')
-\end{equation}
-inspired by one used in Gaussian processes, where $k$ and $k'$ are indices ranging from $1$ to $K$ and $q=1.0$, $e=100.0$, and $t=q\cdot10^{-5}$ are constants chosen to permit draws from this prior distribution to produce shapes similar to that of a true $\tilde{\ndphi}$.
+\Sigma_{k,k'} = q\ \exp\left[-\frac{\new{e}{\varepsilon}}{2}\ (\bar{z}_{k}-\bar{z}_{k'})^{2}\right]\ +\ t\delta(k,k')
+\end{equation}
+inspired by one used in Gaussian processes, where $k$ and $k'$ are indices ranging from $1$ to $K$ and $q = 1.0$, $\new{e}{\varepsilon} = 100.0$, and $t = q \cdot 10^{-5}$ are constants chosen to permit draws from this prior distribution to produce shapes similar to that of a true $\tilde{\ndphi}$.
 We adapt the full log-posterior of \Eq{eqn:final} to the chosen binning of redshift space.
 
-The sampler is initialized with $W=100$ walkers each with a value chosen from a Gaussian distribution of identity covariance around a sample from the hyperprior distribution.
+The sampler is initialized with $W = 100$ walkers each with a value chosen from a Gaussian distribution of identity covariance around a sample from the hyperprior distribution.  
+\new{}{The computational expense is dominated by the sampler's burn-in phase and driven predominantly by the number of parameters $K$;
+for the tests presented in this study, with $J = 10^{4}$ galaxies, $K = 35$ parameters, $W = 100$ walkers, a Gelman-Rubin convergence threshold of $R = 1.5$, a thinning factor of $s = 1$, and $\frac{I_{0}}{s} = 10^{3}$ accepted post-burn samples, the single-threaded computation time on a 2017-era laptop is on the order of a few hours.}
 
@@
 The experiments of \Sect{sec:alldata} quantify the influence on each estimator of \nz\ due to each of the canonical types of \pz\ error one at a time in isolation.
-Now, we stress-test \Chippr\ by exploring the impact of the implicit prior, which has thus far not received much attention in the literature.
-%and propagate these \nz\ estimates to forecasted constraints on the cosmological parameters in a tomographic analysis (\Sect{sec:lsstdemo})
+Now, we \new{}{forecast the influence of these \nz\ estimators on cosmological parameter constraints in a tomographic analysis and} stress-test \Chippr\ by exploring the impact of the implicit prior, which has thus far not received attention in the literature.
+%(\Sect{sec:lsstdemo})
 \Sect{sec:interim} demonstrates the sensitivity of \nz\ estimation methods to realistically complex implicit priors, and \Sect{sec:violations} demonstrates the consequences of mischaracterization of the implicit prior used to generate the \pzip\ catalog.
+\new{}{\Sect{sec:lsstdemo} presents the impact of \nz\ estimation methods on forecasted cosmological parameter constraints.}
 These results provide compelling motivation for the \pz\ community to prioritize the study of implicit priors of existing and developing \pzpdf\ techniques.
@@
 	The cosmological parameter covariance matrices are those used for \desc\ forecasting with the ground truth \nz\ in the same four bins.
-	\aim{TODO: put all relevant files on GitHub.}
+%	\aim{TODO: put all relevant files on GitHub.}
 	The true \nz\ in each pre-defined bin is provided in the form of an evaluation of the function on a fine grid of $350$ redshifts $0.0101 < z < 3.5001$.
@@
 	The original, binned, and drawn \nz\ are shown in the top panel of \Fig{fig:per-bin-scatter}.
-	We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd}
-	at their given levels.
+	We emulate \pzpdf s for the $10^{4}$ true redshifts drawn from the true \nz\ in each bin using the procedure of \Fig{fig:flowchart} with all three effects of Table~\ref{tab:lsstsrd} at their given levels.
 	The drawn $(z_{spec}, z_{phot})$ pairs are shown in the main panel of \Fig{fig:per-bin-scatter}, and the side panel shows their marginal histograms.
@@
 \item Both the \Chippr\ \mmle\ and the mean of \chippr\ samples are good point estimators of \nz, whereas the histogram of modes is very sensitive to outliers and the stacked estimator is always excessively broad.
-\item The error bars o\new{n}{f} the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that \new{may be}{are sometimes} assumed for the conventional point estimators.
+\item The error bars \new{on}{of} the posterior distribution over \nz\ hyperparameters are interpretable and arise naturally under \Chippr, unlike those that \new{may be}{are sometimes} assumed for the conventional point estimators.
 \item \new{}{When the \nz\ estimates for LSST-like data are propagated through a cosmological forecast, \Chippr's information loss in the $(\Omega_{m}, \Omega_{b}, w_{a}, w_{0}, n_{s}, S_{8}, H_{0})$ parameter space is 2-3 times lower than that corresponding to traditional \nz\ estimators.}
@@
 
-\section{Derivation}
-\label{app:math}
+\section{}%Derivation}
+%\label{app:math}
 
 We perform the derivation of \Eq{eqn:fullpost} using log-probabilities.  
-What we wish to estimate is then the full log-posterior probability distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ given the catalog of photometry $\{\data_{j}\}$.
+What we wish to estimate is \new{then}{} the full log-posterior probability distribution (hereafter the full log-posterior) of the hyperparameters $\ndphi$ given the catalog of photometry $\{\data_{j}\}$.
 
